
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)


library(ggplot2)
library(sf)
library(terra)

library(tmap)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = "C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//Scratch", progress=10)
```


##Introduction
In this sccript I do EDA to determine areas of interest for proof of concept study of ecological resilience thresholds.
For Brazil, I use Rosan et al., 2019 Woody Encroachment study as a basis to select for areas of interest (Fig 1)
Rosan et al., emailed to me the EVImax trend and statistical significance rasters to extract areas where there is woody encroachment in the Cerrado. These areas were defined to be those where EVImax slope>0.002 with p<0.1, which I process in the chunk below.

```{r}
evimax_slope<- rast("C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//Data//Rosanetal2019_WoodyEncroachment//trendEVImax.tif")
p<- rast("C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//Data//Rosanetal2019_WoodyEncroachment//pvalueEVImax.tif")

Sys.time();evimax_slope[evimax_slope<=0.002]<- NA
evimax_slope[evimax_slope>0]<-1
p[p>0.1]<- NA
p[p<1]<-1;Sys.time()

aoi<-p+evimax_slope
writeRaster(aoi, "C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//Data//Rosanetal2019_WoodyEncroachment//woodyencroachment.tif",overwrite=T)

```

I 