```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I apply the 20% native savannas mask at 1km from indices_processing > 5_nativesavanna_delineation.Rmd
to compile final results 
(1) Read all model selection results (annual and monthly) and 20% thresholded raster
(2) Prepare df of model selection results
(3) Make raster of above, apply the 20% threshold raster as a mask to above results rasters & write out results as raster

#Step 1- read all model selection results and 20% thresholded raster
```{r}
overall_file_path <- here("Outputs", "ModelSelection")

read_files_function <- function (index_name, annual_or_monthly_folder, file_name_with_rds){
  rds_file<- read_rds(paste0(overall_file_path,"/", index_name,"/", annual_or_monthly_folder, "/", file_name_with_rds))
}

#NDVI
Sys.time(); annual_ndvi_swin7 <- read_files_function ("NDVI", "annual", "modelselection_swindow_7.rds"); Sys.time()
Sys.time(); annual_ndvi_swin11 <- read_files_function ("NDVI","annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); annual_ndvi_swinperiodic <- read_files_function ("NDVI","annual", "modelselection_swindow_periodic.rds"); Sys.time()
Sys.time(); annual_ndvi_swin_nostl <- read_files_function ("NDVI","annual", "modelselection_value_int.rds"); Sys.time()

Sys.time(); monthly_ndvi_swin7 <- read_files_function ("NDVI", "monthly", "modelselection_swindow_7.rds"); Sys.time()
Sys.time(); monthly_ndvi_swin11 <- read_files_function ("NDVI","monthly", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_ndvi_swinperiodic <- read_files_function ("NDVI","monthly", "modelselection_swindow_periodic.rds"); Sys.time()
Sys.time(); monthly_ndvi_swin_nostl <- read_files_function ("NDVI","monthly", "modelselection_value_int.rds"); Sys.time()

#EVI
Sys.time(); annual_evi_swin7 <- read_files_function ("EVI", "annual", "modelselection_swindow_7.rds"); Sys.time()
Sys.time(); annual_evi_swin11 <- read_files_function ("EVI", "annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); annual_evi_swinperiodic <- read_files_function ("EVI", "annual", "modelselection_swindow_periodic.rds"); Sys.time()
Sys.time(); annual_evi_swin_nostl <- read_files_function ("EVI", "annual", "modelselection_value_int.rds"); Sys.time()

Sys.time(); monthly_evi_swin7 <- read_files_function ("EVI", "monthly", "modelselection_swindow_7.rds"); Sys.time()
Sys.time(); monthly_evi_swin11 <- read_files_function ("EVI", "monthly", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_evi_swinperiodic <- read_files_function ("EVI", "monthly", "modelselection_swindow_periodic.rds"); Sys.time()
Sys.time(); monthly_evi_swin_nostl <- read_files_function ("EVI", "monthly", "modelselection_value_int.rds"); Sys.time()

#anisoEVI
Sys.time(); annual_anisoevi_swin7 <- read_files_function ("anisoEVI", "annual", "modelselection_swindow_7.rds"); Sys.time()
Sys.time(); annual_anisoevi_swin11 <- read_files_function ("anisoEVI", "annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); annual_anisoevi_swinperiodic <- read_files_function ("anisoEVI", "annual", "modelselection_swindow_periodic.rds"); Sys.time()
Sys.time(); annual_anisoevi_swin_nostl <- read_files_function ("anisoEVI", "annual", "modelselection_value_int.rds"); Sys.time()

Sys.time(); monthly_anisoevi_swin7 <- read_files_function ("anisoEVI", "monthly", "modelselection_swindow_7.rds"); Sys.time()
Sys.time(); monthly_anisoevi_swin11 <- read_files_function ("anisoEVI", "monthly", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_anisoevi_swinperiodic <- read_files_function ("anisoEVI", "monthly", "modelselection_swindow_periodic.rds"); Sys.time()
Sys.time(); monthly_anisoevi_swin_nostl <- read_files_function ("anisoEVI", "monthly", "modelselection_value_int.rds"); Sys.time()

threshold20_1km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_20.tif"))

#annual anisoEVI has no x, y. Extracting this information from monthly anisoEVI
month_swin7 <- monthly_anisoevi_swin7 %>% select(c(cell, x, y))
annual_anisoevi_swin7 <- annual_anisoevi_swin7 %>% left_join(month_swin7)
month_swin11 <- monthly_anisoevi_swin11 %>% select(c(cell, x, y))
annual_anisoevi_swin11 <- annual_anisoevi_swin11 %>% left_join(month_swin11)
month_swinperiodic <- monthly_anisoevi_swinperiodic %>% select(c(cell, x, y))
annual_anisoevi_swinperiodic <- annual_anisoevi_swinperiodic %>% left_join(month_swinperiodic)
month_swinnostl <- monthly_anisoevi_swin_nostl %>% select(c(cell, x, y))
annual_anisoevi_swin_nostl <- annual_anisoevi_swin_nostl %>% left_join(month_swinnostl)

remove(month_swin11, month_swin7, month_swinnostl, month_swinperiodic)
```


#Step 2- make df of results
```{r}
results_df_function<- function (modelselection_df) {
  results_df <- modelselection_df %>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~5, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                )) %>% 
    dplyr::select(c(x,y,pixelvalue))
}

#NDVI
Sys.time(); annual_ndvi_swin7 <- results_df_function(annual_ndvi_swin7); Sys.time()
Sys.time(); annual_ndvi_swin11 <- results_df_function(annual_ndvi_swin11); Sys.time()
Sys.time(); annual_ndvi_swinperiodic <- results_df_function(annual_ndvi_swinperiodic); Sys.time()
Sys.time(); annual_ndvi_swin_nostl <- results_df_function(annual_ndvi_swin_nostl); Sys.time()

Sys.time(); monthly_ndvi_swin7 <- results_df_function(monthly_ndvi_swin7); Sys.time()
Sys.time(); monthly_ndvi_swin11 <- results_df_function(monthly_ndvi_swin11); Sys.time()
Sys.time(); monthly_ndvi_swinperiodic <- results_df_function(monthly_ndvi_swinperiodic); Sys.time()
Sys.time(); monthly_ndvi_swin_nostl <- results_df_function(monthly_ndvi_swin_nostl); Sys.time()

#EVI
Sys.time(); annual_evi_swin7 <- results_df_function(annual_evi_swin7); Sys.time()
Sys.time(); annual_evi_swin11 <- results_df_function(annual_evi_swin11); Sys.time()
Sys.time(); annual_evi_swinperiodic <- results_df_function(annual_evi_swinperiodic); Sys.time()
Sys.time(); annual_evi_swin_nostl <- results_df_function(annual_evi_swin_nostl); Sys.time()

Sys.time(); monthly_evi_swin7 <- results_df_function(monthly_evi_swin7); Sys.time()
Sys.time(); monthly_evi_swin11 <- results_df_function(monthly_evi_swin11); Sys.time()
Sys.time(); monthly_evi_swinperiodic <- results_df_function(monthly_evi_swinperiodic); Sys.time()
Sys.time(); monthly_evi_swin_nostl <- results_df_function(monthly_evi_swin_nostl); Sys.time()

#anisoEVI
Sys.time(); annual_anisoevi_swin7 <- results_df_function(annual_anisoevi_swin7); Sys.time()
Sys.time(); annual_anisoevi_swin11 <- results_df_function(annual_anisoevi_swin11); Sys.time()
Sys.time(); annual_anisoevi_swinperiodic <- results_df_function(annual_anisoevi_swinperiodic); Sys.time()
Sys.time(); annual_anisoevi_swin_nostl <- results_df_function(annual_anisoevi_swin_nostl); Sys.time()

Sys.time(); monthly_anisoevi_swin7 <- results_df_function(monthly_anisoevi_swin7); Sys.time()
Sys.time(); monthly_anisoevi_swin11 <- results_df_function(monthly_anisoevi_swin11); Sys.time()
Sys.time(); monthly_anisoevi_swinperiodic <- results_df_function(monthly_anisoevi_swinperiodic); Sys.time()
Sys.time(); monthly_anisoevi_swin_nostl <- results_df_function(monthly_anisoevi_swin_nostl); Sys.time()
```


#Step 3- make results rasters, apply mask and write out final results 
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()

results_raster_function<- function (results_df, annual_or_monthly_folder, outfile_name) {

  cellindex_vector<- terra::vect(results_df, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi , "pixelvalue", fun="max")
  masked_cellindex_raster<- terra::mask(cellindex_raster, threshold20_1km)
  
  writeRaster(masked_cellindex_raster, here("Outputs", "TrendsResults", "results_rasters", annual_or_monthly_folder, paste0(outfile_name, ".tif")))
}

#NDVI
Sys.time(); results_raster_function(annual_ndvi_swin7,"annual", "ndvi_annual_swin7"); Sys.time()
Sys.time(); results_raster_function(annual_ndvi_swin11, "annual", "ndvi_annual_swin11"); Sys.time()
Sys.time(); results_raster_function(annual_ndvi_swinperiodic, "annual", "ndvi_annual_swinperiodic"); Sys.time()
Sys.time(); results_raster_function(annual_ndvi_swin_nostl,"annual", "ndvi_annual_swinnostl"); Sys.time()

Sys.time(); results_raster_function(monthly_ndvi_swin7, "monthly", "ndvi_monthly_swin7"); Sys.time()
Sys.time(); results_raster_function(monthly_ndvi_swin11, "monthly", "ndvi_monthly_swin11"); Sys.time()
Sys.time(); results_raster_function(monthly_ndvi_swinperiodic, "monthly", "ndvi_monthly_swinperiodic"); Sys.time()
Sys.time(); results_raster_function(monthly_ndvi_swin_nostl, "monthly", "ndvi_monthly_swinnostl"); Sys.time()

#EVI
Sys.time(); results_raster_function(annual_evi_swin7, "annual", "evi_annual_swin7"); Sys.time()
Sys.time(); results_raster_function(annual_evi_swin11, "annual", "evi_annual_swin11"); Sys.time()
Sys.time(); results_raster_function(annual_evi_swinperiodic, "annual", "evi_annual_swinperiodic"); Sys.time()
Sys.time(); results_raster_function(annual_evi_swin_nostl, "annual", "evi_annual_swinnostl"); Sys.time()

Sys.time(); results_raster_function(monthly_evi_swin7,"monthly", "evi_monthly_swin7"); Sys.time()
Sys.time(); results_raster_function(monthly_evi_swin11, "monthly", "evi_monthly_swin11"); Sys.time()
Sys.time(); results_raster_function(monthly_evi_swinperiodic, "monthly", "evi_monthly_swinperiodic"); Sys.time()
Sys.time(); results_raster_function(monthly_evi_swin_nostl, "monthly", "evi_monthly_swinnostl"); Sys.time()

#anisoEVI
Sys.time(); results_raster_function(annual_anisoevi_swin7, "annual", "anisoevi_annual_swin7"); Sys.time()
Sys.time(); results_raster_function(annual_anisoevi_swin11, "annual", "anisoevi_annual_swin11"); Sys.time()
Sys.time(); results_raster_function(annual_anisoevi_swinperiodic, "annual", "anisoevi_annual_swinperiodic"); Sys.time()
Sys.time(); results_raster_function(annual_anisoevi_swin_nostl, "annual", "anisoevi_annual_swinnostl"); Sys.time()

Sys.time(); results_raster_function(monthly_anisoevi_swin7,"monthly", "anisoevi_monthly_swin7"); Sys.time()
Sys.time(); results_raster_function(monthly_anisoevi_swin11, "monthly", "anisoevi_monthly_swin11"); Sys.time()
Sys.time(); results_raster_function(monthly_anisoevi_swinperiodic, "monthly", "anisoevi_monthly_swinperiodic"); Sys.time()
Sys.time(); results_raster_function(monthly_anisoevi_swin_nostl, "monthly", "anisoevi_monthly_swinnostl"); Sys.time()

```

