```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I compile results for the stl decomposition (sensitivity) for annual and monthly time step results. 
(1) Read in all results rasters for each index (4 tif each) for annual 
(2) Convert to df and compile
(3) Make barplot of results for stl parameters within annual or monthly category
(4) Make barplot of results for each stl parameter but comparing annual and monthly 

#Step 1- read in all results (monthly and annual timestep results) rasters
```{r}
input_file_path <- here("Outputs", "TrendsResults", "results_rasters")

read_results_rasters_function <- function (annual_or_monthly_folder, index_name){
   file_list<- list.files(path = paste0(input_file_path, "/",annual_or_monthly_folder, "/"), 
                          pattern = paste0("*",index_name,"*"), all.files = T, full.names = T)
   file_list <- gtools::mixedsort(file_list)
   
   results_raster<- lapply(file_list, rast)
   for (i in 1:length(results_raster)){
     names(results_raster[[i]])<- str_split(str_split(str_split(file_list[i], "/")[[1]][11], ".tif")[[1]][1], "_")[[1]][3]
   }
   results_raster
}

Sys.time(); annual_ndvi_results<- read_results_rasters_function ("annual","ndvi"); Sys.time()
Sys.time(); monthly_ndvi_results<- read_results_rasters_function ("monthly","ndvi"); Sys.time()
Sys.time(); annual_evi_results<- read_results_rasters_function ("annual","evi"); Sys.time()
Sys.time(); monthly_evi_results<- read_results_rasters_function ("monthly","evi"); Sys.time()
#Sys.time(); aniso_results<- read_results_rasters_function ("aniso"); Sys.time()

#rename rasters in list 


```

#Step 2- convert to df and compile 
```{r}
df_prep_function<- function (results_raster){
  x_df<- terra::as.data.frame(results_raster, cell = TRUE, xy= TRUE)
  x_df<- x_df %>% mutate(stl_setting = names(x_df)[4])
  names(x_df)[4]<- "value"
  percentage_df <- x_df %>% group_by(value, stl_setting) %>% summarise(count= n()) %>% mutate(percentage= (count/nrow(x_df))*100) 
  percentage_df
}

plot_df_prep_function <- function (results_raster_list){
  percentage_results_df_list<- list()
  for ( i in 1:length(results_raster_list)){
    percentage_results_df_list[[i]]<- df_prep_function(results_raster_list[[i]])
  }
  all_percentage_results <- bind_rows(percentage_results_df_list)
  all_percentage_results <- all_percentage_results %>% mutate(value = case_when(value== 1~ "Linear decrease",
                                                                                value==2~ "Linear increase",
                                                                                value==4~ "Step decrease",
                                                                                value==5~"Step increase",
                                                                                value==6~"Quadratic decrease (accelerated)",
                                                                                value==7~"Quadratic decrease (decelerated)",
                                                                                value==8~"Quadratic increase (accelerated)",
                                                                                value==9~"Quadratic increase (decelerated)",
                                                                                TRUE~"No trend"))
  all_percentage_results
}  


Sys.time(); annual_ndvi_percentage_results_df <- plot_df_prep_function (annual_ndvi_results); Sys.time()
Sys.time(); monthly_ndvi_percentage_results_df <- plot_df_prep_function (monthly_ndvi_results); Sys.time()
Sys.time(); annual_evi_percentage_results_df <- plot_df_prep_function (annual_evi_results); Sys.time()
Sys.time(); monthly_evi_percentage_results_df <- plot_df_prep_function (monthly_evi_results); Sys.time()
#Sys.time(); aniso_percentage_results_df <- plot_df_prep_function (aniso_results); Sys.time()

```

#Step 3- barplot comparing results of all parameters within annual and monthly
```{r}
Set1palette<- c("#44AA99", "#117733", "lightgrey", "#CC6677", "#AA4499", "#332288", "#DDCC77" , "#882255",  "#88CCEE")
 
output_file_path <- here("Outputs", "TrendsResults")
               
barplot_function <- function (compiled_df, file_name){
  x_plot<- compiled_df %>% 
  ggplot(aes(value, percentage, fill = value, group=stl_setting, alpha=as.factor(stl_setting))) +
  geom_col(position = position_dodge(), color = "black") +
  theme_classic() +
  xlab("Trajectory Shape")+ ylab ("% of native savannas")+
  theme(strip.placement  = "outside",
        panel.spacing    = unit(0, "points"),
        strip.background = element_blank(),
        strip.text       = element_text(size = 12, angle = 90),
        axis.text.x = element_text(size = 12, angle = 90))+
  xlab("Model Type")
  x_plot<- x_plot + scale_fill_manual(values=Set1palette) + theme(legend.title=element_blank())
  ggsave(here(output_file_path , paste0(file_name,".png")),
       x_plot,dpi = 700, height = 20, width=20, units = "cm")
}

Sys.time(); barplot_function(annual_ndvi_percentage_results_df, "ndvi_annual_stl_sensitivity"); Sys.time()
Sys.time(); barplot_function(monthly_ndvi_percentage_results_df, "ndvi_monthly_stl_sensitivity"); Sys.time()

Sys.time(); barplot_function(annual_evi_percentage_results_df, "evi_annual_stl_sensitivity"); Sys.time()
Sys.time(); barplot_function(monthly_evi_percentage_results_df, "evi_monthly_stl_sensitivity"); Sys.time()

#Sys.time(); barplot_function(aniso_percentage_results_df, "aniso_stl_sensitivity"); Sys.time()

```

#Step 4- barplot comparing results of annual and monthly timestep within each stl parameter 
```{r}
mutate_annual_or_monthly_col <- function (results_df, annual_or_monthly){
  results_df %>% mutate(timestep = annual_or_monthly)
}

annual_ndvi_percentage_results_df <- mutate_annual_or_monthly_col(annual_ndvi_percentage_results_df, "annual")
monthly_ndvi_percentage_results_df <- mutate_annual_or_monthly_col(monthly_ndvi_percentage_results_df, "monthly")

annual_evi_percentage_results_df <- mutate_annual_or_monthly_col(annual_evi_percentage_results_df, "annual")
monthly_evi_percentage_results_df <- mutate_annual_or_monthly_col(monthly_evi_percentage_results_df, "monthly")

ndvi_percentage_results_df<- rbind(annual_ndvi_percentage_results_df, monthly_ndvi_percentage_results_df)
evi_percentage_results_df<- rbind(annual_evi_percentage_results_df, monthly_evi_percentage_results_df)

stl_settings <- c("swin7", "swin11", "swinnostl", "swinperiodic")

barplot_function <- function (results_df, output_file_name){
   x_plot <- results_df %>% 
    ggplot(aes(value, percentage, fill = value, group= timestep, alpha=as.factor(timestep))) +
    geom_col(position = position_dodge(), color = "black") + 
     theme_linedraw(base_size = 18) +
    theme(axis.text.x=element_blank()) + 
    xlab("Trajectory Shape")+ ylab ("% of native savannas") +
    scale_fill_manual(values=Set1palette) + 
    theme(legend.title=element_blank()) +
    facet_grid(.~stl_setting, scales='free')
   
   ggsave(here(output_file_path, paste0(output_file_name,".png")),
       x_plot,dpi = 700, height = 50, width=50, units = "cm")
}
 
Sys.time(); barplot_function(ndvi_percentage_results_df, "ndvi_monthly_annual_comparison_trendresults"); Sys.time()
Sys.time(); barplot_function(evi_percentage_results_df, "evi_monthly_annual_comparison_trendresults"); Sys.time()

```
