```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)

library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I compile results for the stl decomposition (sensitivity) for annual and monthly time step results. 
(1) Read in all results rasters for each index (4 tif each) for annual 
(2) Convert to df and compile
(3) Make barplot of results for stl parameters within annual or monthly category
(4) Make barplot of results for each stl parameter but comparing annual and monthly 
(5) Make map of results for each stl parameter comparing annual and monthly

#Step 1- read in all results (monthly and annual timestep results) rasters
```{r}
input_file_path <- here("Outputs", "TrendsResults", "results_rasters")

read_results_rasters_function <- function (annual_or_monthly_folder, index_name){
   file_list<- list.files(path = paste0(input_file_path, "/",annual_or_monthly_folder, "/"), 
                          pattern = paste0("*",index_name,"*"), all.files = T, full.names = T)
   file_list <- gtools::mixedsort(file_list)
   
   results_raster<- lapply(file_list, rast)
   for (i in 1:length(results_raster)){
     names(results_raster[[i]])<- str_split(str_split(str_split(file_list[i], "/")[[1]][11], ".tif")[[1]][1], "_")[[1]][3]
   }
   results_raster
}

Sys.time(); annual_ndvi_results<- read_results_rasters_function ("annual","ndvi"); Sys.time()
Sys.time(); monthly_ndvi_results<- read_results_rasters_function ("monthly","ndvi"); Sys.time()
Sys.time(); annual_evi_results<- read_results_rasters_function ("annual","evi"); Sys.time()
Sys.time(); monthly_evi_results<- read_results_rasters_function ("monthly","evi"); Sys.time()
Sys.time(); annual_anisoevi_results<- read_results_rasters_function ("annual","aniso"); Sys.time()
Sys.time(); monthly_anisoevi_results<- read_results_rasters_function ("monthly","aniso"); Sys.time()

```

#Step 2- convert to df and compile 
```{r}
df_prep_function<- function (results_raster){
  x_df<- terra::as.data.frame(results_raster, cell = TRUE, xy= TRUE)
  x_df<- x_df %>% mutate(stl_setting = names(x_df)[4])
  names(x_df)[4]<- "value"
  percentage_df <- x_df %>% group_by(value, stl_setting) %>% summarise(count= n()) %>% mutate(percentage= (count/nrow(x_df))*100) 
  percentage_df
}

plot_df_prep_function <- function (results_raster_list){
  percentage_results_df_list<- list()
  for ( i in 1:length(results_raster_list)){
    percentage_results_df_list[[i]]<- df_prep_function(results_raster_list[[i]])
  }
  all_percentage_results <- bind_rows(percentage_results_df_list)
  all_percentage_results <- all_percentage_results %>% mutate(value = case_when(value== 1~ "Linear decrease",
                                                                                value==2~ "Linear increase",
                                                                                value==4~ "Step decrease",
                                                                                value==5~"Step increase",
                                                                                value==6~"Quadratic decrease (accelerated)",
                                                                                value==7~"Quadratic decrease (decelerated)",
                                                                                value==8~"Quadratic increase (accelerated)",
                                                                                value==9~"Quadratic increase (decelerated)",
                                                                                TRUE~"No trend"))
  all_percentage_results
}  


Sys.time(); annual_ndvi_percentage_results_df <- plot_df_prep_function (annual_ndvi_results); Sys.time()
Sys.time(); monthly_ndvi_percentage_results_df <- plot_df_prep_function (monthly_ndvi_results); Sys.time()
Sys.time(); annual_evi_percentage_results_df <- plot_df_prep_function (annual_evi_results); Sys.time()
Sys.time(); monthly_evi_percentage_results_df <- plot_df_prep_function (monthly_evi_results); Sys.time()
Sys.time(); annual_anisoevi_percentage_results_df <- plot_df_prep_function (annual_anisoevi_results); Sys.time()
Sys.time(); monthly_anisoevi_percentage_results_df <- plot_df_prep_function (monthly_anisoevi_results); Sys.time()

```

#Step 3- barplot comparing results of all parameters within annual and monthly
```{r}
Set1palette<- c("#44AA99", "#117733", "lightgrey", "#CC6677", "#AA4499", "#332288", "#DDCC77" , "#882255",  "#88CCEE")
 
output_file_path <- here("Outputs", "TrendsResults")
               
barplot_function <- function (compiled_df, file_name){
  x_plot<- compiled_df %>% 
  ggplot(aes(value, percentage, fill = value, group=stl_setting, alpha=as.factor(stl_setting))) +
  geom_col(position = position_dodge(), color = "black") +
  theme_classic() +
  xlab("Trajectory Shape")+ ylab ("% of native savannas")+
  theme(strip.placement  = "outside",
        panel.spacing    = unit(0, "points"),
        strip.background = element_blank(),
        strip.text       = element_text(size = 12, angle = 90),
        axis.text.x = element_text(size = 12, angle = 90))+
  xlab("Model Type")
  x_plot<- x_plot + scale_fill_manual(values=Set1palette) + theme(legend.title=element_blank())
  ggsave(here(output_file_path , paste0(file_name,".png")),
       x_plot,dpi = 700, height = 20, width=20, units = "cm")
}

Sys.time(); barplot_function(annual_ndvi_percentage_results_df, "ndvi_annual_stl_sensitivity"); Sys.time()
Sys.time(); barplot_function(monthly_ndvi_percentage_results_df, "ndvi_monthly_stl_sensitivity"); Sys.time()

Sys.time(); barplot_function(annual_evi_percentage_results_df, "evi_annual_stl_sensitivity"); Sys.time()
Sys.time(); barplot_function(monthly_evi_percentage_results_df, "evi_monthly_stl_sensitivity"); Sys.time()

Sys.time(); barplot_function(annual_anisoevi_percentage_results_df, "anisoevi_annual_stl_sensitivity"); Sys.time()
Sys.time(); barplot_function(monthly_anisoevi_percentage_results_df, "anisoevi_monthly_stl_sensitivity"); Sys.time()


```

#Step 4- barplot comparing results of annual and monthly timestep within each stl parameter 
```{r}
mutate_annual_or_monthly_col <- function (results_df, annual_or_monthly){
  results_df %>% mutate(timestep = annual_or_monthly)
}

annual_ndvi_percentage_results_df <- mutate_annual_or_monthly_col(annual_ndvi_percentage_results_df, "annual")
monthly_ndvi_percentage_results_df <- mutate_annual_or_monthly_col(monthly_ndvi_percentage_results_df, "monthly")

annual_evi_percentage_results_df <- mutate_annual_or_monthly_col(annual_evi_percentage_results_df, "annual")
monthly_evi_percentage_results_df <- mutate_annual_or_monthly_col(monthly_evi_percentage_results_df, "monthly")

annual_anisoevi_percentage_results_df <- mutate_annual_or_monthly_col(annual_anisoevi_percentage_results_df, "annual")
monthly_anisoevi_percentage_results_df <- mutate_annual_or_monthly_col(monthly_anisoevi_percentage_results_df, "monthly")

ndvi_percentage_results_df<- rbind(annual_ndvi_percentage_results_df, monthly_ndvi_percentage_results_df)
evi_percentage_results_df<- rbind(annual_evi_percentage_results_df, monthly_evi_percentage_results_df)
aniso_percentage_results_df<- rbind(annual_anisoevi_percentage_results_df, monthly_anisoevi_percentage_results_df)

stl_settings <- c("swin7", "swin11", "swinnostl", "swinperiodic")

barplot_function <- function (results_df, output_file_name){
   x_plot <- results_df %>% 
    ggplot(aes(value, percentage, fill = value, group= timestep, alpha=as.factor(timestep))) +
    geom_col(position = position_dodge(), color = "black") + 
     theme_linedraw(base_size = 18) +
    theme(axis.text.x=element_blank()) + 
    xlab("Trajectory Shape")+ ylab ("% of native savannas") +
    scale_fill_manual(values=Set1palette) + 
    theme(legend.title=element_blank()) +
    facet_grid(.~stl_setting, scales='free')
   
   ggsave(here(output_file_path, "annual_monthly_comparison", paste0(output_file_name,".png")),
       x_plot,dpi = 700, height = 40, width=80, units = "cm")
}
 
Sys.time(); barplot_function(ndvi_percentage_results_df, "ndvi_monthly_annual_comparison_trendresults"); Sys.time()
Sys.time(); barplot_function(evi_percentage_results_df, "evi_monthly_annual_comparison_trendresults"); Sys.time()
Sys.time(); barplot_function(aniso_percentage_results_df, "aniso_monthly_annual_comparison_trendresults"); Sys.time()

```

#Step 5- maps comparing trajectory results of annual and monthly for each stl parameter
```{r}
annual_ndvi_results
monthly_ndvi_results

annual_evi_results
monthly_evi_results

annual_anisoevi_results
monthly_anisoevi_results


d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

swin7_rasters_function <- function(annual_results_list, monthly_results_list){
  swin7_raster <- c(annual_results_list[[1]], monthly_results_list[[1]])
  names(swin7_raster)<- c("annual-swin7", "monthly-swin7")
  swin7_raster
}
swin11_rasters_function <- function(annual_results_list, monthly_results_list){
  swin11_raster <- c(annual_results_list[[2]], monthly_results_list[[2]])
  names(swin11_raster)<- c("annual-swin11", "monthly-swin11")
  swin11_raster
}
swinnostl_rasters_function <- function(annual_results_list, monthly_results_list){
  swinnostl_raster <- c(annual_results_list[[3]], monthly_results_list[[3]])
  names(swinnostl_raster)<- c("annual-nostl", "monthly-nostl")
  swinnostl_raster
}
swinperiodic_rasters_function <- function(annual_results_list, monthly_results_list){
  swinperiodic_raster <- c(annual_results_list[[4]], monthly_results_list[[4]])
  names(swinperiodic_raster)<- c("annual-swinperiodic", "monthly-swinperiodic")
  swinperiodic_raster
}

#NDVI
ndvi_swin7<- swin7_rasters_function(annual_ndvi_results, monthly_ndvi_results)
ndvi_swin11<- swin11_rasters_function(annual_ndvi_results, monthly_ndvi_results)
ndvi_swinnostl<- swinnostl_rasters_function(annual_ndvi_results, monthly_ndvi_results)
ndvi_swinperiodic<- swinperiodic_rasters_function(annual_ndvi_results, monthly_ndvi_results)

#EVI
evi_swin7<- swin7_rasters_function(annual_evi_results, monthly_evi_results)
evi_swin11<- swin11_rasters_function(annual_evi_results, monthly_evi_results)
evi_swinnostl<- swinnostl_rasters_function(annual_evi_results, monthly_evi_results)
evi_swinperiodic<- swinperiodic_rasters_function(annual_evi_results, monthly_evi_results)

#anisoEVI
aniso_swin7<- swin7_rasters_function(annual_anisoevi_results, monthly_anisoevi_results)
aniso_swin11<- swin11_rasters_function(annual_anisoevi_results, monthly_anisoevi_results)
aniso_swinnostl<- swinnostl_rasters_function(annual_anisoevi_results, monthly_anisoevi_results)
aniso_swinperiodic<- swinperiodic_rasters_function(annual_anisoevi_results, monthly_anisoevi_results)


animation_function<- function (raster,index_name, stlparameter_string){
  animation_map<-
    tm_shape(cerrado)+ tm_borders()+
    tm_shape (raster)+
    tm_raster(col.scale = tm_scale_categorical(n=9, values = c("#44AA99", "#117733", "lightgrey", "#882255", "#88CCEE", "#CC6677", "#AA4499", "#332288", "#DDCC77" )),
              col.legend = tm_legend_hide())+
    tm_add_legend(fill = c("#44AA99", "#117733", "lightgrey", "#882255", "#88CCEE", "#CC6677", "#AA4499", "#332288", "#DDCC77" ),
      labels = c("Linear decrease", 
                  "Linear increase",
                  "No trend",
                  "Step decrease",
                  "Step increase",
                  "Quad decrease (accelerated)",
                  "Quad decrease (decelerated)",
                  "Quad increase (accelerated)",
                  "Quad incrase (decelerated)"))+
    tm_facets(nrow=1, ncol=1, free.coords = TRUE) 
  
  tmap_animation(animation_map, filename = here("Outputs", "Trendsresults", "annual_monthly_comparison",
                                                  paste0(index_name,"_", "animation_annual_monthly","_", stlparameter_string, ".gif")),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
}

#NDVI
animation_function(ndvi_swin7,"ndvi", "swin7")
animation_function(ndvi_swin11,"ndvi", "swin11")
animation_function(ndvi_swinnostl,"ndvi", "swinnostl")
animation_function(ndvi_swinperiodic,"ndvi", "swinperiodic")

#EVI
animation_function(evi_swin7,"evi", "swin7")
animation_function(evi_swin11,"evi", "swin11")
animation_function(evi_swinnostl,"evi", "swinnostl")
animation_function(evi_swinperiodic,"evi", "swinperiodic")

#anisoEVI
animation_function(aniso_swin7,"anisoevi", "swin7")
animation_function(aniso_swin11,"anisoevi", "swin11")
animation_function(aniso_swinnostl,"anisoevi", "swinnostl")
animation_function(aniso_swinperiodic,"anisoevi", "swinperiodic")


```
