```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I compile final trend results for each index only for swindow = 11 and estimate "stable areas"
(1) Read in only swin11 rasters for the three indices
(2) Map plot
(3) Make barplot of results across the three indices
(4) Make "stable areas" rasters for all three indices

#Step 1- reading in only swin11 trend results in native savannas
```{r}
input_file_path <- here("Outputs", "TrendsResults")
file_list<- list.files(path = paste0(input_file_path, "/"), pattern = paste0("*","swin11","*"), all.files = T, full.names = T)
file_list<- file_list[2:3]
file_list <- gtools::mixedsort(file_list)

rds_list<- lapply(file_list, rast)
```

#Step 2- maps of swin11 trend results
```{r}
Set1palette<- c("#44AA99", "#117733", "lightgrey", "#CC6677", "#AA4499", "#332288", "#DDCC77" , "#882255",  "#88CCEE")

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mapping_function <- function (trendresults_raster, title){
  x_map<- tm_shape(cerrado)+ tm_borders()+
    tm_shape (trendresults_raster) + 
    tm_raster(col.scale = tm_scale_categorical(values = Set1palette), col.legend = tm_legend_hide())+
    tm_layout(legend.text.size = 1, title.size = 4) +
    tm_add_legend( labels = c("Linear decrease", 
                           "Linear increase",
                           "No trend",
                           "Step decrease",
                           "Step increase",
                           "Quad decrease (accelerated)",
                           "Quad decrease (decelerated)",
                           "Quad increase (accelerated)",
                           "Quad incrase (decelerated)"),
                fill = Set1palette,
                tm_legend(legend.text.size =10)) + 
    tm_title(title)
  x_map
}
Sys.time();ndvi_results_map <- mapping_function (rds_list[[2]], "NDVI"); Sys.time()
Sys.time();evi_results_map <- mapping_function (rds_list[[1]], "EVI"); Sys.time()
#Sys.time();aniso_results_map <- mapping_function (rds_list[[<<include>>]], "anisoEVI"); Sys.time()

trendresult_map<- tmap_arrange( evi_results_map, ndvi_results_map,
                               ncol =2, nrow= 1) #inlcude aniso_results_map when you have results
tmap_save(trendresult_map, paste0(input_file_path, "/", "map1km_trendresults.png"),
        height = 50, width = 50, units = "cm", dpi=700)

remove(aniso_results_map, evi_results_map, ndvi_results_map, trendresult_map)
```

#Step 3- barplot of trend results across three indices
```{r}
df_prep_function<- function (trendresults_raster, index_name){
  x_df<- terra::as.data.frame(trendresults_raster, cell = TRUE, xy= TRUE)
  x_df<- x_df %>% mutate(index = index_name)
  names(x_df)[4]<- "value"
  percentage_df <- x_df %>% group_by(value, index) %>% summarise(count= n()) %>% mutate(percentage= (count/nrow(x_df))*100) 
  percentage_df
}

Sys.time(); ndvi_trend_results <- df_prep_function(rds_list[[2]], "NDVI"); Sys.time()
Sys.time(); evi_trend_results <- df_prep_function(rds_list[[1]], "EVI"); Sys.time()
#Sys.time(); aniso_trend_results <- df_prep_function(rds_list[[<<include>>]], "anisoEVI"); Sys.time()

plot_df <- bind_rows(ndvi_trend_results, evi_trend_results)#inlcude aniso_results_map when you have results

plot_df <- plot_df %>% mutate(value = case_when(value== 1~ "Linear decrease",
                                                value==2~ "Linear increase",
                                                value==4~ "Step decrease",
                                                value==5~"Step increase",
                                                value==6~"Quadratic decrease (accelerated)",
                                                value==7~"Quadratic decrease (decelerated)",
                                                value==8~"Quadratic increase (accelerated)",
                                                value==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))

trendresults_barplot<- plot_df %>% 
  ggplot(aes(value, percentage, fill = value, group=index, alpha=as.factor(index))) +
  geom_col(position = position_dodge(), color = "black") +
  theme_classic() +
  xlab("Trajectory Shape")+ ylab ("% of native savannas")+
  theme(strip.placement  = "outside",
        panel.spacing    = unit(0, "points"),
        strip.background = element_blank(),
        strip.text       = element_text(size = 12, angle = 90),
        axis.text.x = element_text(size = 12, angle = 90))+
  scale_fill_manual(values=Set1palette) + theme(legend.title=element_blank())+
  xlab("Model Type")
ggsave(here(input_file_path, paste0("barplotswin11_trendresults.png")),
       trendresults_barplot,dpi = 700, height = 50, width=50, units = "cm")

remove(plot_df, trendresults_barplot)
remove(ndvi_trend_results, evi_trend_results)
```

#Step 4- stable area estimation for all three indices
(1) consider trajectory results 1km results such that no trend trajectory and remainder 
trajectories are two different pixel values, say 1 & 2 (before applying anthropogenic mask)
(2) aggregate to 5km by considering the "mean" value of the 1km pixels of
the no trend trajectory within the larger 5km pixel. 5km raster crs, extent and alignment
is from mean annual temperature processing in drivers_processing.
(3) apply anthropogenic mask at 5km resolution
```{r}
#Reading 1km resolution results (before anthropic mask is applied)
overall_file_path <- here("Outputs", "ModelSelection")

read_files_function <- function (index_name, file_name_with_rds){
  rds_file<- read_rds(paste0(overall_file_path,"/", index_name,"/", file_name_with_rds))
}
Sys.time(); ndvi_swin11 <- read_files_function ("NDVI", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); evi_swin11 <- read_files_function ("EVI", "modelselection_swindow_11.rds"); Sys.time()

#results preparation such that no trend cells are classified to be 1 and remainder are 2
results_df_function<- function (modelselection_df) {
  results_df <- modelselection_df %>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~2,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~1,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~1,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~1,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~1,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~2,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~2, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~2,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~2,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~2,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~2
                                )) %>% 
    dplyr::select(c(x,y,pixelvalue))
}

Sys.time(); ndvi_swin11 <- results_df_function(ndvi_swin11); Sys.time()
Sys.time(); evi_swin11 <- results_df_function(evi_swin11); Sys.time()

#raster preparation
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()

results_raster_function<- function (results_df) {
  cellindex_vector<- terra::vect(results_df, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi , "pixelvalue", fun="max")
  cellindex_raster
}

Sys.time(); ndvi1km_stable<- results_raster_function(ndvi_swin11); Sys.time()
Sys.time(); evi1km_stable<- results_raster_function(evi_swin11); Sys.time()

#coarsen to 5km & apply anthropogenic mask at 5km
threshold20_5km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "5km_thresholded_mapbiomas_mask_20.tif"))
mat<- rast(here("Outputs", "OtherVariables", "Climate", "mat_1981_2021.tif")) #from GEE

stable_area_function <- function (stablearea_1km_raster){
  x_5km<- terra::project(
    terra::segregate(stablearea_1km_raster), mat,
    method = "average", res = res(mat)[1])
 
  x_5km_mask <- terra::mask(x_5km, threshold20_5km)
  x_5km_mask
}
  
Sys.time(); ndvi_5km_stablearea <- stable_area_function(ndvi1km_stable); Sys.time()
writeRaster(ndvi_5km_stablearea[[1]], here("Outputs", "TrendsResults", "ndvi_stablearea.tif"))
Sys.time(); evi_5km_stablearea <- stable_area_function(evi1km_stable); Sys.time()
writeRaster(evi_5km_stablearea [[1]], here("Outputs", "TrendsResults", "evi_stablearea.tif"))
#aniso_stability <<include>>

map_stable_area_function<- function(stablearea_raster, title){
  tm_shape(cerrado)+ tm_borders()+
    tm_shape(stablearea_raster)+
    tm_raster(col.scale = tm_scale_continuous(values = "brewer.greens")) +
     tm_title(title)+
    tm_layout(legend.text.size = 1, title.size = 4)
}

map_ndvi_stability <- map_stable_area_function(ndvi_5km_stablearea [[1]], "NDVI")
map_evi_stability <- map_stable_area_function(evi_5km_stablearea[[1]], "EVI")
#map_aniso_stability <- stable_area_function(<<include>>, "anisoEVI")

stablearea_results_map<- tmap_arrange(map_ndvi_stability, map_evi_stability,
                               ncol =2, nrow= 1)#include map_aniso_stability when you have results
tmap_save(stablearea_results_map, paste0(input_file_path, "/", "map5km_stablearearesults.png"),
        height = 50, width = 50, units = "cm", dpi=700)


```