```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I compile final trend results for each index only for swindow = 11 for monthly timestep and estimate "stable areas"
(1) Read in only swin11 monthly rasters for the three indices
(2) Map plot
(3) Make barplot of results across the three indices
(4) Make "stable areas" rasters for all three indices at 5km resolution
(5) Make binary rasters of respective trajectory/not that trajectory at 1km resolution

#Step 1- reading in only swin11 trend results (monthly and annual) in native savannas
```{r}
monthly_input_file_path <- here("Outputs", "TrendsResults", "results_rasters", "monthly")
monthly_file_list<- list.files(path = paste0(monthly_input_file_path, "/"), pattern = paste0("*","_monthly_swin11","*"), all.files = T, full.names = T)
monthly_file_list <- gtools::mixedsort(monthly_file_list)

monthly_rds_list<- lapply(monthly_file_list, rast)
names(monthly_rds_list[[1]])<- "evi_swin11_monthly"
names(monthly_rds_list[[2]])<- "ndvi_swin11_monthly"

annual_input_file_path <- here("Outputs", "TrendsResults", "results_rasters", "annual")
annual_file_list<- list.files(path = paste0(annual_input_file_path, "/"), pattern = paste0("*","_annual_swin11","*"), all.files = T, full.names = T)
annual_file_list <- gtools::mixedsort(annual_file_list)

annual_rds_list<- lapply(annual_file_list, rast)
names(annual_rds_list[[1]])<- "evi_swin11_annual"
names(annual_rds_list[[2]])<- "ndvi_swin11_annual"

remove(annual_file_list, annual_input_file_path, monthly_file_list, monthly_input_file_path)

```

#Step 2- maps of swin11 trend results
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mapping_function <- function (trendresults_raster, title){
  x_map<- tm_shape(cerrado)+ tm_borders()+
    tm_shape (trendresults_raster) + 
    tm_raster(col.scale = tm_scale_categorical(n=9, values = c("#44AA99", "#117733", "lightgrey", "#882255", "#88CCEE", "#CC6677", "#AA4499", "#332288", "#DDCC77" )), col.legend = tm_legend_hide()) +
    tm_add_legend( fill = c("#44AA99", "#117733", "lightgrey", "#882255", "#88CCEE", "#CC6677", "#AA4499", "#332288", "#DDCC77" ),
      labels = c("Linear decrease", 
                           "Linear increase",
                           "No trend",
                           "Step decrease",
                           "Step increase",
                           "Quad decrease (accelerated)",
                           "Quad decrease (decelerated)",
                           "Quad increase (accelerated)",
                           "Quad incrase (decelerated)")) + 
    tm_layout(legend.text.size = 2) + tm_title(title)
  x_map
}
Sys.time();monthly_ndvi_results_map <- mapping_function (monthly_rds_list[[2]], "NDVI (monthly)"); Sys.time()
Sys.time();monthly_evi_results_map <- mapping_function (monthly_rds_list[[1]], "EVI (monthly)"); Sys.time()
Sys.time();annual_ndvi_results_map <- mapping_function (annual_rds_list[[2]], "NDVI (annual)"); Sys.time()
Sys.time();annual_evi_results_map <- mapping_function (annual_rds_list[[1]], "EVI (annual)"); Sys.time()

trendresult_map<- tmap_arrange(monthly_ndvi_results_map, annual_ndvi_results_map, monthly_evi_results_map, annual_evi_results_map,  
                               ncol =2, nrow= 2)
output_file_path <- here("Outputs", "TrendsResults")
tmap_save(trendresult_map, paste0(output_file_path, "/", "map1km_trendresults_swin11_monthlyannual.png"),
        height = 80, width = 80, units = "cm", dpi=700)

remove(monthly_evi_results_map, monthly_ndvi_results_map, annual_evi_results_map, annual_ndvi_results_map, trendresult_map)
```

#Step 3- barplot of trend results across three indices
```{r}
df_prep_function<- function (trendresults_raster, index_name){
  x_df<- terra::as.data.frame(trendresults_raster, cell = TRUE, xy= TRUE)
  x_df<- x_df %>% mutate(index = index_name)
  names(x_df)[4]<- "value"
  percentage_df <- x_df %>% group_by(value, index) %>% summarise(count= n()) %>% mutate(percentage= (count/nrow(x_df))*100) 
  percentage_df
}

Sys.time(); monthly_ndvi_trend_results <- df_prep_function(monthly_rds_list[[2]], "NDVI (monthly)"); Sys.time()
Sys.time(); monthly_evi_trend_results <- df_prep_function(monthly_rds_list[[1]], "EVI(monthly)"); Sys.time()
Sys.time(); annual_ndvi_trend_results <- df_prep_function(annual_rds_list[[2]], "NDVI (annual)"); Sys.time()
Sys.time(); annual_evi_trend_results <- df_prep_function(annual_rds_list[[1]], "EVI(annual)"); Sys.time()



plot_df <- bind_rows(monthly_ndvi_trend_results, monthly_evi_trend_results,
                     annual_ndvi_trend_results, annual_evi_trend_results)

plot_df <- plot_df %>% mutate(value = case_when(value== 1~ "Linear decrease",
                                                value==2~ "Linear increase",
                                                value==4~ "Step decrease",
                                                value==5~"Step increase",
                                                value==6~"Quadratic decrease (accelerated)",
                                                value==7~"Quadratic decrease (decelerated)",
                                                value==8~"Quadratic increase (accelerated)",
                                                value==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))

Set1palette<- c("#44AA99", "#117733", "lightgrey", "#CC6677", "#AA4499", "#332288", "#DDCC77" , "#882255",  "#88CCEE")

trendresults_barplot<- plot_df %>% 
  ggplot(aes(value, percentage, fill = value, group=index, alpha=as.factor(index))) +
  geom_col(position = position_dodge(), color = "black") +
  theme_classic() +
  xlab("Trajectory Shape")+ ylab ("% of native savannas")+
  theme(strip.placement  = "outside",
        panel.spacing    = unit(0, "points"),
        strip.background = element_blank(),
        strip.text       = element_text(size = 12, angle = 90),
        axis.text.x = element_text(size = 12, angle = 90)) +
  scale_fill_manual(values = Set1palette) + theme(legend.title=element_blank()) +
  xlab("Model Type")
ggsave(here(output_file_path, paste0("barplot_monthlyswin11_trendresults.png")),
       trendresults_barplot,dpi = 700, height = 20, width=20, units = "cm")

remove(plot_df, trendresults_barplot)
remove(monthly_ndvi_trend_results, monthly_evi_trend_results, annual_ndvi_trend_results, annual_evi_results_map)
```

#Step 4- stable area estimation for all three indices
(1) consider trajectory results 1km results such that no trend trajectory and remainder 
trajectories are two different pixel values, say 1 & 2 (before applying anthropogenic mask)
(2) aggregate to 5km by considering the "mean" value of the 1km pixels of
the no trend trajectory within the larger 5km pixel. 5km raster crs, extent and alignment
is from mean annual temperature processing in drivers_processing.
(3) apply anthropogenic mask at 5km resolution
```{r}
#Reading 1km resolution results (before anthropic mask is applied)
overall_file_path <- here("Outputs", "ModelSelection")

read_files_function <- function (index_name,annual_or_monthly_folder, file_name_with_rds){
  rds_file<- read_rds(paste0(overall_file_path,"/", index_name,"/", annual_or_monthly_folder,"/", file_name_with_rds))
}
Sys.time(); annual_ndvi_swin11 <- read_files_function ("NDVI","annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_ndvi_swin11 <- read_files_function ("NDVI","monthly", "modelselection_swindow_11.rds"); Sys.time()

Sys.time(); annual_evi_swin11 <- read_files_function ("EVI","annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_evi_swin11 <- read_files_function ("EVI", "monthly",  "modelselection_swindow_11.rds"); Sys.time()

#results preparation such that no trend cells are classified to be 1 and remainder are 2
results_df_function<- function (modelselection_df) {
  results_df <- modelselection_df %>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~2,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~1,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~1,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~1,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~1,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~2,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~2, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~2,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~2,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~2,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~2
                                )) %>% 
    dplyr::select(c(x,y,pixelvalue))
}

Sys.time(); annual_ndvi_swin11 <- results_df_function(annual_ndvi_swin11); Sys.time()
Sys.time(); monthly_ndvi_swin11 <- results_df_function(monthly_ndvi_swin11); Sys.time()

Sys.time(); annual_evi_swin11 <- results_df_function(annual_evi_swin11); Sys.time()
Sys.time(); monthly_evi_swin11 <- results_df_function(monthly_evi_swin11); Sys.time()

#raster preparation
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()

results_raster_function<- function (results_df) {
  cellindex_vector<- terra::vect(results_df, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi , "pixelvalue", fun="max")
  cellindex_raster
}

Sys.time(); annual_ndvi1km_stable<- results_raster_function(annual_ndvi_swin11); Sys.time()
Sys.time(); monthly_ndvi1km_stable<- results_raster_function(monthly_ndvi_swin11); Sys.time()

Sys.time(); annual_evi1km_stable<- results_raster_function(annual_evi_swin11); Sys.time()
Sys.time(); monthly_evi1km_stable<- results_raster_function(monthly_evi_swin11); Sys.time()

#coarsen to 5km & apply anthropogenic mask at 5km
threshold20_5km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "5km_thresholded_mapbiomas_mask_20.tif"))
mat<- rast(here("Outputs", "OtherVariables", "Climate", "mat_1981_2021.tif")) #from GEE

stable_area_function <- function (stablearea_1km_raster){
  x_5km<- terra::project(
    terra::segregate(stablearea_1km_raster), mat,
    method = "average", res = res(mat)[1])
 
  x_5km_mask <- terra::mask(x_5km, threshold20_5km)
  x_5km_mask
}
  
Sys.time(); annual_ndvi_5km_unstablearea <- stable_area_function(annual_ndvi1km_stable); Sys.time()
#writeRaster(annual_ndvi_5km_unstablearea [[2]], here("Outputs", "TrendsResults", "results_rasters" , "annual" ,"annualswin11_ndvi_unstablearea.tif"))
Sys.time(); monthly_ndvi_5km_unstablearea <- stable_area_function(monthly_ndvi1km_stable); Sys.time()
#writeRaster(monthly_ndvi_5km_unstablearea [[2]], here("Outputs", "TrendsResults", "results_rasters" , "monthly" , "monthlyswin11_ndvi_unstablearea.tif"))

Sys.time(); annual_evi_5km_unstablearea <- stable_area_function(annual_evi1km_stable); Sys.time()
#writeRaster(annual_evi_5km_unstablearea [[2]], here("Outputs", "TrendsResults", "results_rasters" , "annual", "annualswin11_evi_unstablearea.tif"))
Sys.time(); monthly_evi_5km_unstablearea <- stable_area_function(monthly_evi1km_stable); Sys.time()
#writeRaster(monthly_evi_5km_unstablearea [[2]], here("Outputs", "TrendsResults", "results_rasters" , "monthly" ,"monthlyswin11_evi_unstablearea.tif"))

#map making
annual_ndvi_5km_unstablearea <- rast(here("Outputs", "TrendsResults", "results_rasters" , "annual" ,"annualswin11_ndvi_unstablearea.tif"))
monthly_ndvi_5km_unstablearea <- rast(here("Outputs", "TrendsResults", "results_rasters" , "monthly" ,"monthlyswin11_ndvi_unstablearea.tif"))
ndvi_5km_unstablearea <- c(annual_ndvi_5km_unstablearea, monthly_ndvi_5km_unstablearea)
names(ndvi_5km_unstablearea) <- c("annual_swin11_unstablearea", "monthly_swin11_unstablearea")

annual_evi_5km_unstablearea <- rast(here("Outputs", "TrendsResults", "results_rasters" , "annual" ,"annualswin11_evi_unstablearea.tif"))
monthly_evi_5km_unstablearea <- rast(here("Outputs", "TrendsResults", "results_rasters" , "monthly" ,"monthlyswin11_evi_unstablearea.tif"))
evi_5km_unstablearea <- c(annual_evi_5km_unstablearea, monthly_evi_5km_unstablearea)
names(evi_5km_unstablearea) <- c("annual_swin11_unstablearea", "monthly_swin11_unstablearea")

animation_function<- function (raster,index_name){
  animation_map<-
    tm_shape(cerrado)+ tm_borders()+
    tm_shape (raster)+
    tm_raster(col.scale = tm_scale_continuous(values ="Greens")) +
    tm_facets(nrow=1, ncol=1, free.coords = TRUE) 
  
  tmap_animation(animation_map, filename = here("Outputs", "Trendsresults", "annual_monthly_comparison",
                                                  paste0(index_name,"_", "animation_annual_monthly","_", "unstablearea", ".gif")),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
}

animation_function(ndvi_5km_unstablearea, "ndvi")
animation_function(evi_5km_unstablearea, "evi")

static_map_function<- function (raster,index_name){
  s_map<-
    tm_shape(cerrado)+ tm_borders()+
    tm_shape (raster)+
    tm_raster(col.scale = tm_scale_continuous(values ="Greens")) +tm_title (index_name)
  
  tmap_save(s_map, filename = here("Outputs", "Trendsresults",paste0(index_name,"_","_", "unstablearea", ".png")),
                width = 900, height = 900, dpi = 150)
}

static_map_function(annual_evi_5km_unstablearea, "EVI_annual")
static_map_function(monthly_evi_5km_unstablearea, "EVI_monthly")
static_map_function(annual_ndvi_5km_unstablearea, "NDVI_annual")
static_map_function(monthly_ndvi_5km_unstablearea, "NDVI_monthly")

```

#Step 5- binary trajectory rasters of respective trajectory/not that trajectory at 1km resolution
for further analyses (regression analyses)
```{r}
#Read code chunk 1 for all data
output_file_path <- output_file_path <- here("Outputs", "TrendsResults", "results_rasters")

lin_dec_classification <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(1,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

lin_inc_classification <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(2,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

stable_classification <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(3,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

step_dec_classification <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(4,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

step_inc_classification <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(5,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

quad_dec_classification1 <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(6,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

quad_dec_classification2 <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(7,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

quad_inc_classification1 <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(8,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly, "binary_trajectories", file_name_with_tif))
}

quad_inc_classification2 <- function (trajectory_results_raster, annual_or_monthly, file_name_with_tif){
  x<- terra::classify(trajectory_results_raster, cbind(9,1), others =0)
  writeRaster(x, here(output_file_path, annual_or_monthly,"binary_trajectories", file_name_with_tif))
}

Sys.time(); lin_dec_classification(monthly_rds_list[[1]], "monthly", "evi_monthly_lin_dec.tif"); Sys.time()
Sys.time(); lin_inc_classification(monthly_rds_list[[1]], "monthly", "evi_monthly_lin_inc.tif"); Sys.time()
Sys.time(); stable_classification(monthly_rds_list[[1]], "monthly", "evi_monthly_stable.tif"); Sys.time()
Sys.time(); step_dec_classification(monthly_rds_list[[1]], "monthly", "evi_monthly_step_dec.tif"); Sys.time()
Sys.time(); step_inc_classification(monthly_rds_list[[1]], "monthly", "evi_monthly_step_inc.tif"); Sys.time()
Sys.time(); quad_dec_classification1(monthly_rds_list[[1]], "monthly", "evi_monthly_quad_dec_acc.tif"); Sys.time()
Sys.time(); quad_dec_classification2(monthly_rds_list[[1]], "monthly", "evi_monthly_quad_dec_decc.tif"); Sys.time()
Sys.time(); quad_inc_classification1(monthly_rds_list[[1]], "monthly", "evi_monthly_quad_inc_acc.tif"); Sys.time()
Sys.time(); quad_inc_classification2(monthly_rds_list[[1]], "monthly", "evi_monthly_quad_inc_decc.tif"); Sys.time()

Sys.time(); lin_dec_classification(annual_rds_list[[1]], "annual", "evi_annual_lin_dec.tif"); Sys.time()
Sys.time(); lin_inc_classification(annual_rds_list[[1]], "annual", "evi_annual_lin_inc.tif"); Sys.time()
Sys.time(); stable_classification(annual_rds_list[[1]], "annual", "evi_annual_stable.tif"); Sys.time()
Sys.time(); step_dec_classification(annual_rds_list[[1]], "annual", "evi_annual_step_dec.tif"); Sys.time()
Sys.time(); step_inc_classification(annual_rds_list[[1]], "annual", "evi_annual_step_inc.tif"); Sys.time()
Sys.time(); quad_dec_classification1(annual_rds_list[[1]], "annual", "evi_annual_quad_dec_acc.tif"); Sys.time()
Sys.time(); quad_dec_classification2(annual_rds_list[[1]], "annual", "evi_annual_quad_dec_decc.tif"); Sys.time()
Sys.time(); quad_inc_classification1(annual_rds_list[[1]], "annual", "evi_annual_quad_inc_acc.tif"); Sys.time()
Sys.time(); quad_inc_classification2(annual_rds_list[[1]], "annual", "evi_annual_quad_inc_decc.tif"); Sys.time()

Sys.time(); lin_dec_classification(monthly_rds_list[[2]], "monthly", "ndvi_monthly_lin_dec.tif"); Sys.time()
Sys.time(); lin_inc_classification(monthly_rds_list[[2]], "monthly", "ndvi_monthly_lin_inc.tif"); Sys.time()
Sys.time(); stable_classification(monthly_rds_list[[2]], "monthly", "ndvi_monthly_stable.tif"); Sys.time()
Sys.time(); step_dec_classification(monthly_rds_list[[2]], "monthly", "ndvi_monthly_step_dec.tif"); Sys.time()
Sys.time(); step_inc_classification(monthly_rds_list[[2]], "monthly", "ndvi_monthly_step_inc.tif"); Sys.time()
Sys.time(); quad_dec_classification1(monthly_rds_list[[2]], "monthly", "ndvi_monthly_quad_dec_acc.tif"); Sys.time()
Sys.time(); quad_dec_classification2(monthly_rds_list[[2]], "monthly", "ndvi_monthly_quad_dec_decc.tif"); Sys.time()
Sys.time(); quad_inc_classification1(monthly_rds_list[[2]], "monthly", "ndvi_monthly_quad_inc_acc.tif"); Sys.time()
Sys.time(); quad_inc_classification2(monthly_rds_list[[2]], "monthly", "ndvi_monthly_quad_inc_decc.tif"); Sys.time()

Sys.time(); lin_dec_classification(annual_rds_list[[2]], "annual", "ndvi_annual_lin_dec.tif"); Sys.time()
Sys.time(); lin_inc_classification(annual_rds_list[[2]], "annual", "ndvi_annual_lin_inc.tif"); Sys.time()
Sys.time(); stable_classification(annual_rds_list[[2]], "annual", "ndvi_annual_stable.tif"); Sys.time()
Sys.time(); step_dec_classification(annual_rds_list[[1]], "annual", "ndvi_annual_step_dec.tif"); Sys.time()
Sys.time(); step_inc_classification(annual_rds_list[[1]], "annual", "ndvi_annual_step_inc.tif"); Sys.time()
Sys.time(); quad_dec_classification1(annual_rds_list[[1]], "annual", "ndvi_annual_quad_dec_acc.tif"); Sys.time()
Sys.time(); quad_dec_classification2(annual_rds_list[[1]], "annual", "ndvi_annual_quad_dec_decc.tif"); Sys.time()
Sys.time(); quad_inc_classification1(annual_rds_list[[1]], "annual", "ndvi_annual_quad_inc_acc.tif"); Sys.time()
Sys.time(); quad_inc_classification2(annual_rds_list[[1]], "annual", "ndvi_annual_quad_inc_decc.tif"); Sys.time()
```
