```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)

library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I evaluate the composition, in terms of veg formations from Mapbiomas, in each trajectory type 
at 1km resolution 
(1) Read final trajectory type results for all indices and mapbiomass annual lulc product
Note mapbiomas 1km %formation product was processed on GEE
(2) Process mapbiomass annual product by cropping/masking, aligning projection, resolution against trajectory result rasters
and average composition across years
(3) Preliminary data analyses 
(4) Visualization of above results
(5) Proportion of each trajectory covered by more than half of a veg composition & visualization

#Step 1- read data
```{r}
monthly_input_file_path <- here("Outputs", "TrendsResults", "results_rasters", "monthly")
monthly_file_list<- list.files(path = paste0(monthly_input_file_path, "/"), pattern = paste0("*","_monthly_swin11","*"), all.files = T, full.names = T)
monthly_file_list <- gtools::mixedsort(monthly_file_list)

monthly_rds_list<- lapply(monthly_file_list, rast)
names(monthly_rds_list[[2]])<- "evi_swin11_monthly"
names(monthly_rds_list[[3]])<- "ndvi_swin11_monthly"
names(monthly_rds_list[[1]])<- "aniso_swin11_monthly"

mapbiomas_input_file_path <- here("Data", "Mapbiomas_formations")
mapbiomas_file_list <- list.files(path = paste0(mapbiomas_input_file_path , "/"), pattern = paste0("*","_1km","*"), all.files = T, full.names = T)

mapbiomas_rds_list<- lapply(mapbiomas_file_list, rast)

```

#Step 2- data processing
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

#Crop/mask
crop_mask_function <- function (raster){
  x_crop <- terra::crop(raster, vect(cerrado))
  x_mask <- terra::mask (x_crop, vect(cerrado))
  x_mask
}

Sys.time(); forest_percentage <- crop_mask_function(mapbiomas_rds_list[[1]]); Sys.time()
Sys.time(); grass_percentage <- crop_mask_function(mapbiomas_rds_list[[2]]); Sys.time()
Sys.time(); savanna_percentage <- crop_mask_function(mapbiomas_rds_list[[3]]); Sys.time()

#Time period 2002 - 2021 
forest_percentage<- forest_percentage[[18:37]]
#writeRaster(forest_percentage, here("Outputs", "OtherVariables", "Formations_Heterogeniety", "forest_percentage_1km_2002_2021.tif"))
grass_percentage<- grass_percentage[[18:37]]
#writeRaster(grass_percentage, here("Outputs", "OtherVariables", "Formations_Heterogeniety", "grass_percentage_1km_2002_2021.tif"))
savanna_percentage<- savanna_percentage[[18:37]]
#writeRaster(savanna_percentage, here("Outputs", "OtherVariables", "Formations_Heterogeniety", "savanna_percentage_1km_2002_2021.tif"))

#Map
animation_map_function <- function (timeseries_raster){
  animation_x_time_map<-
   tm_shape(cerrado)+ tm_borders()+
   tm_shape (timeseries_raster)+
   tm_raster(col.scale = tm_scale_intervals(n=5, style= "quantile", values = "Greens")) +
   tm_facets(nrow=1, ncol=1, free.coords = TRUE) 
  animation_x_time_map
}

map_forest<- animation_map_function(forest_percentage)
tmap_animation(map_forest, filename = here("Outputs", "OtherVariables", "Formations_Heterogeniety", "1km_annual_forestformationpercentage_2002_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
map_grass<- animation_map_function(grass_percentage)
tmap_animation(map_grass, filename = here("Outputs", "OtherVariables", "Formations_Heterogeniety", "1km_annual_grassformationpercentage_2002_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
map_savanna<- animation_map_function(savanna_percentage)
tmap_animation(map_savanna, filename = here("Outputs", "OtherVariables", "Formations_Heterogeniety", "1km_annual_savannaformationpercentage_2002_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
remove(crop_mask_function, mapbiomas_file_list, mapbiomas_input_file_path)
remove(animation_map_function, map_forest, map_grass, map_savanna)
remove(monthly_file_list, monthly_input_file_path)

#Average % veg formation across 20 years
mean_formation <-  function (veg_formation){
  x_mean <- terra::app(veg_formation, fun = "mean")
  x_mean
}
mean_forest <- mean_formation(forest_percentage)
mean_grass <- mean_formation(grass_percentage)
mean_savanna <- mean_formation(savanna_percentage)

#Reprojection to align veg formation rasters with trajectory results rasters
resample_mean_vegformation <- function (veg_formation){
  x_resample <- terra::resample(veg_formation, monthly_rds_list[[3]], method = "bilinear")
  x_resample
}
resample_mean_forest <- resample_mean_vegformation(mean_forest)
resample_mean_grass <- resample_mean_vegformation(mean_grass)
resample_mean_savanna <- resample_mean_vegformation(mean_savanna)

remove(mean_forest, mean_savanna, mean_grass, forest_percentage, savanna_percentage, grass_percentage, mean_formation, resample_mean_vegformation)
remove(mapbiomas_rds_list)

```

#Step 3- analyses of veg composition in each trajectory at 1km
```{r}
resample_veg_formation <- c( resample_mean_grass, resample_mean_savanna, resample_mean_forest)
names(resample_veg_formation)<- c( "average_grass_percentage", "average_savanna_percentage", "average_forest_percentage")

studyarea_veg_formation <- terra::mask (resample_veg_formation, monthly_rds_list[[3]]) #crop and mask to trajectory result rasters

formation_trajectory <- function(trajectory_results_raster) {
  split_trajectory <- terra::segregate(trajectory_results_raster, other = NA)
  trajectory_names <- c("Lin_dec", "Lin_inc", "No_trend", "Step_dec", "Step_inc", "Quad_dec_acc", "Quad_dec_dec", "Quad_inc_acc", "Quad_inc_dec")
  df_for_histogram<- list()
  for (i in 1:9){
    stack_raster <- c(split_trajectory[[i]], resample_veg_formation)
    names(stack_raster)<- c(trajectory_names[[i]], "average_grass_percentage", "average_savanna_percentage", "average_forest_percentage")
    stack_df <- terra::as.data.frame(stack_raster, na.rm = T)
    df_for_histogram[[i]]<- stack_df
  }
  df_for_histogram
}

Sys.time(); composition_ndvitrajectories <- formation_trajectory(monthly_rds_list[[3]]); Sys.time()

```

#Step 4- visualization of results
```{r}
df_prep_function <- function (df_veg_trajectory){
  prepped_df_list <- list()
  for(i in 1:length(df_veg_trajectory)){
    x_pivot <- pivot_longer(df_veg_trajectory[[i]], 2:4)
    x_pivot <- x_pivot %>% mutate(TrajType = (names(x_pivot)[1])) %>%
      dplyr::select(c(-1))
    x_pivot <- x_pivot %>% group_by (name, TrajType) %>%
      mutate(mean = mean(value),
                median = median (value))
    prepped_df_list[[i]] <- x_pivot
  }
  prepped_data <- bind_rows(prepped_df_list)
  prepped_data
}

Sys.time(); df_trial <- df_prep_function(composition_ndvitrajectories); Sys.time()

y <- ggplot(df_trial, aes(x = value)) + 
   geom_histogram(position = "identity", bins = 10) + 
  theme_classic(base_size = 14) + scale_x_continuous(c(0,1)) +
  facet_grid(name ~ TrajType) + 
  geom_vline(aes(xintercept = mean, colour = 'mean')) +
  geom_text(data = dplyr::distinct(df_trial, mean), 
            aes(x = (mean + 0.2), y = 200000, label = round(mean,2), angle =90), position = position_dodge(0.5)) +
    geom_vline(aes(xintercept = median, colour = 'median')) +
  geom_text(data = dplyr::distinct(df_trial, median), 
            aes(x = (median - 0.2), y = 200000, label = round(median,2), angle =90), position = position_dodge(0.5)) +
    scale_color_manual(name = "statistics", values = c(median = "blue", mean = "red")) +
  theme(axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) 
ggsave(here("Outputs", "TrendsResults", "veg_composition_NDVItrajectories.png"),
       y,dpi = 700, height = 50, width = 50, units = "cm")

```

#Step 5- proportion of each trajectory with more than the mean proportion of the three veg formations
```{r}
df_trial

x<- df_trial %>% 
  dplyr::select(-c(median)) %>%
  count(value > mean) %>%
  group_by(TrajType) %>% #total is the number of all pixels of each trajectory type
  mutate(pct= prop.table(n) * 100) %>%
  filter(`value > mean` == TRUE) %>%
  dplyr::select(-c(`value > mean`)) %>%
  ggplot() + aes(TrajType, pct, fill = name) +
  scale_fill_manual(values = c("#336600", "#CCFFFF", "#993300"))+
  geom_bar(stat="identity") +
  ylab("Proportion of pixels covered by more than average of a veg formation") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  coord_flip() +
  theme_bw(base_size = 14)
ggsave(here("Outputs", "TrendsResults", "more_than_average_veg_composition_NDVItrajectories.png"),
      x,dpi = 700, height = 50, width = 50, units = "cm")


```

