```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)

library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I evaluate the composition, in terms of veg formations from Mapbiomas, in each trajectory type 
at 1km resolution and across aggregated trajectories across all indices
(1) Read final trajectory type results for all indices, aggregated trajectories for all indices and mapbiomass annual lulc product
Note mapbiomas 1km %formation product was processed on GEE
(2) Process mapbiomass annual product by cropping/masking, aligning projection, resolution against trajectory result rasters and average composition across years
(3) Preliminary data analyses of the distribution of mean veg formation (#2) across seperate trajectories and aggregated trajectories across all indices
(4) Visualization of above results
(5) Proportion of each trajectory covered by more than half of a veg composition & visualization

#Step 1- read data- seperate trajectories, aggregated trajectories for all indices at annual and monthly time steps
```{r}
#separate trajectories 
monthly_input_file_path <- here("Outputs", "TrendsResults", "results_rasters", "monthly")
monthly_file_list<- list.files(path = paste0(monthly_input_file_path, "/"), pattern = paste0("*","_monthly_swin11","*"), all.files = T, full.names = T)
monthly_file_list <- gtools::mixedsort(monthly_file_list)

monthly_rds_list<- lapply(monthly_file_list, rast)
names(monthly_rds_list[[2]])<- "evi_swin11_monthly"
names(monthly_rds_list[[3]])<- "ndvi_swin11_monthly"
names(monthly_rds_list[[1]])<- "aniso_swin11_monthly"

annual_input_file_path <- here("Outputs", "TrendsResults", "results_rasters", "annual")
annual_file_list<- list.files(path = paste0(annual_input_file_path, "/"), pattern = paste0("*","_annual_swin11","*"), all.files = T, full.names = T)
annual_file_list <- gtools::mixedsort(annual_file_list)

annual_rds_list<- lapply(annual_file_list, rast)
names(annual_rds_list[[2]])<- "evi_swin11_annual"
names(annual_rds_list[[3]])<- "ndvi_swin11_annual"
names(annual_rds_list[[1]])<- "aniso_swin11_annual"

remove(monthly_input_file_path, monthly_file_list, annual_input_file_path, annual_file_list)

#aggregate trajectories
aggregate_results_input_file_path <- here("Outputs", "TrendsResults", "aggregate_trajectory_results")
aggregate_file_list <- list.files(path = paste0(aggregate_results_input_file_path, "/"), pattern = ".tif", all.files = T, full.names = T)

aggregate_rds_list <- lapply(aggregate_file_list, rast)
names(aggregate_rds_list[[1]])<- "annual_aniso1km_agg"
names(aggregate_rds_list[[2]])<- "annual_evi1km_agg"
names(aggregate_rds_list[[3]])<- "annual_ndvi1km_agg"
names(aggregate_rds_list[[4]])<- "monthly_aniso1km_agg"
names(aggregate_rds_list[[5]])<- "monthly_evi1km_agg"
names(aggregate_rds_list[[6]])<- "monthly_ndvi1km_agg"

remove(aggregate_results_input_file_path , aggregate_file_list)

#mapbiomas %area of 1km of savanna, forest and grassland
mapbiomas_input_file_path <- here("Data", "Mapbiomas_formations")
mapbiomas_file_list <- list.files(path = paste0(mapbiomas_input_file_path , "/"), pattern = paste0("*","_1km","*"), all.files = T, full.names = T)

mapbiomas_rds_list<- lapply(mapbiomas_file_list, rast)
remove(mapbiomas_input_file_path, mapbiomas_file_list )

```

#Step 2- data processing
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

#Crop/mask
crop_mask_function <- function (raster){
  x_crop <- terra::crop(raster, vect(cerrado))
  x_mask <- terra::mask (x_crop, vect(cerrado))
  x_mask
}

Sys.time(); forest_percentage <- crop_mask_function(mapbiomas_rds_list[[1]]); Sys.time()
Sys.time(); grass_percentage <- crop_mask_function(mapbiomas_rds_list[[2]]); Sys.time()
Sys.time(); savanna_percentage <- crop_mask_function(mapbiomas_rds_list[[3]]); Sys.time()

#Time period 2002 - 2021 
forest_percentage<- forest_percentage[[18:37]]
#writeRaster(forest_percentage, here("Outputs", "OtherVariables", "Formations_Heterogeniety", "forest_percentage_1km_2002_2021.tif"))
grass_percentage<- grass_percentage[[18:37]]
#writeRaster(grass_percentage, here("Outputs", "OtherVariables", "Formations_Heterogeniety", "grass_percentage_1km_2002_2021.tif"))
savanna_percentage<- savanna_percentage[[18:37]]
#writeRaster(savanna_percentage, here("Outputs", "OtherVariables", "Formations_Heterogeniety", "savanna_percentage_1km_2002_2021.tif"))

#Map
animation_map_function <- function (timeseries_raster){
  animation_x_time_map<-
   tm_shape(cerrado)+ tm_borders()+
   tm_shape (timeseries_raster)+
   tm_raster(col.scale = tm_scale_intervals(n=5, style= "quantile", values = "Greens")) +
   tm_facets(nrow=1, ncol=1, free.coords = TRUE) 
  animation_x_time_map
}

map_forest<- animation_map_function(forest_percentage)
#tmap_animation(map_forest, filename = here("Outputs", "OtherVariables", "Formations_Heterogeniety", "1km_annual_forestformationpercentage_2002_2021.gif"),
#                delay = 50, loop = TRUE, restart.delay = 500,
#                width = 900, height = 900, dpi = 150)
map_grass<- animation_map_function(grass_percentage)
#tmap_animation(map_grass, filename = here("Outputs", "OtherVariables", "Formations_Heterogeniety", "1km_annual_grassformationpercentage_2002_2021.gif"),
#                delay = 50, loop = TRUE, restart.delay = 500,
#                width = 900, height = 900, dpi = 150)
map_savanna<- animation_map_function(savanna_percentage)
#tmap_animation(map_savanna, filename = here("Outputs", "OtherVariables", "Formations_Heterogeniety", "1km_annual_savannaformationpercentage_2002_2021.gif"),
#                delay = 50, loop = TRUE, restart.delay = 500,
#                width = 900, height = 900, dpi = 150)
remove(crop_mask_function, mapbiomas_file_list, mapbiomas_input_file_path)
remove(animation_map_function, map_forest, map_grass, map_savanna)
remove(monthly_file_list, monthly_input_file_path)

#Average % veg formation across 20 years
mean_formation <-  function (veg_formation){
  x_mean <- terra::app(veg_formation, fun = "mean")
  x_mean
}
mean_forest <- mean_formation(forest_percentage)
mean_grass <- mean_formation(grass_percentage)
mean_savanna <- mean_formation(savanna_percentage)

#Reprojection to align veg formation rasters with trajectory results rasters
resample_mean_vegformation <- function (veg_formation){
  x_resample <- terra::resample(veg_formation, monthly_rds_list[[3]], method = "bilinear")
  x_resample
}
resample_mean_forest <- resample_mean_vegformation(mean_forest)
resample_mean_grass <- resample_mean_vegformation(mean_grass)
resample_mean_savanna <- resample_mean_vegformation(mean_savanna)

remove(mean_forest, mean_savanna, mean_grass, forest_percentage, savanna_percentage, grass_percentage, mean_formation, resample_mean_vegformation)
remove(mapbiomas_rds_list)

resample_veg_formation <- c( resample_mean_grass, resample_mean_savanna, resample_mean_forest)
names(resample_veg_formation)<- c( "average_grass_percentage", "average_savanna_percentage", "average_forest_percentage")

studyarea_veg_formation <- terra::mask (resample_veg_formation, monthly_rds_list[[3]]) #crop and mask to trajectory result rasters
```

#Step 3- analyses of veg composition in each seperate trajectory at 1km(annual and monthly) and for aggregate trajectories at 1km (annual and monthly)
```{r}
studyarea_veg_formation

seperate_trajectory_veg <- function(trajectory_results_raster) {
  split_trajectory <- terra::segregate(trajectory_results_raster, other = NA)
  trajectory_names <- c("Lin_dec", "Lin_inc", "No_trend", "Step_dec", "Step_inc", "Quad_dec_acc", "Quad_dec_dec", "Quad_inc_acc", "Quad_inc_dec")
  df_for_histogram<- list()
  for (i in 1:9){
    stack_raster <- c(split_trajectory[[i]], resample_veg_formation)
    names(stack_raster)<- c(trajectory_names[[i]], "average_grass_percentage", "average_savanna_percentage", "average_forest_percentage")
    stack_df <- terra::as.data.frame(stack_raster, na.rm = T)
    df_for_histogram[[i]]<- stack_df
  }
  df_for_histogram
}

Sys.time(); monthly_aniso_seperate <- seperate_trajectory_veg(monthly_rds_list[[1]]); Sys.time()
Sys.time(); monthly_evi_seperate <- seperate_trajectory_veg(monthly_rds_list[[2]]); Sys.time()
Sys.time(); monthly_ndvi_seperate <- seperate_trajectory_veg(monthly_rds_list[[3]]); Sys.time()

Sys.time(); annual_aniso_seperate <- seperate_trajectory_veg(annual_rds_list[[1]]); Sys.time()
Sys.time(); annual_evi_seperate <- seperate_trajectory_veg(annual_rds_list[[2]]); Sys.time()
Sys.time(); annual_ndvi_seperate <- seperate_trajectory_veg(annual_rds_list[[3]]); Sys.time()

aggregate_trajectory_veg <- function(agg_trajectory_results_raster) {
  split_trajectory <- terra::segregate(agg_trajectory_results_raster, other = NA)
  trajectory_names <- c("Decrease", "Increase", "No_trend")
  df_for_histogram<- list()
  for (i in 1:3){
    stack_raster <- c(split_trajectory[[i]], resample_veg_formation)
    names(stack_raster)<- c(trajectory_names[[i]], "average_grass_percentage", "average_savanna_percentage", "average_forest_percentage")
    stack_df <- terra::as.data.frame(stack_raster, na.rm = T)
    df_for_histogram[[i]]<- stack_df
  }
  df_for_histogram
}

Sys.time(); monthly_aniso_agg <- aggregate_trajectory_veg(aggregate_rds_list[[4]]); Sys.time()
Sys.time(); monthly_evi_agg <- aggregate_trajectory_veg(aggregate_rds_list[[5]]); Sys.time()
Sys.time(); monthly_ndvi_agg <- aggregate_trajectory_veg(aggregate_rds_list[[6]]); Sys.time()

Sys.time(); annual_aniso_agg <- aggregate_trajectory_veg(aggregate_rds_list[[1]]); Sys.time()
Sys.time(); annual_evi_agg <- aggregate_trajectory_veg(aggregate_rds_list[[2]]); Sys.time()
Sys.time(); annual_ndvi_agg <- aggregate_trajectory_veg(aggregate_rds_list[[3]]); Sys.time()

```

#Step 4- visualization of results - violin plots
```{r}
df_prep_function <- function (df_veg_trajectory){
  prepped_df_list <- list()
  for(i in 1:length(df_veg_trajectory)){
    x_pivot <- pivot_longer(df_veg_trajectory[[i]], 2:4)
    x_pivot <- x_pivot %>% mutate(TrajType = (names(x_pivot)[1])) %>%
      dplyr::select(c(-1))
   x_pivot <- x_pivot %>% mutate(Vegformation = case_when(name == "average_grass_percentage"~"Grass",
                                                          name == "average_savanna_percentage"~ "Savanna",
                                                          TRUE ~"Forest"))
    prepped_df_list[[i]] <- x_pivot
  }
  prepped_data <- bind_rows(prepped_df_list)
  prepped_data
}

Sys.time(); df_monthly_aniso_seperate <- df_prep_function(monthly_aniso_seperate); Sys.time()
Sys.time(); df_monthly_evi_seperate <- df_prep_function(monthly_evi_seperate); Sys.time()
Sys.time(); df_monthly_ndvi_seperate <- df_prep_function(monthly_ndvi_seperate); Sys.time()

Sys.time(); df_annual_aniso_seperate <- df_prep_function(annual_aniso_seperate); Sys.time()
Sys.time(); df_annual_evi_seperate <- df_prep_function(annual_evi_seperate); Sys.time()
Sys.time(); df_annual_ndvi_seperate <- df_prep_function(annual_ndvi_seperate); Sys.time()

Sys.time(); df_monthly_aniso_agg <- df_prep_function(monthly_aniso_agg); Sys.time()
Sys.time(); df_monthly_evi_agg <- df_prep_function(monthly_evi_agg ); Sys.time()
Sys.time(); df_monthly_ndvi_agg <- df_prep_function(monthly_ndvi_agg); Sys.time()

Sys.time(); df_annual_aniso_agg <- df_prep_function(annual_aniso_agg); Sys.time()
Sys.time(); df_annual_evi_agg <- df_prep_function(annual_evi_agg); Sys.time()
Sys.time(); df_annual_ndvi_agg <- df_prep_function(annual_ndvi_agg ); Sys.time()

vegformation_palette<- c("#117733", "#44AA99", "#CC9933" )
boxplot_function <- function (df){
  y_boxplot <- ggplot(df, aes(y = value, x = Vegformation)) +
  geom_boxplot(aes(fill = Vegformation)) + 
  facet_wrap(.~TrajType) +
  theme_classic() +
  theme(legend.title=element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = vegformation_palette) +
  xlab("Vegetation Formation") + ylab("Proportion of 1km pixel")
  y_boxplot

}

monthly_ndvi_agg <- boxplot_function(df_monthly_ndvi_agg)
ggsave(here("Outputs", "TrendsResults", "boxplot_agg_ndvi_monthly.png"),
      monthly_ndvi_agg ,dpi = 700, height = 10, width = 10, units = "cm")
annual_ndvi_agg <- boxplot_function(df_annual_ndvi_agg)
ggsave(here("Outputs", "TrendsResults", "boxplot_agg_ndvi_annual.png"),
      annual_ndvi_agg ,dpi = 700, height = 10, width = 10, units = "cm")
monthly_ndvi_seperate <- boxplot_function(df_monthly_ndvi_seperate)
ggsave(here("Outputs", "TrendsResults", "boxplot_separate_ndvi_monthly.png"),
      monthly_ndvi_seperate ,dpi = 700, height = 20, width = 20, units = "cm")
annual_ndvi_seperate <- boxplot_function(df_annual_ndvi_seperate)
ggsave(here("Outputs", "TrendsResults", "boxplot_separate_ndvi_annual.png"),
      annual_ndvi_seperate ,dpi = 700, height = 20, width = 20, units = "cm")

#aggregated, cross timesteps and indices
Sys.time(); df_monthly_aniso_agg <- df_monthly_aniso_agg %>% mutate(TimeStep ="monthly", Index = "anisotropic EVI"); Sys.time()
Sys.time(); df_monthly_evi_agg <- df_monthly_evi_agg %>% mutate(TimeStep ="monthly", Index = "EVI"); Sys.time()
Sys.time(); df_monthly_ndvi_agg <- df_monthly_ndvi_agg %>% mutate(TimeStep ="monthly", Index = "NDVI") ; Sys.time()

Sys.time(); df_annual_aniso_agg <- df_annual_aniso_agg %>% mutate(TimeStep ="annual", Index = "anisotropic EVI"); Sys.time()
Sys.time(); df_annual_evi_agg <- df_annual_evi_agg %>% mutate(TimeStep ="annual", Index = "EVI"); Sys.time()
Sys.time(); df_annual_ndvi_agg <- df_annual_ndvi_agg %>% mutate(TimeStep ="annual", Index = "NDVI") ; Sys.time()

df_agg <- bind_rows(df_monthly_aniso_agg, df_monthly_evi_agg, df_monthly_ndvi_agg, df_annual_aniso_agg, df_annual_evi_agg, df_annual_ndvi_agg)
agg_boxplot <- ggplot(df_agg, aes(y = value, x = Vegformation)) +
  geom_boxplot(aes(fill = Vegformation, alpha = factor(Index))) + 
  scale_alpha_discrete(range = c(0.3, 0.6, 0.9), guide = guide_legend(override.aes = list(fill = "black"))) +
  facet_grid(TimeStep~TrajType) +
  theme_classic() +
  theme(legend.title=element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())+ 
  scale_fill_manual(values = vegformation_palette) +
  xlab("Vegetation Formation") + ylab("Proportion of 1km pixel")
ggsave(here("Outputs", "TrendsResults", "boxplot_agg_formations_indices_timestep.png"),
      agg_boxplot  ,dpi = 700, height = 20, width = 20, units = "cm")


#y_hist <- ggplot(df_trial, aes(x = value)) + 
#   geom_histogram(position = "identity", bins = 10) + 
#  theme_classic(base_size = 14) + scale_x_continuous(c(0,1)) +
#  facet_grid(name ~ TrajType) + 
#  geom_vline(aes(xintercept = mean, colour = 'mean')) +
#  geom_text(data = dplyr::distinct(df_trial, mean), 
#            aes(x = (mean + 0.2), y = 200000, label = round(mean,2), angle =90), position = position_dodge(0.5)) +
#    geom_vline(aes(xintercept = median, colour = 'median')) +
#  geom_text(data = dplyr::distinct(df_trial, median), 
#            aes(x = (median - 0.2), y = 200000, label = round(median,2), angle =90), position = position_dodge(0.5)) +
#    scale_color_manual(name = "statistics", values = c(median = "blue", mean = "red")) +
#  theme(axis.title.x = element_blank(),
#    axis.ticks.x = element_blank(),
#    axis.title.y = element_blank(),
#    axis.text.y = element_blank(),
#    axis.ticks.y = element_blank()) 
#ggsave(here("Outputs", "TrendsResults", "hist_veg_composition_NDVItrajectories.png"),
#       y_hist,dpi = 700, height = 50, width = 50, units = "cm")



```
