```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I prepare "ground truth"  data  using Stevens et al., 2017.
In the SI of Stevens et al, there is information about the reviewed papers including key information 
(such as lat/long of plots). I manually made a csv by copy pasting from the SI word doc to excel.

7 papers reviewed in Stevens et al., basically have the same lat/long. These plots are in Assis, SP and are 
mostly Durigan plots. 

I consider the final trajectory (after excluding 20% anthropic area) for all three using the swin =11 stl setting
indices at 1km scale. 
(1) Reading data and making a point vector of the lat/long of the plots from Stevens et al 
(2) Finding the trajectory of the pixels in which the plots from Stevens et al fall and the distance to the closest trajectory 
(3) Finding the second closest trajectory in case lat/long of the plots are outside the study area of our project
(4) Results compilation

#Step 1- reading data, processign plots into point vector and map making
```{r}
stevens_si<- read_csv(here("Data","Stevensetal_woodyencroachment_metaanalyses_data", "si_plotinformation.csv"))
names(stevens_si)[14]<-"y"
names(stevens_si)[15]<-"x"
stevens_si_removeNA<- stevens_si %>% filter(!is.na(x))
stevens_si_removeNA<- stevens_si_removeNA %>% filter(Country!="Venezuela")

#Duplicate plots resulted in multiple papers eg Durigan et al 1987 and 2006 are from the same plots
#So I "merge" rows and take the mean of the "annual change" that was calculated in Stevens et al 
x<-stevens_si_removeNA %>% 
  group_by(x,y) %>%
  summarise(mean_annual_rate= mean(Annual_change)) %>%
           ungroup()

x_join<- stevens_si_removeNA %>% left_join(x)
x_join2<- x_join %>% filter(!duplicated(mean_annual_rate))

stevens_si_vector<- sf::st_as_sf(x_join2, coords = c("x", "y"), crs="epsg:4326")
stevens_si_vector<- stevens_si_vector %>% mutate(uniqueID= 1:nrow(stevens_si_vector))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)

#tmap_mode("view")
#tm_shape(d_trans)+ tm_fill()+ tm_shape(stevens_si_vector) + tm_dots(size="Annual_change")
#st_write(stevens_si_vector, here("Data","Stevensetal_woodyencroachment_metaanalyses_data", "plotinformation.shp"), overwrite = T)

input_file_path <- here("Outputs", "TrendsResults")
file_list<- list.files(path = paste0(input_file_path, "/"), pattern = paste0("*","swin11","*"), all.files = T, full.names = T)
file_list <- gtools::mixedsort(file_list)

rds_list<- lapply(file_list, rast)

#1~ "Linear decrease", 2~ "Linear increase",3 ="No trend", 4~ "Step decrease", 5~"Step increase",
#6~"Quadratic decrease (accelerated)", 7~"Quadratic decrease (decelerated)", 8~"Quadratic increase (accelerated)",
#9~"Quadratic increase (decelerated)"

## To make a map of where field points fall in our study
Set1palette<- c("#44AA99", "#117733", "lightgrey", "#CC6677", "#AA4499", "#332288", "#DDCC77" , "#882255",  "#88CCEE")

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

points_map_withindex_function<- function (trendsresults_raster, title){
  x_map <- tm_shape(st_as_sf(cerrado))+ tm_borders()+ 
  tm_shape(trendsresults_raster) +
  tm_raster(col.scale = tm_scale_categorical(values = Set1palette), col.legend = tm_legend_hide()) +
  tm_shape(stevens_si_vector) + 
  tm_dots(size="Annual_change")+
  tm_add_legend( labels = c("Linear decrease", 
                           "Linear increase",
                           "No trend",
                           "Step decrease",
                           "Step increase",
                           "Quad decrease (accelerated)",
                           "Quad decrease (decelerated)",
                           "Quad increase (accelerated)",
                           "Quad incrase (decelerated)"),
                fill = Set1palette,
                tm_legend(legend.text.size =10)) + tm_title(title)  
  x_map
}

Sys.time(); map_ndvi_points<- points_map_withindex_function(rds_list[[1]], "NDVI"); Sys.time()
Sys.time(); map_evi_points<- points_map_withindex_function(rds_list[[2]], "EVI"); Sys.time()
Sys.time(); map_aniso_points<- points_map_withindex_function(rds_list[[3]], "anisoEVI"); Sys.time()

points_trajectory_map<- tmap_arrange(map_ndvi_points, map_evi_points, map_aniso_points, ncol = 3, nrow = 1)
tmap_save(points_trajectory_map, here("Outputs", "AnalysesResults", "Relationship_to_fieldplots", "points_trajectories_map.jpeg"),
         height = 40, width = 80, units = "cm", dpi=700)

remove(stevens_si, x, x_join, x_join2, mappalette)
remove(map_ndvi_points, map_evi_points, map_aniso_points, points_trajectory_map)

```

#Step 2 - finding the closest trajectory shape (distance and type of trajectory) to each of the plots in the Stevens et al paper. 
```{r}
trajectory_segregate <- function(x) {
    x_split<- terra::segregate(x, other=999)
    x_split_crop<- terra::crop(x_split, vect(d_trans))
    x_split_mask<- terra::mask(x_split_crop, vect(d_trans))
    x_split_mask
}
NDVItrajectories_split<- trajectory_segregate(rds_list[[1]])
EVItrajectories_split<- trajectory_segregate(rds_list[[2]])
anisoEVItrajectories_split<- trajectory_segregate(rds_list[[3]])

distance_function<- function(split_trajectory_raster){
  x_dist<- terra::gridDist(split_trajectory_raster, target= 1)
}


calculate_distances<- function (split_raster){
  list_distance_raster<- list()
  for (i in 1:nlyr(split_raster)){
    x_distance<- distance_function(split_raster[[i]])
    list_distance_raster[[i]]<- x_distance
}
  distance_raster<- rast(list_distance_raster)
  names(distance_raster)<- 1:nlyr(split_raster)
  distance_raster
}

Sys.time(); NDVI_distances<- calculate_distances(NDVItrajectories_split); Sys.time() 
Sys.time(); EVI_distances<- calculate_distances(EVItrajectories_split); Sys.time() 
Sys.time(); anisoEVI_distances<- calculate_distances(anisoEVItrajectories_split); Sys.time() 


shortestdistance_trajectory<- function (distance_raster){
  distance_df<- terra::extract(distance_raster, stevens_si_vector)
  pivot_dist_df<- distance_df %>% pivot_longer(-1)
  pivot_dist_df2<-pivot_dist_df  %>% group_by(as.factor(ID)) %>% summarise(min(value), name[which.min(value)])
  names(pivot_dist_df2)<- c("uniqueID", "min_dist_m", "respective_trajshape")
  pivot_dist_df2
}
Sys.time(); NDVI_closest_trajectory_df<- shortestdistance_trajectory(NDVI_distances); Sys.time()
Sys.time(); EVI_closest_trajectory_df<- shortestdistance_trajectory(EVI_distances); Sys.time()
Sys.time(); anisoEVI_closest_trajectory_df<- shortestdistance_trajectory(anisoEVI_distances); Sys.time()


```
Only 5 out of 17 of the plots fall within the study area of this study. Hence for the remaining
12 plots I find the closest pixel that has a trajectory and the distance.


#Step 3- finding the second closest pixel (distance and trajectory) for plots that are outside
of the study area of this study
```{r}
overall_file_path <- here("Outputs", "ModelSelection")

read_files_function <- function (index_name, file_name_with_rds){
  rds_file<- read_rds(paste0(overall_file_path,"/", index_name,"/", file_name_with_rds))
}

Sys.time(); ndvi_complete_cerrado_trendresults <- read_files_function("NDVI", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); evi_complete_cerrado_trendresults <- read_files_function("EVI", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); aniso_complete_cerrado_trendresults <- read_files_function("anisoEVI", "modelselection_swindow_11.rds"); Sys.time()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()
remove(mosaic_ndvi)

alltrends_map_shortestdistance_function <- function (completecerrado_trendresults){
  completecerrado_trendresults<- completecerrado_trendresults%>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~5, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                )) %>% 
    dplyr::select(c(x,y,pixelvalue))
  
  cellindex_vector<- terra::vect(completecerrado_trendresults, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi , "pixelvalue", fun="max")
  
  indextrajectories_split<- trajectory_segregate(cellindex_raster)
  trajectory_distances<- calculate_distances(indextrajectories_split); Sys.time() 
  
  second_closest_trajectories_distance_df <- shortestdistance_trajectory(trajectory_distances)
  second_closest_trajectories_distance_df
}

Sys.time();ndvi_second_closest_distance_df<- alltrends_map_shortestdistance_function(ndvi_complete_cerrado_trendresults); Sys.time()
Sys.time();evi_second_closest_distance_df<- alltrends_map_shortestdistance_function(evi_complete_cerrado_trendresults); Sys.time()
Sys.time();aniso_second_closest_distance_df<- alltrends_map_shortestdistance_function(aniso_complete_cerrado_trendresults); Sys.time()

```

#Step 4 - Results presentation 
 
```{r}


results_withinstudyarea <- left_join()



df_prep<- function (closest_distance_df, index_name){
  x<- closest_distance_df %>% 
    dplyr::select(-c(min_dist_m))
  names(x)[2]<- index_name
x
}

closest_ndvi<- df_prep(NDVI_closest_trajectory_df, "NDVI") 
closest_evi<- df_prep(EVI_closest_trajectory_df, "EVI")
closest_aniso<- df_prep(anisoEVI_closest_trajectory_df, "anisoEVI")

closest_df<- left_join(closest_aniso, closest_evi)
closest_df<- left_join (closest_df, closest_ndvi)
closest_df<- closest_df %>% mutate(proximity= "closest")    
closest_df$uniqueID<-as.numeric(closest_df$uniqueID)
pivot_closest_df<- pivot_longer (closest_df, 2:4)
        
second_closest_aniso<- df_prep(sec_anisoEVI_closest_trajectory_df, "anisoEVI")
second_closest_evi<- df_prep(sec_EVI_closest_trajectory_df, "EVI")
second_closest_ndvi<- df_prep(sec_NDVI_closest_trajectory_df, "NDVI")

sec_closest_df<- left_join(second_closest_aniso, second_closest_evi)
sec_closest_df<- left_join (sec_closest_df, second_closest_ndvi)
sec_closest_df<- sec_closest_df %>% mutate(proximity= "second_closest")
pivot_sec_closest_df<- pivot_longer (sec_closest_df, 2:4)

plot_df<- rbind(pivot_closest_df, pivot_sec_closest_df)
stackedbarplot_df2<- plot_df %>% group_by(proximity, name, value) %>% 
  summarise(sum=n())
stackedbarplot_df2<- stackedbarplot_df2 %>% mutate(value = case_when(value == 1~ "Linear decrease",
                                                                     value == 2~ "Linear increase",
                                                                     value == 3~ "No trend",
                                                                     value == 4~ "Step decrease",
                                                                     value == 5~ "Step increase") )


mappalette<- brewer.pal(n=7, "Set1")
mappalette[[3]]<- "darkgrey"
mappalette[[4]]<-  "#984EA3"
mappalette[[5]]<-  "#FF7F00"
mappalette[[6]]<- "#FFFF33"
mappalette[[7]]<- "#A65628"


distance_stackedbarplot<- ggplot(stackedbarplot_df2, aes( x= proximity, y = sum, fill = value)) + 
    geom_bar(stat = 'identity', position = 'stack') + theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_fill_manual (values = mappalette)+ 
  facet_grid(~ name) 

sec_anisoEVI_closest_trajectory_df <- sec_anisoEVI_closest_trajectory_df %>% mutate(index= "anisoEVI")
sec_EVI_closest_trajectory_df <- sec_EVI_closest_trajectory_df %>% mutate(index= "EVI")
sec_NDVI_closest_trajectory_df <- sec_NDVI_closest_trajectory_df %>% mutate(index= "NDVI")

sec_dist_df<- rbind(sec_anisoEVI_closest_trajectory_df, sec_EVI_closest_trajectory_df, sec_NDVI_closest_trajectory_df)
sec_dist_df<- sec_dist_df %>% mutate(respective_trajshape= case_when(respective_trajshape == 1~ "Linear decrease",
                                                      respective_trajshape == 2~ "Linear increase",
                                                      respective_trajshape == 3~ "No trend",
                                                      respective_trajshape == 4~ "Step decrease",
                                                      respective_trajshape == 5~ "Step increase"))


distance_barplot<- ggplot(sec_dist_df, aes(x= respective_trajshape, y= min_dist_m, fill= respective_trajshape)) +
  geom_boxplot()+ theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_fill_manual (values = mappalette)+
  facet_grid(~index) 

distance<- ggarrange(distance_stackedbarplot, distance_barplot)
ggsave(here("Outputs", "AnalysesResults", "Relationship_to_fieldplots", "fieldplots_trajectoryshape_results.png"),
       distance,dpi = 700, height = 30, width=30, units = "cm")



```













