```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I prepare "ground truth"  data  using Stevens et al., 2017.
In the SI of Stevens et al, there is information about the reviewed papers including key information 
(such as lat/long of plots). I manually made a csv by copy pasting from the SI word doc to excel.

7 papers reviewed in Stevens et al., basically have the same lat/long. These plots are in Assis, SP and are 
mostly Durigan plots. 

I consider the final trajectory results using monthly time steps (1km reslution after excluding 20% anthropic area) for all three indices 
using the swin =11 stl setting
(1) Reading data and making a point vector of the lat/long of the plots from Stevens et al 
(2) Finding the trajectory of the pixels in which the plots from Stevens et al fall and the distance to the closest trajectory 
(3) Finding the second closest trajectory in case lat/long of the plots are outside the study area of our project
(4) Results compilation

#Step 1- reading data, processign plots into point vector and map making
```{r}
stevens_si<- read_csv(here("Data","Stevensetal_woodyencroachment_metaanalyses_data", "si_plotinformation.csv"))
names(stevens_si)[14]<-"y"
names(stevens_si)[15]<-"x"
stevens_si_removeNA<- stevens_si %>% filter(!is.na(x))
stevens_si_removeNA<- stevens_si_removeNA %>% filter(Country!="Venezuela")

#Duplicate plots resulted in multiple papers eg Durigan et al 1987 and 2006 are from the same plots
#So I "merge" rows and take the mean of the "annual change" that was calculated in Stevens et al 
x<-stevens_si_removeNA %>% 
  group_by(x,y) %>%
  summarise(mean_annual_rate= mean(Annual_change)) %>%
           ungroup()

x_join<- stevens_si_removeNA %>% left_join(x)
x_join2<- x_join %>% filter(!duplicated(mean_annual_rate))

stevens_si_vector<- sf::st_as_sf(x_join2, coords = c("x", "y"), crs="epsg:4326")
stevens_si_vector<- stevens_si_vector %>% mutate(uniqueID= 1:nrow(stevens_si_vector))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)

#tmap_mode("view")
#tm_shape(d_trans)+ tm_fill()+ tm_shape(stevens_si_vector) + tm_dots(size="Annual_change")
#st_write(stevens_si_vector, here("Data","Stevensetal_woodyencroachment_metaanalyses_data", "plotinformation.shp"), overwrite = T)

input_file_path <- here("Outputs", "TrendsResults", "results_rasters", "monthly")
file_list<- list.files(path = paste0(input_file_path, "/"), pattern = paste0("*","monthly_swin11","*"), all.files = T, full.names = T)
file_list <- gtools::mixedsort(file_list)

rds_list<- lapply(file_list, rast)

#1~ "Linear decrease", 2~ "Linear increase",3 ="No trend", 4~ "Step decrease", 5~"Step increase",
#6~"Quadratic decrease (accelerated)", 7~"Quadratic decrease (decelerated)", 8~"Quadratic increase (accelerated)",
#9~"Quadratic increase (decelerated)"

## To make a map of where field site points fall in our study region/results
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

points_map_withindex_function<- function (trendsresults_raster, title){
  x_map <- tm_shape(st_as_sf(cerrado))+ tm_borders()+ 
  tm_shape(trendsresults_raster) +
  tm_raster(col.scale = tm_scale_categorical(n=9, 
                                             values = c("#44AA99", "#117733", "lightgrey", "#882255", "#88CCEE", "#CC6677", "#AA4499", "#332288", "#DDCC77" )), 
            col.legend = tm_legend_hide()) +
  tm_shape(stevens_si_vector) + 
  tm_dots(size="Annual_change")+
  tm_add_legend( fill = c("#44AA99", "#117733", "lightgrey", "#882255", "#88CCEE", "#CC6677", "#AA4499", "#332288", "#DDCC77" ),
    labels = c("Linear decrease", 
                           "Linear increase",
                           "No trend",
                           "Step decrease",
                           "Step increase",
                           "Quad decrease (accelerated)",
                           "Quad decrease (decelerated)",
                           "Quad increase (accelerated)",
                           "Quad incrase (decelerated)")) +
                tm_layout(legend.text.size = 2) +tm_title(title)  
  x_map
}

Sys.time(); map_ndvi_points<- points_map_withindex_function(rds_list[[1]], "NDVI"); Sys.time()
Sys.time(); map_evi_points<- points_map_withindex_function(rds_list[[2]], "EVI"); Sys.time()
#Sys.time(); map_aniso_points<- points_map_withindex_function(rds_list[[3]], "anisoEVI"); Sys.time()

points_trajectory_map<- tmap_arrange(map_ndvi_points, map_evi_points, ncol = 2, nrow = 1)
tmap_save(points_trajectory_map, here("Outputs", "AnalysesResults", "Relationship_to_fieldplots", "points_trajectories_map.jpeg"),
         height = 40, width = 80, units = "cm", dpi=700)

remove(stevens_si, x, x_join, x_join2, mappalette)
remove(map_ndvi_points, map_evi_points, map_aniso_points, points_trajectory_map)

```

#Step 2 - finding the closest trajectory shape (distance and type of trajectory) to each of the plots in the Stevens et al paper. 
```{r}
trajectory_segregate <- function(x) {
    x_split<- terra::segregate(x, other=999)
    x_split_crop<- terra::crop(x_split, vect(d_trans))
    x_split_mask<- terra::mask(x_split_crop, vect(d_trans))
    x_split_mask
}
NDVItrajectories_split<- trajectory_segregate(rds_list[[1]])
EVItrajectories_split<- trajectory_segregate(rds_list[[2]])
#anisoEVItrajectories_split<- trajectory_segregate(rds_list[[3]])

distance_function<- function(split_trajectory_raster){
  x_dist<- terra::gridDist(split_trajectory_raster, target= 1)
}


calculate_distances<- function (split_raster){
  list_distance_raster<- list()
  for (i in 1:nlyr(split_raster)){
    x_distance<- distance_function(split_raster[[i]])
    list_distance_raster[[i]]<- x_distance
}
  distance_raster<- rast(list_distance_raster)
  names(distance_raster)<- 1:nlyr(split_raster)
  distance_raster
}

Sys.time(); NDVI_distances<- calculate_distances(NDVItrajectories_split); Sys.time() 
Sys.time(); EVI_distances<- calculate_distances(EVItrajectories_split); Sys.time() 
#Sys.time(); anisoEVI_distances<- calculate_distances(anisoEVItrajectories_split); Sys.time() 


shortestdistance_trajectory<- function (distance_raster){
  distance_df<- terra::extract(distance_raster, stevens_si_vector)
  pivot_dist_df<- distance_df %>% pivot_longer(-1)
  pivot_dist_df2<-pivot_dist_df  %>% 
    group_by(as.factor(ID)) %>% 
    summarise(min(value, na.rm= TRUE), name[which.min(value)])
  names(pivot_dist_df2)<- c("uniqueID", "min_dist_m", "respective_trajshape")
  pivot_dist_df2
}
Sys.time(); NDVI_closest_trajectory_df<- shortestdistance_trajectory(NDVI_distances); Sys.time()
Sys.time(); EVI_closest_trajectory_df<- shortestdistance_trajectory(EVI_distances); Sys.time()
#Sys.time(); anisoEVI_closest_trajectory_df<- shortestdistance_trajectory(anisoEVI_distances); Sys.time()


```
Only 5 out of 17 of the plots fall within the study area of this study. Hence for the remaining
12 plots I find the closest pixel that has a trajectory and the distance.
Of the 5 studies, when using NDVI all 5 studies fall in "Step increase" trajectories,
but when using EVI, one of the 5 studies Aquino et al., 2007 is in "Step decrease" trajectory.


#Step 3- finding the second closest pixel (distance and trajectory) for plots that are outside
of the study area of this study
```{r}
NDVItrajectories_split
EVItrajectories_split

convert_NA_to_999_function <- function (split_trajectory_raster){
  classified_raster <- terra::classify(split_trajectory_raster, cbind (NA, 999))
  crop_classified_raster <- terra::crop (classified_raster, vect(cerrado))
  mask_classified_raster <- terra::mask (crop_classified_raster, vect(cerrado))
  mask_classified_raster
}

converted_NDVItrajectories_split <- convert_NA_to_999_function(NDVItrajectories_split)
converted_EVItrajectories_split <- convert_NA_to_999_function(EVItrajectories_split)

Sys.time(); converted_NDVI_distances<- calculate_distances(converted_NDVItrajectories_split); Sys.time() 
Sys.time(); converted_EVI_distances<- calculate_distances(converted_EVItrajectories_split); Sys.time() 

Sys.time(); converted_NDVI_closest_trajectory_df<- shortestdistance_trajectory(converted_NDVI_distances); Sys.time()
Sys.time(); converted_EVI_closest_trajectory_df<- shortestdistance_trajectory(converted_EVI_distances); Sys.time()

```
Most of the remaining 12 studies are also closest to "Step increase" trajectory pixels, though there are 3 studies 
closest to Quadratic Increase(decelerated) and 1 study closest to Quadratic decrease (accelerated) when using NDVI.
With EVI, its the same with one additional study that is Quadratic Increase (decelerated) (which is Step increase in NDVI) and
the same study that is closest to Quadratic decrease (accelerated).

#Step 4 - Results presentation 
```{r}
NDVI_closest_trajectory_df
EVI_closest_trajectory_df

#Studies/field sites that are within my study region
df_prep<- function (closest_distance_df, index_name){
  x<- closest_distance_df %>% 
    dplyr::select(-c(min_dist_m))
  names(x)[2]<- index_name
x
}

closest_ndvi<- df_prep(NDVI_closest_trajectory_df, "NDVI") 
closest_evi<- df_prep(EVI_closest_trajectory_df, "EVI")

closest_df<- left_join(closest_ndvi, closest_evi)
closest_df$uniqueID<-as.numeric(closest_df$uniqueID)
pivot_closest_df <- pivot_longer (closest_df, 2:3)
pivot_closest_df <- pivot_closest_df %>% mutate(value = case_when(value== 1~ "Linear decrease",
                                                value==2~ "Linear increase",
                                                value==4~ "Step decrease",
                                                value==5~"Step increase",
                                                value==6~"Quadratic decrease (accelerated)",
                                                value==7~"Quadratic decrease (decelerated)",
                                                value==8~"Quadratic increase (accelerated)",
                                                value==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))  

within_studyarea_plot <- ggplot(pivot_closest_df, aes(x = name, fill = value)) + 
  geom_bar() +
  scale_fill_manual(values = c("#882255",  "#88CCEE")) +
  theme_classic(base_size = 14) +
  theme(legend.title=element_blank()) +xlab("Index") +ylab ("Count") +
  ggtitle ("Field bases studies within our study area")

#Studies/field sites that are outside the study area of this project
converted_NDVI_closest_trajectory_df
names(converted_NDVI_closest_trajectory_df)<- c("uniqueID", "NDVI_min_dist", "NDVI")
converted_EVI_closest_trajectory_df
names(converted_EVI_closest_trajectory_df)<- c("uniqueID", "EVI_min_dist", "EVI")

closest_df2 <- left_join(converted_NDVI_closest_trajectory_df, converted_EVI_closest_trajectory_df)
closest_df2 <- closest_df2 %>% filter(NDVI_min_dist != 0)
closest_df2$uniqueID<-as.numeric(closest_df2$uniqueID)
pivot_closest_df2 <- pivot_longer (closest_df2, c(3,5))
pivot_closest_df2 <- pivot_closest_df2 %>% mutate(value = case_when(value== 1~ "Linear decrease",
                                                value==2~ "Linear increase",
                                                value==4~ "Step decrease",
                                                value==5~"Step increase",
                                                value==6~"Quadratic decrease (accelerated)",
                                                value==7~"Quadratic decrease (decelerated)",
                                                value==8~"Quadratic increase (accelerated)",
                                                value==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))  
pivot_closest_df2 <- pivot_closest_df2 %>% dplyr::select(-c(EVI_min_dist))
names(pivot_closest_df2)[2] <- ("min_distance_m")
pivot

outside_studyarea_plot <- ggplot(pivot_closest_df2, aes(x = name, fill = value)) + 
  geom_bar() +
  scale_fill_manual(values = c("#CC6677", "#DDCC77" ,  "#882255",  "#88CCEE")) +
  scale_y_continuous(limits=c(0,12), breaks=seq(0,12,2)) +
  theme_classic(base_size = 14) +
  theme(legend.title=element_blank()) +xlab("Index") +ylab ("Count") +
  ggtitle ("Field bases studies outside our study area")


p <- ggplot(pivot_closest_df2, aes(x = value, y= min_distance_m, fill = value)) + 
  geom_violin() + 
  coord_flip() +
  scale_y_continuous(limits=c(0,18000), breaks=seq(0,18000,2000)) +
  scale_fill_manual(values = c("#CC6677", "#DDCC77" ,  "#882255",  "#88CCEE")) +
  theme_classic(base_size = 14) + ylab ("Trajectory type") + xlab ("Closest distance to (m)") +
  theme(legend.title=element_blank()) +ggtitle ("Distribution of distance of field study to closest trajectory")

library(ggpubr)
x<- ggarrange(within_studyarea_plot, outside_studyarea_plot, p, ncol = 2, nrow =2)
ggsave(here("Outputs", "AnalysesResults", "Relationship_to_fieldplots", "fieldsites_trajectories_results.jpeg"),
       x,dpi = 700, height = 50, width=80, units = "cm")
```













