```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
This script is about verification of the trajetories using asdetect. I do the verification only on a sample of the total pixels (generated
in 1_sampling.Rmd) 
(1) Install asdetect, read sampled pixels df which contains time series, monthly and annual trajectories results for different indices
(2) Run asdetect function for stltrend component and original timeseries in monthly time series derived results 
for sample points for all indices
(3) Run asdetect function for stltrend component and original timeseries in annual time series derived results 
for sample points for all indices
(4) Plotting results of as detect
(5) Plotting asdetect dt_th distribution

#Step 1- install package, read files
```{r}
#install.packages('devtools')
#devtools::install_github("caboulton/asdetect")
library(asdetect)

ndvi_evi_ts_complete_tr <- read.csv(here("Outputs", "TrajectoryPlotting", "ndvi_evi_ts_trajectories.csv"))

monthly_df <- ndvi_evi_ts_complete_tr %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time, 
                                                          NDVI_monthly_tr, EVI_monthly_tr, NDVI_original, NDVI_trend,
                                                          EVI_original, EVI_trend))

annual_df <- ndvi_evi_ts_complete_tr %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time, 
                                                          NDVI_annual_tr, EVI_annual_tr, NDVI_original, NDVI_trend,
                                                          EVI_original, EVI_trend))
annual_df <- annual_df %>%  
  separate_wider_delim(cols = time, delim = "_", names = c("Year","Month")) %>%
  dplyr::select(-c(Month))

annual_df2 <- annual_df %>% 
  group_by(NDVI_ID, climate_zone, NDVI_annual_tr, EVI_annual_tr, Year) %>%
  summarise(x = mean(x), y= mean(y),
    mean_NDVI_original = mean(NDVI_original), 
            mean_NDVI_trend = mean(NDVI_trend),
            mean_EVI_original = mean(EVI_original),
            mean_EVI_trend = mean(EVI_trend))
  
remove(annual_df)
remove(ndvi_evi_ts_complete_tr)
annual_df2 <- annual_df2 %>% rename("NDVI_original" = "mean_NDVI_original", 
                                    "NDVI_trend" = "mean_NDVI_trend",
                                    "EVI_original" = "mean_EVI_original", 
                                    "EVI_trend" = "mean_EVI_trend")

annual_df2 <- annual_df2 %>% rename("NDVI_tr" = "NDVI_annual_tr", 
                                    "EVI_tr" = "EVI_annual_tr")

monthly_df <- monthly_df %>% rename("NDVI_tr" = "NDVI_monthly_tr",
                                    "EVI_tr" = "EVI_monthly_tr")
```

#Step 2- asdetect function for stltrend component and original timeseries in time series derived results 
for sample points for all indices
```{r}
generate_asdetectvalues<- function (x_df){
 x_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
   mutate(NDVI_original_asdetect = as_detect(NDVI_original),
          NDVI_trend_asdetect = as_detect (NDVI_trend),
          EVI_original_asdetect = as_detect(EVI_original),
          EVI_trend_asdetect = as_detect(EVI_trend)
          )
  x_df
  }
Sys.time(); monthly_asdetect_df <- generate_asdetectvalues(monthly_df); Sys.time() #7-8 min
Sys.time(); annual_asdetect_df <- generate_asdetectvalues(annual_df2); Sys.time() 

generate_position_dtthresh <- function(asdetect_df){
  x_df <- asdetect_df %>% dplyr::select(NDVI_ID, climate_zone, NDVI_tr, EVI_tr,
                                        NDVI_original_asdetect, NDVI_trend_asdetect, EVI_original_asdetect, EVI_trend_asdetect) 
  
  ndvi_where_as_as_pos_function1 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(NDVI_original_as_pos = where_as(NDVI_original_asdetect)[1])
    as_df
  }
  
  final_df <- ndvi_where_as_as_pos_function1(x_df)
  
  ndvi_where_as_as_pos_function2 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(NDVI_trend_as_pos = where_as(NDVI_trend_asdetect)[1])
    as_df
  }
  
  final_df2 <- ndvi_where_as_as_pos_function2(final_df)
  
  evi_where_as_as_pos_function1 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(EVI_original_as_pos = where_as(EVI_original_asdetect)[1])
    as_df
  }
  
  final_df3 <- evi_where_as_as_pos_function1(final_df2)
  
  evi_where_as_as_pos_function2 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(EVI_trend_as_pos = where_as(EVI_trend_asdetect)[1])
    as_df
  }
  
  final_df4 <- evi_where_as_as_pos_function2(final_df3)
  
  ndvi_where_as_dt_value_function1 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(NDVI_original_dt_value = where_as(NDVI_original_asdetect)[2])
    as_df
  }
  
  final_df5 <- ndvi_where_as_dt_value_function1(final_df4)
  
  ndvi_where_as_dt_value_function2 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(NDVI_trend_dt_value = where_as(NDVI_trend_asdetect)[2])
    as_df
  }
  
  final_df6 <- ndvi_where_as_dt_value_function2(final_df5)
  
  evi_where_as_dt_value_function1 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(EVI_original_dt_value = where_as(EVI_original_asdetect)[2])
    as_df
  }
  
  final_df7 <- evi_where_as_dt_value_function1(final_df6)
  
  evi_where_as_dt_value_function2 <- function (x_df){
    as_df <- x_df %>% group_by(NDVI_ID, climate_zone, NDVI_tr, EVI_tr) %>%
      mutate(EVI_trend_dt_value = where_as(EVI_trend_asdetect)[2])
    as_df
  }
  
  final_df8 <- evi_where_as_dt_value_function2(final_df7)
  
  final_df8 <- final_df8 %>% dplyr::select(-c(NDVI_original_asdetect, NDVI_trend_asdetect, EVI_original_asdetect, EVI_trend_asdetect)) %>%
    unnest(c(NDVI_original_as_pos, NDVI_trend_as_pos, 
             EVI_original_as_pos, EVI_trend_as_pos, 
             NDVI_original_dt_value, NDVI_trend_dt_value, 
             EVI_original_dt_value, EVI_trend_dt_value))
  
  final_df8 <- final_df8 %>% distinct()
  
  final_df8
  }

Sys.time(); monthly_asdetect_df2 <- generate_position_dtthresh(monthly_asdetect_df); Sys.time() #1min
Sys.time(); annual_asdetect_df <- generate_position_dtthresh(annual_asdetect_df);
```

#Step 3- asdetect function for stltrend component and original timeseries in annual time series derived results 
for sample points for all indices

#Step 4- plotting results of asdetect
```{r}
input_file_path <- here("Outputs", "TrajectoryPlotting", "NDVI_trajectories")

read_sampledpoints_shp <- function (cz){
  x_shp <- st_read(here(input_file_path, paste0(cz, "/", cz, "_monthlytimeseries_NDVIsample.shp")))
  x_shp
}

central_shp <- read_sampledpoints_shp("central")
eastern_shp <- read_sampledpoints_shp("eastern")
southern_shp <- read_sampledpoints_shp("southern")
remove(input_file_path, read_sampledpoints_shp)
NDVIsample<- bind_rows(central_shp, eastern_shp, southern_shp)
remove(central_shp, eastern_shp, southern_shp)

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

# df preparation and plotting function for line plots of original & trend values from NDVI & EVI
df_preparation1 <- function (asdetect_df){
  x_df <- asdetect_df %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time, 
                                                          NDVI_monthly_tr, EVI_monthly_tr, NDVI_original, NDVI_trend,
                                                          EVI_original, EVI_trend))

  pivot_x_df <- pivot_longer(x_df, 8:11)
  pivot_x_df<- pivot_x_df %>% separate_wider_delim(cols = name, delim = "_", names = c("IndexName", "trend_original"))
  pivot_x_df2<- pivot_x_df %>% 
    pivot_wider(names_from = trend_original, values_from = value)
  pivot_x_df2 <- pivot_x_df2 %>% mutate(NDVI_monthly_tr = case_when(NDVI_monthly_tr == 1~ "Linear decrease",
                                                NDVI_monthly_tr ==2~ "Linear increase",
                                                NDVI_monthly_tr ==4~ "Step decrease",
                                                NDVI_monthly_tr ==5~"Step increase",
                                                NDVI_monthly_tr ==6~"Quadratic decrease (accelerated)",
                                                NDVI_monthly_tr ==7~"Quadratic decrease (decelerated)",
                                                NDVI_monthly_tr ==8~"Quadratic increase (accelerated)",
                                                NDVI_monthly_tr ==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))
  pivot_x_df2 <- pivot_x_df2 %>% mutate(EVI_monthly_tr = case_when(EVI_monthly_tr == 1~ "Linear decrease",
                                                EVI_monthly_tr ==2~ "Linear increase",
                                                EVI_monthly_tr ==4~ "Step decrease",
                                                EVI_monthly_tr ==5~"Step increase",
                                                EVI_monthly_tr ==6~"Quadratic decrease (accelerated)",
                                                EVI_monthly_tr ==7~"Quadratic decrease (decelerated)",
                                                EVI_monthly_tr ==8~"Quadratic increase (accelerated)",
                                                EVI_monthly_tr ==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))
  pivot_x_df2
}

Sys.time(); monthly_originaltrend_plot_df <- df_preparation1(monthly_asdetect_df); Sys.time()

# df preparation and plotting function for line plots of asdetect output values for the timeseries
df_preparation2 <- function (asdetect_df){
  x_df <- asdetect_df %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time, 
                                                          NDVI_original_asdetect, NDVI_trend_asdetect,
                                                          EVI_original_asdetect, EVI_trend_asdetect))

  pivot_x_df <- pivot_longer(x_df, 8:11)
  pivot_x_df<- pivot_x_df %>% separate_wider_delim(cols = name, delim = "_", names = c("IndexName", "trend_original", "asdetect")) %>% dplyr::select(-c(asdetect))
  pivot_x_df2<- pivot_x_df %>% 
    pivot_wider(names_from = trend_original, values_from = value)
  pivot_x_df2
}

Sys.time(); monthly_asdetectoutput_plot_df <- df_preparation2(monthly_asdetect_df); Sys.time()

plotting1 <- function (prepared_df, threshold_df, asdetectoutput_df ){
  cz <- c("central", "eastern", "southern")
  for (i in 1:length(cz)){
    mappalette<- c("red", "black")
  
    x_df <- prepared_df %>% filter(climate_zone == cz[[i]])
    y_df <- asdetectoutput_df %>% filter(climate_zone == cz[[i]])
    z_df <- threshold_df %>% filter(climate_zone == cz[[i]])
    unique_ID<- unique(x_df$NDVI_ID)
  
    for (j in 1:length(unique_ID)){
  
      df<- x_df %>% filter(NDVI_ID == unique_ID[[j]])
      ndvi_trajectory <- unique(df$NDVI_monthly_tr)
  
      lp_fig1<- df %>%  #lineplot #1 stl trend and original faceted by index
        ggplot(aes(x = as.factor(time)))+
        geom_line(aes(y = original, alpha = 0.4, group = IndexName )) + 
        geom_line(aes(y = trend, group = IndexName), lwd= 1.5, linetype="longdash")+
        theme_classic(base_size = 14)+
        theme (axis.text.x = element_blank())+
        ylab("index") + 
        xlab("Time (Year + Months)") + scale_alpha(guide = 'none')+
          ggtitle (paste0("monthly EVI = ", unique(df$EVI_monthly_tr), ";",
                    " monthly NDVI = ", unique(df$NDVI_monthly_tr))) + facet_grid(.~IndexName)
    
      lp_fig2<- df %>%  #lineplot #2 only stl trend faceted by index
        ggplot(aes(x = as.factor(time)))+
        geom_line(aes(y = trend, group = IndexName), lwd= 1.5, linetype="longdash")+
        theme_classic(base_size = 14)+
        theme (axis.text.x = element_blank())+
        ylab("index") + 
        xlab("Time (Year + Months)") + scale_alpha(guide = 'none')+ facet_grid(.~IndexName)
      
      df2<- y_df %>% filter(NDVI_ID == unique_ID[[j]])
      df3<- z_df %>% filter(NDVI_ID == unique_ID[[j]])
      
      lp_fig3<- df2 %>% #lineplot #3 of the asdetect stl trend with title being the dt_th and position
        ggplot(aes(x = as.factor(time)))+
        geom_line(aes(y = trend, group = IndexName )) + 
        theme_classic(base_size = 14)+
        theme (axis.text.x = element_blank())+
        ylab("index") + 
        xlab("Time (Year + Months)") + scale_alpha(guide = 'none')+
          ggtitle (paste0(" monthly EVI trend dtvalue = ", round(df3$EVI_trend_dt_value, digits = 2), ";",
                        " monthly NDVI trend dtvalue = ", round(df3$NDVI_trend_dt_value, digits = 2), ";",
                        " EVI position = ", df3$EVI_trend_as_pos, ";",
                        " NDVI position = ", df3$NDVI_trend_as_pos)) + 
        facet_grid(.~IndexName)
      
      lp_fig4<- df2 %>% #lineplot #4 of the asdetect original with title being the dt_th and position
        ggplot(aes(x = as.factor(time)))+
        geom_line(aes(y = original, group = IndexName )) + 
        theme_classic(base_size = 14)+
        theme (axis.text.x = element_blank())+
        ylab("index") + 
        xlab("Time (Year + Months)") + scale_alpha(guide = 'none')+
          ggtitle (paste0(" monthly EVI original dtvalue = ", round(df3$EVI_original_dt_value, digits = 2), ";",
                        " monthly NDVI original dtvalue = ", round(df3$NDVI_original_dt_value, digits = 2), ";",
                        " EVI position = ", df3$EVI_original_as_pos, ";",
                        " NDVI position = ", df3$NDVI_original_as_pos)) + 
        facet_grid(.~IndexName)
      
      lineplot_fig <- ggarrange(lp_fig1, lp_fig2, lp_fig3, lp_fig4, nrow = 4, ncol = 1)
    
      #map
      NDVIsample<- NDVIsample %>% 
        mutate(colorcode = case_when(NDVI_ID==unique_ID[[j]] ~"1", TRUE ~ "2"))
    
      map_points<-   
        tm_shape(cerrado) + tm_borders()+
        tm_shape (NDVIsample)+
        tm_dots(size =0.5, fill = "colorcode", 
              fill.scale = tm_scale_categorical(values = mappalette)) +
        tm_layout(legend.show = FALSE) 
      grobobject<- tmap_grob(map_points)
    
      final_fig<- ggarrange(grobobject, lineplot_fig, widths = c(1,2))
      ggsave(here("Outputs", "TrajectoryPlotting", "monthly_asdetect_results",
                paste0(round(df3$NDVI_trend_dt_value, digits = 2), "_", ndvi_trajectory,"_", unique_ID[[j]], ".png")), final_fig,
           dpi= 700, height = 80, width = 60, units= "cm")
      print (j)
    }
    print (i)
  }  
}

Sys.time(); plotting1(monthly_originaltrend_plot_df, monthly_asdetect_df2, monthly_asdetectoutput_plot_df); Sys.time()

```


#Step 5 - plotting distribution of detected threshold values from as detect in sample
```{r}
df_preparation3 <- function (threshold_df){
  x_df <- threshold_df %>% dplyr::select(c(NDVI_ID, NDVI_original_dt_value, NDVI_trend_dt_value,
                                                      EVI_original_dt_value, EVI_trend_dt_value))

  pivot_x_df <- pivot_longer(x_df, 5:8)
  pivot_x_df<- pivot_x_df %>% separate_wider_delim(cols = name, delim = "_", names = c("IndexName", "trend_original", "dt", "value2")) %>%
    dplyr::select(-c(dt, value2))
  pivot_x_df
    
}

Sys.time(); monthly_dtthreshold_df <- df_preparation3(monthly_asdetect_df2); Sys.time()

plotting2<- function (prepared_threshold_df){
  p <- ggplot(prepared_threshold_df, aes(value, fill = trend_original)) +
    geom_density(alpha = 0.5) +
    geom_vline(aes(xintercept = 0), color = "blue", linetype ="dashed", size =1) +
     facet_grid(.~IndexName)
  p <- p + scale_fill_brewer(palette = "Dark2") + 
    theme_classic(base_size = 14) + xlab ("detection value") +
    theme(legend.title=element_blank()) 
  p
}

Sys.time(); monthly_dt_distribution <- plotting2(monthly_dtthreshold_df); Sys.time()
monthly_dt_distribution <- monthly_dt_distribution + labs (title = "Monthly detection values distribution")
ggsave(here("Outputs", "TrajectoryPlotting", "monthly_asdetect_results","monthly_detection_distribution.png"), monthly_dt_distribution,
           dpi= 700, height = 50, width = 50, units= "cm")


```