```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)
library(pak)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = here("Scratch"), progress=10)
```


##Introduction
In this script I learn how to use the chewie package to download GEDI LiDAR. 
I follow the tutorial on the chewie github page https://github.com/Permian-Global-Research/chewie
First, I load the library, log in to my earthdata profile (NASA has the GEDI prints) and then set a cache folder, which I think is like a Scratch folder
```{r}
#pak::pkg_install("Permian-Global-Research/chewie")
library (chewie)

chewie_creds(force = T) # to set up your credentials
chewie_health_check() # to check your credentials and cache setup.
```

Then I load the climate zones & chapada national park shp and search for gedi footprints that intersect with them. 
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
plot(d_trans)
d_trans<- st_transform(d_trans, crs = 4326)

chapada<- st_read(here("Data", "Admin", "WDPA_WDOECM_Jan2024_Public_59_shp", "WDPA_WDOECM_Jan2024_Public_59_shp_0", "WDPA_WDOECM_Jan2024_Public_59_shp-polygons.shp"))
plot(chapada)

gedi_1b_search <- find_gedi(chapada,
  gedi_product = "1B",
  date_start = "2019-05-31"
)

print(gedi_1b_search) #134 footprints in chapada area
plot(gedi_1b_search)

chewie_show(
  gedi_1b_search,
  time_group = "year",
  zoom = 8
) #shows the footprints and the year in which the satellite went over the footprint

```


```{r}
gedi_1b_sf <- grab_gedi(gedi_1b_search) |>
  select(
   beam, rxwaveform, shot_number,txwaveform, longitude_bin0, latitude_bin0
  ) |>
  collect_gedi(gedi_find = gedi_1b_search)

```