```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(tidyverse)
library(here)
library(tictoc)
library(ggplot2)
library(ggpubr)
library(sf)
library(terra)
library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I create the Mapbiomas transitions that I am interested in and also LULC of interest

#Data input
I created a csv of the mapbiomas legend-https://brasil.mapbiomas.org/wp-content/uploads/sites/4/2024/08/ATBD-Collection-9-v2.docx.pdf
```{r}
mapbiomas<- read_csv(here("Data", "mapbiomas_legend_collection9.csv"))

```

#LULC interested
```{r}
lulc_interest<- mapbiomas %>% mutate(lulc_class= case_when(`Pixel value`==3~3,
                                                           `Pixel value`==4~4,
                                                           (`Pixel value`==5 |`Pixel value`==6| `Pixel value`==49|
                                                            `Pixel value`==11|`Pixel value`==32|
                                                              `Pixel value`==29|`Pixel value`==50|`Pixel value`==5|`Pixel value`==23)~1, #natural
                                                            `Pixel value`==12~12,
                                                           `Pixel value`==39~39,
                                                           `Pixel value`==15~15,
                                                           TRUE~2 #anthropic
                                                           ))
write_csv(lulc_interest, here("Data", "lulc_interest.csv"))

lulc_interestGEE<-lulc_interest %>% dplyr::select(c(`Pixel value`, lulc_class))
names(lulc_interestGEE)<- c("from", "to")
write_csv(lulc_interestGEE, here("Data", "lulc_GEEclassification.csv"))

```

#Transitions
First transition- forest formation to (i) savanna (ii) grassland (iii) pasture (iv) soy (v) plantation (vi) remaining
```{r}
forestformation_transition<- mapbiomas %>% filter(`Pixel value`!=3)
forestformation_transition<- forestformation_transition %>% mutate(from= "forest_formation", pixel_from=3)
names(forestformation_transition)<- c("pixel_to", "to","N_A", "from", "pixel_from")
forestformation_transition<- forestformation_transition %>% mutate(pixel_transition= (pixel_from*100)+pixel_to)

#Of interest
forestformation_transition %>% filter(to=="savanna_formation"|to=="grassland_formation"| to=="pasture"| to=="soybean"| to=="forest_plantation")

forestformation_transition<- forestformation_transition %>% mutate(transition_type= case_when(pixel_transition==304~ "forest_savanna",
                                                                                              pixel_transition==312~ "forest_grassland",
                                                                                              pixel_transition==315~ "forest_pasture",
                                                                                              pixel_transition==339~ "forest_soybean",
                                                                                              pixel_transition==309~ "forest_plantation",
                                                                                              TRUE~ "forest_to_other"))

other_forestformation_transition<- forestformation_transition %>% filter(transition_type=="forest_to_other")
other_forestformation_transition<- other_forestformation_transition %>% mutate(transition_type= case_when(N_A=="N"~"forest_to_natural",
                                                                                                          TRUE~"forest_to_anthropic"))

forestformation_transition<- forestformation_transition %>% filter(transition_type!="forest_to_other")
forestformation_transition <- forestformation_transition %>% rbind(other_forestformation_transition)

remove(other_forestformation_transition)

```

Second transition- savanna formation to (i)  forest (ii) grassland (iii) pasture (iv) soy (v) plantation (vi) remaining
```{r}

savannaformation_transition<- mapbiomas %>% filter(`Pixel value`!=4) 
savannaformation_transition<- savannaformation_transition %>% mutate(from= "savanna_formation", pixel_from=4)
names(savannaformation_transition)<- c("pixel_to", "to","N_A", "from", "pixel_from")
savannaformation_transition<- savannaformation_transition %>% mutate(pixel_transition= (pixel_from*100)+pixel_to)

#Of interest
savannaformation_transition %>% filter(to=="forest_formation"|to=="grassland_formation"| to=="pasture"| to=="soybean"| to=="forest_plantation")

savannaformation_transition<- savannaformation_transition %>% mutate(transition_type= case_when(pixel_transition==403~ "savanna_forest",
                                                                                              pixel_transition==412~ "savanna_grassland",
                                                                                              pixel_transition==415~ "savanna_pasture",
                                                                                              pixel_transition==439~ "savanna_soybean",
                                                                                              pixel_transition==409~ "savanna_plantation",
                                                                                              TRUE~ "savanna_to_other"))

other_savannaformation_transition<- savannaformation_transition %>% filter(transition_type=="savanna_to_other")
other_savannaformation_transition<- other_savannaformation_transition %>% mutate(transition_type= case_when(N_A=="N"~"savanna_to_natural",
                                                                                                          TRUE~"savanna_to_anthropic"))

savannaformation_transition<- savannaformation_transition %>% filter(transition_type!="savanna_to_other")
savannaformation_transition <- savannaformation_transition %>% rbind(other_savannaformation_transition)

remove(other_savannaformation_transition)

```


Third transition- pasture to (i) forest (ii) grassland (iii) savanna (iv)plantation (v) remaining. Note no transition to soy because I am not interested in inter-farming conversion
```{r}
pastureformation_transition<- mapbiomas %>% filter(`Pixel value`!=15) 
pastureformation_transition<- pastureformation_transition %>% mutate(from= "pasture", pixel_from=15)
names(pastureformation_transition)<- c("pixel_to", "to","N_A", "from", "pixel_from")
pastureformation_transition<- pastureformation_transition %>% mutate(pixel_transition= (pixel_from*100)+pixel_to)

#Of interest
pastureformation_transition %>% filter(to=="savanna_formation"|to=="grassland_formation"| to=="forest_formation"| to=="soybean"| to=="forest_plantation") #decided not to look at transtiion to soy

pastureformation_transition<- pastureformation_transition %>% mutate(transition_type= case_when(pixel_transition==1503~ "pasture_forest",
                                                                                              pixel_transition==1512~ "pasture_grassland",
                                                                                              pixel_transition==1504~ "pasture_savanna",
                                                                                              pixel_transition==1509~ "pasture_plantation",
                                                                                              TRUE~ "pasture_to_other"))

other_pastureformation_transition<- pastureformation_transition %>% filter(transition_type=="pasture_to_other")
other_pastureformation_transition<- other_pastureformation_transition %>% mutate(transition_type= case_when(N_A=="N"~"pasture_to_natural",
                                                                                                          TRUE~"pasture_to_anthropic"))

pastureformation_transition<- pastureformation_transition %>% filter(transition_type!="pasture_to_other")
pastureformation_transition <- pastureformation_transition %>% rbind(other_pastureformation_transition)

remove(other_pastureformation_transition)


```

Fourth transition- soy to (i) forest (ii) savanna (iii) grassland (iv) plantation and (v) remaining. Note not interested in inter farming conversion
```{r}
soyformation_transition<- mapbiomas %>% filter(`Pixel value`!=39) 
soyformation_transition<- soyformation_transition %>% mutate(from= "soy", pixel_from=39)
names(soyformation_transition)<- c("pixel_to", "to","N_A", "from", "pixel_from")
soyformation_transition<- soyformation_transition %>% mutate(pixel_transition= (pixel_from*100)+pixel_to)

#Of interest
soyformation_transition %>% filter(to=="savanna_formation"|to=="grassland_formation"| to=="forest_formation"| to=="soybean"| to=="forest_plantation")

soyformation_transition<- soyformation_transition %>% mutate(transition_type= case_when(pixel_transition==3903~ "soy_forest",
                                                                                              pixel_transition==3912~ "soy_grassland",
                                                                                              pixel_transition==3904~ "soy_savanna",
                                                                                              pixel_transition==3909~ "soy_plantation",
                                                                                              TRUE~ "soy_to_other"))

other_soyformation_transition<- soyformation_transition %>% filter(transition_type=="soy_to_other")
other_soyformation_transition<- other_soyformation_transition %>% mutate(transition_type= case_when(N_A=="N"~"soy_to_natural",
                                                                                                          TRUE~"soy_to_anthropic"))

soyformation_transition<- soyformation_transition %>% filter(transition_type!="soy_to_other")
soyformation_transition <- soyformation_transition %>% rbind(other_soyformation_transition)

remove(other_soyformation_transition)


```


Compile
```{r}
transition_df<- rbind(forestformation_transition,savannaformation_transition, pastureformation_transition,soyformation_transition)

transition_df<- transition_df %>% mutate(pixel_final= as.integer(as.factor(transition_type))) #26 unique transitions mapped
write_csv(transition_df, here("Data", "transitions_classification_complete.csv"))

#I changed the columns names to match what is needed in GEE
transition_dfGEE<- transition_df %>% dplyr::select(c(pixel_transition, pixel_final))
names(transition_dfGEE)<- c("from", "to")
write_csv(transition_dfGEE, here("Data", "transitions_GEEclassification.csv"))
```



