```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(tidyverse)
library(here)
library(tictoc)
library(ggplot2)
library(ggpubr)
library(sf)
library(terra)
library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I process the Mapbiomas anthropic and coarse vegetation types within Cerrado (forest, savanna and grass formation)

#Anthropic area
I processed Mapbiomass such that for each year, pixels are attributed as natural or anthropic using the definition of natural/anthropic from Mapbiomas in GEE (LULCChanges>Mapbiomas_NaturalAnthropic_Mask). I also create similar layers for forest, savanna and grass formations (LULCChanges>Mapbiomas_VegClassification_Mask). In GEE I also coarsen it to 1km. 

```{r}
#Border input
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

#Anthropic mask from GEE
anthropic<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask_LATEST.tif"))
Sys.time(); anthropic_crop<- terra::crop(anthropic, cerrado)
anthropic_mask<- terra::mask(anthropic_crop, vect(cerrado)); Sys.time()

library(magick) #for animation

neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))

map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()

tic(); agg_anthropic<- terra::aggregate(anthropic_mask, fact = 3 , fun= "median");toc()

tmap_mode("plot")
anthropic_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(agg_anthropic)+
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="Anthropic %") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()
  

tic();tmap_animation(anthropic_map, filename = here("Data", "Masks_to_define_pixelsofinterest","Mapbiomas_anthropic_mask", "annual_anthropic.gif"),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150) ;toc() 


remove(agg_anthropic)


#Grassland, forest and savanna from GEE
grass<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_natural_mask", "Mapbiomass_grass_LULC_mask.tif"))
Sys.time(); grass_crop<- terra::crop(grass, cerrado)
grass_mask<- terra::mask(grass_crop, vect(cerrado)); Sys.time()

savanna<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_natural_mask", "Mapbiomass_savanna_LULC_mask.tif"))
Sys.time(); savanna_crop<- terra::crop(savanna, cerrado)
savanna_mask<- terra::mask(savanna_crop, vect(cerrado)); Sys.time()

forest<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_natural_mask", "Mapbiomass_forest_LULC_mask.tif"))
Sys.time(); forest_crop<- terra::crop(forest, cerrado)
forest_mask<- terra::mask(forest_crop, vect(cerrado)); Sys.time()

remove(anthropic, grass, savanna, forest, anthropic_crop, grass_crop, savanna_crop, forest_crop)
```


##Annual rate of change
I fit a linear model to the time series of %anthropic area. 
In this chunk I write a lmfunction, which I then apply to understand trends in Mapbiomas anthropic, grass, tree and savanna formations. 
Details of the lm_func: pvalue (and rsq) do not exist when all pixel values have either (1)  values 0 ie remained 0% anthropic through time
(2) are all NAs as I masked all water and unobserved in GEE and (3) if there are a couple of NAs and then a couple of non NAs (this is weirdly
the case in some pixels that seem to be water based on my brief qualitative EDA in QGIS) and (4) the pixel value did not change through time
ie remained the exact same value ie 10% anthropic of 91% anthropic through time ie 0 slope. 

To deal with (3), I cam up with the rule that if a pixel has more than 50% NA values then I exclude the pixel ie resulting slope, rsq and pvalue
are all NA. And to deal with (4) I added an ifelse statement that if the slope is 0, then the pixel's pvalue and rsq are NA (because
it is rechnically non significant and of no interest)
```{r}
#install.packages("RobustLinearReg")
library(RobustLinearReg)

lm_func<- function(y) { 
  if(sum(is.na(y)) > 0.5 * length(y)){ 
    rep(NA, 3)} else {
        
        year<- 1:21
        mod <- RobustLinearReg::theil_sen_regression(y ~ year) 
        resid_slope <- coef(mod)[2]
        if(resid_slope==0){
          resid_r_sq <- NA
          resid_p_value <-NA
        } else{
          resid_r_sq <- broom::glance(mod)$r.squared[1]
          resid_p_value <- broom::tidy(mod)$p.value[2]
        }
    c(resid_slope, resid_r_sq, resid_p_value)
  } 
}

#Subsetting raster stack to only 2001-2021
Sys.time(); anthropic_rate<- terra::app(anthropic_mask[[17:37]], fun=lm_func); Sys.time()#1.5 hrs
#writeRaster(anthropic_rate, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_anthropic_mask", "anthropic_rate_slope_pvalue_rsq.tif"))

Sys.time(); grass_rate<- terra::app(grass_mask[[17:37]], fun=lm_func); Sys.time()#1 hour
#writeRaster(grass_rate, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "grass_rate_slope_pvalue_rsq.tif"))

Sys.time(); savanna_rate<- terra::app(savanna_mask[[17:37]], fun=lm_func); Sys.time()#2 hours
#writeRaster(savanna_rate, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "savanna_rate_slope_pvalue_rsq.tif"))

Sys.time(); forest_rate<- terra::app(forest_mask[[17:37]], fun=lm_func); Sys.time() #1.5 hours
#writeRaster(forest_rate, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "forest_rate_slope_pvalue_rsq.tif"))

remove(anthropic_mask, grass_mask, savanna_mask, forest_mask)

#maps of above
#I chose to make final maps of rate of change of each only when significant (p<0.05)

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

m <- c(-Inf, 0.05, 1,
       0.05, Inf, NA)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

anthropic_pvalue<- anthropic_rate[[3]] #third raster in stack
anthropic_significantpixels<- terra::classify(anthropic_pvalue, rclmat, include.lowest= TRUE)
#writeRaster(anthropic_significantpixels, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_anthropic_mask", "anthropic_signficantpvalue_mask.tif"))

sig_anthropicslope<- terra::mask(anthropic_rate[[1]], anthropic_significantpixels)
#writeRaster(sig_anthropicslope, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_anthropic_mask", "significant_anthropicslope.tif"))

remove(anthropic_pvalue, anthropic_significantpixels)

grass_pvalue<- grass_rate[[3]] #third raster in stack
grass_significantpixels<- terra::classify(grass_pvalue, rclmat, include.lowest= TRUE)
#writeRaster(grass_significantpixels, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "grass_signficantpvalue_mask.tif"))

sig_grassslope<- terra::mask(grass_rate[[1]], grass_significantpixels)
#writeRaster(sig_grassslope, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "significant_grassslope.tif"))

remove(grass_pvalue, grass_significantpixels)

savanna_pvalue<- savanna_rate[[3]] #third raster in stack
savanna_significantpixels<- terra::classify(savanna_pvalue, rclmat, include.lowest= TRUE)
#writeRaster(savanna_significantpixels, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "savanna_signficantpvalue_mask.tif"))

sig_savannaslope<- terra::mask(savanna_rate[[1]], savanna_significantpixels)
#writeRaster(sig_savannaslope, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "significant_savannaslope.tif"))

remove(savanna_pvalue, savanna_significantpixels)

forest_pvalue<- forest_rate[[3]] #third raster in stack
forest_significantpixels<- terra::classify(forest_pvalue, rclmat, include.lowest= TRUE)
#writeRaster(forest_significantpixels, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "forest_signficantpvalue_mask.tif"))

sig_forestslope<- terra::mask(forest_rate[[1]], forest_significantpixels)
#writeRaster(sig_forestslope, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "significant_forestslope.tif"))

remove(forest_pvalue, forest_significantpixels)

tmap_mode("plot")
anthropic_slope_sig_map<-#tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  #tm_shape(d_trans)+ tm_fill()+
  tm_shape (sig_anthropicslope)+
  tm_raster(style = "quantile", palette = "PiYG",n=6, legend.show = T) +
  tm_layout(legend.text.size = 1)

tmap_save(anthropic_slope_sig_map, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_anthropic_mask", "significant_anthropicslope.png"),
        height = 30, width = 30, units = "cm", dpi=700)

grass_slope_sig_map<-#tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  #tm_shape(d_trans)+ tm_fill()+
  tm_shape (sig_grassslope)+
  tm_raster(style = "quantile", palette = "Purples",n=3, legend.show = T) +
  tm_layout(legend.text.size = 1)

tmap_save(grass_slope_sig_map, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "significant_grassslope.png"),
        height = 30, width = 30, units = "cm", dpi=700)

savanna_slope_sig_map<-#tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  #tm_shape(d_trans)+ tm_fill()+
  tm_shape (sig_savannaslope)+
  tm_raster(style = "quantile", palette = "Oranges",n=4, legend.show = T) +
  tm_layout(legend.text.size = 1)

tmap_save(savanna_slope_sig_map, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "significant_savannaslope.png"),
        height = 30, width = 30, units = "cm", dpi=700)

forest_slope_sig_map<-#tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  #tm_shape(d_trans)+ tm_fill()+
  tm_shape (sig_forestslope)+
  tm_raster(style = "quantile", palette = "Greens",n=4, legend.show = T) +
  tm_layout(legend.text.size = 1)

tmap_save(forest_slope_sig_map, here("Data","Masks_to_define_pixelsofinterest","Mapbiomas_natural_mask", "significant_forestslope.png"),
        height = 30, width = 30, units = "cm", dpi=700)

remove(forest_slope_sig_map, savanna_slope_sig_map, grass_slope_sig_map, anthropic_slope_sig_map)
remove(m, rclmat, lm_func)
remove(d_trans, evi, cerrado)

```

##EDA of trajectory shapes with above rate of change of anthropic, tree, grass and savanna systems 
##using boxplots
The trajectory shapes that I use is the result using annual mean after STL.

There is too much data ie too many pixels aka sample size per trajectory shape class to analyse
And from the boxplot, I see absolute no difference in the %anthropic change, %grass change etc
between the different trajectory shapes. 

```{r}
#From lines 769-784 in 5_trajectoryshapeanalyses.Rmd
amtrend_map_results<- rast(here("Outputs", "Final_TrajShape_Models", "Annual", "am_trends11_trajresults.tif"))
#1- Linear decrease, 2- Linear increase, 3- Stable classes, 4- Step decrease, 5- step increase

df_amtrend_results<- terra::as.data.frame(amtrend_map_results, xy= TRUE, cells= TRUE, na.rm= TRUE)
df_anthropic_rate<- terra:: as.data.frame(anthropic_rate, xy=TRUE, cells= TRUE)
names(df_anthropic_rate)<- c("cell", "x", "y", "slope", "rsq", "pvalue")


#join_amtrend_siganthropicslope<- full_join(df_amtrend_results, df_anthropic_rate, by=join_by(cell))
#The n of df is an odd number ie 2079283, which is more than the usual number of pixels from
#trend results (by a lot)
#blank_amtrendresults<- join_amtrend_siganthropicslope %>% filter(is.na(x.x))
#shp_blankamtrends<- st_as_sf(blank_amtrendresults, coords = c("x.y", "y.y"), crs=crs(amtrend_map_results))
#st_write(shp_blankamtrends, here("Scratch", "blank_amtrends.shp"))
#I visually analysed the blank points in QGIS against the amtrends and anthropic rate (MapBiomas). 
#It seems like a lot of the NAs are either those that have been completely removed due to more than
#3 continuous NAs in a year, water bodies but are still in anthropic rate from Mapbiomas. So I have
#decided to trust the am trends results and focus on those pixels only. 
#remove(blank_amtrendresults, shp_blankamtrends)

join_amtrend_siganthropicslope<- left_join(df_amtrend_results, df_anthropic_rate, by=join_by(cell))
#blank_anthropicresults<- join_amtrend_siganthropicslope %>% filter(is.na(x.y))
#shp_blankanthropicresults<- st_as_sf(blank_anthropicresults, coords = c("x.x", "y.x"), crs=crs(anthropic_rate))
#st_write(shp_blankanthropicresults, here("Scratch", "blank_anthropic.shp"))
#there are 13842 ie 0.069% of pixels for which we have trajectory results, that we do not 
#have anthropic rate info. This is because these 0.06% are in the borders, which might have
#got cut off when processing Mapbiomas in GEE. And there are a couple of pixels in the 
#interior in water body areas which are missing maybe because of differing water masks applied
#by Mapbiomas (classification as water) and that applied in anisoEVI MODIS product. I am 
#willing to let go off this 0.069% instead of try to reprocess now. Even if I do not get to it later
#I can mention in manuscript 
#remove(blank_anthropicresults, shp_blankanthropicresults)

names(join_amtrend_siganthropicslope)<- c("cell", "x.amtrend", "y.amtrend", "trajshape", "x.anthropic", 
                                          "y.anthropic", "anthropic_slope", "anthropic_rsq", "anthropic_pvalue")

join_amtrend_siganthropicslope2<- join_amtrend_siganthropicslope %>% dplyr::select(c(trajshape, anthropic_slope, anthropic_pvalue)) %>%
      mutate(anthropic_pvalue=case_when(anthropic_pvalue<=0.05~"Significant",
                                        anthropic_pvalue>0.05~"NonSignificant",
                                        TRUE~"No Slope"))
join_amtrend_siganthropicslope2<- join_amtrend_siganthropicslope2 %>% mutate(trajshape=case_when(trajshape==1~"Linear_decrease",
                                                                                                 trajshape==2~"Linear_increase",
                                                                                                 trajshape==4~"Step decrease",
                                                                                                 trajshape==5~"Step increase",
                                                                                                 TRUE~"No trend"))
plot1a <- ggplot(join_amtrend_siganthropicslope2 %>% filter(anthropic_pvalue!="No Slope" & anthropic_slope>0),
               aes(x=factor(anthropic_pvalue), y=anthropic_slope, color = factor(trajshape)))+
  geom_boxplot(outlier.shape = NA)+
  ggtitle("Positive rate of change of %anthropic (Mapbiomas)")
plot1a<- plot1a + scale_color_brewer(palette = "Dark2")+ 
  theme_classic() +theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape" ,"amtrend_anthropic_postiveslope_boxplot.png"),
       plot1a,dpi = 700, height = 10, width=20, units = "cm")

plot1b <- ggplot(join_amtrend_siganthropicslope2 %>% filter(anthropic_pvalue!="No Slope" & anthropic_slope<0),
               aes(x=factor(anthropic_pvalue), y=anthropic_slope, color = factor(trajshape)))+
  geom_boxplot(outlier.shape = NA)+
  ggtitle("Negative rate of change of %anthropic (Mapbiomas)")
plot1b<- plot1b + scale_color_brewer(palette = "Dark2")+ 
  theme_classic() +theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape" ,"amtrend_anthropic_negativeslope_boxplot.png"),
       plot1b,dpi = 700, height = 10, width=20, units = "cm")

join_amtrend_siganthropicslope2 %>% filter(anthropic_pvalue!="No Slope" & anthropic_pvalue=="Significant") %>% 
  filter(anthropic_slope<0) %>% 
  ggplot(aes(x = anthropic_slope)) + 
  geom_histogram(aes(fill = anthropic_slope))
+ 
  guides(fill=FALSE)

plot1c<- barplot_df %>% filter(anthropic_pvalue=="Significant") %>%
  ggplot(aes(x = abs(anthropic_slope), fill = name)) + 
  stat_bin(aes(y=ifelse(fill == "Negative", -after_stat(count), after_stat(count)))) +
  facet_grid(~trajshape, scales= "free")
plot1c<- plot1c + scale_fill_brewer("Dark2") +theme_classic()


remove(df_anthropic_rate,join_amtrend_siganthropicslope)

###### repeating above for grass, tree and savanna formations
##grass
df_grass_rate<- terra:: as.data.frame(grass_rate, xy=TRUE, cells= TRUE)
names(df_grass_rate)<- c("cell", "x", "y", "slope", "rsq", "pvalue")

join_amtrend_siggrassslope<- left_join(df_amtrend_results, df_grass_rate, by=join_by(cell))
names(join_amtrend_siggrassslope)<- c("cell", "x.amtrend", "y.amtrend", "trajshape", "x.grass", 
                                          "y.grass", "grass_slope", "grass_rsq", "grass_pvalue")
join_amtrend_siggrassslope2<- join_amtrend_siggrassslope %>% dplyr::select(c(trajshape, grass_slope, grass_pvalue)) %>%
      mutate(grass_pvalue=case_when(grass_pvalue<=0.05~"Significant",
                                        grass_pvalue>0.05~"NonSignificant",
                                        TRUE~"No Slope"))
join_amtrend_siggrassslope2<- join_amtrend_siggrassslope2 %>% mutate(trajshape=case_when(trajshape==1~"Linear_decrease",
                                                                                                 trajshape==2~"Linear_increase",
                                                                                                 trajshape==4~"Step decrease",
                                                                                                 trajshape==5~"Step increase",
                                                                                                 TRUE~"No trend"))
plot2 <- ggplot(join_amtrend_siggrassslope2, aes(x=factor(trajshape), y=grass_slope, color = factor(grass_pvalue)))+
  geom_boxplot()+
  ggtitle("Rate of change of %grass (Mapbiomas)")
plot2<- plot2 + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+ 
  theme_classic() +theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape" ,"amtrend_grassslope_boxplot.png"),
       plot2,dpi = 700, height = 10, width=20, units = "cm")

remove(df_grass_rate,join_amtrend_siggrassslope)

###savanna
df_savanna_rate<- terra:: as.data.frame(savanna_rate, xy=TRUE, cells= TRUE)
names(df_savanna_rate)<- c("cell", "x", "y", "slope", "rsq", "pvalue")

join_amtrend_sigsavannaslope<- left_join(df_amtrend_results, df_savanna_rate, by=join_by(cell))
names(join_amtrend_sigsavannaslope)<- c("cell", "x.amtrend", "y.amtrend", "trajshape", "x.savanna", 
                                          "y.savanna", "savanna_slope", "savanna_rsq", "savanna_pvalue")
join_amtrend_sigsavannaslope2<- join_amtrend_sigsavannaslope %>% dplyr::select(c(trajshape, savanna_slope, savanna_pvalue)) %>%
      mutate(savanna_pvalue=case_when(savanna_pvalue<=0.05~"Significant",
                                        savanna_pvalue>0.05~"NonSignificant",
                                        TRUE~"No Slope"))
join_amtrend_sigsavannaslope2<- join_amtrend_sigsavannaslope2 %>% mutate(trajshape=case_when(trajshape==1~"Linear_decrease",
                                                                                                 trajshape==2~"Linear_increase",
                                                                                                 trajshape==4~"Step decrease",
                                                                                                 trajshape==5~"Step increase",
                                                                                                 TRUE~"No trend"))
plot3 <- ggplot(join_amtrend_sigsavannaslope2, aes(x=factor(trajshape), y=savanna_slope, color = factor(savanna_pvalue)))+
  geom_boxplot()+
  ggtitle("Rate of change of %savanna (Mapbiomas)")
plot3<- plot3 + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+ 
  theme_classic() +theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape" ,"amtrend_savannaslope_boxplot.png"),
       plot3,dpi = 700, height = 10, width=20, units = "cm")

remove(df_savanna_rate,join_amtrend_sigsavannaslope)

###forest
df_forest_rate<- terra:: as.data.frame(forest_rate, xy=TRUE, cells= TRUE)
names(df_forest_rate)<- c("cell", "x", "y", "slope", "rsq", "pvalue")

join_amtrend_sigforestslope<- left_join(df_amtrend_results, df_forest_rate, by=join_by(cell))
names(join_amtrend_sigforestslope)<- c("cell", "x.amtrend", "y.amtrend", "trajshape", "x.forest", 
                                          "y.forest", "forest_slope", "forest_rsq", "forest_pvalue")
join_amtrend_sigforestslope2<- join_amtrend_sigforestslope %>% dplyr::select(c(trajshape, forest_slope, forest_pvalue)) %>%
      mutate(forest_pvalue=case_when(forest_pvalue<=0.05~"Significant",
                                        forest_pvalue>0.05~"NonSignificant",
                                        TRUE~"No Slope"))
join_amtrend_sigforestslope2<- join_amtrend_sigforestslope2 %>% mutate(trajshape=case_when(trajshape==1~"Linear_decrease",
                                                                                                 trajshape==2~"Linear_increase",
                                                                                                 trajshape==4~"Step decrease",
                                                                                                 trajshape==5~"Step increase",
                                                                                                 TRUE~"No trend"))
plot4 <- ggplot(join_amtrend_sigforestslope2, aes(x=factor(trajshape), y=forest_slope, color = factor(forest_pvalue)))+
  geom_boxplot()+
  ggtitle("Rate of change of %forest (Mapbiomas)")
plot4 <- plot4 + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))+ 
  theme_classic() +theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape" ,"amtrend_forestslope_boxplot.png"),
       plot4, dpi = 700, height = 10, width=20, units = "cm")

remove(df_forest_rate,join_amtrend_sigforestslope)
```

##One way ANOVA and Tukey's tests 
Since the boxplots do not provide any prelim results, I do formal one way anova tests
to see if there is any significant difference between the trajectory shapes when 
considering %anthropic change etc. First I only consider pixels where there are
significant trends in %anthropic, %grass change etc

```{r}
one.way <- aov(anthropic_slope ~ as.factor(trajshape), data = significant)

```






