```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I complete analyses to understand where are the step changes and the step change date. 
(1) Data input of all trajectory results
(2) Processing above data- filtering to step change pixels only and processing date
(3) Fruther processing to exclude anthropic mask
(4) Results compilation in the form of barplot of area of total step pixels (increase/decrease) in each date of step change


#Step 1- data input
```{r}
threshold20_1km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_20.tif"))

input_file_path <- here("Outputs", "ModelSelection")
read_files_function <- function (index_name, annual_or_monthly_folder, file_name_with_rds){
  rds_file<- read_rds(paste0(input_file_path,"/", index_name,"/", annual_or_monthly_folder, "/", file_name_with_rds))
}

#NDVI
Sys.time(); annual_ndvi_swin11 <- read_files_function ("NDVI","annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_ndvi_swin11 <- read_files_function ("NDVI","monthly", "modelselection_swindow_11.rds"); Sys.time()

#EVI
Sys.time(); annual_evi_swin11 <- read_files_function ("EVI", "annual", "modelselection_swindow_11.rds"); Sys.time()
Sys.time(); monthly_evi_swin11 <- read_files_function ("EVI", "monthly", "modelselection_swindow_11.rds"); Sys.time()

remove(read_files_function)
```

#Step 2- processing step change pixels only across indices and date of step change
```{r}
results_df_function<- function (modelselection_df) {
  results_df <- modelselection_df %>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone, loc_brk)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~5, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                )) %>% 
    dplyr::select(c(x,y,pixelvalue,loc_brk))
}
Sys.time(); filtered_annual_ndvi_swin11 <- results_df_function(annual_ndvi_swin11) %>%
  filter((!is.na(loc_brk))); Sys.time()
Sys.time(); filtered_monthly_ndvi_swin11 <- results_df_function(monthly_ndvi_swin11) %>%
  filter((!is.na(loc_brk))); Sys.time()
Sys.time(); filtered_annual_evi_swin11 <- results_df_function(annual_evi_swin11) %>%
  filter((!is.na(loc_brk))); Sys.time()
Sys.time(); filtered_monthly_evi_swin11 <- results_df_function(monthly_evi_swin11) %>%
  filter((!is.na(loc_brk))); Sys.time()

remove(annual_ndvi_swin11, annual_evi_swin11, monthly_evi_swin11, monthly_ndvi_swin11)
remove(results_df_function)

insert_step_date_function1 <- function (filtered_df){
  filtered_df <- filtered_df %>% mutate(stepyear = (loc_brk + 2001))
  filtered_df
}
filtered_annual_ndvi_swin11_date <- insert_step_date_function1(filtered_annual_ndvi_swin11) 
filtered_annual_evi_swin11_date <- insert_step_date_function1(filtered_annual_evi_swin11)

insert_step_date_function2 <- function (filtered_df){
  filtered_df <- filtered_df %>% mutate(stepyear = (loc_brk %/% 12) + 2002,
                                        stepmonth = (loc_brk %% 12))
  filtered_df
}
filtered_monthly_ndvi_swin11_date <- insert_step_date_function2(filtered_monthly_ndvi_swin11) 
filtered_monthly_evi_swin11_date <- insert_step_date_function2(filtered_monthly_evi_swin11) 

remove(insert_step_date_function1, insert_step_date_function2)

```

#Step 3- rasterizing above filtered step pixels to extract date of step change, cropping/masking to CVNP
```{r}
#Step increase- pixelvalue = 5, step decrease- pixelvalue = 4
inc_annual_ndvi_swin11_date <- filtered_annual_ndvi_swin11_date %>% filter(pixelvalue ==5)
dec_annual_ndvi_swin11_date <- filtered_annual_ndvi_swin11_date %>% filter(pixelvalue ==4)
inc_annual_evi_swin11_date <- filtered_annual_evi_swin11_date %>% filter(pixelvalue ==5)
dec_annual_evi_swin11_date <- filtered_annual_evi_swin11_date %>% filter(pixelvalue ==4)
inc_monthly_ndvi_swin11_date <- filtered_monthly_ndvi_swin11_date %>% filter(pixelvalue ==5)
dec_monthly_ndvi_swin11_date <- filtered_monthly_ndvi_swin11_date %>% filter(pixelvalue ==4)
inc_monthly_evi_swin11_date <- filtered_monthly_evi_swin11_date %>% filter(pixelvalue ==5)
dec_monthly_evi_swin11_date <- filtered_monthly_evi_swin11_date %>% filter(pixelvalue ==4)

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()

rasterize_stepdf_function_year <- function (filered_date_df){
  cellindex_vector<- terra::vect(filered_date_df, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi , "stepyear", fun="max")
  masked_cellindex_raster<- terra::mask(cellindex_raster, threshold20_1km)
  masked_cellindex_raster 
}
#increases
Sys.time(); raster_inc_annual_ndvi_swin11_date <- rasterize_stepdf_function_year(inc_annual_ndvi_swin11_date); Sys.time()
Sys.time(); raster_inc_annual_evi_swin11_date <- rasterize_stepdf_function_year(inc_annual_evi_swin11_date); Sys.time()
Sys.time(); raster_inc_monthly_ndvi_swin11_date  <- rasterize_stepdf_function_year(inc_monthly_ndvi_swin11_date); Sys.time()
Sys.time(); raster_inc_monthly_evi_swin11_date  <- rasterize_stepdf_function_year(inc_monthly_evi_swin11_date); Sys.time()
#decreases
Sys.time(); raster_dec_annual_ndvi_swin11_date <- rasterize_stepdf_function_year(dec_annual_ndvi_swin11_date); Sys.time()
Sys.time(); raster_dec_annual_evi_swin11_date <- rasterize_stepdf_function_year(dec_annual_evi_swin11_date); Sys.time()
Sys.time(); raster_dec_monthly_ndvi_swin11_date  <- rasterize_stepdf_function_year(dec_monthly_ndvi_swin11_date); Sys.time()
Sys.time(); raster_dec_monthly_evi_swin11_date  <- rasterize_stepdf_function_year(dec_monthly_evi_swin11_date); Sys.time()

rasterize_stepdf_function_month <- function (filered_date_df){
  cellindex_vector<- terra::vect(filered_date_df, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi , "stepmonth", fun="max")
  masked_cellindex_raster<- terra::mask(cellindex_raster, threshold20_1km)
  masked_cellindex_raster 
}
Sys.time(); raster_inc_monthly_ndvi_swin11_date_month <- rasterize_stepdf_function_month(inc_monthly_ndvi_swin11_date); Sys.time()
Sys.time(); raster_inc_monthly_evi_swin11_date_month <- rasterize_stepdf_function_month(inc_monthly_evi_swin11_date); Sys.time()
Sys.time(); raster_dec_monthly_ndvi_swin11_date_month  <- rasterize_stepdf_function_month(dec_monthly_ndvi_swin11_date); Sys.time()
Sys.time(); raster_dec_monthly_evi_swin11_date_month <- rasterize_stepdf_function_month(dec_monthly_evi_swin11_date); Sys.time()

#CVNP shp
cvnp_shp<- st_read(here("Data", "Admin", "WDPA_WDOECM_Jan2024_Public_59_shp","WDPA_WDOECM_Jan2024_Public_59_shp_0", "WDPA_WDOECM_Jan2024_Public_59_shp-polygons.shp")) 

cvnp_crop_mask <- function(raster){
  x_crop<- terra::crop(raster, vect(cvnp_shp))
  x_mask<- terra::mask(x_crop, vect(cvnp_shp))
  x_mask
}
#increases
cvnp_year_inc_annual_ndvi <- cvnp_crop_mask(raster_inc_annual_ndvi_swin11_date)
cvnp_year_inc_annual_evi <- cvnp_crop_mask(raster_inc_annual_evi_swin11_date)
cvnp_year_inc_monthly_ndvi <- cvnp_crop_mask(raster_inc_monthly_ndvi_swin11_date)
cvnp_year_inc_monthly_evi <- cvnp_crop_mask(raster_inc_monthly_evi_swin11_date)
cvnp_month_inc_monthly_ndvi <- cvnp_crop_mask(raster_inc_monthly_ndvi_swin11_date_month)
cvnp_month_inc_monthly_evi <- cvnp_crop_mask(raster_inc_monthly_evi_swin11_date_month)
#decreases
cvnp_year_dec_annual_ndvi <- cvnp_crop_mask(raster_dec_annual_ndvi_swin11_date)
cvnp_year_dec_annual_evi <- cvnp_crop_mask(raster_dec_annual_evi_swin11_date)
cvnp_year_dec_monthly_ndvi <- cvnp_crop_mask(raster_dec_monthly_ndvi_swin11_date)
cvnp_year_dec_monthly_evi <- cvnp_crop_mask(raster_dec_monthly_evi_swin11_date)
cvnp_month_dec_monthly_ndvi <- cvnp_crop_mask(raster_dec_monthly_ndvi_swin11_date_month)
cvnp_month_dec_monthly_evi <- cvnp_crop_mask(raster_dec_monthly_evi_swin11_date_month)

```


#Step 4- compilation of results of step date- by year and by month
```{r}
year_inc<- c(cvnp_year_inc_annual_ndvi, cvnp_year_inc_annual_evi, cvnp_year_inc_monthly_ndvi, cvnp_year_inc_monthly_evi)
year_dec<- c(cvnp_year_dec_annual_ndvi, cvnp_year_dec_annual_evi, cvnp_year_dec_monthly_ndvi, cvnp_year_dec_monthly_evi)

remove(rasterize_stepdf_function_year,rasterize_stepdf_function_month, cvnp_crop_mask)

df_year_inc <- terra::as.data.frame(year_inc, cells=TRUE, xy=TRUE)
names(df_year_inc)[[4]] <- "NDVI(annual)" 
names(df_year_inc)[[5]] <- "EVI(annual)" 
names(df_year_inc)[[6]] <- "NDVI(monthly)"
names(df_year_inc)[[7]] <- "EVI(monthly)"
df_year_inc <- df_year_inc  %>% mutate(StepType="increase")
df_year_dec <- terra::as.data.frame(year_dec, cells=TRUE, xy=TRUE)
names(df_year_dec)[[4]] <- "NDVI(annual)" 
names(df_year_dec)[[5]] <- "EVI(annual)" 
names(df_year_dec)[[6]] <- "NDVI(monthly)"
names(df_year_dec)[[7]] <- "EVI(monthly)"
df_year_dec <- df_year_dec %>% mutate(StepType="decrease")
df_step<- bind_rows(df_year_inc, df_year_dec)
remove(df_year_dec, df_year_inc)
pivot_step<- df_step %>% pivot_longer(4:7)

barplot_df<- pivot_step %>% group_by(name, value, StepType) %>%
  summarise (count=n())
number_steppixels_df<- pivot_step %>% group_by(name) %>% #Total pixels that are Step in an index irrespective of year
  summarise(total_step_pixels=sum(!is.na(value)))
barplot_df<- barplot_df %>% left_join(number_steppixels_df)
barplot_df<- barplot_df %>% mutate(percentage_totalstep=(count/total_step_pixels)*100)
summary(barplot_df) #there are many Years missing because these pixels were
#classified as inc/dec in various other indices and not the specific one which is NA
barplot_df<- barplot_df %>% filter(!is.na(value))

output_file_path <- here("Outputs", "TrendsResults", "cvnp_results")
cvnp_step_year_barplot<- barplot_df %>% 
  ggplot(aes(x=percentage_totalstep, y=as.factor(value), fill= StepType)) +
  geom_bar(position="dodge", stat="identity") +
  theme_classic() +
  facet_grid(~name, switch = "x") +
  scale_fill_manual(values = c("#882255", "#88CCEE"))+
  xlab("% of step pixels")+ ylab ("Year of step change")+
  theme(strip.placement  = "outside",
        panel.spacing    = unit(0, "points"),
        strip.background = element_blank(),
        strip.text       = element_text(size = 12, angle = 90),
        axis.text.x = element_text(size = 12, angle = 90))
ggsave(paste0(output_file_path, "//", "cvnp_stepchange_year_allindices.png"),
       cvnp_step_year_barplot,dpi = 700, height = 20, width=20, units = "cm")

remove(year_dec, year_inc, df_year_dec, barplot_df, cvnp_step_year_barplot, df_step)


month_inc<- c(cvnp_month_inc_monthly_ndvi, cvnp_month_inc_monthly_evi)
month_dec<- c(cvnp_month_dec_monthly_ndvi, cvnp_month_dec_monthly_evi)
df_month_inc <- terra::as.data.frame(month_inc, cells=TRUE, xy=TRUE)
names(df_month_inc)[[4]] <- "NDVI(monthly)" 
names(df_month_inc)[[5]] <- "EVI(monthly)" 
df_month_inc <- df_month_inc  %>% mutate(StepType="increase")
df_month_dec <- terra::as.data.frame(month_dec, cells=TRUE, xy=TRUE)
names(df_month_dec)[[4]] <- "NDVI(monthly)" 
names(df_month_dec)[[5]] <- "EVI(monthly)" 
df_month_dec <- df_month_dec %>% mutate(StepType="decrease")
df_step<- bind_rows(df_month_inc, df_month_dec)
remove(df_month_dec, df_month_inc)
pivot_step<- df_step %>% pivot_longer(4:5)

barplot_df<- pivot_step %>% group_by(name, value, StepType) %>%
  summarise (count=n())
number_steppixels_df<- pivot_step %>% group_by(name) %>% #Total pixels that are Step in an index irrespective of year
  summarise(total_step_pixels=sum(!is.na(value)))
barplot_df<- barplot_df %>% left_join(number_steppixels_df)
barplot_df<- barplot_df %>% mutate(percentage_totalstep=(count/total_step_pixels)*100)
summary(barplot_df) #there are many Years missing because these pixels were
#classified as inc/dec in various other indices and not the specific one which is NA
barplot_df<- barplot_df %>% filter(!is.na(value))
barplot_df<- barplot_df %>% mutate(month = case_when(value==0 ~"Jan",
                                                     value==1~"Feb",
                                                     value==2~"Mar",
                                                     value==3~"Apr",
                                                     value==4~"May",
                                                     value==5~"Jun",
                                                     value==6~"Jul",
                                                     value==7~"Aug",
                                                     value==8~"Sep",
                                                     value==9~"Oct",
                                                     value==10~"Nov",
                                                     value==11~"Dec"))
barplot_df$month <- as.character(barplot_df$month) 
barplot_df$month <- factor(barplot_df$month, levels= unique(barplot_df$month))

cvnp_step_month_barplot<- barplot_df %>% 
  ggplot(aes(x=percentage_totalstep, y=as.factor(month), fill= StepType)) +
  geom_bar(position="dodge", stat="identity") +
  theme_classic() +
  facet_grid(~name, switch = "x") +
  scale_fill_manual(values = c("#882255", "#88CCEE"))+
  xlab("% of step pixels")+ ylab ("Year of step change")+
  theme(strip.placement  = "outside",
        panel.spacing    = unit(0, "points"),
        strip.background = element_blank(),
        strip.text       = element_text(size = 12, angle = 90),
        axis.text.x = element_text(size = 12, angle = 90))
ggsave(paste0(output_file_path, "//", "cvnp_stepchange_month_allindices.png"),
       cvnp_step_month_barplot,dpi = 700, height = 20, width=20, units = "cm")

```