```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I complete analyses to explore the relationship between the trajectories and vegetation physiognomies from Lewis et al., 2020
(1) Data input of all trajectory results
(2) Lewis et al data input and processing
(3) Relationships between trajectories (across indices) and above veg phyiognomy proportions within 1km pixel

#Step 1- data input of trajectory shapes
```{r}
cvnp_shp<- st_read(here("Data", "Admin", "WDPA_WDOECM_Jan2024_Public_59_shp","WDPA_WDOECM_Jan2024_Public_59_shp_0", "WDPA_WDOECM_Jan2024_Public_59_shp-polygons.shp")) 

input_file_path <- here("Outputs", "TrendsResults", "cvnp_results")

cvnp_annual_evi<- rast(here(input_file_path,"cvnp_evi_annual.tif"))
cvnp_annual_ndvi<- rast(here(input_file_path,"cvnp_ndvi_annual.tif"))
cvnp_monthly_evi<- rast(here(input_file_path,"cvnp_evi_monthly.tif"))
cvnp_monthly_ndvi<- rast(here(input_file_path,"cvnp_ndvi_monthly.tif"))

```

#Step 2 - veg physiognomies input and processing
```{r}
veg_phy<- rast(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "PNCV_vegMap_woText.tif"))
#0= campo umido, 1=campo sujo, 3=campu rupestre, 5=cerrado sensu stricto, 6=cerrado rupestre, 8=verada
#9=cerradao, 10=mata de galeria, 11-pasture, 12- water, 13- plantation, 14- agriculture, 15- nonvegetated

crop_veg_phy<- terra::crop(veg_phy, vect(cvnp_shp))
mask_veg_phy<- terra::mask(crop_veg_phy, vect(cvnp_shp))
remove(crop_veg_phy, veg_phy)

#Reclassifying to level 1 typology as all the campo are grasslands-1, both cerrado and verada are savannas-2
#cerradao and mata de galeria is woodland-3. Remainder are anthropic-4, water is NA
m<- rbind(c(0,1), c(1,1), c(3,1), c(5,2), 
          c(6,2), c(8,2), c(9,3), c(10,3), 
          c(11,4), c(12,NA), c(13,4), c(14,4), c(15,4))

lev1_phy<- terra::classify(mask_veg_phy, m)
remove(m)

coarsening_function <- function (classified_raster){
  x_1km<- terra::project(
    terra::segregate(classified_raster), cvnp_annual_evi,
    method = "average", res = res(cvnp_annual_evi)[1])
 
  x_1km_mask <- terra::mask(x_1km, cvnp_shp)
  x_1km_mask
}

coarse_lev1_phy <- coarsening_function(lev1_phy)
names(coarse_lev1_phy) <- c("grasslands", "savannas", "woodland", "anthropic")
remove(coarsening_function)
```

#Step 3a - eda grass, forest and savanna spread in pixels classified as each of the trajectories
```{r}
formation_trajectory <- function(trajectory_results_raster) {
  split_trajectory <- terra::segregate(trajectory_results_raster, other = NA)
  trajectory_names <- c("Lin_inc", "No_trend", "Step_dec", "Step_inc", "Quad_inc_acc", "Quad_inc_dec")
  df_for_histogram<- list()
  for (i in 1:nlyr(split_trajectory)){
    stack_raster <- c(split_trajectory[[i]], coarse_lev1_phy)
    names(stack_raster)<- c(trajectory_names[[i]], "grasslands", "savannas", "woodland", "anthropic")
    stack_df <- terra::as.data.frame(stack_raster, na.rm = T)
    df_for_histogram[[i]]<- stack_df
  }
  df_for_histogram
}

Sys.time(); monthly_ndvi_composition_trajectories <- formation_trajectory(cvnp_monthly_ndvi); Sys.time()
Sys.time(); monthly_evi_composition_trajectories <- formation_trajectory(cvnp_monthly_evi); Sys.time()
Sys.time(); annual_ndvi_composition_trajectories <- formation_trajectory(cvnp_annual_ndvi); Sys.time()
Sys.time(); annual_evi_composition_trajectories <- formation_trajectory(cvnp_annual_evi); Sys.time()


df_prep_function <- function (df_veg_trajectory){
  prepped_df_list <- list()
  for(i in 1:length(df_veg_trajectory)){
    x_pivot <- pivot_longer(df_veg_trajectory[[i]], 2:5)
    x_pivot <- x_pivot %>% mutate(TrajType = (names(x_pivot)[1])) %>%
      dplyr::select(c(-1))
    prepped_df_list[[i]] <- x_pivot
  }
  prepped_data <- bind_rows(prepped_df_list)
  prepped_data
}

Sys.time(); df_prep_monthly_ndvi_trajcomp <- df_prep_function( monthly_ndvi_composition_trajectories); Sys.time()
Sys.time(); df_prep_monthly_evi_trajcomp <- df_prep_function( monthly_evi_composition_trajectories); Sys.time()
Sys.time(); df_prep_annual_ndvi_trajcomp <- df_prep_function( annual_ndvi_composition_trajectories); Sys.time()
Sys.time(); df_prep_annual_evi_trajcomp <- df_prep_function( annual_evi_composition_trajectories); Sys.time()

output_file_path <- here("Outputs", "TrendsResults", "cvnp_results")
hist_plot_function <- function (plot_df){
  ggplot(plot_df, aes(x = value)) + 
   geom_histogram(position = "identity", bins = 10) + 
  theme_classic(base_size = 14) + scale_x_continuous(c(0,1)) +
  facet_grid(name ~ TrajType) + 
  theme(axis.title.x = element_blank(),
    #axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) + 
  theme(legend.position="none")
}
monthly_ndvi_hist_traj_comp <- hist_plot_function(df_prep_monthly_ndvi_trajcomp)
ggsave(paste0(output_file_path, "//", "monthly_ndvi_traj_composition.png"),
       monthly_ndvi_hist_traj_comp  ,dpi = 700, height = 50, width = 50, units = "cm")

monthly_evi_hist_traj_comp <- hist_plot_function(df_prep_monthly_evi_trajcomp)
ggsave(paste0(output_file_path, "//", "monthly_evi_traj_composition.png"),
       monthly_evi_hist_traj_comp  ,dpi = 700, height = 50, width = 50, units = "cm")

annual_ndvi_hist_traj_comp <- hist_plot_function(df_prep_annual_ndvi_trajcomp)
ggsave(paste0(output_file_path, "//", "annual_ndvi_traj_composition.png"),
       annual_ndvi_hist_traj_comp  ,dpi = 700, height = 50, width = 50, units = "cm")

annual_evi_hist_traj_comp <- hist_plot_function(df_prep_annual_evi_trajcomp)
ggsave(paste0(output_file_path, "//", "annual_evi_traj_composition.png"),
       annual_evi_hist_traj_comp  ,dpi = 700, height = 50, width = 50, units = "cm")

```


#Step 3b - eda of relationship between trajectories and veg physiognomies using boxplots
```{r}
veg_binning_function <- function (df, veg_formation_selected){
  df_veg_selected <- df %>% dplyr::select(c(veg_formation_selected, annual_EVI, annual_NDVI, monthly_EVI, monthly_NDVI))
  df_veg_selected <- df_veg_selected %>% rename("veg_selected"=veg_formation_selected)
  
  df_veg_selected <- df_veg_selected %>% mutate(bins = cut(veg_selected, breaks =seq(0,1,0.02), ordered_result = TRUE))
  
  number_pixels_bin<- df_veg_selected %>% group_by(bins) %>% summarise(count=n())
  
  pivot_df_veg_selected <- df_veg_selected %>% pivot_longer(2:5)
  
  veg_trajectory_df <- pivot_df_veg_selected %>%
    group_by(name, bins, value) %>%
    summarise(total_pixels = n())
  veg_trajectory_df <- inner_join(veg_trajectory_df, number_pixels_bin)
  veg_trajectory_df<- veg_trajectory_df %>% mutate(Proportion= (total_pixels/count)*100) #nrow 
  
  larger_bins<- rep(c("1_20%", "21_40%", "41_60%", "61_80%", "81_100%"),each=10)
  bin_lookup<- data.frame(bins=levels(veg_trajectory_df$bins), larger_bins=larger_bins)
  
   veg_trajectory_df <-  veg_trajectory_df  %>% left_join(bin_lookup)
   veg_trajectory_df <-  veg_trajectory_df  %>% mutate(bins=if_else(is.na(bins), "No specific veg", bins)) #there are pixels where %veg is 0 which I treat as a seperate category.
  veg_trajectory_df <- veg_trajectory_df %>% mutate(larger_bins=if_else(is.na(larger_bins), "No specific veg", larger_bins))

  veg_trajectory_df <- veg_trajectory_df %>% mutate(TrajType = case_when(value == 2~" Linear Increase",
                                                                         value == 3 ~"No trend",
                                                                         value == 4 ~"Step Decrease",
                                                                         value == 5 ~ "Step Increase",
                                                                         value == 8 ~ "Quadratic Increase (acc)",
                                                                         value == 9 ~ "Quadratic Increase (dec)"))
  veg_trajectory_df
  
}

Sys.time(); grass_df<- veg_binning_function(df_data_noNA, "percentage_grass"); Sys.time()
Sys.time(); savanna_df<- veg_binning_function(df_data_noNA, "percentage_savanna"); Sys.time()
Sys.time(); forest_df<- veg_binning_function(df_data_noNA, "percentage_woodland"); Sys.time()

veg_trajectory_plot_function <- function (binned_veg_df, title){
  plot<- ggplot(binned_veg_df, aes(x = larger_bins, y = Proportion, fill = TrajType)) + 
    geom_boxplot() + labs(title = title) +theme_classic(base_size = 14) +
    scale_fill_manual(values=c("#117733", "lightgrey", "#332288", "#DDCC77", "#882255", "#88CCEE"))
  plot
}

annual_evi_grass <- veg_trajectory_plot_function(grass_df %>% filter(name=="annual_EVI"), "Trajectory from EVI (mean annual trend) related to grass%")
annual_ndvi_grass <- veg_trajectory_plot_function(grass_df %>% filter(name=="annual_NDVI"), "Trajectory from NDVI (mean annual trend) related to grass%")
monthly_evi_grass <- veg_trajectory_plot_function(grass_df %>% filter(name=="monthly_EVI"), "Trajectory from EVI (monthly trend) related to grass%")
monthly_ndvi_grass <- veg_trajectory_plot_function(grass_df %>% filter(name=="monthly_NDVI"), "Trajectory from NDVI (monthly trend) related to grass%")

output_file_path <- here("Outputs", "TrendsResults", "cvnp_results")
grass_plots <- ggarrange(annual_evi_grass, annual_ndvi_grass, monthly_evi_grass, monthly_ndvi_grass, ncol =2, nrow=2 )
ggsave(paste0(output_file_path, "//", "trajectory_grassproportion_relationship.png"),
       grass_plots,dpi = 700, height = 80, width = 80, units = "cm")

annual_evi_savanna <- veg_trajectory_plot_function(savanna_df %>% filter(name=="annual_EVI"), "Trajectory from EVI (mean annual trend) related to savanna%")
annual_ndvi_savanna <- veg_trajectory_plot_function(savanna_df %>% filter(name=="annual_NDVI"), "Trajectory from NDVI (mean annual trend) related to savanna%")
monthly_evi_savanna <- veg_trajectory_plot_function(savanna_df %>% filter(name=="monthly_EVI"), "Trajectory from EVI (monthly trend) related to savanna%")
monthly_ndvi_savanna <- veg_trajectory_plot_function(savanna_df %>% filter(name=="monthly_NDVI"), "Trajectory from NDVI (monthly trend) related to savanna%")
savanna_plots <- ggarrange(annual_evi_savanna, annual_ndvi_savanna, monthly_evi_savanna, monthly_ndvi_savanna, ncol =2, nrow=2 )
ggsave(paste0(output_file_path, "//", "trajectory_savannaproportion_relationship.png"),
       savanna_plots,dpi = 700, height = 80, width = 80, units = "cm")

annual_evi_forest<- veg_trajectory_plot_function(forest_df %>% filter(name=="annual_EVI"), "Trajectory from EVI (mean annual trend) related to forest%")
annual_ndvi_forest <- veg_trajectory_plot_function(forest_df %>% filter(name=="annual_NDVI"), "Trajectory from NDVI (mean annual trend) related to forest%")
monthly_evi_forest <- veg_trajectory_plot_function(forest_df %>% filter(name=="monthly_EVI"), "Trajectory from EVI (monthly trend) related to forest%")
monthly_ndvi_forest <- veg_trajectory_plot_function(forest_df %>% filter(name=="monthly_NDVI"), "Trajectory from NDVI (monthly trend) related to forest%")
forest_plots <- ggarrange(annual_evi_forest, annual_ndvi_forest, monthly_evi_forest, monthly_ndvi_forest, ncol =2, nrow=2 )
ggsave(paste0(output_file_path, "//", "trajectory_forestproportion_relationship.png"),
       forest_plots,dpi = 700, height = 80, width = 80, units = "cm")

```


#Step 4- statistical analyses of relationship between percentage savanna and trajectory types (categorical variable) using anova
```{r}
ndvimonthlytraj_savanna_aov<- lm(percentage_savanna ~ as.factor(monthly_NDVI),
                                df_data_noNA %>% filter(monthly_NDVI == 5 | monthly_NDVI == 3)) 
summary(ndvimonthlytraj_savanna_aov)
```

No difference in group means of percentage grass in step increase and no trend trajectories. 
