```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I complete analyses to explore the relationship between the trajectories and pixel heterogeniety based on vegetation physiognomies from Lewis et al., 2020
(1) Data input of all trajectory results
(2) Lewis et al data input and processing to calculate pixel heterogeniety
(3) EDA relationships

#Step 1- data input of trajectory results
```{r}
cvnp_shp<- st_read(here("Data", "Admin", "WDPA_WDOECM_Jan2024_Public_59_shp","WDPA_WDOECM_Jan2024_Public_59_shp_0", "WDPA_WDOECM_Jan2024_Public_59_shp-polygons.shp")) 

input_file_path <- here("Outputs", "TrendsResults", "cvnp_results")

cvnp_annual_evi<- rast(here(input_file_path,"cvnp_evi_annual.tif"))
cvnp_annual_ndvi<- rast(here(input_file_path,"cvnp_ndvi_annual.tif"))
cvnp_monthly_evi<- rast(here(input_file_path,"cvnp_evi_monthly.tif"))
cvnp_monthly_ndvi<- rast(here(input_file_path,"cvnp_ndvi_monthly.tif"))

```

#Step 2 - veg physiognomies input and processing to calcualte pixel heterogeniety
```{r}
veg_phy<- rast(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "PNCV_vegMap_woText.tif"))
#0= campo umido, 1=campo sujo, 3=campu rupestre, 5=cerrado sensu stricto, 6=cerrado rupestre, 8=verada
#9=cerradao, 10=mata de galeria, 11-pasture, 12- water, 13- plantation, 14- agriculture, 15- nonvegetated

crop_veg_phy<- terra::crop(veg_phy, vect(cvnp_shp))
mask_veg_phy<- terra::mask(crop_veg_phy, vect(cvnp_shp))
remove(crop_veg_phy, veg_phy)

m_nativephysiognomies_only<- rbind(c(0,0), c(1,1), c(3,3), c(5,5), c(6,6), c(8,8),
                                   c(9,9), c(10,10), c(11,NA), c(12,NA), c(13,NA), c(14,NA), c(15, NA)) # Excluding anthropic physiognomies and calculating richness of native physiognomies
reclass_veg_phy <- terra::classify(mask_veg_phy, m_nativephysiognomies_only)
remove(m_nativephysiognomies_only)

coarsening_function <- function (classified_raster){
  x_1km<- terra::project(
    terra::segregate(classified_raster), cvnp_annual_evi,
    method = "average", res = res(cvnp_annual_evi)[1])
 
  x_1km_mask <- terra::mask(x_1km, cvnp_shp)
  x_1km_mask
}

coarse_reclass_veg_phy <- coarsening_function(reclass_veg_phy)
remove(coarsening_function)

library(vegan) ##Shannon's diversity index
shannon_diversity<- function (reclass_coarse_raster) {
    x_df<- as.data.frame(reclass_coarse_raster, cells=TRUE, xy=TRUE)
    x_diversity<- vegan::diversity(x_df %>% dplyr::select(-c(cell,x,y)), index="shannon")
    x_diversity<- as.data.frame(x_diversity)
    x_df<- bind_cols(x_df, x_diversity)
    trial_vector<- terra::vect(x_df %>% dplyr::select(c("cell", "x", "y", "x_diversity")), geom=c("x", "y"), crs="epsg:4326")
    trial_raster<- terra::rasterize(trial_vector, reclass_coarse_raster, "x_diversity", fun="max")
    trial_raster
}
Sys.time(); heterogen <- shannon_diversity (coarse_reclass_veg_phy);Sys.time()
remove(shannon_diversity)

```

#Step 3- eda relationships between trajectories and heterigeneity
```{r}

```