```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)

library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I complete analyses to explore the relationship between the trajectories and fire ie burned area 
(1) Data input of all trajectory results
(2) Burned area processing in CVNP
(3) EDA relationships

#Step 1- data input of trajectory results
```{r}
cvnp_shp<- st_read(here("Data", "Admin", "WDPA_WDOECM_Jan2024_Public_59_shp","WDPA_WDOECM_Jan2024_Public_59_shp_0", "WDPA_WDOECM_Jan2024_Public_59_shp-polygons.shp")) 

input_file_path <- here("Outputs", "TrendsResults", "cvnp_results")

cvnp_annual_evi<- rast(here(input_file_path,"cvnp_evi_annual.tif"))
cvnp_annual_ndvi<- rast(here(input_file_path,"cvnp_ndvi_annual.tif"))
cvnp_monthly_evi<- rast(here(input_file_path,"cvnp_evi_monthly.tif"))
cvnp_monthly_ndvi<- rast(here(input_file_path,"cvnp_ndvi_monthly.tif"))

```

#Step 2- processing burned area in CVNP using 1_fireprocessing.Rmd in drivres_processing
```{r}
burnedarea<- rast(here("Outputs", "OtherVariables", "Fire", "burnedarea_1km_2002_2021.tif"))

cvnp_burnedarea <- terra::crop(burnedarea, cvnp_shp)
cvnp_burnedarea <- terra::mask(cvnp_burnedarea, cvnp_shp)

#burned area and trajectory results in cvnp do not have the same extent and resolution, hence resampling
cvnp_burnedarea2 <- terra::resample(cvnp_burnedarea, cvnp_monthly_ndvi, method = "bilinear")
#writeRaster(cvnp_burnedarea2, here("Scratch", "cvnp_burnedarea.tif"))
```

#Step 3a - relationship between average burned area through period and trajectories
```{r}
mean_cvnp_burnedarea<- terra::app(cvnp_burnedarea2, fun ="mean") 

data_stack <- c(cvnp_annual_evi, cvnp_annual_ndvi,cvnp_monthly_evi,cvnp_monthly_ndvi, mean_cvnp_burnedarea)
data_df <- terra::as.data.frame(data_stack)

meanburnedarea_trajtype_boxplot_function <- function (traj_column, plot_title){
  df_filter <- data_df %>% dplyr::select(mean, traj_column)
  df_filter <- df_filter %>% mutate(TrajType = case_when(df_filter[[traj_column]] == 2~" Linear Increase",
                                                                         df_filter[[traj_column]] == 3 ~"No trend",
                                                                         df_filter[[traj_column]] == 4 ~"Step Decrease",
                                                                         df_filter[[traj_column]] == 5 ~ "Step Increase",
                                                                         df_filter[[traj_column]] == 8 ~ "Quadratic Increase (acc)",
                                                                         df_filter[[traj_column]] == 9 ~ "Quadratic Increase (dec)"))
  df_filter <- df_filter %>% filter(!is.na(df_filter[[traj_column]]))
  plot<- ggplot(df_filter, aes(x = TrajType, y = mean, fill = TrajType)) + 
    geom_boxplot() + ylab("average %burnedarea") +
    labs(title = plot_title) +
    theme_classic(base_size = 16) +
     theme(axis.text.x =element_blank(),
        axis.ticks.x =element_blank()) +
    scale_fill_manual(values=c("#117733", "lightgrey", "#332288", "#DDCC77", "#882255", "#88CCEE"))
  plot 
}

box_ndvi_monthly <- meanburnedarea_trajtype_boxplot_function("ndvi_swin11_monthly", "NDVI(monthly)")
box_evi_monthly <- meanburnedarea_trajtype_boxplot_function("evi_swin11_monthly", "EVI(monthly)")
box_ndvi_annual <- meanburnedarea_trajtype_boxplot_function("ndvi_swin11_annual", "NDVI(annual)")
box_evi_annual <- meanburnedarea_trajtype_boxplot_function("evi_swin11_annual", "EVI(annual)")

boxplot_plots <- ggarrange(box_ndvi_monthly, box_evi_monthly, box_ndvi_annual, box_evi_annual, ncol =2, nrow=2 )
output_file_path <- here("Outputs", "TrendsResults", "cvnp_results")
ggsave(paste0(output_file_path, "//", "boxplot_trajectory_meanburnedarea.png"),
       boxplot_plots ,dpi = 700, height = 80, width = 80, units = "cm")

remove(box_ndvi_annual, box_evi_annual, box_evi_monthly, box_ndvi_monthly, boxplot_plots, meanburnedarea_trajtype_boxplot_function)

```

#Step 3b - proportion of pixels in particular bins of burned area and the trajectory that the pixels belong to
```{r}
data_df <- data_df %>% mutate(bins = cut(mean, breaks = quantile(mean, na.rm= TRUE), right = FALSE, labels = FALSE, ordered_result = TRUE))
data_df <- data_df %>% mutate(bins = ifelse(is.na(bins), 4, bins))
#data_df <- data_df %>% mutate(bins = ifelse(is.na(bins), "No heterogeniety", bins))
number_pixels_bin <- data_df %>% group_by(bins) %>% summarise(count=n())

pivot_data_df <- data_df %>% pivot_longer(1:4)
burnedarea_trajectory_df <- pivot_data_df %>%
  group_by(name, bins, value) %>%
  summarise(total_pixels = n())

burnedarea_trajectory_df  <- inner_join(burnedarea_trajectory_df , number_pixels_bin)
burnedarea_trajectory_df <- burnedarea_trajectory_df %>% mutate(Proportion= (total_pixels/count)*100) #nrow 
burnedarea_trajectory_df <- burnedarea_trajectory_df %>% filter(!is.na(value))
burnedarea_trajectory_df <- burnedarea_trajectory_df %>% mutate(TrajType = case_when(value == 2~" Linear Increase",
                                                                         value == 3 ~"No trend",
                                                                         value == 4 ~"Step Decrease",
                                                                         value == 5 ~ "Step Increase",
                                                                         value == 8 ~ "Quadratic Increase (acc)",
                                                                         value == 9 ~ "Quadratic Increase (dec)"))
burnedarea_trajectory_df <- burnedarea_trajectory_df %>% mutate(bins = case_when (bins == 1~"1st quantile",
                                                                        bins == 2 ~ "2nd quantile",
                                                                        bins == 3 ~ "3rd quantile",
                                                                        bins == 4 ~ "4th quantile",
                                                                        TRUE ~"No heterogeniety"))
burnedarea_trajectory_df <- burnedarea_trajectory_df %>% mutate(name = case_when (name == "evi_swin11_annual" ~"EVI(annual)",
                                                                        name == "evi_swin11_monthly" ~"EVI(monthly)",
                                                                        name == "ndvi_swin11_annual" ~"NDVI(annual)",
                                                                        TRUE ~ "NDVI(monthly)"))

plot <- ggplot(data = burnedarea_trajectory_df, aes(x = bins, y= Proportion, fill = TrajType))+
    geom_bar(stat = "identity", position = position_dodge()) + 
  theme_classic(base_size = 16) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Average burned area")+
  scale_fill_manual(values=c("#117733", "lightgrey", "#332288", "#DDCC77", "#882255", "#88CCEE"))+
  facet_grid(~name)
ggsave(paste0(output_file_path, "//", "relationship_trajectory_meanburnedarea.png"),
       plot ,dpi = 700, height = 80, width = 80, units = "cm")

```

#Step 3c - in year/month of step change, the relationship between burned area and trajectories
From 2_stepchangepixel_analyses.Rmd in cvnp_analyses I know that most step increases happened
in 2018 for monthly EVI and NDVI trajectories and 2017 for annual EVI & NDVI trajectories
```{r}
years<- 2002:2021
names(cvnp_burnedarea2) <- years 

year_burnedarea_trajectory <- c(cvnp_annual_evi, cvnp_annual_ndvi,cvnp_monthly_evi,cvnp_monthly_ndvi, cvnp_burnedarea2)
data_df2 <- terra::as.data.frame(year_burnedarea_trajectory)

data_df2 <- data_df2 %>% dplyr::select(c(evi_swin11_annual, ndvi_swin11_annual, evi_swin11_monthly, ndvi_swin11_monthly, '2017', '2018'))


yearburnedarea_trajtype_boxplot_function <- function (traj_column, plot_title, step_changeyear){
  df_filter <- data_df2 %>% dplyr::select(c(step_changeyear, traj_column))
  df_filter <- df_filter %>% mutate(TrajType = case_when(df_filter[[traj_column]] == 2~" Linear Increase",
                                                                         df_filter[[traj_column]] == 3 ~"No trend",
                                                                         df_filter[[traj_column]] == 4 ~"Step Decrease",
                                                                         df_filter[[traj_column]] == 5 ~ "Step Increase",
                                                                         df_filter[[traj_column]] == 8 ~ "Quadratic Increase (acc)",
                                                                         df_filter[[traj_column]] == 9 ~ "Quadratic Increase (dec)"))
  df_filter <- df_filter %>% filter(!is.na(df_filter[[traj_column]]))
  plot<- ggplot(df_filter, aes(x = TrajType, y = step_changeyear, fill = TrajType)) + 
    geom_boxplot() + ylab("average %burnedarea") +
    labs(title = plot_title) +
    theme_classic(base_size = 16) +
     theme(axis.text.x =element_blank(),
        axis.ticks.x =element_blank()) +
    scale_fill_manual(values=c("#117733", "lightgrey", "#332288", "#DDCC77", "#882255", "#88CCEE"))
  plot 
}

year_box_ndvi_monthly <- yearburnedarea_trajtype_boxplot_function ("ndvi_swin11_monthly", "NDVI(monthly)", "2018")
year_box_evi_monthly <- yearburnedarea_trajtype_boxplot_function ("evi_swin11_monthly", "EVI(monthly)", "2018")
year_box_ndvi_annual <- yearburnedarea_trajtype_boxplot_function ("ndvi_swin11_annual", "NDVI(annual)", "2017")
year_box_evi_annual <- yearburnedarea_trajtype_boxplot_function ("evi_swin11_annual", "EVI(annual)", "2017")


```