```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I process the indices exported from GEE. 
(1) I mosaic the multiple files exported from GEE
(2) Apply the respective scaling factor mentioned on GEE page of each index
(3a) Align resolution, extent and crs across all three indices
(3b) Match time series ie start from Mar 2000 to Dec 2021
(4) Crop and mask to the three climate zones of Cattelan et al., 2024

#Step 1- make .vrt from all files exported from GEE
```{r}
vrt_function <- function (folder_address){
  x_vrt <- terra::vrt(folder_address)
  x_vrt 
}

ndvi_gee_folder_filepath <- list.files(path = here("Data", "Indices", "Cerrado_monthlyNDVI_GEE"), pattern= ".tif$", all.files=TRUE, full.names=TRUE)
evi_gee_folder_filepath <- list.files(path = here("Data", "Indices", "Cerrado_monthlyEVI_GEE"), pattern= ".tif$", all.files=TRUE, full.names=TRUE)
anisoevi_gee_folder_filepath <- list.files(path = here("Data", "Indices", "AnisoEVI_Cerrado_GEE"), pattern= ".tif$", all.files=TRUE, full.names=TRUE)

Sys.time(); ndvi_vrt <- vrt_function(ndvi_gee_folder_filepath); Sys.time()
Sys.time(); evi_vrt <- vrt_function(evi_gee_folder_filepath); Sys.time()
Sys.time(); anisoevi_vrt <- vrt_function(anisoevi_gee_folder_filepath); Sys.time()

remove(vrt_function)
remove(ndvi_gee_folder_filepath, evi_gee_folder_filepath, anisoevi_gee_folder_filepath)
```

#Step 2- application of scaling factor from respective veg index pages of GEE 
```{r}
scale_factor <- function( index_vrt, scalefactor){
  x_scaled_vrt <- index_vrt * scalefactor
}

Sys.time(); scaled_ndvi_vrt <- scale_factor(ndvi_vrt, 0.0001); Sys.time() #https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MOD13A3
Sys.time(); scaled_evi_vrt <- scale_factor(evi_vrt, 0.0001); Sys.time() #https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MOD13A3
Sys.time(); scaled_anisoevi_vrt <- scale_factor(anisoevi_vrt, 0.0001); Sys.time() #Documented in anisoEVI asset description

remove(scale_factor)
```

#Step 3a- align resolution, crs, extent
I match NDVI and EVI to anisoEVI. This is because anisoEVI is in WGS 1984 projection
with spatial resolution ~1km 
```{r}
reproject_function <- function (scaled_vrt){
  x_proj_scaled_vrt <- terra::project(scaled_vrt, scaled_anisoevi_vrt, method = "bilinear", align_only = TRUE)
}  

Sys.time() ; proj_scaled_ndvi_vrt <- reproject_function(scaled_ndvi_vrt); Sys.time()
Sys.time() ; proj_scaled_evi_vrt <- reproject_function(scaled_evi_vrt); Sys.time()

remove(reproject_function)
```

#Step 3b- match time series ie start from Mar 2000
```{r}
scaled_anisoevi_vrt #262 layers
proj_scaled_evi_vrt #263 layers Starts from Feb 2000 as per GEE page
proj_scaled_ndvi_vrt #263 layers Starts from Feb 2000 as per GEE page

proj_scaled_ndvi_vrt<- proj_scaled_ndvi_vrt[[2:263]]
proj_scaled_evi_vrt<- proj_scaled_evi_vrt [[2:263]]

```

#Step 4- extract indices in the climate zones from Catellan et al., 2024
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
#Climate zone vector number- 1= Central; 2 = Southern; 3 = Eastern

zone_function <- function (proj_scaled_vrt, index_name, climate_zone_vector_number){
  x_crop<- terra::crop(proj_scaled_vrt, vect(d_trans$geometry[[climate_zone_vector_number]]))
  x_mask<- terra::mask(x_crop, vect(d_trans$geometry[[climate_zone_vector_number]]))
  writeRaster(x_mask, here("Outputs", "Indices", "Climate_Zones_Indices", paste0(index_name, "_scaled_projected_", climate_zone_vector_number, ".tif")))
}


Sys.time(); zone_function(proj_scaled_ndvi_vrt, "NDVI", 1); Sys.time()
Sys.time(); zone_function(proj_scaled_ndvi_vrt, "NDVI", 2); Sys.time()
Sys.time(); zone_function(proj_scaled_ndvi_vrt, "NDVI", 3); Sys.time()

Sys.time(); zone_function(proj_scaled_evi_vrt, "EVI", 1); Sys.time()
Sys.time(); zone_function(proj_scaled_evi_vrt, "EVI", 2); Sys.time()
Sys.time(); zone_function(proj_scaled_evi_vrt, "EVI", 3); Sys.time()

Sys.time(); zone_function(scaled_anisoevi_vrt, "anisoEVI", 1); Sys.time()
Sys.time(); zone_function(scaled_anisoevi_vrt, "anisoEVI", 2); Sys.time()
Sys.time(); zone_function(scaled_anisoevi_vrt, "anisoEVI", 3); Sys.time()

```



