```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

#Introduction
In this script, I process the time series of all indices.
(1) Reading data
(2) Retain time series Jan 2002 - Dec 2021 due to irregularity in index values for NDVI and EVI in 2001.
(3) Exclude pixels which have a max of 4 consecutive missing monthly index values in a year.
(4) Linear interpolation using neighboring time steps for pixels with max one consecutive NA index values.
(5) For pixels with 2 or 3 continuous missing index values, use mean index values for that pixel from 
other months for which the month is missing.

#Step 1- reading in all data written in 1_indices_preparation.Rmd
```{r}
ndvi_files <- list.files(path = here("Outputs", "Indices", "Climate_Zones_Indices"), pattern= "*NDVI*", all.files=TRUE, full.names=TRUE)
ndvi_list <- lapply(ndvi_files, rast)
evi_files <- Sys.glob(here("Outputs", "Indices", "Climate_Zones_Indices", "EVI_scaled_projected_*.tif"))
evi_list <- lapply(evi_files, rast)
aniso_files <- list.files(path = here("Outputs", "Indices", "Climate_Zones_Indices"), pattern= "*aniso*", all.files=TRUE, full.names=TRUE)
aniso_list <- lapply(aniso_files, rast)

remove(ndvi_files, evi_files, aniso_files)

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
#Climate zone vector number- 1= Central; 2 = Southern; 3 = Eastern

```

#Step 2- retain time series Jan 2002 - Dec 2021
```{r}
retain_timeseries <- function (index_timeseries){
  retained_index_timeseries <- index_timeseries[[23:263]] #23 because 10 months of 2000 (starting Mar) and 12 months of 2001 are excluded
  retained_index_timeseries
}

Sys.time(); ndvi_central <- retain_timeseries(ndvi_list[[1]]); Sys.time()
Sys.time(); evi_central <- retain_timeseries(evi_list[[1]]); Sys.time()
Sys.time(); aniso_central <- retain_timeseries(aniso_list[[1]]); Sys.time()

Sys.time(); ndvi_southern <- retain_timeseries(ndvi_list[[2]]); Sys.time()
Sys.time(); evi_southern <- retain_timeseries(evi_list[[2]]); Sys.time()
Sys.time(); aniso_southern <- retain_timeseries(aniso_list[[2]]); Sys.time()

Sys.time(); ndvi_eastern <- retain_timeseries(ndvi_list[[3]]); Sys.time()
Sys.time(); evi_eastern <- retain_timeseries(evi_list[[3]]); Sys.time()
Sys.time(); aniso_eastern <- retain_timeseries(aniso_list[[3]]); Sys.time()

remove(retain_timeseries)
remove(ndvi_list, evi_list, aniso_list)
```


#Step 3-  exclude pixels which have a max of 4 consecutive missing monthly index values in a year
To do this I need to first prepare a mask of pixels with the above condition for each index in each zone
```{r}
### TRIAL- to find highest number of consecutive NA values
#x<- c(0.80, -0.32, NA, NA, 0.60,  0.13, -0.48, -0.09,  0.69, NA,  0.78,  0.89)
#y<- is.na(x)
#max(rle(y)$lengths[rle(y)$values]) #longest consecutive NA months i.e. 2 consectuive months of NA, 3 consecutive months of NA

#1. Calculate number of consecutive missing/NA index values per year
years<- seq(from=2002, to=2021, by=1)

maxconsecutiveNA<- function (index_completeraster, zone_polygon, file_path1, zone_name){
  annual_index_list<- list() #making a list of spatrasters, where each spatrasters is annual with 12 months rasters in the stack
  for(i in 1:length(years)){
  x<- subset(index_completeraster, grep(years[i], names(index_completeraster)))
  annual_index_list[[i]]<-x
  }
  
  no_consecutive_NA<- function (x){
    y<- is.na(x)
      if (sum(y)==0){
        x<-0
    } else {
    x<- max(rle(y)$lengths[rle(y)$values])
  }
  x
  }
  
  consecutive_NA_through_year_list<- list()
  tic(); for (i in 1:length(annual_index_list)){
    x<- terra::app(annual_index_list[[i]], fun= no_consecutive_NA)
    y<- terra::mask(x, vect(zone_polygon))
    consecutive_NA_through_year_list[i]<-y
    remove(y)
  }; toc() #1 hour  
  
  for (i in 1:length(consecutive_NA_through_year_list)){
  names(consecutive_NA_through_year_list[[i]])<- years[[i]]
  writeRaster(consecutive_NA_through_year_list[[i]],
              here("Outputs","Indices", file_path1, "Consecutive_missingNAs", paste0(zone_name, '_', years[[i]],"_throughtime.tif")))
  }  
}

Sys.time(); maxconsecutiveNA(ndvi_central, d_trans$geometry[[1]], file_path1 = "NDVI", zone_name = "central"); Sys.time()
Sys.time(); maxconsecutiveNA(ndvi_southern, d_trans$geometry[[2]], file_path1 = "NDVI", zone_name = "southern"); Sys.time()
Sys.time(); maxconsecutiveNA(ndvi_eastern, d_trans$geometry[[3]], file_path1 = "NDVI", zone_name = "eastern"); Sys.time()

Sys.time(); maxconsecutiveNA(evi_central, d_trans$geometry[[1]], file_path1 = "EVI", zone_name = "central"); Sys.time()
Sys.time(); maxconsecutiveNA(evi_southern, d_trans$geometry[[2]], file_path1 = "EVI", zone_name = "southern"); Sys.time()
Sys.time(); maxconsecutiveNA(evi_eastern, d_trans$geometry[[3]], file_path1 = "EVI", zone_name = "eastern"); Sys.time()

Sys.time(); maxconsecutiveNA(aniso_central, d_trans$geometry[[1]], file_path1 = "anisoEVI", zone_name = "central"); Sys.time()
Sys.time(); maxconsecutiveNA(aniso_southern, d_trans$geometry[[2]], file_path1 = "anisoEVI", zone_name = "southern"); Sys.time()
Sys.time(); maxconsecutiveNA(aniso_eastern, d_trans$geometry[[3]], file_path1 = "anisoEVI", zone_name = "eastern"); Sys.time()

remove(years, maxconsecutiveNA)

#2. Calculate max missing/NA index value across years
central_max_consecutiveNA_function<- function (index_path){
  raster_filepath <- list.files(path = here("Outputs", "Indices", index_path ,"Consecutive_missingNAs"), pattern= "^central", all.files=TRUE, full.names=TRUE)
  raster_list<-lapply(raster_filepath, rast)
  consecutivemissingNA_peryear<-rast(raster_list)
  max_consecutiveNAs<-  terra::app(consecutivemissingNA_peryear, fun= "max")
  writeRaster(max_consecutiveNAs, here("Outputs", "Indices", index_path, "Max_consecutive_missingNA", paste0("central", index_path, "maxconsecutiveNA.tif")))
}

southern_max_consecutiveNA_function<- function (index_path){
  raster_filepath <- list.files(path = here("Outputs", "Indices", index_path ,"Consecutive_missingNAs"), pattern= "^southern", all.files=TRUE, full.names=TRUE)
  raster_list<-lapply(raster_filepath, rast)
  consecutivemissingNA_peryear<-rast(raster_list)
  max_consecutiveNAs<-  terra::app(consecutivemissingNA_peryear, fun= "max")
  writeRaster(max_consecutiveNAs, here("Outputs", "Indices", index_path, "Max_consecutive_missingNA", paste0("southern", index_path, "maxconsecutiveNA.tif")))
}

eastern_max_consecutiveNA_function<- function (index_path){
  raster_filepath <- list.files(path = here("Outputs", "Indices", index_path ,"Consecutive_missingNAs"), pattern= "^eastern", all.files=TRUE, full.names=TRUE)
  raster_list<-lapply(raster_filepath, rast)
  consecutivemissingNA_peryear<-rast(raster_list)
  max_consecutiveNAs<-  terra::app(consecutivemissingNA_peryear, fun= "max")
  writeRaster(max_consecutiveNAs, here("Outputs", "Indices", index_path, "Max_consecutive_missingNA", paste0("eastern", index_path, "maxconsecutiveNA.tif")))
}

Sys.time(); central_max_consecutiveNA_function("NDVI"); Sys.time()
Sys.time(); central_max_consecutiveNA_function("EVI"); Sys.time() 
Sys.time(); central_max_consecutiveNA_function("anisoEVI"); Sys.time()

Sys.time(); southern_max_consecutiveNA_function("NDVI"); Sys.time()
Sys.time(); southern_max_consecutiveNA_function("EVI"); Sys.time()
Sys.time(); southern_max_consecutiveNA_function("anisoEVI"); Sys.time()

Sys.time(); eastern_max_consecutiveNA_function("NDVI"); Sys.time()
Sys.time(); eastern_max_consecutiveNA_function("EVI"); Sys.time()
Sys.time(); eastern_max_consecutiveNA_function("anisoEVI"); Sys.time()

remove(central_max_consecutiveNA_function, eastern_max_consecutiveNA_function, southern_max_consecutiveNA_function)

#3. Classification of max consecutive NA rasters to retain pixels with value < 4 
m <- rbind(c(0, 0), c(1, 1), c(2, 2), c(3, 3))

max_consecutive_mask<- function (index_path){
  raster_filepath <- list.files(path = here("Outputs", "Indices", index_path ,"Max_consecutive_missingNA"), pattern= ".tif$", all.files=TRUE, full.names=TRUE)
  raster_list<-lapply(raster_filepath, rast)
  for ( i in 1:length(raster_filepath)){
    rcx1 <- classify(raster_list[[i]], m, others= NA)
    writeRaster(rcx1, here("Outputs", "Indices", index_path, "Max_consecutive_missingNA", 
                              paste0("reclass_", strsplit(strsplit(strsplit(raster_filepath[[i]], "_")[[1]][3], "/")[[1]][2], ".tif")[[1]][1], ".tif")))
  }
print (paste0("Counter =",i))
} 

Sys.time(); max_consecutive_mask("anisoEVI"); Sys.time()
Sys.time(); max_consecutive_mask("EVI"); Sys.time()
Sys.time(); max_consecutive_mask("NDVI"); Sys.time()

remove(max_consecutive_mask, m)

#4. Extract all indices (for each cliamte zone) within above mask of max 4 consecutive missing pixels
final_mask_function<- function(time_series_raster_list, index_path){
  file_p<- list.files(here("Outputs", "Indices", index_path ,"Max_consecutive_missingNA"), pattern= "^reclass_", all.files=TRUE, full.names=TRUE)
  mask_raster<- lapply(file_p, rast)
  maxconsecutive_masked_list<- list()
  print ("Starting loop")
  for(i in 1:length(mask_raster)){
    maxconsecutive_masked_list[[i]]<- terra::mask(time_series_raster_list[[i]], mask_raster[[i]])
    writeRaster(maxconsecutive_masked_list[[i]], here("Outputs", "Indices", index_path, "Max_consecutive_missingNA",
                                            paste0("masked_",strsplit(strsplit(strsplit(file_p[[i]], "/")[[1]][11], "_")[[1]][2], ".tif")[[1]][1], ".tif"                                                                                                                )))
  }
  print (paste0("Counter=", i))
}

anisoEVI_list<- list(aniso_central, aniso_eastern, aniso_southern)
Sys.time(); final_mask_function(anisoEVI_list, "anisoEVI"); Sys.time()
ndvi_list<- list(ndvi_central, ndvi_eastern, ndvi_southern)
Sys.time(); final_mask_function(ndvi_list, "NDVI"); Sys.time()
evi_list<- list(evi_central, evi_eastern, evi_southern)
Sys.time(); final_mask_function(evi_list, "EVI"); Sys.time()

remove(final_mask_function)
remove(anisoEVI_list, ndvi_list, evi_list)
remove(evi_central, evi_eastern, evi_southern, 
       aniso_central, aniso_eastern, aniso_southern, 
       ndvi_central, ndvi_eastern,ndvi_southern)
```


#Step 4 - linear interpolation using neighboring time steps for pixels with max one consecutive NA index values
```{r}
linear_interpolation_function<- function(index_path){
  file_p<- list.files(here("Outputs", "Indices", index_path ,"Max_consecutive_missingNA"), 
                      pattern= "^masked_", all.files=TRUE, full.names=TRUE)
  masked_raster_list<- lapply(file_p, rast)
  
  impute_oneNA_function<- function(x){
    if (all(is.na(x))){
      return (x)
    } else {
      x<- imputeTS::na_interpolation(x, option = "linear", maxgap = 1)
      return(x)
    }
  }
    for (i in 1:length(masked_raster_list)){
    print ("Starting imputation")
    Sys.time(); imputed_raster<- terra::app(masked_raster_list[[i]], fun = impute_oneNA_function); Sys.time()
    Sys.time();writeRaster(imputed_raster, here("Outputs", "Indices", index_path, "Max_consecutive_missingNA",
                                 paste0("int_", strsplit(strsplit(strsplit(file_p[[i]], "/")[[1]][11], "_")[[1]][2], ".tif")[[1]][1], ".tif"))); Sys.time()
    print ("Check file on disk")
    print (paste0("Counter = ",i))
  }
}

Sys.time(); linear_interpolation_function("anisoEVI"); Sys.time()
Sys.time(); linear_interpolation_function("EVI"); Sys.time()
Sys.time(); linear_interpolation_function("NDVI"); Sys.time()

remove(linear_interpolation_function)
```


#Step 5 - for pixels with 2 or 3 continuous missing index values, use mean index values for that pixel from other months for which the month is missing. For example is April 2005 is NA for pixel A, then for that pixel use mean of remaining Aprils to fill the NA value

```{r}
convert_to_df<- function(raster_stack, index_path, climate_zone){
    trial_df<-terra::as.data.frame(raster_stack,xy=TRUE, cell=TRUE); 
    
    pivot_trial_df<- pivot_longer(trial_df, 4:243); Sys.time() 
    pivot_trial_df<- pivot_trial_df %>% mutate(
      Year = substring (name,1, 4) %>% as.numeric(),
      Month = substring (name, 6) %>% as.numeric());
    
    pivot_trial_df <- pivot_trial_df%>%
      group_by(cell, Month) %>%
      mutate(value_mean = mean(value, na.rm=T)) %>%
      mutate(value_int= ifelse(is.na(value), value_mean, value)) %>%
      select(-value_mean); Sys.time()
    
    write_rds(pivot_trial_df, here("Outputs", "Indices", index_path, "Max_consecutive_missingNA",
                                   paste0("meanfill_",index_path, "_", climate_zone, ".rds")))
}

raster_filepath <- list.files(path = here("Outputs", "Indices", "anisoEVI","Max_consecutive_missingNA"), pattern= "^int_", all.files=TRUE, full.names=TRUE)
raster_list<-lapply(raster_filepath, rast);
Sys.time(); convert_to_df(raster_list[[1]], "anisoEVI", "central"); Sys.time() 
Sys.time(); convert_to_df(raster_list[[2]], "anisoEVI", "eastern"); Sys.time() 
Sys.time(); convert_to_df(raster_list[[3]], "anisoEVI", "southern"); Sys.time() 
remove(raster_filepath, raster_list)


raster_filepath <- list.files(path = here("Outputs", "Indices", "NDVI","Max_consecutive_missingNA"), pattern= "^int_", all.files=TRUE, full.names=TRUE)
raster_list<-lapply(raster_filepath, rast);
Sys.time(); convert_to_df(raster_list[[1]], "NDVI", "central"); Sys.time() 
Sys.time(); convert_to_df(raster_list[[2]], "NDVI", "eastern"); Sys.time() 
Sys.time(); convert_to_df(raster_list[[3]], "NDVI", "southern"); Sys.time()
remove(raster_filepath, raster_list)
gc()

raster_filepath <- list.files(path = here("Outputs", "Indices", "EVI","Max_consecutive_missingNA"), pattern= "^int_", all.files=TRUE, full.names=TRUE)
raster_list<-lapply(raster_filepath, rast);
Sys.time(); convert_to_df(raster_list[[1]], "EVI", "central"); Sys.time() 
Sys.time(); convert_to_df(raster_list[[2]], "EVI", "eastern"); Sys.time() 
Sys.time(); convert_to_df(raster_list[[3]], "EVI", "southern"); Sys.time()
remove(raster_filepath, raster_list)

```



