```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)
library(gganimate)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I delineate the study area within the Cerrado biome. The aim is to 
exclude all anthropic area and focus only on native savannas. However, I need to figure out how to 
delineate anthropic and native areas.

#Data processing
I use the Mapbiomas vegetation product that is commonly used across Brazil and its biomes. 
This product is at 30m spatial resolution, derived from Landsat imagergy. In GEE, I completed 
processing this data such that the resulting product is a multiband image at 1km resolution,
where each band denotes a year (1985-2023) and each pixel value is a proportion. The 
proportion is the % of total Mapbiomas pixels within the 1km & 5km pixel that are anthropic (as per
Mapbiomas typology)
https://code.earthengine.google.com/f7e17500a39018685d022f079335ac51
(1) Read in files and crop/mask to Cerrado biome
(2) EDA and plots

#Step 1- reading all files and crop/mask to biome
```{r}
#Border input
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

crop_mask_function <- function (anthropic_mapbiomas_raster){
  x_crop <- terra::crop (anthropic_mapbiomas_raster, vect(cerrado))
  x_mask <- terra::mask (x_crop, vect(cerrado))
  x_mask
}

anthropic_1km<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask_1km.tif"))
Sys.time(); mask1km<- crop_mask_function (anthropic_1km); Sys.time()
anthropic_5km<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask_5km.tif"))
Sys.time(); mask5km<- crop_mask_function (anthropic_5km); Sys.time()
remove(anthropic_1km, anthropic_5km)
```

#Step 2- eda and plots
```{r}
#1. Animation of anthropic areas through time
library(magick) 
neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))
map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()

anthropic_time_animation_function<- function (anthropic_raster, km1_or_5_string) {
  mapbiomas_anthropic_animation_map<-tm_shape(neighbors, bbox = map_extent) + tm_borders() +
  tm_shape( mapbiomas_mask) +
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="%pixel that is anthropic") + 
              tm_layout (legend.position = c("left", "bottom")) +
              tm_facets(nrow=1, ncol=1, free.coords = FALSE) +
  tm_shape(d_trans) + tm_borders()

  tmap_animation(mapbiomas_anthropic_animation_map, filename = here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", 
                                                          paste0("Mapbiomass_anthropic_",  km1_or_5_string, "km_LULC_mask.gif")),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150) 
}
Sys.time(); anthropic_time_animation_function(mask1km, "1"); Sys.time()
Sys.time(); anthropic_time_animation_function(mask5km, "5"); Sys.time()

remove(anthropic_time_animation_function)

#2. % of climate zone that is anthropic (any proportion) per year in each climate zone
calc_df<- tibble(year= 1985:2023)

spatraster_into_list_function <- function (mapbiomas_mask){
  list_rasters<- list()
    for(i in 1:nlyr(mapbiomas_mask)){
      list_rasters[[i]]<- mapbiomas_mask[[i]]
    }
  list_rasters
}
  
Sys.time(); km1_list_raster<- spatraster_into_list_function(mask1km); Sys.time()
calc_df<- calc_df %>% mutate(anthropic_raster_1km =  km1_list_raster)
Sys.time(); km5_list_raster<- spatraster_into_list_function(mask5km); Sys.time()
calc_df<- calc_df %>% mutate(anthropic_raster_5km =  km5_list_raster)
remove(spatraster_into_list_function, km1_list_raster,km5_list_raster)

climatezone_crop_mask<- function(raster, climate_zone_vector_number){ #Climate zone vector number- 1= Central; 2 = Southern; 3 = Eastern
   x_crop<- terra::crop(raster, vect(d_trans$geometry[[climate_zone_vector_number]]))
   x_mask<- terra::mask(x_crop, vect(d_trans$geometry[[climate_zone_vector_number]]))
   x_mask
}
calc_df<- calc_df %>% mutate(climate_zone_vector_1 = 1)
calc_df<- calc_df %>% mutate(climate_zone_vector_2 = 2)
calc_df<- calc_df %>% mutate(climate_zone_vector_3 = 3)

Sys.time(); calc_df<- calc_df %>% mutate(central1km= purrr::map2(anthropic_raster_1km, climate_zone_vector_1, climatezone_crop_mask)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(central5km= purrr::map2(anthropic_raster_5km, climate_zone_vector_1, climatezone_crop_mask)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern1km= purrr::map2(anthropic_raster_1km, climate_zone_vector_2, climatezone_crop_mask)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern5km= purrr::map2(anthropic_raster_5km, climate_zone_vector_2, climatezone_crop_mask)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern1km= purrr::map2(anthropic_raster_1km, climate_zone_vector_3, climatezone_crop_mask)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern5km= purrr::map2(anthropic_raster_5km, climate_zone_vector_3, climatezone_crop_mask)); Sys.time()

calc_df <- calc_df %>% dplyr::select(-c(climate_zone_vector_1,climate_zone_vector_2,climate_zone_vector_3))

reclass_function<- function(regional_raster){
  reclass_raster <- classify(regional_raster, cbind(0, NA)) #pixels that have 0% anthropic is made NA
}
Sys.time(); calc_df<- calc_df %>% mutate(central_reclass1km= purrr::map(central1km, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(central_reclass5km= purrr::map(central5km, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_reclass1km= purrr::map(southern1km, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_reclass5km= purrr::map(southern5km, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_reclass1km= purrr::map(eastern1km, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_reclass5km= purrr::map(eastern5km, reclass_function)); Sys.time()

area_function<- function(reclass_raster){
    terra::cellSize(reclass_raster, mask=T, unit="km")
}
Sys.time(); calc_df<- calc_df %>% mutate(central_anthropic_area1km= purrr::map(central_reclass1km, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(central_anthropic_area5km= purrr::map(central_reclass5km, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_anthropic_area1km= purrr::map(southern_reclass1km, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_anthropic_area5km= purrr::map(southern_reclass5km, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_anthropic_area1km= purrr::map(eastern_reclass1km, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_anthropic_area5km= purrr::map(eastern_reclass5km, area_function)); Sys.time()

total_area_function<- function (area_raster){
  sum(terra::values(area_raster), na.rm = T)
}
Sys.time(); calc_df<- calc_df %>% mutate(central_anthropic_totalarea1km= purrr::map(central_anthropic_area1km, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(central_anthropic_totalarea5km= purrr::map(central_anthropic_area5km, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_anthropic_totalarea1km= purrr::map(southern_anthropic_area1km, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_anthropic_totalarea5km= purrr::map(southern_anthropic_area5km, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_anthropic_totalarea1km= purrr::map(eastern_anthropic_area1km, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_anthropic_totalarea5km= purrr::map(eastern_anthropic_area5km, total_area_function)); Sys.time()

area_anthropic_1km_df<- calc_df %>% dplyr::select(c(year, central_anthropic_totalarea1km, southern_anthropic_totalarea1km, eastern_anthropic_totalarea1km)) 
area_anthropic_5km_df<- calc_df %>% dplyr::select(c(year, central_anthropic_totalarea5km, southern_anthropic_totalarea5km, eastern_anthropic_totalarea5km))

plot_df_prep<- function(anthropic_area_df){
  anthropic_area_df[[2]] <- unlist(anthropic_area_df[[2]])
  anthropic_area_df[[3]] <- unlist(anthropic_area_df[[3]])
  anthropic_area_df[[4]] <- unlist(anthropic_area_df[[4]])
    
  anthropic_area_df<- anthropic_area_df %>% mutate(central_percent_anthropic= (anthropic_area_df[[2]]/d_trans$area_km[[1]])*100)
  anthropic_area_df<- anthropic_area_df %>% mutate(southern_percent_anthropic= (anthropic_area_df[[3]]/d_trans$area_km[[2]])*100)
  anthropic_area_df<- anthropic_area_df %>% mutate(eastern_percent_anthropic= (anthropic_area_df[[4]]/d_trans$area_km[[3]])*100)
  
  pivot_area_anthropic_df<- pivot_longer(anthropic_area_df %>% dplyr::select(year, central_percent_anthropic, southern_percent_anthropic, eastern_percent_anthropic), 2:4)
  pivot_area_anthropic_df
}

Sys.time(); pivot_area_anthropic_df_1km <- plot_df_prep(area_anthropic_1km_df); Sys.time()
Sys.time(); pivot_area_anthropic_df_5km <- plot_df_prep(area_anthropic_5km_df); Sys.time()
remove(plot_df_prep)

plot_function <- function (pivot_area_anthropic_df, km1_or_5_string){
  p<-ggplot(pivot_area_anthropic_df, aes(x=factor(year), y=value, group=name)) +
  geom_line(aes(color=name))+
  geom_point(aes(color=name))+
  ylab("% of pixels in region that are anthropic (any proportion)") +
  xlab("Year")
  p<- p+scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
  ggsave(here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", paste0("area_anthropic_", km1_or_5_string , "km.png")),p, 
       dpi=700, height = 12, width = 12, units = "cm")
}

Sys.time(); plot_function(pivot_area_anthropic_df_1km, "1"); Sys.time()
Sys.time(); plot_function(pivot_area_anthropic_df_5km, "5"); Sys.time()
remove(plot_function)
remove(pivot_area_anthropic_df_1km, pivot_area_anthropic_df_1km)

#3. Boxplot of non-zero pixels per year per climate zone for both resolutions
calc_df_1km<- calc_df %>% dplyr::select(c(year, central_reclass1km, southern_reclass1km, eastern_reclass1km))
calc_df_5km<- calc_df %>% dplyr::select(c(year, central_reclass5km, southern_reclass5km, eastern_reclass5km))

values_function<- function (reclass_raster){
  x<- terra::values(reclass_raster, dataframe= T, na.rm=T)
  names(x)<- "frac_anthropic"
  x
}
Sys.time(); calc_df_1km <- calc_df_1km %>% mutate(central_values= purrr::map(central_reclass1km, values_function)); Sys.time()
Sys.time(); calc_df_1km <- calc_df_1km %>% mutate(southern_values= purrr::map(southern_reclass1km, values_function)); Sys.time()
Sys.time(); calc_df_1km <- calc_df_1km %>% mutate(eastern_values= purrr::map(eastern_reclass1km, values_function)); Sys.time()

Sys.time(); calc_df_5km <- calc_df_5km %>% mutate(central_values= purrr::map(central_reclass5km, values_function)); Sys.time()
Sys.time(); calc_df_5km <- calc_df_5km %>% mutate(southern_values= purrr::map(southern_reclass5km, values_function)); Sys.time()
Sys.time(); calc_df_5km <- calc_df_5km %>% mutate(eastern_values= purrr::map(eastern_reclass5km, values_function)); Sys.time()

plot_df_prep_function<- function (reclass_df){
  boxplot_df<- reclass_df %>% dplyr::select(c(year, central_values, southern_values, eastern_values))
  pivot_boxplot_df<- pivot_longer(boxplot_df, 2:4)
  pivot_boxplot_df<- pivot_boxplot_df %>% unnest(value)
  pivot_boxplot_df
} 

Sys.time(); pivot_boxplot_df_1km <- plot_df_prep_function(calc_df_1km); Sys.time()
Sys.time(); pivot_boxplot_df_5km <- plot_df_prep_function(calc_df_5km); Sys.time()
remove(plot_df_prep_function)

plot_function<- function (pivot_boxplot_df, km1_or_5_string ){
  q<-ggplot(pivot_boxplot_df, aes(x=factor(year), y=frac_anthropic, group= factor(year))) +
  geom_boxplot(aes(color=name)) + facet_grid(name~.)+
  ylab("Anthropic Fraction (0% removed)") +xlab("Year")
  q<- q+scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
  ggsave(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", paste0("fraction_anthropic_", km1_or_5_string, "km_boxplot.png")),q, 
       dpi=700, height = 20, width = 20, units = "cm")
}

remove(plot_function,pivot_boxplot_df_1km, pivot_boxplot_df_1km)

#4. Histogram of %anthropic through time (animation)
histogram_df_1km<- calc_df_1km %>% dplyr::select(c(year, central_values, southern_values, eastern_values))
histogram_df_5km<- calc_df_5km %>% dplyr::select(c(year, central_values, southern_values, eastern_values))

plot_df_prep_function <- function (histogram_df){
  pivot_histogram_df<- pivot_longer(histogram_df, 2:4)
  pivot_histogram_df<- pivot_histogram_df %>% unnest(value)
  pivot_histogram_df
}

Sys.time(); pivot_histogram_1km <- plot_df_prep_function(histogram_df_1km); Sys.time()
Sys.time(); pivot_histogram_5km <- plot_df_prep_function(histogram_df_5km); Sys.time()
remove(plot_df_prep_function)

plot_function <- function (pivot_histogram_df, km1_or_5_string){
  r<- ggplot(pivot_histogram_df, aes(x=frac_anthropic, color=name, fill=name)) + 
  geom_histogram(fill="white") + 
  transition_time(year)+ facet_grid(name~.)+
  ylab("Pixel count") +xlab("%Anthropic")+
  labs(title = "Year: {frame_time}")
  
  r<- r + scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
  
  anim_save(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", paste0("histogram_%anthropic_", km1_or_5_string, "km_no0&1.gif")),r)
}

remove(plot_function, histogram_df_1km, histogram_df_5km, pivot_histogram_df_1km, pivot_histogram_5km)

```

According to Mattos et al., 2023, "we use an 80% valid threshold to avoid small 
sample sizes inside each larger pixel."I interpret this as 80% is the cut-off i.e. 
more than 80% of larger pixel is natural. 

I can make a decision, like Mattos et al, and retain pixels across all 3 zones
such that 20% is the maximum anthropic area. For now I chose three thresholds- 10, 20 & 30% 

```{r}
mask1km
mask5km

#1. Keep layers 2002-2021
names(mask1km)
names(mask5km)
tp_mask1km<- mask1km[[18:37]]
tp_mask5km<- mask5km[[18:37]]

#2. Exclude all pixels where pixel value in any year with pixel value   >0.1,0.2 & 0.3

threshold_function0.1<- function (x){
      ifelse(max(x)>0.1, NA, x) 
}

threshold_function0.2<- function (x){
      ifelse(max(x)>0.2, NA, x) 
}

threshold_function0.3<- function (x){
      ifelse(max(x)>0.3, NA, x) 
}

Sys.time(); thresh_10_1km <- terra::app (tp_mask1km, fun = threshold_function0.1); Sys.time()
#writeRaster(thresh_10_1km, here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_10.tif"))
Sys.time(); thresh_20_1km <- terra::app (tp_mask1km, fun = threshold_function0.2); Sys.time()
#writeRaster(thresh_20_1km, here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_20.tif"))
Sys.time(); thresh_30_1km <- terra::app (tp_mask1km, fun = threshold_function0.3); Sys.time()
#writeRaster(thresh_30_1km, here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_30.tif"))
Sys.time(); thresh_10_5km <- terra::app (tp_mask5km, fun = threshold_function0.1); Sys.time()
#writeRaster(thresh_10_5km, here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", "5km_thresholded_mapbiomas_mask_10.tif"))
Sys.time(); thresh_20_5km <- terra::app (tp_mask5km, fun = threshold_function0.2); Sys.time()
#writeRaster(thresh_20_5km, here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", "5km_thresholded_mapbiomas_mask_20.tif"))
Sys.time(); thresh_30_5km <- terra::app (tp_mask5km, fun = threshold_function0.3); Sys.time()
#writeRaster(thresh_30_5km, here("Data", "Masks_to_define_pixelsofinterest", "Mapbiomas_anthropic_mask", "5km_thresholded_mapbiomas_mask_30.tif"))

#Map of above thresholded rasters
neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))
map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()

threshold_map_function<- function (threshold_raster, title){
  tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(threshold_raster)+
  tm_raster(style = 'cont',
            palette= get_brewer_pal("Reds"), 
            n=5, breaks = seq(0,0.3,0.05), 
            title=title) + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()
}

map_thresh10_1km<- threshold_map_function(thresh_10_1km, "10% threshold")
map_thresh20_1km<- threshold_map_function(thresh_20_1km, "20% threshold")
map_thresh30_1km<- threshold_map_function(thresh_30_1km, "30% threshold")
all_maps_1km<- tmap_arrange(map_thresh10_1km, map_thresh20_1km, map_thresh30_1km)
tic();tmap_save(all_maps_1km, 
filename = here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "1km_threshold_Mapbiomass_anthropic_LULC_mask.png"),
               width = 900, height = 900, dpi = 150) ;toc() 
map_thresh10_5km<- threshold_map_function(thresh_10_5km, "10% threshold")
map_thresh20_5km<- threshold_map_function(thresh_20_5km, "20% threshold")
map_thresh30_5km<- threshold_map_function(thresh_30_5km, "30% threshold")
all_maps_5km<- tmap_arrange(map_thresh10_5km, map_thresh20_5km, map_thresh30_5km)
tic();tmap_save(all_maps_5km, 
filename = here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "5km_threshold_Mapbiomass_anthropic_LULC_mask.png"),
               width = 900, height = 900, dpi = 150) ;toc() 

remove(threshold_map_function, neighbors, map_extent, map_thresh10_1km, map_thresh20_1km, map_thresh30_1km,
       map_thresh10_5km, map_thresh20_5km, map_thresh30_5km, all_maps_1km, all_maps_5km)

```
