```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script I complete the STL decomposition to detrend the time series of all indices for 
each climate zone, for different stl parameters. This script is written such that it can be run
on the server.
(1) Read prepared rds files (prepared from indices_processing > 4_indices_processing.Rmd)
(2) Check of number of unique cells and number of rows per cell (idealy 240) 
(3) Setup core parallelization
(4) stl decomposition funciton
(5) Application of function

#Step 1 & 2- read prepared files and check unique pixels in each read file
```{r}
unique_cell_df<- data.frame(matrix(ncol = 2, nrow = 0))
x <- c("rds_file_number", "number_uniquecells")
colnames(unique_cell_df) <- x
remove(x)
  
index_file_path <- "/dungbeetle/home/tg505/Trisha/ecothresholds/Outputs/Indices/"

read_rds_check_function <- function (index_name, climate_zone_folder){
  file_list<- list.files (path = paste0(index_file_path, index_name,"/", "Max_consecutive_missingNA/", climate_zone_folder, "/"), pattern =".rds$", all.files = T, full.names = T)
  for (i in 1:length(file_list)){
    rds_file <- read_rds (file_list[[i]])
    unique_cell_df[i,1]<-i
    unique_cells<-length(unique(rds_file$cell))
    unique_cell_df[i,2]<- unique_cells
  }
  write_csv(unique_cell_df, paste0(index_file_path, index_name,"/", "Max_consecutive_missingNA/", climate_zone_folder, "/", "unique_cell_df.csv"))
}

Sys.time(); read_rds_check_function("NDVI", "meanfill_central_dfs"); Sys.time()
Sys.time(); read_rds_check_function("NDVI", "meanfill_eastern_dfs"); Sys.time()
Sys.time(); read_rds_check_function("NDVI", "meanfill_southern_dfs"); Sys.time()

Sys.time(); read_rds_check_function("EVI", "meanfill_central_dfs"); Sys.time()
Sys.time(); read_rds_check_function("EVI", "meanfill_eastern_dfs"); Sys.time()
Sys.time(); read_rds_check_function("EVI", "meanfill_southern_dfs"); Sys.time()

Sys.time(); read_rds_check_function("anisoEVI", "meanfill_central_dfs"); Sys.time()
Sys.time(); read_rds_check_function("anisoEVI", "meanfill_eastern_dfs"); Sys.time()
Sys.time(); read_rds_check_function("anisoEVI", "meanfill_southern_dfs"); Sys.time()

```

#Step 3- core parallelization setup
```{r}
numCores<- 16
library(foreach)
library(doParallel)

my.cluster<- parallel::makeCluster(
  numCores,
  type = "FORK",
  outfile = ""
)

doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered() #should be TRUE
foreach::getDoParWorkers() #should give numCores
```

#Step 4- stl decomposition function
```{r}

stl_function <- function (index_name, input_rds_climate_zone_folder, swin, swinfolder, climate_zone){
  file_list<- list.files (path = paste0(index_file_path, index_name,"/", "Max_consecutive_missingNA/", input_rds_climate_zone_folder, "/"), pattern =".rds$", all.files = T, full.names = T)
  file_list <- gtools::mixedsort(file_list)
  x<- split(file_list, ceiling(seq_along(file_list)/1000))
  
  for (i in 1:length(x)){
    rds_list<- lapply(x[[i]], read_rds)
    stl_decomposition<- function (df, swin){
    df_ts<- ts(df$value_int, start=c(2002,1), end=c(2021,12), frequency=12)
    stl_df<- stlplus::stlplus(df_ts, s.window = swin, s.degree = 1, t.degree = 1)
    df_needed<- (stl_df$data)
    df_needed<- df_needed %>% dplyr::select(-c(weights, sub.labels ))
    df_needed
  }
    message ("Stl function defined")
  
    results <- foreach(
      i= 1:length(rds_list),
      .combine = "rbind") %dopar% {
        rds_list[[i]] %>%
          group_by(cell) %>%
          nest() %>%
          mutate(stl_data = purrr::map(data, stl_decomposition, swin= swin)) %>%
          unnest(c(data,stl_data)) %>%
          dplyr::select(-c(Year, Month, raw))
      }
    message("Stl decomposition complete")
  
    write_rds(results, paste0(index_file_path, index_name, "/", "STL_Decomposition","/", swindowfolder,"/", "stl_", climate_zone, "_", i, ".rds"))#changed output file address manually
    message ("rds written out")
    }
}

Sys.time(); stl_function("NDVI", "meanfill_central_dfs", 7, "swindow_7", "central"); Sys.time()
Sys.time(); stl_function("NDVI", "meanfill_eastern_dfs", 7, "swindow_7", "eastern"); Sys.time()
Sys.time(); stl_function("NDVI", "meanfill_southern_dfs", 7, "swindow_7", "southern"); Sys.time()

Sys.time(); stl_function("EVI", "meanfill_central_dfs", 7, "swindow_7", "central"); Sys.time()
Sys.time(); stl_function("EVI", "meanfill_eastern_dfs", 7, "swindow_7", "eastern"); Sys.time()
Sys.time(); stl_function("EVI", "meanfill_southern_dfs", 7, "swindow_7", "southern"); Sys.time()

Sys.time(); stl_function("anisoEVI", "meanfill_central_dfs", 7, "swindow_7", "central"); Sys.time()
Sys.time(); stl_function("anisoEVI", "meanfill_eastern_dfs", 7, "swindow_7", "eastern"); Sys.time()
Sys.time(); stl_function("anisoEVI", "meanfill_southern_dfs", 7, "swindow_7", "southern"); Sys.time()

Sys.time(); stl_function("NDVI", "meanfill_central_dfs", "periodic", "swindow_periodic", "central"); Sys.time()
Sys.time(); stl_function("NDVI", "meanfill_eastern_dfs", "periodic", "swindow_periodic", "eastern"); Sys.time()
Sys.time(); stl_function("NDVI", "meanfill_southern_dfs", "periodic", "swindow_periodic", "southern"); Sys.time()

Sys.time(); stl_function("EVI", "meanfill_central_dfs", "periodic", "swindow_periodic", "central"); Sys.time()
Sys.time(); stl_function("EVI", "meanfill_eastern_dfs", "periodic", "swindow_periodic", "eastern"); Sys.time()
Sys.time(); stl_function("EVI", "meanfill_southern_dfs", "periodic", "swindow_periodic", "southern"); Sys.time()

Sys.time(); stl_function("anisoEVI", "meanfill_central_dfs", "periodic", "swindow_periodic", "central"); Sys.time()
Sys.time(); stl_function("anisoEVI", "meanfill_eastern_dfs", "periodic", "swindow_periodic", "eastern"); Sys.time()
Sys.time(); stl_function("anisoEVI", "meanfill_southern_dfs", "periodic", "swindow_periodic", "southern"); Sys.time()

Sys.time(); stl_function("NDVI", "meanfill_central_dfs", 11, "swindow_11", "central"); Sys.time()
Sys.time(); stl_function("NDVI", "meanfill_eastern_dfs", 11, "swindow_11", "eastern"); Sys.time()
Sys.time(); stl_function("NDVI", "meanfill_southern_dfs", 11, "swindow_11", "southern"); Sys.time()

Sys.time(); stl_function("EVI", "meanfill_central_dfs", 11, "swindow_11", "central"); Sys.time()
Sys.time(); stl_function("EVI", "meanfill_eastern_dfs", 11, "swindow_11", "eastern"); Sys.time()
Sys.time(); stl_function("EVI", "meanfill_southern_dfs", 11, "swindow_11", "southern"); Sys.time()

Sys.time(); stl_function("anisoEVI", "meanfill_central_dfs", 11, "swindow_11", "central"); Sys.time()
Sys.time(); stl_function("anisoEVI", "meanfill_eastern_dfs", 11, "swindow_11", "eastern"); Sys.time()
Sys.time(); stl_function("anisoEVI", "meanfill_southern_dfs", 11, "swindow_11", "southern"); Sys.time()


```