```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I extract sample points across the study area to be used for plotting time series and for "verification"
using asdetect R package.
(1) Read in final results of NDVI (rds format - 1km results, for stl=11 setting, only consider monthly time series results for now)
(2) Since above is rds format of all results, convert to raster and apply anthropogenic mask
(3) Sample 10 points in each trajectory result type in each zone
(4) Matching up above sampled points with other index results (monthly only)


#Step 1- read rds format NDVI results - stl setting swin =11, 1km resolution results of trajectories of monthly time series
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()
remove(mosaic_ndvi)

threshold20_1km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_20.tif"))

ndvi_swin11_monthly_results <- read_rds(here("Outputs", "ModelSelection", "NDVI", "monthly", "modelselection_swindow_11.rds"))
results_df <- ndvi_swin11_monthly_results %>% 
    dplyr::select(c(cell,x,y, model_order, shape_class, trend, loc_brk, climate_zone))

results_df <- results_df %>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~5, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                )) %>% 
    dplyr::select(c(cell, x,y,pixelvalue, climate_zone))


```

#Step 2- convert to raster format and apply anthropogenic mask
```{r}
climate_zones <- unique(results_df$climate_zone)

Sys.time(); for (i in 1:length(climate_zones)){
  x_climate_zone <- results_df %>% filter(climate_zone == climate_zones[[i]])
  unique_traj <- c(unique(x_climate_zone$pixelvalue))
  for ( j in 1:length(unique_traj)){
    selected_traj<- x_climate_zone %>% filter(pixelvalue == unique_traj[[j]])
    selected_traj_vector<- terra::vect(selected_traj, geom=c("x", "y"), crs="epsg:4326")
    selected_traj_raster<- terra::rasterize(selected_traj_vector, ndvi, "cell", fun="max")
    masked_selected_traj_raster<- terra::mask(selected_traj_raster, threshold20_1km)
    writeRaster(masked_selected_traj_raster, here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", climate_zones[[i]], paste0(unique_traj[[j]], "_traj_cellnumber.tif")))
  }
}; Sys.time()

remove(i, j, unique_traj)
remove(climate_zones)
remove(masked_selected_traj_raster, selected_traj, selected_traj_raster, selected_traj_vector)
```

#Step 3- sampling 10 points in each trajectory type per climate zone
```{r}

match_df_sample_function <- function (cz){
  cellraster_files<- list.files(path = here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", cz), pattern =".tif", all.files= TRUE, full.names = TRUE)
  cellraster_files<- lapply (cellraster_files, rast)
  
  pixels_studyarea_eachtraj <- list()
  for ( i in 1:length(cellraster_files)){
    raster_to_df <- terra::as.data.frame(cellraster_files[[i]], xy=TRUE)
    names(raster_to_df)[3]<- "cell"
    
    x_climate_zone <- results_df %>% filter(climate_zone == cz)
    pixels_studyarea_eachtraj[[i]] <- left_join (raster_to_df,  x_climate_zone, by= c("cell"= "cell"))
  }
  
  randomsample_studyarea_eachtraj <- list() #Sampling 10 random pixels in each trajctory ie each of the dfs in the list of dfs from previous step
  for (i in 1:length(pixels_studyarea_eachtraj)){
    randomsample_studyarea_eachtraj[[i]]<- pixels_studyarea_eachtraj[[i]] %>% 
                                            ungroup() %>%
                                              slice_sample(n = 10)
  }
  
  randomsample_studyarea_eachtraj<- bind_rows(randomsample_studyarea_eachtraj)
  randomsample_studyarea_eachtraj<- randomsample_studyarea_eachtraj %>% 
      select(x.x, y.x, cell, climate_zone, pixelvalue) %>% rename("x"="x.x", "y"="y.x")
  write.csv(randomsample_studyarea_eachtraj, here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", cz, "monthly_sampled_df.csv" ))
}

Sys.time(); match_df_sample_function("central"); Sys.time()
Sys.time(); match_df_sample_function("southern"); Sys.time()
Sys.time(); match_df_sample_function("eastern"); Sys.time()

remove(match_df_sample_function, ndvi_swin11_monthly_results, results_df)
remove(cellraster_files, pixels_studyarea_eachtraj, raster_to_df, i, cz, x_climate_zone)

#shp of sampled point
ndvi_sample_central<- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "central", "sampled_df.csv"))
ndvi_sample_southern<- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "southern", "sampled_df.csv"))
ndvi_sample_eastern<- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "eastern", "sampled_df.csv"))

rename_ndvisample<- function (ndvi_sample_df){ #renaming to avoid confusion with other indices
  ndvi_sample_df<- ndvi_sample_df %>% dplyr::select(-c(X))
  names(ndvi_sample_df)[3]<- "NDVI_cellID"
  names(ndvi_sample_df)[5]<- "NDVI_traj"
  ndvi_sample_df
}

ndvi_sample_central<- rename_ndvisample(ndvi_sample_central)
ndvi_sample_southern<- rename_ndvisample(ndvi_sample_southern)
ndvi_sample_eastern<- rename_ndvisample(ndvi_sample_eastern)


NDVIsample_pointshp<- function(ndvi_sample_df, climate_zone){
  point_vector<- st_as_sf(terra::vect(ndvi_sample_df, geom=c("x", "y"), crs="epsg:4326"))
  st_write(point_vector, here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", climate_zone, paste0(climate_zone, "_monthlytimeseries_NDVIsample.shp")))
} 

NDVIsample_pointshp(ndvi_sample_central, "central")
NDVIsample_pointshp(ndvi_sample_southern, "southern")
NDVIsample_pointshp(ndvi_sample_eastern, "eastern")
remove(ndvi_sample_central, ndvi_sample_eastern, ndvi_sample_southern, rename_ndvisample, NDVIsample_pointshp)
```

#Step 4- extracting cell index for above sampled NDVI points but in EVI monthly time series results
```{r}
step1<- function (modelselection_df, index_name, cz, ndvisample_pointshp){
  #df prep
  modelselection_df<- modelselection_df %>% 
    dplyr::select(c(cell,x,y, model_order, shape_class, trend, loc_brk, climate_zone))
  
  x <- modelselection_df %>% filter(climate_zone == cz)
  
  x<- x%>% ungroup() %>%
    dplyr::select(c(cell, x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                 model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                 model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                 model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                 model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                 model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                 model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                 model_order=="Step" & is.na(shape_class) & trend == "increase" ~5,
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                 )) %>%
    dplyr::select(-c(model_order, shape_class,trend)) #note Linear & Null stable is same category
  
  #raster of cell index
  cellindex_vector<- terra::vect(x, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi, "cell", fun="max")
  masked_cellindex_raster<- terra::mask(cellindex_raster, threshold20_1km)
  
  #sample NDVI points in above raster toi get cellID
  newindex_sample_cellID<- terra::extract(masked_cellindex_raster, ndvisample_pointshp, method ="simple", bind = TRUE)
  
  #for above cellID get trajectory from new index
  newindex_sample_cellID<- as.data.frame(newindex_sample_cellID)
  
    
  newindex_sample_celltraj<- left_join(newindex_sample_cellID,x, by= c("max"="cell"))
 
  
  names(newindex_sample_celltraj)[4]<-  paste0(index_name, "_ID")
  names(newindex_sample_celltraj)[8]<- paste0(index_name, "_tr")
  
  write.csv(newindex_sample_celltraj, here("Outputs", "TrajectoryPlotting", paste0("NDVI_", index_name, "_", cz,  "_monthlytimeseries_sampledpoints.csv")))
}  

#Application of above function  
ndvi_sample_centralshp<- st_read(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "central", "central_monthlytimeseries_NDVIsample.shp"))
ndvi_sample_southernshp<- st_read(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "southern", "southern_monthlytimeseries_NDVIsample.shp"))
ndvi_sample_easternshp<- st_read(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "eastern", "eastern_monthlytimeseries_NDVIsample.shp"))

#EVI
modelselection_evi<- read_rds(here("Outputs", "ModelSelection", "EVI", "monthly", "modelselection_swindow_11.rds"))

Sys.time(); step1(modelselection_evi, "EVI", "central", ndvi_sample_centralshp); Sys.time()
Sys.time(); step1(modelselection_evi, "EVI", "southern", ndvi_sample_southernshp); Sys.time()
Sys.time(); step1(modelselection_evi, "EVI", "eastern", ndvi_sample_easternshp); Sys.time()

```



