```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I plot the time series of sampled plots for different indices, monthly & annual and for stl setting swin=11. 
(1) Read sampled points shp and extract trajectories for respective index using annual time series (after stl decomposition)
(2) Compile dataframe with monthly, annual trajectories and time series for all indices for sampled points
(3) Plotting monthly time series results
(4) For same points, plotting annual time series results

#Step 1- read sampled points and extract results of trajectories obtained from annual time series (after stl decomposition)
```{r}
input_file_path <- here("Outputs", "TrajectoryPlotting", "NDVI_trajectories")

read_sampledpoints_shp <- function (cz){
  x_shp <- st_read(here(input_file_path, paste0(cz, "/", cz, "_monthlytimeseries_NDVIsample.shp")))
  x_shp
}

central_shp <- read_sampledpoints_shp("central")
eastern_shp <- read_sampledpoints_shp("eastern")
southern_shp <- read_sampledpoints_shp("southern")

#annual trajectory rasters
annual_ndvi <- rast(here("Outputs", "TrendsResults", "results_rasters", "annual", "ndvi_annual_swin11.tif"))
annual_evi <- rast(here("Outputs", "TrendsResults", "results_rasters", "annual", "evi_annual_swin11.tif"))

annual_extract_function <- function (annual_index_raster, index_name){
  list_shp <- list()
  sampled_points <- list(central_shp, eastern_shp, southern_shp)
  cz <- c("central", "eastern", "southern")
  
  for (i in 1:length(sampled_points)){
  annual_extract <- terra::extract(annual_index_raster, sampled_points[[i]], xy= TRUE, bind = TRUE)
  df_annual_extract <- annual_extract %>% st_drop_geometry() %>% as.data.frame()
  list_shp[[i]] <- df_annual_extract
  }
  
  list_shp
}

Sys.time(); annual_ndvi_list <- annual_extract_function(annual_ndvi, "NDVI"); Sys.time()
annual_ndvi_list <- do.call("rbind", annual_ndvi_list)
annual_ndvi_list  <- annual_ndvi_list %>% rename(c("NDVI_monthly_tr"= "NDVI_tr", "NDVI_annual_tr"= "max"))
Sys.time(); annual_evi_list <- annual_extract_function(annual_evi, "EVI"); Sys.time()
annual_evi_list <- do.call("rbind", annual_evi_list)
annual_evi_list  <- annual_evi_list %>% rename(c("NDVI_monthly_tr"= "NDVI_tr", "EVI_annual_tr"= "max"))

#join complete annual trajectory results df for sampled points
annual_trajectories_sampled_points <- left_join (annual_ndvi_list, annual_evi_list)

#join above annual complete df to monthly trajectory results of EVI and anisoEVI (stl setting swin =11) 
monthly_evi_swin11 <- rast (here("Outputs", "TrendsResults", "results_rasters", "monthly", "evi_monthly_swin11.tif"))

monthly_extract_function <- function (monthly_index_raster, index_name){
  list_shp <- list()
  sampled_points <- list(central_shp, eastern_shp, southern_shp)
  cz <- c("central", "eastern", "southern")
  
  for (i in 1:length(sampled_points)){
  monthly_extract <- terra::extract(monthly_index_raster, sampled_points[[i]], xy= TRUE, bind = TRUE)
  df_monthly_extract <- monthly_extract %>% st_drop_geometry() %>% as.data.frame()
  list_shp[[i]] <- df_monthly_extract
  }
  
  list_shp
}
Sys.time(); monthly_evi_list <- monthly_extract_function(monthly_evi_swin11, "EVI"); Sys.time()
monthly_evi_list <- do.call("rbind",  monthly_evi_list)
monthly_evi_list  <-  monthly_evi_list %>% dplyr::select(-c(NDVI_tr, x, y)) %>%
  rename(c("EVI_monthly_tr"= "max"))


#complete trajectory results
annual_trajectories_sampled_points <- left_join(annual_trajectories_sampled_points, monthly_evi_list, by = join_by(NDVI_ID))
annual_trajectories_sampled_points <- annual_trajectories_sampled_points %>% dplyr::select (-c(clmt_zn.y))


remove(annual_ndvi_list, annual_evi_list, annual_extract_function, annual_ndvi, annual_evi)
remove(input_file_path)
remove(monthly_evi_list, monthly_extract_function, monthly_evi_list, monthly_evi_swin11)
remove(read_sampledpoints_shp)
```

#Step 2- compile df
```{r}
################################################# Preparation of NDVI df with NDVI time series with monthly & annual trajectory results from above chunk

input_file_path <- here("Outputs", "TrajectoryPlotting")
ndvi_file_list <- list.files(path = paste0(input_file_path, "/", "NDVI_trajectories"), pattern = paste0("*.csv*"), all.files = T, full.names = T)
ts_ndvi_sampled <- lapply(ndvi_file_list, read.csv)
cz <- c("central", "eastern", "southern")
include_cz <- function (ts_df_list){
  ts_cz_list <- list()
  for (i in 1: length(cz)){
    x <- ts_df_list[[i]] %>% mutate(climate_zone = cz[[i]])
    ts_cz_list[[i]] <- x
  }
  ts_cz_list
}

ts_cz_ndvi_sampled <- include_cz(ts_ndvi_sampled)
remove(ts_ndvi_sampled)
ts_cz_ndvi_sampled <- do.call("rbind", ts_cz_ndvi_sampled)

ts_cz_ndvi_sampled <- ts_cz_ndvi_sampled  %>% dplyr::select(-c(x, y, value, seasonal, remainder)) %>%
    rename(c("NDVI_ID" = "cell",
             "NDVI_original" = "value_int", 
             "NDVI_trend" = "trend",
             "time"= "name"))

#join above time series to complete trajectory results
cz_list<- list()
for (i in 1:length(cz)){
  cz_complete_traj <- annual_trajectories_sampled_points %>% filter(clmt_zn.x == cz[[i]])
  cz_ts <- ts_cz_ndvi_sampled  %>% filter(climate_zone == cz[[i]])
  unique_cells <- unique(cz_ts$NDVI_ID)
  joined_list <- list()
    for (j in 1:length(unique_cells)){
      x <- cz_complete_traj %>% filter(NDVI_ID == unique_cells[[j]])
      y <- cz_ts %>% filter(NDVI_ID == unique_cells[[j]])
      df_join <- left_join (x,y, by = join_by("NDVI_ID"))
      joined_list[[j]]<- df_join
    }
  joined_df <- do.call("rbind", joined_list)
  joined_df <- joined_df %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, 
                                             time, NDVI_monthly_tr, EVI_monthly_tr, NDVI_annual_tr, EVI_annual_tr, NDVI_original, NDVI_trend))
  cz_list[[i]]<- joined_df
}

remove(x, y, df_join, i, j, joined_list, joined_df, unique_cells, cz_complete_traj, cz_ts)
remove(input_file_path, ndvi_file_list)

ndvi_ts_complete_tr <- do.call("rbind", cz_list)
remove(ts_cz_ndvi_sampled)
remove(cz_list)
################################################# Including time series of remaining indices
input_file_path <- here("Outputs", "TrajectoryPlotting")
ndvi_evi_file_list <- list.files(path = input_file_path, pattern = paste0("*.csv*"), all.files = T, full.names = T)
ndvi_evi_file_list <- lapply(ndvi_evi_file_list, read.csv)
ndvi_evi_df <- do.call("rbind", ndvi_evi_file_list)
ndvi_evi_df <- ndvi_evi_df %>% dplyr::select(c("NDVI_ID", "EVI_ID", "climate_zone"))

remove(ndvi_evi_file_list)
remove(input_file_path)

input_file_path <- here("Outputs", "TrajectoryPlotting")
evi_file_list <- list.files(path = paste0(input_file_path, "/", "EVI_trajectories"), pattern = paste0("*.csv*"), all.files = T, full.names = T)
ts_evi_sampled <- lapply(evi_file_list, read.csv)

ts_cz_evi_sampled <- include_cz(ts_evi_sampled)
remove(ts_evi_sampled)
ts_cz_evi_sampled <- do.call("rbind", ts_cz_evi_sampled)

ts_cz_evi_sampled <- ts_cz_evi_sampled  %>% dplyr::select(-c(x, y, value, seasonal, remainder, name, X)) %>%
    rename(c("EVI_ID" = "cell",
             "EVI_original" = "value_int", 
             "EVI_trend" = "trend"))

remove(evi_file_list, input_file_path)

#getting NDVI_ID & EVI ID for the df immediately above that has the EVI ts
cz_list<- list()
for (i in 1:length(cz)){
  cz_ts <- ts_cz_evi_sampled  %>% filter(climate_zone == cz[[i]])
  id_df <- ndvi_evi_df %>% filter(climate_zone == cz[[i]])
  unique_cells <- unique(cz_ts$EVI_ID)
  joined_list <- list()
    for (j in 1:length(unique_cells)){
      x <- cz_ts %>% filter(EVI_ID == unique_cells[[j]])
      y <- id_df %>% filter(EVI_ID == unique_cells[[j]])
      df_join <- left_join (x,y, by = join_by("EVI_ID"))
      joined_list[[j]]<- df_join
    }
  joined_df <- do.call("rbind", joined_list)
  joined_df <- joined_df %>% dplyr::select(c(NDVI_ID, EVI_ID, climate_zone.x, EVI_original, EVI_trend))
  cz_list[[i]]<- joined_df
}

remove(cz_ts, id_df, unique_cells, joined_list, x, y, df_join, i, j, joined_list, joined_df)
evi_ts <- do.call("rbind", cz_list)
remove(ts_cz_evi_sampled)
remove(cz_list)

cz_list<- list()
for (i in 1:length(cz)){
  cz_complete_traj <- ndvi_ts_complete_tr %>% filter(climate_zone == cz[[i]])
  cz_evi_ts <- evi_ts  %>% filter(climate_zone.x == cz[[i]])
  unique_cells <- unique(cz_evi_ts$NDVI_ID)
  joined_list <- list()
    for (j in 1:length(unique_cells)){
      x <- cz_complete_traj %>% filter(NDVI_ID == unique_cells[[j]])
      y <- cz_evi_ts %>% filter(NDVI_ID == unique_cells[[j]])
      df_join <- cbind (x,y)
      joined_list[[j]]<- df_join
    }
  joined_df <- do.call("rbind", joined_list)
  names(joined_df)[1] <- "dup_NDVI_ID"
  joined_df <- joined_df %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time,
                                             NDVI_monthly_tr, EVI_monthly_tr, NDVI_annual_tr, EVI_annual_tr, 
                                             NDVI_original, NDVI_trend, EVI_original, EVI_trend))
  cz_list[[i]]<- joined_df
}

remove(cz_complete_traj, cz_evi_ts, unique_cells, joined_list, x, y, df_join, i, j, joined_list, joined_df)
ndvi_evi_ts_complete_tr <- do.call("rbind", cz_list)
remove(cz_list)

write.csv(ndvi_evi_ts_complete_tr, here("Outputs", "TrajectoryPlotting", "ndvi_evi_ts_trajectories.csv"))

remove(cz, central_shp, eastern_shp, southern_shp, ndvi_evi_df, ndvi_ts_complete_tr, annual_trajectories_sampled_points, evi_ts)
remove(include_cz)
```

#Step 3- plotting using monthly time series results
```{r}
ndvi_evi_ts_complete_tr <- read.csv(here("Outputs", "TrajectoryPlotting", "ndvi_evi_ts_trajectories.csv"))

input_file_path <- here("Outputs", "TrajectoryPlotting", "NDVI_trajectories")

read_sampledpoints_shp <- function (cz){
  x_shp <- st_read(here(input_file_path, paste0(cz, "/", cz, "_monthlytimeseries_NDVIsample.shp")))
  x_shp
}

central_shp <- read_sampledpoints_shp("central")
eastern_shp <- read_sampledpoints_shp("eastern")
southern_shp <- read_sampledpoints_shp("southern")
remove(input_file_path, read_sampledpoints_shp)
NDVIsample<- bind_rows(central_shp, eastern_shp, southern_shp)
remove(central_shp, eastern_shp, southern_shp)

monthly_df <- ndvi_evi_ts_complete_tr %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time, 
                                                          NDVI_monthly_tr, EVI_monthly_tr, NDVI_original, NDVI_trend,
                                                          EVI_original, EVI_trend))

pivot_monthly_df <- pivot_longer(monthly_df, 8:11)
pivot_monthly_df<- pivot_monthly_df %>% separate_wider_delim(cols = name, delim = "_", names = c("IndexName", "trend_original"))
pivot_monthly_df2<- pivot_monthly_df %>% 
  pivot_wider(names_from = trend_original, values_from = value)
pivot_monthly_df2 <- pivot_monthly_df2 %>% mutate(NDVI_monthly_tr = case_when(NDVI_monthly_tr == 1~ "Linear decrease",
                                                NDVI_monthly_tr ==2~ "Linear increase",
                                                NDVI_monthly_tr ==4~ "Step decrease",
                                                NDVI_monthly_tr ==5~"Step increase",
                                                NDVI_monthly_tr ==6~"Quadratic decrease (accelerated)",
                                                NDVI_monthly_tr ==7~"Quadratic decrease (decelerated)",
                                                NDVI_monthly_tr ==8~"Quadratic increase (accelerated)",
                                                NDVI_monthly_tr ==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))
pivot_monthly_df2 <- pivot_monthly_df2 %>% mutate(EVI_monthly_tr = case_when(EVI_monthly_tr == 1~ "Linear decrease",
                                                EVI_monthly_tr ==2~ "Linear increase",
                                                EVI_monthly_tr ==4~ "Step decrease",
                                                EVI_monthly_tr ==5~"Step increase",
                                                EVI_monthly_tr ==6~"Quadratic decrease (accelerated)",
                                                EVI_monthly_tr ==7~"Quadratic decrease (decelerated)",
                                                EVI_monthly_tr ==8~"Quadratic increase (accelerated)",
                                                EVI_monthly_tr ==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))


d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

#plotting
cz <- c("central", "eastern", "southern")

for (i in 1:length(cz)){
  mappalette<- c("red", "black")
  
  x_df <- pivot_monthly_df2 %>% filter(climate_zone == cz[[i]])
  unique_ID<- unique(x_df$NDVI_ID)
  
  
  for (j in 1:length(unique_ID)){
  
    df<- x_df %>% filter(NDVI_ID == unique_ID[[j]])
    ndvi_trajectory <- unique(df$NDVI_monthly_tr)
  
    lp_fig1<- df %>%  #lineplot #1 stl trend and original faceted by index
      ggplot(aes(x = as.factor(time)))+
      geom_line(aes(y = original, alpha = 0.4, group = IndexName )) + 
      geom_line(aes(y = trend, group = IndexName), lwd= 1.5, linetype="longdash")+
      theme_classic()+
      theme (axis.text.x = element_blank())+
      ylab("index") + 
      xlab("Time (Year + Months)") + scale_alpha(guide = 'none')+
        ggtitle (paste0("monthly NDVI = ", unique(df$NDVI_monthly_tr), ";",
                    " monthly EVI = ", unique(df$EVI_monthly_tr))) + facet_grid(.~IndexName)
    
    lp_fig2<- df %>%  #lineplot #2 only stl trend faceted by index
      ggplot(aes(x = as.factor(time)))+
      geom_line(aes(y = trend, group = IndexName), lwd= 1.5, linetype="longdash")+
      theme_classic()+
      theme (axis.text.x = element_blank())+
      ylab("index") + 
      xlab("Time (Year + Months)") + scale_alpha(guide = 'none')+ facet_grid(.~IndexName)
  
    #map
    NDVIsample<- NDVIsample %>% 
      mutate(colorcode = case_when(NDVI_ID==unique_ID[[j]] ~"1", TRUE ~ "2"))
    
    map_points<-   
      tm_shape(cerrado) + tm_borders()+
      tm_shape (NDVIsample)+
      tm_dots(size =0.5, fill = "colorcode", 
              fill.scale = tm_scale_categorical(values = mappalette)) +
      tm_layout(legend.show = FALSE) 
  
    grobobject<- tmap_grob(map_points)
    fig<- ggarrange(lp_fig1, lp_fig2, grobobject, nrow = 3, ncol = 1, widths = c(1,2))
    ggsave(here("Outputs", "TrajectoryPlotting", "monthly_timeseries_results",
                paste0(ndvi_trajectory,"_", unique_ID[[j]], ".png")), fig,
           dpi= 700, height = 70, width = 40, units= "cm")
    print (j)
  }
  print (i)
}  


remove(df, fig, grobobject, lp_fig1, lp_fig2, map_points, monthly_df, pivot_monthly_df, pivot_monthly_df2, x_df, i, j, mappalette, ndvi_trajectory, unique_ID)
```

#Step 4- plotting using annual time series results
```{r}
annual_df <- ndvi_evi_ts_complete_tr %>% dplyr::select(c(NDVI_ID, climate_zone, x, y, time, 
                                                          NDVI_annual_tr, EVI_annual_tr, NDVI_original, NDVI_trend,
                                                          EVI_original, EVI_trend))
annual_df <- annual_df %>%  
  separate_wider_delim(cols = time, delim = "_", names = c("Year","Month")) %>%
  dplyr::select(-c(Month))

annual_df2 <- annual_df %>% 
  group_by(NDVI_ID, climate_zone, NDVI_annual_tr, EVI_annual_tr, Year) %>%
  summarise(x = mean(x), y= mean(y),
    mean_NDVI_original = mean(NDVI_original), 
            mean_NDVI_trend = mean(NDVI_trend),
            mean_EVI_original = mean(EVI_original),
            mean_EVI_trend = mean(EVI_trend))
  

pivot_annual_df <- pivot_longer(annual_df2, 8:11)
pivot_annual_df<- pivot_annual_df %>% separate_wider_delim(cols = name, delim = "_", names = c("mean","IndexName", "trend_original"))
pivot_annual_df2<- pivot_annual_df %>% 
  dplyr::select(-c(mean)) %>%
  pivot_wider(names_from = trend_original, values_from = value)
pivot_annual_df2 <- pivot_annual_df2 %>% mutate(NDVI_annual_tr = case_when(NDVI_annual_tr == 1~ "Linear decrease",
                                                NDVI_annual_tr ==2~ "Linear increase",
                                                NDVI_annual_tr ==4~ "Step decrease",
                                                NDVI_annual_tr ==5~"Step increase",
                                                NDVI_annual_tr ==6~"Quadratic decrease (accelerated)",
                                                NDVI_annual_tr ==7~"Quadratic decrease (decelerated)",
                                                NDVI_annual_tr ==8~"Quadratic increase (accelerated)",
                                                NDVI_annual_tr ==9~"Quadratic increase (decelerated)",
                                                TRUE~"No trend"))
pivot_annual_df2 <- pivot_annual_df2 %>% mutate(EVI_annual_tr = case_when(EVI_annual_tr == 1~ "Linear decrease",
                                                EVI_annual_tr ==2~ "Linear increase",
                                                EVI_annual_tr ==4~ "Step decrease",
                                                EVI_annual_tr ==5~"Step increase",
                                                EVI_annual_tr ==6~"Quadratic decrease (accelerated)",
                                                EVI_annual_tr ==7~"Quadratic decrease (decelerated)",
                                                EVI_annual_tr ==8~"Quadratic increase (accelerated)",
                                                EVI_annual_tr ==9~"Quadratic increase (decelerated)",

                                                                                                TRUE~"No trend"))

for (i in 1:length(cz)){
  mappalette<- c("red", "black")
  
  x_df <- pivot_annual_df2 %>% filter(climate_zone == cz[[i]])
  unique_ID<- unique(x_df$NDVI_ID)
  
  
  for (j in 1:length(unique_ID)){
  
    df<- x_df %>% filter(NDVI_ID == unique_ID[[j]])
    ndvi_trajectory <- unique(df$NDVI_annual_tr)
  
    lp_fig1<- df %>%  #lineplot #1 stl trend and original faceted by index
      ggplot(aes(x = as.factor(Year)))+
      geom_line(aes(y = original, alpha = 0.4, group = IndexName )) + 
      geom_line(aes(y = trend, group = IndexName), lwd= 1.5, linetype="longdash")+
      theme_classic()+
      theme (axis.text.x = element_blank())+
      ylab("index") + 
      xlab("Time (Year)") + scale_alpha(guide = 'none')+
        ggtitle (paste0("annual NDVI = ", unique(df$NDVI_annual_tr), ";",
                    " annual EVI = ", unique(df$EVI_annual_tr))) + facet_grid(.~IndexName)
    
    lp_fig2<- df %>%  #lineplot #2 only stl trend faceted by index
      ggplot(aes(x = as.factor(Year)))+
      geom_line(aes(y = trend, group = IndexName), lwd= 1.5, linetype="longdash")+
      theme_classic()+
      theme (axis.text.x = element_blank())+
      ylab("index") + 
      xlab("Time (Year)") + scale_alpha(guide = 'none')+ facet_grid(.~IndexName)
  
    #map
    NDVIsample<- NDVIsample %>% 
      mutate(colorcode = case_when(NDVI_ID==unique_ID[[j]] ~"1", TRUE ~ "2"))
    
    map_points<-   
      tm_shape(cerrado) + tm_borders()+
      tm_shape (NDVIsample)+
      tm_dots(size =0.5, fill = "colorcode", 
              fill.scale = tm_scale_categorical(values = mappalette)) +
      tm_layout(legend.show = FALSE) 
  
    grobobject<- tmap_grob(map_points)
    fig<- ggarrange(lp_fig1, lp_fig2, grobobject, nrow = 3, ncol = 1, widths = c(1,2))
    ggsave(here("Outputs", "TrajectoryPlotting", "annual_timeseries_results",
                paste0(ndvi_trajectory,"_", unique_ID[[j]], ".png")), fig,
           dpi= 700, height = 70, width = 40, units= "cm")
    print (j)
  }
  print (i)
}  


```