```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I explore the results of this project against %annual woody cover change estimated by Hunter et al 2025 
"Global 30-m annual median vegetation height maps (2000–2022) based on ICESat-2 data and Machine Learning". 
(1) Data input and mosaicing
(2) Process input data to match resolution of 1km (this project)
(3) Mask resampled input data against results of this project
(4) Results compilation

#Step 1 - significant trends in median height (30m) resolution was exported out of
GEE, which is the data input below
```{r}
#Data input
raster_filepath <- list.files(path = here("Data", "Hunteretal2025_MedianVegHeight", "SignificantVegHeight"), pattern= paste0("*.tif"), all.files=TRUE, full.names=TRUE)
raster_list<-lapply(raster_filepath, rast)

d_trans<- st_read(here("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

#Mosiac
Sys.time(); mosaiced_sigveg <- terra::vrt(raster_filepath); Sys.time()

#Crop/mask
crop_mask_function <- function (raster){
  x_crop <- terra::crop(raster, vect(cerrado))
  x_mask <- terra::mask (x_crop, vect(cerrado))
  x_mask
}

Sys.time(); cerrado_sigveg <- crop_mask_function(mosaiced_sigveg); Sys.time() #half hour

#Scaling from GEE
scaled_cerrado_sigveg <- cerrado_sigveg/10^5
#writeRaster(scaled_cerrado_sigveg, here("Data", "Hunteretal2025_MedianVegHeight", "SignificantVegHeight", "scaled_cerrado_sigveg.tif"))
```

#Step 2 - Coarsen sig veg height trend to 1km and mask to results
```{r}
scaled_cerrado_sigveg<- rast(here("Data", "Hunteretal2025_MedianVegHeight", "SignificantVegHeight", "scaled_cerrado_sigveg.tif"))

monthly_NDVI_disagg_results <- rast(here("Outputs", "TrendsResults", "results_rasters", "monthly", "ndvi_monthly_swin11.tif"))
#1-Linear decrease 2-Linear increase 3- No trend 4-Step decrease 5- Step increase 6- Quad dec acc 7-Quad dec decc 8- Quad inc acc 9-Quad inc decc

Sys.time(); coarse_scaled_sigveg <- terra::resample(scaled_cerrado_sigveg, monthly_NDVI_disagg_results, method ="mean"); Sys.time() #20 min

Sys.time(); mask_coarse_scaled_sigveg <- terra::mask(coarse_scaled_sigveg, monthly_NDVI_disagg_results); Sys.time()
#writeRaster(mask_coarse_scaled_sigveg, here("Data", "Hunteretal2025_MedianVegHeight", "SignificantVegHeight", "mask_coarse_scaled_sigveg.tif"))

mask_coarse_scaled_sigveg<- rast(here("Data", "Hunteretal2025_MedianVegHeight", "SignificantVegHeight", "mask_coarse_scaled_sigveg.tif"))

together <- c(mask_coarse_scaled_sigveg, monthly_NDVI_disagg_results)
Sys.time(); sigveg_traj_df <- terra::as.data.frame(together); Sys.time()

names(sigveg_traj_df)<- c("percentage_annualchange", "trajectory_number")
sigveg_traj_df <- sigveg_traj_df %>% mutate(TrajectoryType= case_when(trajectory_number==1 ~"Linear_decrease",
                                trajectory_number==2 ~"Linear_increase",
                                trajectory_number==3 ~"No_trend",
                                trajectory_number==4 ~"Step_decrease",
                                trajectory_number==5 ~"Step_increase", 
                                trajectory_number==6 ~"Quad_dec_acc",
                                trajectory_number==7 ~"Quad_dec_decc",
                                trajectory_number==8 ~"Quad_inc_acc",
                                trajectory_number==9 ~"Quad_inc_decc")) 

```

#Step 3- Results compilation
```{r}

p <- ggplot(sigveg_traj_df, aes(x=TrajectoryType, y=percentage_annualchange)) + 
  geom_boxplot(outlier.shape = NA) +
  theme_minimal()
p

```

