```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
#Introduction
In this script, I extract sample points across the study area and plot the respective time series of all
indices with trend for one stl setting (swindow =11) and include the final trajectory result for each index
(1) Sample n = 10 random points per trajectory in the NDVI results 
(2) 

#Step 1- sample 10 random points in each trajectory consideirng NDVI results only
Note that all outputs from the chunk below have been written to disk
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()
remove(mosaic_ndvi)

threshold20_1km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_20.tif"))

ndvi_swin11_results <- read_rds(here("Outputs", "ModelSelection", "NDVI", "modelselection_swindow_11.rds"))
results_df <- ndvi_swin11_results %>% 
    dplyr::select(c(cell,x,y, model_order, shape_class, trend, loc_brk, climate_zone))
  
results_df <- results_df %>% ungroup() %>% 
    dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~5, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                )) %>% 
    dplyr::select(c(cell, x,y,pixelvalue, climate_zone))

climate_zones <- unique(results_df$climate_zone)

Sys.time(); for (i in 1:length(climate_zones)){
  x_climate_zone <- results_df %>% filter(climate_zone == climate_zones[[i]])
  unique_traj <- c(unique(x_climate_zone$pixelvalue))
  for ( j in 1:length(unique_traj)){
    selected_traj<- x_climate_zone %>% filter(pixelvalue == unique_traj[[j]])
    selected_traj_vector<- terra::vect(selected_traj, geom=c("x", "y"), crs="epsg:4326")
    selected_traj_raster<- terra::rasterize(selected_traj_vector,ndvi, "cell", fun="max")
    masked_selected_traj_raster<- terra::mask(selected_traj_raster, threshold20_1km)
    writeRaster(masked_selected_traj_raster, here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", climate_zones[[i]], paste0(j, "_traj_cellnumber.tif")))
  }
}; Sys.time()

match_df_sample_function <- function (cz){
  cellraster_files<- list.files(path = here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", cz), pattern =".tif", all.files= TRUE, full.names = TRUE)
  cellraster_files<- lapply (cellraster_files, rast)
  
  pixels_studyarea_eachtraj <- list()
  for ( i in 1:length(cellraster_files)){
    raster_to_df <- terra:: as.data.frame(cellraster_files[[i]], xy=TRUE)
    names(raster_to_df)[3]<- "cell"
    
    x_climate_zone <- results_df %>% filter(climate_zone == cz)
    pixels_studyarea_eachtraj[[i]] <- left_join (raster_to_df,  x_climate_zone, by= c("cell"= "cell"))
  }
  
  randomsample_studyarea_eachtraj <- list() #Sampling 10 random pixels in each trajctory ie each of the dfs in the list of dfs from previous step
  for (i in 1:length(pixels_studyarea_eachtraj)){
    randomsample_studyarea_eachtraj[[i]]<- pixels_studyarea_eachtraj[[i]] %>% 
                                            ungroup() %>%
                                              slice_sample(n = 10)
  }
  
  randomsample_studyarea_eachtraj<- bind_rows(randomsample_studyarea_eachtraj)
  randomsample_studyarea_eachtraj<- randomsample_studyarea_eachtraj %>% 
      select(x.x, y.x, cell, climate_zone, pixelvalue) %>% rename("x"="x.x", "y"="y.x")
  write.csv(randomsample_studyarea_eachtraj, here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", cz, "sampled_df.csv" ))
}

Sys.time(); match_df_sample_function("central"); Sys.time()
Sys.time(); match_df_sample_function("southern"); Sys.time()
Sys.time(); match_df_sample_function("eastern"); Sys.time()

remove(match_df_sample_function, ndvi_swin11_results, results_df)

#Sampled points from NDVI index, per trajectory and in each climate zone
ndvi_sample_central<- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "central", "sampled_df.csv"))
ndvi_sample_southern<- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "southern", "sampled_df.csv"))
ndvi_sample_eastern<- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "eastern", "sampled_df.csv"))

rename_ndvisample<- function (ndvi_sample_df){ #renaming to avoid confusion with other indices
  ndvi_sample_df<- ndvi_sample_df %>% dplyr::select(-c(X))
  names(ndvi_sample_df)[3]<- "NDVI_cellID"
  names(ndvi_sample_df)[5]<- "NDVI_traj"
  ndvi_sample_df
}

ndvi_sample_central<- rename_ndvisample(ndvi_sample_central)
ndvi_sample_southern<- rename_ndvisample(ndvi_sample_southern)
ndvi_sample_eastern<- rename_ndvisample(ndvi_sample_eastern)


NDVIsample_pointshp<- function(ndvi_sample_df, climate_zone){
  point_vector<- st_as_sf(terra::vect(ndvi_sample_df, geom=c("x", "y"), crs="epsg:4326"))
  st_write(point_vector, here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", paste0(climate_zone, "_NDVIsample.shp")))
} 

NDVIsample_pointshp(ndvi_sample_central, "central")
NDVIsample_pointshp(ndvi_sample_southern, "southern")
NDVIsample_pointshp(ndvi_sample_eastern, "eastern")
remove(ndvi_sample_central, ndvi_sample_eastern, ndvi_sample_southern, rename_ndvisample, NDVIsample_pointshp)
```

#Step 2- extracting cell index for above sampled NDVI points but in EVI and anisoEVI
```{r}
threshold20_1km <- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "1km_thresholded_mapbiomas_mask_20.tif"))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mosaic_ndvi <- rast(here("Data", "Indices", "Cerrado_monthlyNDVI_GEE", "buffer_ndvi_proj.tif"))
Sys.time(); ndvi<- terra::crop(mosaic_ndvi[[262]], vect(cerrado)); Sys.time() #selecting random layer
Sys.time(); ndvi<- terra::mask(ndvi, vect(cerrado)); Sys.time()
remove(mosaic_ndvi)

#Extracting cell index and traj from EVI and anisoEVI for above NDVI sampled points
step1<- function (modelselection_df, index_name, cz, ndvisample_pointshp){
  #df prep
  modelselection_df<- modelselection_df %>% 
    dplyr::select(c(cell,x,y, model_order, shape_class, trend, loc_brk, climate_zone))
  
  x <- modelselection_df %>% filter(climate_zone == cz)
  
  x<- x%>% ungroup() %>%
    dplyr::select(c(cell, x, y, model_order, shape_class, trend, climate_zone)) %>%
    mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                                 model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                 model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                 model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                 model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                 model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                 model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                 model_order=="Step" & is.na(shape_class) & trend == "increase" ~5,
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                 )) %>%
    dplyr::select(-c(model_order, shape_class,trend)) #note Linear & Null stable is same category
  
  #raster of cell index
  cellindex_vector<- terra::vect(x, geom=c("x", "y"), crs="epsg:4326")
  cellindex_raster<- terra::rasterize(cellindex_vector, ndvi, "cell", fun="max")
  masked_cellindex_raster<- terra::mask(cellindex_raster, threshold20_1km)
  
  #sample NDVI points in above raster toi get cellID
  newindex_sample_cellID<- terra::extract(masked_cellindex_raster, ndvisample_pointshp, method ="simple", bind = TRUE)
  
  #for above cellID get trajectory from new index
  newindex_sample_cellID<- as.data.frame(newindex_sample_cellID)
  
    
  newindex_sample_celltraj<- left_join(newindex_sample_cellID,x, by= c("max"="cell"))
 
  
  names(newindex_sample_celltraj)[4]<-  paste0(index_name, "_ID")
  names(newindex_sample_celltraj)[8]<- paste0(index_name, "_tr")
  
  write.csv(newindex_sample_celltraj, here("Outputs", "TrajectoryPlotting", paste0("NDVI_", index_name, "_", cz,  "_sampledpoints.csv")))
}  


#Application of above function  
ndvi_sample_centralshp<- st_read(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "central_NDVIsample.shp"))
ndvi_sample_southernshp<- st_read(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "southern_NDVIsample.shp"))
ndvi_sample_easternshp<- st_read(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "eastern_NDVIsample.shp"))

#EVI
modelselection_evi<- read_rds(here("Outputs", "ModelSelection", "EVI", "modelselection_swindow_11.rds"))

Sys.time(); step1(modelselection_evi, "EVI", "central", ndvi_sample_centralshp); Sys.time()
Sys.time(); step1(modelselection_evi, "EVI", "southern", ndvi_sample_southernshp); Sys.time()
Sys.time(); step1(modelselection_evi, "EVI", "eastern", ndvi_sample_easternshp); Sys.time()

#anisoEVI
modelselection_aniso<- read_rds(here("Outputs", "ModelSelection", "anisoEVI", "modelselection_swindow11.rds"))

Sys.time(); step1(modelselection_aniso, "EVI", "central", ndvi_sample_centralshp); Sys.time()
Sys.time(); step1(modelselection_aniso, "EVI", "southern", ndvi_sample_southernshp); Sys.time()
Sys.time(); step1(modelselection_aniso, "EVI", "eastern", ndvi_sample_easternshp); Sys.time()

```

#Step 3- extracting time series from all indices for above sampled points
Note that I ran the below commands on the server, since the time series stl 
decomposition results for each index are on the server
```{r}
step2<- function (sample_df, stl_df){
  cell_sample<- unique(sample_df$cell)
  x<- stl_df %>% filter(cell %in% cell_sample)
}
```

#Step 4- df preparation for plotting
```{r}
central_sampled_points <- read.csv(here("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "central", "central_complete_sampled.csv"))
southern_sampled_points <- read.csv(here ("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "southern", "southern_complete_sampled.csv"))
eastern_sampled_points <- read.csv(here ("Outputs", "TrajectoryPlotting", "NDVI_trajectories", "eastern", "eastern_complete_sampled.csv"))

 
#Join to ndvi model selection results
join_timeseries_trajectory_function <- function (results_df, zone_sampled_points, cz){
  unique_cells <- unique(zone_sampled_points$cell)
  zone_sampled_points <- zone_sampled_points %>% mutate(climate_zone = cz)
  
  results_df <- ndvi_swin11_results %>% 
    dplyr::select(c(cell,x,y, model_order, shape_class, trend, loc_brk, climate_zone))
  
  results_df <- results_df %>% ungroup() %>% 
      dplyr::select(c(cell,x, y, model_order, shape_class, trend, climate_zone)) %>%
      mutate(pixelvalue= case_when(model_order=="Lin" & shape_class== "decrease_constant"& is.na(trend)~1,
                               model_order=="Lin" & shape_class== "increase_constant" & is.na(trend)~2,
                                model_order=="Lin" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Null" & shape_class== "stable_constant" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_concave" & is.na(trend)~3,
                                model_order=="Quad" & shape_class== "stable_convex" & is.na(trend)~3,
                                model_order=="Step" & is.na(shape_class) & trend == "decrease" ~4,
                                model_order=="Step" & is.na(shape_class) & trend == "increase" ~5, 
                                model_order=="Quad" & shape_class=="decrease_accelerated" & is.na (trend)~6,
                                model_order=="Quad" & shape_class=="decrease_decelerated" & is.na(trend)~7,
                                model_order=="Quad" & shape_class=="increase_accelerated" & is.na (trend)~8,
                                model_order=="Quad" & shape_class=="increase_decelerated" & is.na(trend)~9
                                )) %>% 
    dplyr::select(c(cell, x,y,pixelvalue, climate_zone))
  
  results_df <- results_df  %>% filter(climate_zone == cz)
  results_sampled <- results_df %>% filter(cell %in% unique_cells)
  x_join <- left_join (results_sampled, zone_sampled_points, by= c("cell"= "cell", "climate_zone" = "climate_zone"))
  x_join
}

ndvi_swin11_results <- read_rds(here("Outputs", "ModelSelection", "NDVI", "modelselection_swindow_11.rds"))

Sys.time(); central_plot_df<-  join_timeseries_trajectory_function(ndvi_swin11_results, central_sampled_points, "central"); Sys.time()
Sys.time(); southern_plot_df<-  join_timeseries_trajectory_function(ndvi_swin11_results, southern_sampled_points, "southern"); Sys.time()
Sys.time(); eastern_plot_df<-  join_timeseries_trajectory_function(ndvi_swin11_results, eastern_sampled_points, "eastern"); Sys.time()

ndvi_sampled_df<- bind_rows(central_plot_df, southern_plot_df, eastern_plot_df)
ndvi_sampled_df<- ndvi_sampled_df %>% 
  dplyr::select(-c(x.y, y.y, X, value, seasonal, remainder)) %>% 
  rename(c("NDVI_ID"="cell","x"="x.x", "y"="y.x", "NDVI_original"="value_int", "time"="name", "NDVI_trend"="trend"))

remove(central_sampled_points, southern_sampled_points, eastern_sampled_points, join_timeseries_trajectory_function,
       ndvi_swin11_results, central_plot_df, southern_plot_df, eastern_plot_df)

##Read anisoEVI and EVI sample df (outputs of Step 2 that are transferred to server)
##and anisoEVI and EVI complete dfs (outputs from Step 3 which we get on server and transfer
##back to laptop)

input_file_path <- here("Outputs", "TrajectoryPlotting")
read_df_join_function <- function (index_name_with_NDVI, index_considered_for_join, just_index){
  complete_df_list <-  list.files(path = paste0(input_file_path, "/"), pattern = ("*complete*"), all.files = T, full.names = T)
  complete_df_list <- lapply(complete_df_list, read.csv)
  
  df_list<- list.files(path = paste0(input_file_path, "/"), pattern = paste0("*",index_name_with_NDVI,"*"), all.files = T, full.names = T)
  df_list <- lapply(df_list, read.csv)
  
  df_list<- df_list[4:6]
  
  join_list<- list()
  for (i in 1:3){
    x_join <- left_join(complete_df_list[[i]], df_list[[i]], by =c("cell"=index_considered_for_join)) #can be "EVI_ID" or "aniso_ID"
    x_join<- x_join %>% dplyr::select(-c(X.x, value, seasonal, remainder, X.y, x.y, y.y, clmt_zn))
    names(x_join)[1]<-index_considered_for_join
    names(x_join)[2]<- "x"
    names(x_join)[3]<- "y"
    names(x_join)[5]<- paste0(just_index, "_", "original")
    names(x_join)[6]<- paste0(just_index, "_", "trend")
    join_list[[i]] <- x_join
  }
  df<- bind_rows(join_list)
  df
}

Sys.time(); ndvi_evi_sample_df <- read_df_join_function ("NDVI_EVI_", "EVI_ID", "EVI"); Sys.time()
#Sys.time(); ndvi_aniso_sample_df <- read_complete_stl_df_function(<<include>>); Sys.time()
remove(read_df_join_function)

#Joining EVI & NDVI together (do EVI & anisoEVI join first and then to NDVI when you have anisoEVI results)
cz <- c("central", "southern", "eastern")
complete_list<- list()
for (i in 1:3){
  x_ndvi_sample<- ndvi_sampled_df %>% filter(climate_zone == cz[[i]])
  x_evi_sample <- ndvi_evi_sample_df %>% filter(climate_zone == cz[[i]])
  
  unique_cell<- unique(x_ndvi_sample$NDVI_ID)
  
  x_join_list<- list()
  for (j in 1:length(unique_cell)){
    cell_ndvi<- x_ndvi_sample %>% filter(NDVI_ID ==unique_cell[[j]])
    cell_ndvi<- cell_ndvi %>% separate_wider_delim(cols = time, delim = "_", names = c("year","month"))
    cell_evi<- x_evi_sample %>% filter(NDVI_ID == unique_cell[[j]])
    cell_evi<- cell_evi %>% separate_wider_delim(cols = name, delim = "_", names = c("year","month"))
    
    annual_cell_ndvi <- cell_ndvi %>% 
      group_by(NDVI_ID, year) %>% 
      mutate(NDVI_mean_original = mean (NDVI_original),NDVI_mean_trend = mean (NDVI_trend))
    annual_cell_ndvi <- annual_cell_ndvi %>% 
      filter(!duplicated(year)) %>% 
      dplyr::select(-c(NDVI_original, NDVI_trend))
    
    annual_cell_evi <- cell_evi %>% 
      group_by(NDVI_ID, year) %>% 
      mutate(EVI_mean_original = mean (EVI_original),EVI_mean_trend = mean (EVI_trend))
    annual_cell_evi <- annual_cell_evi %>% 
      filter(!duplicated(year)) %>% 
      dplyr::select(-c(EVI_original, EVI_trend))
    
    cell_df <- bind_cols (annual_cell_ndvi, annual_cell_evi)
    cell_df <- cell_df %>% 
      dplyr::select(c(NDVI_ID...1, year...6, NDVI_mean_original, 
                      NDVI_mean_trend, NDVI_tr, EVI_mean_original, EVI_mean_trend, EVI_tr, x...11, y...12, climate_zone...17))
    names(cell_df)[1]<- "NDVI_ID"
    names(cell_df)[2]<- "year"
    names(cell_df)[9]<- "x"
    names(cell_df)[10]<- "y"
    names(cell_df)[11]<- "climate_zone"
    x_join_list[[j]]<- cell_df
  }
  x_df <- bind_rows(x_join_list)
  complete_list[[i]]<- x_df
}

df_ndvi_evi <- bind_rows(complete_list)
remove(complete_list, x_join_list, cell_df, cell_evi, cell_ndvi, x_ndvi_sample, x_evi_sample, x_df, annual_cell_evi, annual_cell_ndvi)
remove(i, j, unique_cell)

df_ndvi_evi<- df_ndvi_evi%>% mutate(NDVI_tr = case_when(NDVI_tr == 1~ "Linear decrease",
                                          NDVI_tr ==2~ "Linear increase",
                                          NDVI_tr ==3 ~ "No trend",
                                          NDVI_tr ==4~ "Step decrease",
                                          NDVI_tr ==5~"Step increase",
                                          NDVI_tr ==6~"Quadratic decrease (accelerated)",
                                          NDVI_tr ==7~"Quadratic decrease (decelerated)",
                                          NDVI_tr ==8~"Quadratic increase (accelerated)",
                                          NDVI_tr ==9~"Quadratic increase (decelerated)"))
  

df_ndvi_evi<- df_ndvi_evi%>% mutate(EVI_tr = case_when(EVI_tr == 1~ "Linear decrease",
                                          EVI_tr ==2~ "Linear increase",
                                          EVI_tr ==3 ~ "No trend",
                                          EVI_tr ==4~ "Step decrease",
                                          EVI_tr ==5~"Step increase",
                                          EVI_tr ==6~"Quadratic decrease (accelerated)",
                                          EVI_tr ==7~"Quadratic decrease (decelerated)",
                                          EVI_tr ==8~"Quadratic increase (accelerated)",
                                          EVI_tr ==9~"Quadratic increase (decelerated)"))
```

#Step 5- plotting 
```{r}
final_df<- df_ndvi_evi %>% dplyr::select(c(climate_zone, x, y, year, NDVI_ID, NDVI_tr, EVI_tr, 
                                         NDVI_mean_original, NDVI_mean_trend, EVI_mean_original, 
                                         EVI_mean_trend))
pivot_df<- pivot_longer(final_df, 8:11)
pivot_df<- pivot_df %>% separate_wider_delim(cols = name, delim = "_", names = c("IndexName","mean", "value1"))

pivot_df2<- pivot_df %>% dplyr::select(-c(mean)) %>% 
  pivot_wider(names_from = value1, values_from = value)

centralNDVIsample<- st_read(here("Outputs","TrajectoryPlotting", "NDVI_trajectories", "central_NDVIsample.shp"))
southernNDVIsample<- st_read(here("Outputs","TrajectoryPlotting", "NDVI_trajectories", "southern_NDVIsample.shp"))
easternNDVIsample<- st_read(here("Outputs","TrajectoryPlotting", "NDVI_trajectories", "eastern_NDVIsample.shp"))
NDVIsample<- bind_rows(centralNDVIsample, southernNDVIsample, easternNDVIsample)
remove(centralNDVIsample, southernNDVIsample, easternNDVIsample)

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

for (i in 1:length(cz)){
  mappalette<- c("red", "black")
  
  x_df <- pivot_df2 %>% filter(climate_zone == cz[[i]])
  unique_ID<- unique(x_df$NDVI_ID)
  
  #lineplot
  for (j in 1:length(unique_ID)){
  
    df<- x_df %>% filter(NDVI_ID == unique_ID[[j]])
    ndvi_trajectory <- unique(df$NDVI_tr)
  
    lp_fig<- df %>%  
      ggplot(aes(x = as.factor(year), group = IndexName))+
      geom_line(aes(y = original, alpha = 0.4, color = IndexName )) + 
      geom_line(aes(y = trend, color = IndexName), lwd= 1.5, linetype="longdash")+
      theme_classic()+
      ylab("index") + 
      xlab("Time") + scale_alpha(guide = 'none')+
        ggtitle (paste0("NDVI = ", unique(df$NDVI_tr), ";",
                    " EVI = ", unique(df$EVI_tr)))
    lp_fig <- lp_fig + scale_color_brewer(palette = "Dark2")
    
    #map
    NDVIsample<- NDVIsample %>% 
      mutate(colorcode = case_when(NDVI_ID==unique_ID[[j]] ~"1", TRUE ~ "2"))
    
    map_points<-   
      tm_shape(cerrado) + tm_borders()+
      tm_shape (NDVIsample)+
      tm_dots(size =0.5, fill = "colorcode", 
              fill.scale = tm_scale_categorical(values = mappalette)) +
      tm_layout(legend.show = FALSE) 
  
    grobobject<- tmap_grob(map_points)
    fig<- ggarrange(grobobject, lp_fig, widths= c(1,2))
    ggsave(here("Outputs", "TrajectoryPlotting", "plots",
                paste0(ndvi_trajectory,"_", unique_ID[[j]], ".png")), fig,
           dpi= 700, height = 10, width = 30, units= "cm")
    print (j)
  }
  print (i)
}  

```