```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(tidyverse)
library(here)
library(tictoc)
library(ggplot2)
library(ggpubr)
library(sf)
library(terra)
library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I complete attribution of the trajectory shape analyses. This means that- all pixels have a categorical trajector shape assigend to them (based on the anisoEVI time series of the pixel). I now look at these pixels and figure out which ones have had (i) no transitions --> if no transitions, then what lulc it has remained and (ii) transitions 

The lulc time series data I have used is Mapbiomas. Because of the fine granularity of lulc classes (ie.there are too many classes), I have decided to focus on only pixels that are savanna, grass, forest, soy, pasture. The remaining classes are then treated as anthropic or natural (based on the categorization by Mpabiomas). I also focus on the transtions between these classes (see 7_mapbiomastransition_metadata.Rmd).

##No transitions analyses 
All data was prepared in GEE AbruptChanges_Cerrado>LULCChanges>Mapbiomas Transitions.

#Data entry
```{r}
#Mapbiomas notransitions rom GEE
Sys.time()
file_list <- list.files(path = here("Data", "Mapbiomas_transitions", "Correct"), 
                       pattern='.tif$', all.files=TRUE, full.names=TRUE)
raster_list<-lapply(file_list, rast)
Sys.time()

notransitions<-rast(raster_list)
names(notransitions)<- c("alwaysforest","alwaysotheranthropic","alwaysgrassland", "alwaysothernatural",  "alwayspasture", "alwayssavanna", "alwayssoy","notransitions")
remove(raster_list, file_list)

#trajectoryshapeanalyses
#From lines 769-784 in 5_trajectoryshapeanalyses.Rmd
amtrend_map_results<- rast(here("Outputs", "Final_TrajShape_Models", "Annual", "am_trends11_trajresults.tif"))
#1- Linear decrease, 2- Linear increase, 3- Stable classes, 4- Step decrease, 5- step increase

x<- terra::crop(notransitions, amtrend_map_results)
notransitions_masked<- terra::mask(x, amtrend_map_results)
remove(x)

```

#Attribution - No transitions
```{r}
Sys.time(); df_notransitions_masked<- terra::as.data.frame(notransitions_masked, xy=TRUE, cells=TRUE, na.rm=NA ); Sys.time()#few seconds
df_trajshapes<- terra::as.data.frame(amtrend_map_results, xy=TRUE, cells=TRUE, na.rm=NA)

trial_join<- left_join(df_trajshapes, (df_notransitions_masked %>% dplyr::select(-c(x,y))) , by=join_by("cell"))
summary(trial_join) #NAs in the mapbiomas lulc product. 

#NA_notransitions<- trial_join %>% filter(is.na(notransitions))
#NA_notransitions<- NA_notransitions %>% mutate(condition=ifelse((is.na(alwaysforest)|is.na(alwaysotheranthropic)|is.na(alwaysgrassland)|is.na(alwaysothernatural)| is.na(alwayspasture)| is.na(alwayssavanna)| is.na(alwayssoy)),1,0))
#sum(NA_notransitions$condition) #235 basically all NA no transitions have no values in the individual classes, which makes sense
#NAnotransitions_map_vector<- terra::vect(NA_notransitions, geom=c("x", "y"), crs="epsg:4326")
#terra::writeVector(NAnotransitions_map_vector, here("Scratch", "NAnotransitions_map_vector.shp"))
#evi<- rast(here("Data", "Indices", "AnisoEVI_Cerrado_GEE", "mosaiced_anisoEVI.tif"))
#Sys.time(); NAnotransitions_map_results<- terra::rasterize(NAnotransitions_map_vector, evi, "max", fun="max"); Sys.time()
#writeRaster(NAnotransitions_map_results, here("Scratch", "NA_notransitions_trajshape2.tif"))
#The NA no transitions is all in water bodies or around water bodies! So I do not understand why I have anisoEVI traj shape results
#for these pixels because anisoEVI should have masked out all the water bodies. 

################################Decision made- all pixels with NA no transitions column are excluded from the attribution analyses
trial_join_excldueNAtransitions<- trial_join %>% filter(!is.na(notransitions))


#check 1- All NAs in individual lulc rasters should be 0% notransition
#check1<- trial_join %>% filter(notransitions==0) #12307 rows
#check1<- check1 %>% mutate(condition=ifelse((is.na(alwaysforest)|is.na(alwaysotheranthropic)|is.na(alwaysgrassland)|is.na(alwaysothernatural)| is.na(alwayspasture)| is.na(alwayssavanna)| is.na(alwayssoy)),1,0))
#sum(check1$condition) #12307 passed check
#We need to figure out why there are NAs still remaining in the individual classes after accounting for 0% no transition
#check2<- trial_join %>%  filter(!is.na(notransitions))
#summary(check2)
#check2<- trial_join %>% filter(notransitions!=0)
#xcheck2<- check2 %>% filter(is.na(alwaysforest))
#hist(xcheck2$notransitions)
#xcheck2_map_vector<- terra::vect(xcheck2, geom=c("x", "y"), crs="epsg:4326")
#terra::writeVector(xcheck2_map_vector, here("Scratch", "xcheck4.shp"))
#xcheck2_map_results<- terra::rasterize(xcheck2_map_vector, evi, "max", fun="max")
#writeRaster(xcheck2_map_results, here("Scratch", "xcheck2.tif"))
#I manually checked in QGIS. All pixels that have exteremly small %no transitions is basically 0% no transitions ie all of the pixel transitioned. I think
#the way GEE aggregates to 1 km and stores value is the problem ie when aggregarting it keeps even the smallest value, but it stores (at some step) the very small value
# as 0. And for the pixels with close to 40% no tranistions (which is like 4-5 pixels see histogram and df) its all water. For some reason, the no transition
# raster has no water masked out (even though I have masked it int he very first step). So in the individual rclass rasters, it is NA but it is some value in the no transitions
#raster. 
#remove(check1, check2, NA_notransitions, xcheck2, xcheck2_map_vector)

################################Decision made- all pixels with NA values in individual classes is 0% no transitions. 
trial_join2 <- trial_join_excldueNAtransitions %>% mutate(notransitions=ifelse((notransitions!=0 & is.na(alwaysforest) ),0, notransitions))

###############################Final attribution
Sys.time(); df_transitions1<- trial_join2 %>% group_by(max) %>% summarise(across(alwaysforest:notransitions, ~sum(.x,na.rm=TRUE)),); Sys.time()
Sys.time(); area_trajshapes<- trial_join2 %>% group_by(max) %>% summarise(trajshape_totalrea_sqkm=n())

df_transitions1<- df_transitions1 %>% full_join(area_trajshapes)
remove(area_trajshapes)
  
##Total no transitions
df_totalnotransition<- df_transitions1 %>% dplyr::select(max, notransitions, trajshape_totalrea_sqkm)
names(df_totalnotransition)<- c("TrajShape", "notransition_totalarea_sqkm", "trajshape_totalarea_sqkm")
df_totalnotransition <- df_totalnotransition %>% mutate(transition_totalarea_sqkm=(trajshape_totalarea_sqkm-notransition_totalarea_sqkm))
df_totalnotransition<-df_totalnotransition %>% mutate(percentage_notransition=(notransition_totalarea_sqkm/trajshape_totalarea_sqkm)*100)
pivot_df_totalnotransition<- df_totalnotransition %>% dplyr::select(-c(trajshape_totalarea_sqkm)) %>% pivot_longer(2:3)
pivot_df_totalnotransition<- pivot_df_totalnotransition %>% mutate(TrajShape=case_when(TrajShape==1~"Linear decrease",
                                                                                       TrajShape==2~"Linear increase",
                                                                                       TrajShape==3~"Stable",
                                                                                       TrajShape==4~"Step decrease",
                                                                                       TrajShape==5~"Step increase"))
pivot_df_totalnotransition<- pivot_df_totalnotransition %>% mutate(percentage_notransition=ifelse(name=="transition_totalarea_sqkm",NA, percentage_notransition))


x<-pivot_df_totalnotransition %>%
ggplot(aes(x=reorder(as.factor(TrajShape),-value), y=(value), group= as.factor(TrajShape), fill= as.factor(TrajShape), alpha=name)) +
  geom_bar(stat = 'identity', position = 'stack')+ 
  geom_text(aes(label=round(percentage_notransition,digits=2)), stat = "identity", vjust=1.2)+
  scale_fill_manual(values=c('#414451','#414451', '#414451', '#414451',"#414451"))+
  scale_alpha_manual(values=c(1,0.5))+
  theme_pubr(base_size=16, border= F) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(), 
        strip.text.x = element_text(size = 16),
        panel.border = element_rect(colour = "black", fill = NA))+
  ylab("Total area for which \n trajectory shape results are present (sqkm)")+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x = element_blank())
#ggsave(here("Outputs", "LULCAttribution_TrajShape", "notransitions_trajshape.png"),
#       x,dpi = 700, height = 20, width=20, units = "cm")
remove(df_transitions1,area_trajshapes,df_totalnotransition,pivot_df_totalnotransition,x)

##Total no transitions attributed further to individual classes
trial_join3<- trial_join2 
names(trial_join3)[12]<- "NoTransitions_TotalArea_sqkm" #each pixel is 1*1 sqkm. So notransitions is a proportion, which means proportion* (1*1 sqkm)=no transitions area

#the individual class columns alwaysforest etc is the proportion of the no transitions area. So converting to area
trial_join3<- trial_join3 %>% mutate(AlwaysForest_NoTransition_areasqkm=(alwaysforest*NoTransitions_TotalArea_sqkm),
                                     AlwaysOtherAnthropic_NoTransition_areasqkm=(alwaysotheranthropic*NoTransitions_TotalArea_sqkm),
                                     AlwaysGrassland_NoTransition_areasqkm=(alwaysgrassland*NoTransitions_TotalArea_sqkm),
                                     AlwaysOtherNatural_NoTransition_areasqkm=(alwaysothernatural*NoTransitions_TotalArea_sqkm),
                                     AlwaysPasture_NoTransition_areasqkm=(alwayspasture*NoTransitions_TotalArea_sqkm),
                                     AlwaysSavanna_NoTransition_areasqkm=(alwayssavanna*NoTransitions_TotalArea_sqkm),
                                     AlwaysSoy_NoTransition_areasqkm=(alwayssoy*NoTransitions_TotalArea_sqkm))
trial_join3<- trial_join3 %>% dplyr::select(-c(alwaysforest, alwaysotheranthropic, alwaysgrassland, alwaysothernatural, alwayspasture, alwayssavanna, alwayssoy))
Sys.time(); df_notransitionattribution<- trial_join3 %>% group_by(max) %>%
  summarise(across(NoTransitions_TotalArea_sqkm:AlwaysSoy_NoTransition_areasqkm, ~sum(.x,na.rm=TRUE)),); Sys.time()


#############################################################Anthropic/Natural
pivot_df_notransitionattribution<- df_notransitionattribution %>% pivot_longer(3:9)
pivot_df_notransitionattribution<- pivot_df_notransitionattribution %>% mutate(N_A= case_when(name=="AlwaysForest_NoTransition_areasqkm"~"Always_natural_notransition_sqkm",
                                                                                              name=="AlwaysGrassland_NoTransition_areasqkm"~"Always_natural_notransition_sqkm",
                                                                                              name=="AlwaysOtherNatural_NoTransition_areasqkm"~"Always_natural_notransition_sqkm",
                                                                                              name=="AlwaysSavanna_NoTransition_areasqkm"~"Always_natural_notransition_sqkm",
                                                                                              TRUE~"Always_anthropic_notransition_sqkm"))

pivot_df_notransitionattribution2<- pivot_df_notransitionattribution%>% dplyr::select(-c(name)) %>%
  group_by(max, N_A) %>% summarise(NoTransitions_TotalArea_sqkm=max(NoTransitions_TotalArea_sqkm), value=sum(value))
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(percent=(value/NoTransitions_TotalArea_sqkm)*100)
#see details of check3 below- the sum of anthropic and natural within each max group=100, but it is not.
#Hence taking off the discrepancy evenly from the two groups
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% pivot_wider(names_from = N_A, values_from = value)
pivot_df_notransitionattribution2<-  pivot_df_notransitionattribution2 %>% mutate(percentAnthropic=(Always_anthropic_notransition_sqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                                  percentNatural=(Always_natural_notransition_sqkm/NoTransitions_TotalArea_sqkm)*100)
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(realsumpercent= percentAnthropic+percentNatural)
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(percentAnthropic=percentAnthropic-abs((100-realsumpercent)/2),
                                                                                 percentNatural=percentNatural-abs((100-realsumpercent)/2)
                                                                                 )
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% dplyr::select(c(max, percentAnthropic, percentNatural))
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% pivot_longer(2:3)


pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(max=case_when(max==1~"Linear decrease",
                                                                                       max==2~"Linear increase",
                                                                                       max==3~"Stable",
                                                                                       max==4~"Step decrease",
                                                                                       max==5~"Step increase"))
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(name=ifelse(name=="percentAnthropic","Anthropic","Natural"))

z<- pivot_df_notransitionattribution2%>%
  ggplot() + aes(max, value, fill= name) +
  geom_bar(stat="identity") +
  ylab("% of area that did not transition") +
  geom_text(aes(label=paste0(sprintf("%1.1f", value),"%")),
            position=position_stack(vjust=0.5)) +
  scale_fill_manual(values=c( '#414451','#666600'))+
  theme_pubr(base_size=16, border= F)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x = element_blank())+
  theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape", "notransitions_N_A_attribution.png"),
       z,dpi = 700, height = 20, width=20, units = "cm")
remove(pivot_df_notransitionattribution,pivot_df_notransitionattribution2,z)

##############################################################Fine LULC classes that remained
df_notransitionattribution<- df_notransitionattribution %>% mutate(percentage_AlwaysForest=(AlwaysForest_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                   percentage_AlwaysOtherAnthropic=(AlwaysOtherAnthropic_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                   percentage_AlwaysGrassland=(AlwaysGrassland_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                   percentage_AlwaysOtherNatural=(AlwaysOtherNatural_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                   percentage_AlwaysPasture=(AlwaysPasture_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                   percentage_AlwaysSavanna=(AlwaysSavanna_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100,
                                                                   percentage_AlwaysSoy=(AlwaysSoy_NoTransition_areasqkm/NoTransitions_TotalArea_sqkm)*100)


check3<- df_notransitionattribution %>% dplyr::select(c(max,percentage_AlwaysForest,percentage_AlwaysOtherAnthropic,
                                                        percentage_AlwaysGrassland,percentage_AlwaysOtherNatural,percentage_AlwaysPasture,percentage_AlwaysSavanna,percentage_AlwaysSoy ))

check3<- check3 %>% rowwise() %>%
  mutate(sumpercent=sum(c_across(percentage_AlwaysForest:percentage_AlwaysSoy)))
#Unfortunately this check fails. All the percentages should be equal to 100. I do not know understand why the percent sum is not 100
#Maybe the discrepenccy is in GEE while aggregating the 30m pixels to 1000m ie taking percentage.
#So for now- I take off the discrepancy equally across all individual columns such that it is a 100%

df_notransitionattribution2<- df_notransitionattribution %>% dplyr::select(c(max,percentage_AlwaysForest,
                                                                             percentage_AlwaysOtherAnthropic,percentage_AlwaysGrassland,percentage_AlwaysOtherNatural,
                                                                             percentage_AlwaysPasture,percentage_AlwaysSavanna,percentage_AlwaysSoy))
df_notransitionattribution2<- df_notransitionattribution2 %>% rowwise() %>%
  mutate(sumpercent=sum(c_across(percentage_AlwaysForest:percentage_AlwaysSoy)))


df_notransitionattribution2<- df_notransitionattribution2 %>% mutate(percentage_AlwaysForest=percentage_AlwaysForest-abs((100-sumpercent)/7),
                                                                     percentage_AlwaysOtherAnthropic= percentage_AlwaysOtherAnthropic- abs((100-sumpercent)/7),
                                                                     percentage_AlwaysGrassland=percentage_AlwaysGrassland -abs((100-sumpercent)/7),
                                                                     percentage_AlwaysOtherNatural=percentage_AlwaysOtherNatural- abs((100-sumpercent)/7),
                                                                     percentage_AlwaysPasture= percentage_AlwaysPasture- abs((100-sumpercent)/7),
                                                                     percentage_AlwaysSavanna= percentage_AlwaysSavanna - abs((100-sumpercent)/7),
                                                                     percentage_AlwaysSoy= percentage_AlwaysSoy- abs((100-sumpercent)/7)
                                      )


pivot_df_notransitionattribution2<- df_notransitionattribution2 %>% dplyr::select(-c(sumpercent)) %>%
  pivot_longer(2:8)
pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(max=case_when(max==1~"Linear decrease",
                                                                                       max==2~"Linear increase",
                                                                                       max==3~"Stable",
                                                                                       max==4~"Step decrease",
                                                                                       max==5~"Step increase"))

pivot_df_notransitionattribution2<- pivot_df_notransitionattribution2 %>% mutate(name=case_when(name=="percentage_AlwaysForest"~"Always forest",
                                                                                                name=="percentage_AlwaysOtherAnthropic"~"Always other anthropic",
                                                                                                name=="percentage_AlwaysPasture"~"Always pasture",
                                                                                                name=="percentage_AlwaysSoy"~"Always soy",
                                                                                                name=="percentage_AlwaysGrassland"~"Always grassland",
                                                                                                name=="percentage_AlwaysOtherNatural"~"Always other natural",
                                                                                                name=="percentage_AlwaysSavanna"~"Always savanna"))




y<- pivot_df_notransitionattribution2%>%
  ggplot() + aes(max, value, fill= name) +
  geom_bar(stat="identity") +
  ylab("% of area that did not transition") +
  geom_text(aes(label=paste0(sprintf("%1.1f", value),"%")),
            position=position_stack(vjust=0.5)) +
  scale_fill_manual(values=c('#006600','#66ff00', '#414451', '#66ffcc',"#ff6600", "#666600", "#ffccff"))+
  theme_pubr(base_size=16, border= F)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.title.x = element_blank())+
  theme(legend.title=element_blank())
ggsave(here("Outputs", "LULCAttribution_TrajShape", "notransitions_attribution.png"),
       y,dpi = 700, height = 20, width=20, units = "cm")
remove(y, pivot_df_notransitionattribution2, df_notransitionattribution2,df_notransitionattribution, check3 )


```

My hypothesis is that all of the "stable" anisoEVI pixels should have no transitions in Mapbiomas. But no transition pixels according to Mapbiomas
is only ~72% of that "stable" pixels according to anisoEVI. Interestingly, this pattern holds across the remaining trajectory classes i.e the other
trajectory classes (which means increase or decrease- linear step) is no transitions according to anisoEVI. This means two things- (1) maybe anisoEVI
is not the best index to pick up changes or (2) the products are very separate ie derived from different satellites, methods etc. So they cannot be
be matched up. To show that the first point is true, I consider one year of the anisoEVI results for pixels that have always been savanna, forest, 
pasture.... all classes. If indeed anisoEVI is a good indicator of these different LULCs, there should be a significant difference in the anisoEVI across
for all the pixels in these LULCs. 

I consider the individual LULC rasters and consider only puixels in each raster with value>90%. This means that >90% of the pixel is savanna, soy etc
which gives me some certainty in Mapbiomas product. Then I match it up with the annual mean (after stl) anisoEVI for each pixel in the certainly savanna, soy...
pixels.

```{r}
#rerun 32-40 line
df_notransitions<- terra::as.data.frame(notransitions,xy=TRUE, cells=TRUE, na.rm=NA)
df_notransitions<- df_notransitions %>% dplyr::select(-c(notransitions))
df_notransitions2<- df_notransitions %>% drop_na()

certainforest<- df_notransitions2 %>% dplyr::select(c(cell, x,y, alwaysforest)) %>% filter(alwaysforest>0.9)
certaingrassland<- df_notransitions2 %>% dplyr::select(c(cell, x,y, alwaysgrassland)) %>% filter(alwaysgrassland>0.9)
certainsavanna<- df_notransitions2 %>% dplyr::select(c(cell, x,y, alwayssavanna)) %>% filter(alwayssavanna>0.9)
certainothernatural<- df_notransitions2 %>% dplyr::select(c(cell, x,y, alwaysothernatural)) %>% filter(alwaysothernatural>0.9)

 




```







#Attribution - Transitions
```{r}
#Rerun lines 42-45 from snippet 1

#Reading only from forest and savanna- each raster is multiband (with )
file_list <- list.files(path = here("Data", "Mapbiomas_transitions"), 
                       pattern='Transition_forest_', all.files=TRUE, full.names=TRUE)
fromforest_list<-lapply(file_list, rast)
file_list2 <- list.files(path = here("Data", "Mapbiomas_transitions"), 
                       pattern='Transition_savanna_', all.files=TRUE, full.names=TRUE)
fromsavanna_list<-lapply(file_list2, rast)



```



