```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

#Introduction 
In this script I process MODIS fire data for various analyses. 

I processed MODIS fire data in GEE-
https://code.earthengine.google.com/?scriptPath=users%2Ftrishagopalakrishna%2Fdefault%3AAbruptChanges_Cerrado%2FDisturbances%2FMODIS_firefrequency
The output from GEE is an image collection, where each band is a year 2002-2021 and pixel
value is number a times a pixel burned in the year. Firstly, reading the exported data and mosaicing.
(1) Data processing - mosaic, rename layers, crop/mask to cerrado
(2) Burned area processing- coarsening to 5x5km and determining burned area fraction
(3) Anomaly calculation- linear regression trends, kendal tau trends and variance/std

#Step 1- data processing
```{r}
## 1- Mosaic
# raster_filepath <- list.files(path = here("Data", "Fire", "MODIS", "OG_binaryburn_MODIS"), pattern= ".tif$", all.files=TRUE, full.names=TRUE)
# raster_list<-lapply(raster_filepath, rast)
# tic(); modis_fire<- do.call(mosaic,raster_list); toc() #9.5 hours
# writeRaster(modis_fire, here("Data", "Fire", "MODIS", "binaryburn_monthly_500m_2002_2021.tif"))

## 2- Crop, mask to Cerrado  and rename raster layers to years
modis_fire<- rast(here("Data", "Fire", "MODIS", "binaryburn_monthly_500m_2002_2021.tif"))
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

modis_fire_crop<- terra::crop(modis_fire, vect(cerrado))
modis_fire_mask<- terra::mask(modis_fire_crop, vect(cerrado))
remove(modis_fire_crop)

years<-  rep(2002:2021,each=12) 
months<- rep(1:12,20)
raster_names<- vector()
for (i in 1:length(years)){
  name<- paste0(years[[i]],"_", months[[i]])
  raster_names[i]<-name
}
names(modis_fire_mask)<- raster_names
##writeRaster(modis_fire_mask, here("Outputs", "OtherVariables", "Fire", "cerrado_binaryburn_monthly_500m_2002_2021.tif"))
remove(modis_fire)

```

#Step 2- burned area processing
```{r}
binaryburn<-rast(here("Outputs", "OtherVariables", "Fire", "cerrado_binaryburn_monthly_500m_2002_2021.tif"))

annual_years<- 2002:2021
annual_df<- tibble(Year= annual_years) 

subset<- function (Year){
  terra::subset(binaryburn, grep(Year, names(binaryburn)))
}

annual_df<-annual_df %>% mutate(annual_burn= map(Year,subset))

frequency<- function (monthly_burn_raster){
   terra::app(monthly_burn_raster, fun="sum")
}
annual_df<-annual_df %>% mutate(frequency= map(annual_burn, frequency))

binaryburn<- function (annual_frequency_raster){
  as.numeric(annual_frequency_raster>0)
}
Sys.time(); annual_df<-annual_df %>% mutate(binaryburn= map(frequency, binaryburn)); Sys.time()

mat<- rast(here("Outputs", "OtherVariables", "Climate", "mat_1981_2021.tif"))
prop_aggregate <- function(binaryburn_raster) {
  terra::project(
    terra::segregate(binaryburn_raster), mat,# segregate()- binary (0/1) of classes 0,3,4,12
    method = "average", res = res(mat)[1]
  )
}

Sys.time(); annual_df<-annual_df %>% mutate(coarsened_binaryburn= map(binaryburn, prop_aggregate)); Sys.time()

burnarea<- function (coarsened_binary_raster){
  coarsened_binary_raster[[2]]
}
Sys.time(); annual_df<-annual_df %>% mutate(burnedarea= map(coarsened_binaryburn, burnarea)); Sys.time()

burnedarea_list<- annual_df$burnedarea
burnedarea<- rast(burnedarea_list)
writeRaster(burnedarea, here("Outputs", "OtherVariables", "Fire", "burnedarea_5km_2002_2021.tif"))

burnedarea<- rast(here("Outputs", "OtherVariables", "Fire", "burnedarea_5km_2002_2021.tif"))
mean_burnarea<- terra::app(burnedarea, fun= "mean")

#Map making
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

map_burnedarea<-   
  tm_shape(cerrado)+ tm_fill(title="")+
  tm_shape (mean_burnarea)+
  tm_raster(n=5, style = "quantile",palette = "YlOrBr", title="Burned Area (%)" ) 
tmap_save(map_burnedarea, here("Outputs", "OtherVariables", "Fire", "burnearea_2002_2021.png"),
        height = 30, width = 30, units = "cm", dpi=700)

```

#Step 3- trend calculation using theil sen regression that in turn uses kendall tau rho to determine the sen slope
```{r}
burnedarea<- rast(here("Outputs", "OtherVariables", "Fire", "burnedarea_5km_2002_2021.tif"))

#1- std deviation
Sys.time(); std_burnedarea <- terra::app (burnedarea, fun = sd); Sys.time()

#2- linear regression through time
lm_function <- function (x){
  Y<- 1:length(x)
  
  if(all(is.na(x))){
    return (NA)
  } else{
    lm_model<- lm(x ~ Y )
    slope_coefficient<- lm_model$coefficients[2]
    slope_coefficient
  }
}
Sys.time(); lm_slope_burnedarea <- terra::app (burnedarea, fun = lm_function); Sys.time()

#3- mann kendal correlation through time
#install.packages("Kendall")
library(Kendall)
MKtau <- function(z){
  if(all(is.na(z))){
    return (NA)
  } else{
    as.numeric(Kendall::MannKendall(z)$tau)
  }
} 
Sys.time(); kendall_tau_burnedarea<- terra::app (burnedarea, fun = MKtau); Sys.time()

#4- theil sen regression using kendall tau rho correlation coefficient- https://projects.itrcweb.org/gsmc-1/Content/GW%20Stats/5%20Methods%20in%20indiv%20Topics/5%205%20Trend%20Tests.htm
#install.packages("mblm")
library(mblm)
theilsen_function <- function(z){
  Y<- 1:length(z)
  
  if(all(is.na(z))){
    return (NA)
  } else{
    theilsen_model<- mblm::mblm(z ~ Y, repeated = FALSE)
    slope_coefficient <- theilsen_model$coefficients[2]
    slope_coefficient
  }
} 
Sys.time(); theilsen_burnedarea <- terra::app (burnedarea, fun = theilsen_function); Sys.time()
#writeRaster(theilsen_burnedarea, here("Outputs", "OtherVariables", "Fire", "theilsen_burnedarea.tif"))

#Map making
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

std_burnedarea
lm_slope_burnedarea
theilsen_burnedarea

mapping_function <- function (anomoly_raster, title){
  x_map<- tm_shape(cerrado)+ tm_borders()+
    tm_shape (anomoly_raster) + 
    tm_raster(col.scale = tm_scale_continuous(values = "YlOrRd"),
              col.legend = tm_legend (title = title)) +
    tm_layout(legend.text.size = 1, title.size = 4) 
  x_map
}
Sys.time();std_map <- mapping_function (std_burnedarea, title = "Std"); Sys.time()
Sys.time();lm_map <- mapping_function (lm_slope_burnedarea, "Linear Regression"); Sys.time()
Sys.time();kendaltau_map <- mapping_function (kendall_tau_burnedarea, "MannKendall"); Sys.time()
Sys.time();theilsen_map <- mapping_function (theilsen_burnedarea , "Theil-Sen Regression"); Sys.time()

anomoly_results_map<- tmap_arrange(std_map, lm_map, kendaltau_map, theilsen_map, ncol = 4, nrow= 1) 
tmap_save(anomoly_results_map, here("Outputs", "OtherVariables", "Fire", "anomoly_methods_results2.png"),
        height = 50, width = 50, units = "cm", dpi=700)
```


<<<<<<<<<<<<<<<<<<<<<<< ARCHIVE - I need to deal with the below code chunks later
#CVNP analyses
```{r}
modis_fire<- rast(here("Data", "Fire", "MODIS", "modis_fire_2002_2021.tif"))
cvnp_shp<- st_read(here("Data", "Admin", "WDPA_WDOECM_Jan2024_Public_59_shp","WDPA_WDOECM_Jan2024_Public_59_shp_0", "WDPA_WDOECM_Jan2024_Public_59_shp-polygons.shp")) #cannot remember where this polygon comes from, maybe GEE?

#CVNP MODIS fire 
cvnp_modis_fire_crop<- terra::crop(modis_fire, vect(cvnp_shp))
cvnp_modis_fire_mask<- terra::mask(cvnp_modis_fire_crop, vect(cvnp_shp))
remove(cvnp_modis_fire_crop)
names(cvnp_modis_fire_mask)<- c("2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010",
                                "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019",
                                "2020", "2021")
writeRaster(cvnp_modis_fire_mask, here("Outputs", "OtherVariables", "CVNP", "cvnp_modis_fire2002_2021.tif"))

#Anmation map of fires 2002- 2021
animation_x_time_map<-
  tm_shape(cvnp_shp)+ tm_fill()+
  tm_shape (cvnp_modis_fire_mask)+
  tm_raster(style = "cont", get_brewer_pal("Oranges"),title="Fire occurence") +
  tm_layout (legend.position = c("left", "top"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(cvnp_shp)+ tm_borders() 
tmap_animation(animation_x_time_map, filename = here("Outputs", "Final_TrajShape_Models", "CVNP_results", "modisfire_2002_2021.gif"),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150)


#Map of pixels that burned before 2017, in 2017 and after 2017
before2017_cvnp_fire<- terra::app(cvnp_modis_fire_mask[[1:15]], fun="sum", na.rm=T)
before2017_cvnp_fire[before2017_cvnp_fire>=1]<-1
#writeRaster(before2017_cvnp_fire, here("Scratch", "before2017fire.tif"))
year2017_cvnp_fire<- cvnp_modis_fire_mask[[16]]
#writeRaster(year2017_cvnp_fire, here("Outputs", "Final_TrajShape_Models", "CVNP_results", "modis_fire_only2017.tif"))
after2017_cvnp_fire<- terra::app(cvnp_modis_fire_mask[[17:20]], fun="sum", na.rm=T)
after2017_cvnp_fire[after2017_cvnp_fire>=1]<-1
#writeRaster(after2017_cvnp_fire, here("Scratch", "after2017fire.tif"))

cvnp_2017beforeafter_map<-
  tm_shape(cvnp_shp)+ tm_fill()+
  tm_shape (before2017_cvnp_fire)+
  tm_raster(col = "sum", style = "cat", palette ="grey", legend.show = F) +
  tm_shape (year2017_cvnp_fire)+
  tm_raster(col = "15_BurnDate_sum", style = "cat", palette ="red", alpha=0.5, legend.show = F) +
  tm_shape (after2017_cvnp_fire)+
  tm_raster(col = "sum", style = "cat", palette ="blue", alpha=0.3,  legend.show = F) +
  tm_layout(main.title="CVNP MODIS fire occurence",legend.text.size = 0.8)+
  tm_add_legend(type = "fill", 
                labels = c("before 2017", 
                           "year 2017",
                           "after 2017"),
                col = c("grey", "red", "blue"))   
tmap_save(cvnp_2017beforeafter_map, here("Outputs", "Final_TrajShape_Models", "CVNP_results", "cvnp_fire_beforeafteryear2017.png"),
        height = 30, width = 30, units = "cm", dpi=700)

cvnp_2017fire_map<-
  tm_shape(cvnp_shp)+ tm_fill()+
  tm_shape (year2017_cvnp_fire)+
  tm_raster(col = "15_BurnDate_sum", style = "cat", palette ="red", alpha=0.5, legend.show = F) +
  tm_layout(main.title="CVNP MODIS fire occurence 2017 only",legend.text.size = 0.8)
tmap_save(cvnp_2017fire_map, here("Outputs", "Final_TrajShape_Models", "CVNP_results", "cvnp_fire_year2017.png"),
        height = 30, width = 30, units = "cm", dpi=700)




```





