```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I calcualte the distance to the closest anthropic land use pixel.
I use the same inputs from 1_studyarea_delineation.Rmd ie I use the output
from the GEE script https://code.earthengine.google.com/f7e17500a39018685d022f079335ac51

I use the Mapbiomas vegetation product that is commonly used across Brazil and its biomes. 
This product is at 30m spatial resolution, derived from Landsat imagergy. In GEE, I completed 
processing this data such that the resulting product is a multiband image at 1km and 5km resolution,
where each band denotes a year (1985-2023) and each pixel value is a proportion. The 
proportion is the % of total Mapbiomas pixels within the 1km & 5km pixel that are anthropic (as per
Mapbiomas typology)
https://code.earthengine.google.com/f7e17500a39018685d022f079335ac51

Data input- crop/mask to Cerrado and set time period 2002-2021 (rename layers to 2002,2003...2021)
```{r}
#Border input
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

#Mapbiomas mask
# mapbiomas_anthropic_1<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask.tif"))
# Sys.time(); mapbiomas_crop<- terra::crop(mapbiomas_anthropic, cerrado)
# mapbiomas_mask<- terra::mask(mapbiomas_crop, vect(cerrado)); Sys.time()
# remove(mapbiomas_anthropic, mapbiomas_crop)

mapbiomas_anthropic_5<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask_5km.tif"))
Sys.time(); mapbiomas_crop<- terra::crop(mapbiomas_anthropic_5, cerrado)
mapbiomas_mask<- terra::mask(mapbiomas_crop, vect(cerrado)); Sys.time()
remove(mapbiomas_anthropic_5, mapbiomas_crop)
mapbiomas_mask

#Timepriod 2002-2021
mapbiomas_mask_timeperiod<- mapbiomas_mask[[18:37]]
remove(mapbiomas_mask)
names(mapbiomas_mask_timeperiod)<- c("2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010",
                                 "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019",
                                 "2020", "2021")
mapbiomas_mask_timeperiod
```

Anthropic mask processing
The above chunk output has annual raster where each pixel value is the proportion of the pixel
that is anthropic. Currently, I have used the thresholds such that anything more than
10%, 20% and 30% anthropic is excluded. So I calculate distances to any pixel with greater than
these 3 thresholds 

```{r}
m1 <- c(0.1 ,1, NA)
m2<- c(0.2,1,NA)
m3<- c(0.3,1,NA)

classify_function<- function(thresh_classification){
  rclmat <- matrix(thresh_classification, ncol=3, byrow=TRUE)
  terra::classify(mapbiomas_mask_timeperiod, rclmat, others=0)
}

reclass_anthropic_10<- classify_function(m1)
reclass_anthropic_20<- classify_function(m2)
reclass_anthropic_30<- classify_function(m3)

```

Distance to closest anthropic land use pixel
```{r}
distance_function<- function(reclass_anthropic_raster){
  x_dist<- terra::gridDist(reclass_anthropic_raster, target=NA)
  x_dist_crop<- terra::crop(x_dist, cerrado)
  x_dist_mask<- terra::mask(x_dist_crop, vect(cerrado))
  meandistanthropic<- terra::app(x_dist_mask, fun="mean")
}

meandistanthropic_10_5km<- distance_function(reclass_anthropic_10)
writeRaster(meandistanthropic_10_5km, here("Outputs", "OtherVariables", "AnthropicDist", "5km_meandist_10thresholdsanthropicpixel.tif"), overwrite=T)
meandistanthropic_20_5km<- distance_function(reclass_anthropic_20)
writeRaster(meandistanthropic_20_5km, here("Outputs", "OtherVariables", "AnthropicDist", "5km_meandist_20thresholdsanthropicpixel.tif"), overwrite=T)
meandistanthropic_30_5km<- distance_function(reclass_anthropic_30)
writeRaster(meandistanthropic_30_5km, here("Outputs", "OtherVariables", "AnthropicDist", "5km_meandist_30thresholdsanthropicpixel.tif"), overwrite=T)

map_distanthropic_function<- function (distance_raster, title){
  tm_shape(cerrado)+ tm_fill(title="")+
  tm_shape (distance_raster)+
  tm_raster(style = "cont", palette ="Reds", title=title)  
}   

map_distanthropic_10<- map_distanthropic_function(meandistanthropic_10_5km, "Distance to >10%anthropic pixel(m)")
map_distanthropic_20<- map_distanthropic_function(meandistanthropic_20_5km, "Distance to >20%anthropic pixel(m)")
map_distanthropic_30<- map_distanthropic_function(meandistanthropic_30_5km, "Distance to >30%anthropic pixel(m)")

all_maps<- tmap_arrange(map_distanthropic_10,map_distanthropic_20,map_distanthropic_30)  
tmap_save(all_maps, here("Outputs", "OtherVariables", "AnthropicDist", "map_5km_meandist_thresholdsanthropicpixel.png"),
        height = 40, width = 40, units = "cm", dpi=700)
```
