```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I calcualte the distance to the closest anthropic land use pixel.
I use the same inputs from 1_studyarea_delineation.Rmd ie I use the output
from the GEE script https://code.earthengine.google.com/f7e17500a39018685d022f079335ac51

I use the Mapbiomas vegetation product that is commonly used across Brazil and its biomes. 
This product is at 30m spatial resolution, derived from Landsat imagergy. In GEE, I completed 
processing this data such that the resulting product is a multiband image at 1km resolution,
where each band denotes a year (1985-2023) and each pixel value is a proportion. The 
proportion is the % of total Mapbiomas pixels within the 1km pixel that are anthropic (as per
Mapbiomas typology)
https://code.earthengine.google.com/f7e17500a39018685d022f079335ac51

Data input- crop/mask to Cerrado and set time period 2002-2021 (rename layers to 2002,2003...2021)
```{r}
#Border input
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

#Mapbiomas mask
mapbiomas_anthropic<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask.tif"))
Sys.time(); mapbiomas_crop<- terra::crop(mapbiomas_anthropic, cerrado)
mapbiomas_mask<- terra::mask(mapbiomas_crop, vect(cerrado)); Sys.time()
remove(mapbiomas_anthropic, mapbiomas_crop)

#Timepriod 2002-2021
mapbiomas_mask_timeperiod<- mapbiomas_mask[[18:37]]
remove(mapbiomas_mask)
names(mapbiomas_mask_timeperiod)<- c("2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010",
                                 "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019",
                                 "2020", "2021")
mapbiomas_mask_timeperiod
```

Anthropic mask processing
The above chunk output has annual raster where each pixel value is the proportion of the pixel
that is anthropic. Since the natural area mask threshold is 10% (0.1), then any pixel
with >0.1 value is considered anthropic and hence distance can be calculated to these pixels
```{r}
m <- c(0.1 ,1, NA)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
reclass_anthropic<- terra::classify(mapbiomas_mask_timeperiod, rclmat, others=0)
```

Distance to closest anthropic land use pixel
```{r}
Sys.time(); trial<- terra::gridDist(reclass_anthropic, target=NA); Sys.time()

trial_crop<- terra::crop(trial, cerrado)
trial_mask<- terra::mask(trial_crop, vect(cerrado))
#writeRaster(trial_mask, here("Scratch", "distanthropic1.tif"))

meandistanthropic<- terra::app(trial_mask, fun="mean")
#writeRaster(meandistanthropic, here("Outputs", "OtherVariables", "AnthropicDist", "meandist_anthropicpixel.tif"))

map_distanthropic<-   
  tm_shape(cerrado)+ tm_fill(title="")+
  tm_shape (meandistanthropic)+
  tm_raster(style = "cont", palette ="Reds", title="Mean distance to anthropic use (m)")  
tmap_save(map_distanthropic, here("Outputs", "OtherVariables", "AnthropicDist", "map_meandist_anthropicpixel.png"),
        height = 30, width = 30, units = "cm", dpi=700)
```
