```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script I calculate relative entropy. 

Firstly, using the analyses that Marina's group of Luis Catellan and Caio Mattos
have done, I know that for the Cerrado IMERG and CHIRPS rainfall products
are the best; howver CHIRPS does not chracterise interannual rainfall variability.
I use CHIRPS data because, at this stage, I want to take advantage of the time period
ie data from 1981 as opposed to IMERGE which is from 2000. Also, CHRIPS has finer
spatial resolution that IMERG. I processed monthly rainfall for every month 1981-2021
in GEE-https://code.earthengine.google.com/63b7aa4828e2d7046774a2f12c85a831

##Data input- mosaicing- DO NOT RUN, OUTPUT WRITTEN OUT
```{r}


gee_chirps<- rast(here("Data", "Climate", "Rainfall", "CHIRPS", "CHIRPS_monthlyrainfall.tif"))
gee_chirps
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

crop_rainfall<- terra::crop(gee_chirps, vect(cerrado))
Sys.time(); mask_rainfall<- terra::mask(crop_rainfall, vect(cerrado)); Sys.time()#half hour
writeRaster(mask_rainfall, here("Data", "Climate", "Rainfall", "cerrado_chirps_rainfall_1981_2021.tif"))

remove(crop_rainfall, gee_chirps)
```


```{r}
years<-  rep(1981:2021,each=12) 
months<- rep(1:12,41)
raster_names<- vector()
for (i in 1:length(years)){
  name<- paste0(years[[i]],"_", months[[i]])
  raster_names[i]<-name
}
names(mask_rainfall)<- raster_names
remove(i, months, name, raster_names, years)
```

##Processing-
Adapting relative entropy code from Schwartz et al., 2020 
```{r}
#Feng's seasonality index requires a value for ARmax, or maximum annual rainfall in the dataset. This should be the wettest site if analyzing sites across space, or the wettest year, if analyzing multiple years of the same site.
annual_years<- 1981:2021
annual_df<- tibble(Year= annual_years) 
rainfall_subset_stack<- function (Year){
  terra::subset(mask_rainfall, grep(Year, names(mask_rainfall)))
}

annual_df<-annual_df %>% mutate(monthlyrainfall_by_year= map(Year,rainfall_subset_stack))

annual_rainfall_stack<- function (rs){
  terra::app(rs,sum)
}
Sys.time();annual_df<-annual_df %>% mutate(annual_rainfall=map(monthlyrainfall_by_year, annual_rainfall_stack)); Sys.time() 

annual_rainfall_list<- annual_df$annual_rainfall
r<-rast(annual_rainfall_list)
armax<- terra::app(r, max)
remove(annual_rainfall_list,r)

seasonalityIndexFeng <- function(rainSeries){  
  AR <- sum(rainSeries)
  pm <- rainSeries/AR
  D <- sum(pm*log2((pm+0.0001)/(1/12)))
  seas <- D*AR/armax
  return(seas)
}

Sys.time();annual_df<-annual_df %>% mutate(relative_entropy=map(monthlyrainfall_by_year, seasonalityIndexFeng)); Sys.time() #27 min

annual_relative_entropy<- annual_df$relative_entropy
annual_relative_entropy<- rast(annual_relative_entropy)

remove(annual_rainfall_stack, rainfall_subset_stack, seasonalityIndexFeng)
```


Map of relative entropy and output annual relative entropy
```{r}
crop_relative_entropy<- terra::crop(annual_relative_entropy, vect(cerrado))
mask_relative_entropy<- terra::mask(crop_relative_entropy, vect(cerrado))
remove(crop_relative_entropy)
#writeRaster(mask_relative_entropy, here("Outputs", "OtherVariables", "Climate","annual_relative_entropy_1981_2021.tif"))
raster_names<- vector()
for (i in 1:length(annual_years)){
  name<- paste0("RE_",annual_years[[i]])
  raster_names[i]<-name
}
names(mask_relative_entropy)<- raster_names

#Anmation map of relative entropy 1981-2021
animation_x_time_map<-
   tm_shape(cerrado)+ tm_fill()+
   tm_shape (mask_relative_entropy)+
   tm_raster(style = "quantile", get_brewer_pal("Oranges"),title="Annual relative entropy") +
   tm_layout (legend.position = c("left", "top"))+
               tm_facets(nrow=1, ncol=1, free.coords = TRUE)+
   tm_shape(cerrado)+ tm_borders()
tmap_animation(animation_x_time_map, filename = here("Outputs", "OtherVariables", "Climate", "annual_relative_entropy_1981_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
remove(animation_x_time_map)

```

Mean RE and CV of RE
```{r}
mean_re<- terra::app(mask_relative_entropy, fun="mean")
writeRaster(mean_re, here("Outputs", "OtherVariables", "Climate","annual_relative_entropy_1981_2021.tif"))
```
