```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script I calculate relative entropy. 

Firstly, using the analyses that Marina's group of Luis Catellan and Caio Mattos
have done, I know that for the Cerrado IMERG and CHIRPS rainfall products
are the best; howver CHIRPS does not chracterise interannual rainfall variability.
I use CHIRPS data because, at this stage, I want to take advantage of the time period
ie data from 1981 as opposed to IMERGE which is from 2000. Also, CHRIPS has finer
spatial resolution that IMERG. I processed monthly rainfall for every month 1981-2021
in GEE-https://code.earthengine.google.com/63b7aa4828e2d7046774a2f12c85a831
(1) Reading data, mosaicing, cropping and masking to Cerrado
(2) Data cleaning and renaming layers
(3) Relative entropy calculation
(4) Anomaly calculation

#Step 1- Reading data, mosaicing, cropping and masking DO NOT RUN, OUTPUT WRITTEN OUT
```{r}
gee_chirps<- rast(here("Data", "Climate", "Rainfall", "CHIRPS", "CHIRPS_monthlyrainfall.tif"))
gee_chirps
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

crop_rainfall<- terra::crop(gee_chirps, vect(cerrado))
Sys.time(); mask_rainfall<- terra::mask(crop_rainfall, vect(cerrado)); Sys.time()#half hour
writeRaster(mask_rainfall, here("Data", "Climate", "Rainfall", "cerrado_chirps_rainfall_1981_2021.tif"))

remove(crop_rainfall, gee_chirps)
```

#Step 2- Processing ie renaming of data layers
```{r}
mask_rainfall<- rast(here("Data", "Climate", "Rainfall", "cerrado_chirps_rainfall_1981_2021.tif"))

years<-  rep(1981:2021,each=12) 
months<- rep(1:12,41)
raster_names<- vector()
for (i in 1:length(years)){
  name<- paste0(years[[i]],"_", months[[i]])
  raster_names[i]<-name
}
names(mask_rainfall)<- raster_names
remove(i, months, name, raster_names, years)
```

#Step 3- Adapting relative entropy code from Schwartz et al., 2020
```{r}
#Feng's seasonality index requires a value for ARmax, or maximum annual rainfall in the dataset. This should be the wettest site if analyzing sites across space, or the wettest year, if analyzing multiple years of the same site.
annual_years<- 1981:2021
annual_df<- tibble(Year= annual_years) 
rainfall_subset_stack<- function (Year){
  terra::subset(mask_rainfall, grep(Year, names(mask_rainfall)))
}

annual_df<-annual_df %>% mutate(monthlyrainfall_by_year= map(Year,rainfall_subset_stack))

annual_rainfall_stack<- function (rs){
  terra::app(rs,sum)
}
Sys.time();annual_df<-annual_df %>% mutate(annual_rainfall=map(monthlyrainfall_by_year, annual_rainfall_stack)); Sys.time() 

annual_rainfall_list<- annual_df$annual_rainfall
r<-rast(annual_rainfall_list)
armax<- terra::app(r, max)
remove(annual_rainfall_list,r)

seasonalityIndexFeng <- function(rainSeries){  
  AR <- sum(rainSeries)
  pm <- rainSeries/AR
  D <- sum(pm*log2((pm+0.0001)/(1/12)))
  seas <- D*AR/armax
  return(seas)
}

Sys.time();annual_df<-annual_df %>% mutate(relative_entropy=map(monthlyrainfall_by_year, seasonalityIndexFeng)); Sys.time() 

annual_relative_entropy<- annual_df$relative_entropy
annual_relative_entropy<- rast(annual_relative_entropy)

crop_relative_entropy<- terra::crop(annual_relative_entropy, vect(cerrado))
mask_relative_entropy<- terra::mask(crop_relative_entropy, vect(cerrado))
remove(crop_relative_entropy)
#writeRaster(mask_relative_entropy, here("Outputs", "OtherVariables", "Climate","annual_relative_entropy_1981_2021.tif"))

remove(annual_rainfall_stack, rainfall_subset_stack, seasonalityIndexFeng)

#map
raster_names<- vector()
for (i in 1:length(annual_years)){
  name<- paste0("RE_",annual_years[[i]])
  raster_names[i]<-name
}
names(mask_relative_entropy)<- raster_names

#Anmation map of relative entropy 1981-2021
animation_x_time_map<-
   tm_shape(cerrado)+ tm_borders()+
   tm_shape (mask_relative_entropy)+
   tm_raster(col.scale = tm_scale_intervals(n=5, style= "quantile", values = "Oranges")) +
   tm_facets(nrow=1, ncol=1, free.coords = TRUE) 

tmap_animation(animation_x_time_map, filename = here("Outputs", "OtherVariables", "Climate", "annual_relative_entropy_1981_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)
remove(animation_x_time_map, raster_names)
```

#Step 4- anomoly calculation
```{r}
annual_relative_entropy<-rast(here("Outputs", "OtherVariables", "Climate","annual_relative_entropy_1981_2021.tif"))

#1- std deviation
Sys.time(); std_re <- terra::app (annual_relative_entropy, fun = sd); Sys.time()

#2- linear regression through time
lm_function <- function (x){
  Y<- 1:length(x)
  
  if(all(is.na(x))){
    return (NA)
  } else{
    lm_model<- lm(x ~ Y )
    slope_coefficient<- lm_model$coefficients[2]
    slope_coefficient
  }
}
Sys.time(); lm_slope_re <- terra::app (annual_relative_entropy, fun = lm_function); Sys.time()

#3- mann kendal correlation through time
#install.packages("Kendall")
library(Kendall)
MKtau <- function(z){
  if(all(is.na(z))){
    return (NA)
  } else{
    as.numeric(Kendall::MannKendall(z)$tau)
  }
} 
Sys.time(); kendall_tau_re <- terra::app (annual_relative_entropy, fun = MKtau); Sys.time()

#4- theilsen regression
library(mblm)
theilsen_function <- function(z){
  Y<- 1:length(z)
  
  if(all(is.na(z))){
    return (NA)
  } else{
    theilsen_model<- mblm::mblm(z ~ Y, repeated = FALSE)
    slope_coefficient <- theilsen_model$coefficients[2]
    slope_coefficient
  }
} 
Sys.time(); theilsen_re <- terra::app (annual_relative_entropy, fun = theilsen_function); Sys.time()
writeRaster(theilsen_re, here("Outputs", "OtherVariables", "Climate", "theilsen_re.tif"))

#Map making
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mapping_function <- function (anomoly_raster, title){
  x_map<- tm_shape(cerrado)+ tm_borders()+
    tm_shape (anomoly_raster) + 
    tm_raster(col.scale = tm_scale_continuous(values = "Oranges"),
              col.legend = tm_legend(title = title)) + 
    tm_layout(legend.text.size = 1, title.size = 4)
  x_map
}
Sys.time();std_map <- mapping_function (std_re, "Std"); Sys.time()
Sys.time();lm_map <- mapping_function (lm_slope_re, "Linear Regression"); Sys.time()
Sys.time();tau_map <- mapping_function (kendall_tau_re , "Mann-Kendall"); Sys.time()
Sys.time();theilsen_map <- mapping_function (theilsen_re , "Theil-Sen Regression"); Sys.time()

anomoly_results_map<- tmap_arrange(std_map, lm_map, tau_map, theilsen_map, ncol = 4, nrow= 1) 
tmap_save(anomoly_results_map, here("Outputs", "OtherVariables", "Climate", "anomoly_re_methods_results.png"), height = 50, width = 50, units = "cm", dpi=700)
```
