```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I process height above nearest drainage (HAND)

HAND is a normalized terrain model which is a good proxy for water table depth. 
It was used in Heinreich et al., 2023 (carbon stocks from secondary regrowth). 
As per Nobre et al., 2011, HAND is advantageous to other terrain metrics such as 
Topographic Wetness Index because unline TWI, HAND normalizes by topography 
consideirng soil water dynamics. 

I primarily used the Global HAND product that Heinreich used ie
https://www.researchgate.net/profile/Gennadiy-Donchyts/publication/301559649_Global_30m_Height_Above_the_Nearest_Drainage/links/5719ed8f08ae30c3f9f2cc88/Global-30m-Height-Above-the-Nearest-Drainage.pdf

Information for this product is -https://gee-community-catalog.org/projects/hand/
I use the GEE script- https://code.earthengine.google.com/f4df14a0f175a8b82cad7eeb4e2eafc9 which is mentioned in the
Data Availability statement of Heireich et al. 
The Github page of this product- https://github.com/gena/paper-global-hand?tab=readme-ov-file
This is not a peer reviewed product, but author of the product has given permission to me to use it (via GEE Google Groups query that I posted) and has said to cite the
EGU poster made for this product (on Google scholar)
GEE Google Groups query- https://groups.google.com/g/google-earth-engine-developers/c/RdEEfRt6Pzk

I have decided to use the 90m spatial resolution HAND product because anyway I need
to resample to the resolution of this project ~1 sqkm. In the 90m HAND product, the ~8 sqkm
means that 8 sqkm of upstream area was considered when calculating flow accumulation, flow path etc

My GEE script- https://code.earthengine.google.com/e0927de9dfa81c2732df2d681d6376c5

```{r}
gee_hand<- rast(here("Data", "HAND", "bufferCerrado_HAND.tif"))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

#1. Corp/mask
crop_hand<- terra::crop(gee_hand, cerrado)
crop_mask<- terra::mask(crop_hand, vect(cerrado))
writeRaster(crop_mask, here("Outputs", "OtherVariables", "HAND", "cerrado_hand.tif"))
remove(crop_hand)

#2. Map
formap_hand<- crop_mask
formap_hand[formap_hand<0]<- 0
map_hand<- 
  tm_shape(cerrado)+ tm_fill(title="")+
  tm_shape (formap_hand)+
  tm_raster(n=5, style = "quantile", palette=get_brewer_pal("Purples", n = 5, plot=FALSE),
            title="Height above nearest drainage(m)" ) 
tmap_save(map_hand, here("Outputs", "OtherVariables", "HAND", "map_hand.png"),
        height = 30, width = 30, units = "cm", dpi=700)


```
