```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I process SPEI as a pulse climate disturbance driver. 
I processed SPEI in GEE-https://code.earthengine.google.com/d853defebf48fdf53d12dde88f4b083b
I used the SPEI product based on CRU data and the output that I exported out of GEE
is a multiband raster, in which each band is the monthly SPEI (Jan 2002- Dec 2021) calculated
using 12 month accumulation (of PET etc)
(1) Data processing - mosaic, rename layers, crop/mask to cerrado
(2) Collapsing monthly spei to annual by calculating  mean
(3) Anomaly calculation- linear regression trends, kendal tau trends and variance/std

#Step 1 - data input
```{r}
gee_spei<- rast(here("Data", "Climate", "SPEI", "annualSPEI.tif"))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

cru_spei_crop<- terra::crop(gee_spei, vect(cerrado))
cru_spei_mask<- terra::mask(cru_spei_crop, vect(cerrado))
remove(cru_spei_crop)

years<-  rep(2002:2021,each=12) 
months<- rep(1:12,20)
raster_names<- vector()
for (i in 1:length(years)){
  name<- paste0(years[[i]],"_", months[[i]])
  raster_names[i]<-name
}
names(cru_spei_mask)<- raster_names

writeRaster(cru_spei_mask, here("Data", "Climate", "SPEI" , "cerrado_spei_annual_5km_2002_2021.tif"))

# #Anmation map of spei 2002- 2021
animation_x_time_map<-
  tm_shape(cerrado)+ tm_fill()+
  tm_shape (cru_spei_mask) +
  tm_raster(col.scale = tm_scale_intervals(n=5, style= "quantile", values = "Reds")) +
  tm_layout (legend.position = c("left", "top"))+
  tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(cerrado)+ tm_borders()
  tmap_animation(animation_x_time_map, filename = here("Data", "Climate", "SPEI" , "speiannual_2002_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)

remove(animation_x_time_map)
remove(i, months, name, years, raster_names, gee_spei)
```

#Step 2- calculation of mean of monthly SPEI values to obtain annual spei
```{r}
cru_spei_mask

annual_years<- 2002:2021
annual_df<- tibble(Year= annual_years) 
spei_subset_stack<- function (Year){
  terra::subset(cru_spei_mask, grep(Year, names(cru_spei_mask)))
}

annual_df<-annual_df %>% mutate(monthlyspei_by_year= map(Year,spei_subset_stack))

annual_spei_stack<- function (rs){
  terra::app(rs,mean)
}
Sys.time();annual_df<-annual_df %>% mutate(annual_spei=map(monthlyspei_by_year, annual_spei_stack)); Sys.time()

annual_spei_list<- annual_df$annual_spei
annual_spei<- rast(annual_spei_list)

names(annual_spei) <- annual_years
writeRaster(annual_spei, here("Outputs", "OtherVariables", "Climate", "annualspei_2002_2021_12monthsacc.tif"))

remove(annual_spei_list, annual_spei_stack, spei_subset_stack)
remove(annual_years, annual_df)
```

#Step 3- mean spei and trends calculation
```{r}
annual_spei <- rast(here("Outputs", "OtherVariables", "Climate", "annualspei_2002_2021_12monthsacc.tif"))

mean_annual_spei <- terra::app(annual_spei, fun ="mean")
writeRaster(mean_annual_spei, here("Outputs", "OtherVariables", "Climate", "mean_annualspei.tif"))

#1- std deviation
Sys.time(); std_spei <- terra::app (annual_spei, fun = sd); Sys.time()

#2- linear regression through time
lm_function <- function (x){
  Y<- 1:length(x)
  
  if(all(is.na(x))){
    return (NA)
  } else{
    lm_model<- lm(x ~ Y )
    slope_coefficient<- lm_model$coefficients[2]
    slope_coefficient
  }
}
Sys.time(); lm_slope_spei <- terra::app (annual_spei, fun = lm_function); Sys.time()

#3- mann kendal correlation through time
#install.packages("Kendall")
library(Kendall)
MKtau <- function(z){
  if(all(is.na(z))){
    return (NA)
  } else{
    as.numeric(Kendall::MannKendall(z)$tau)
  }
} 
Sys.time(); kendall_tau_spei <- terra::app (annual_spei , fun = MKtau); Sys.time()

#4- theil sen regression using kendall tau rho correlation coefficient- https://projects.itrcweb.org/gsmc-1/Content/GW%20Stats/5%20Methods%20in%20indiv%20Topics/5%205%20Trend%20Tests.htm
#install.packages("mblm")
library(mblm)
theilsen_function <- function(z){
  Y<- 1:length(z)
  
  if(all(is.na(z))){
    return (NA)
  } else{
    theilsen_model<- mblm::mblm(z ~ Y, repeated = FALSE)
    slope_coefficient <- theilsen_model$coefficients[2]
    slope_coefficient
  }
} 
Sys.time(); theilsen_spei <- terra::app (annual_spei , fun = theilsen_function); Sys.time()
writeRaster(theilsen_spei, here("Outputs", "OtherVariables", "Climate", "theilsen_spei.tif"))

#Map making
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mapping_function <- function (anomoly_raster, title){
  x_map<- tm_shape(cerrado)+ tm_borders()+
    tm_shape (anomoly_raster) + 
    tm_raster(col.scale = tm_scale_continuous(values = "YlOrRd"),
              col.legend = tm_legend (title = title)) +
    tm_layout(legend.text.size = 1, title.size = 4) 
  x_map
}
Sys.time();std_map <- mapping_function (std_spei, title = "Std"); Sys.time()
Sys.time();lm_map <- mapping_function (lm_slope_spei, "Linear Regression"); Sys.time()
Sys.time();kendaltau_map <- mapping_function (kendall_tau_spei, "MannKendall"); Sys.time()
Sys.time();theilsen_map <- mapping_function (theilsen_spei , "Theil-Sen Regression"); Sys.time()

anomoly_results_map<- tmap_arrange(std_map, lm_map, kendaltau_map, theilsen_map, ncol = 4, nrow= 1) 
tmap_save(anomoly_results_map, here("Outputs", "OtherVariables", "Climate", "anomoly_spei_5km.png"),
        height = 50, width = 50, units = "cm", dpi=700)

```

