```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script I process heateaves- across the entire time period and trends in heatwaves using annual product.
I number of heatwaves between 2002 and 2021 as per Libonatti et al., 2024 using ERA4 temperature at 2m product in GEE
https://code.earthengine.google.com/689bdfae966c1a4d69fb31a012632dc3
(1) Number of heatwaves across time period (2002 - 2021)
(2) Trends in heatwaves using number of heatwaves every year in time period

#Step 1- number of heatwaves across time period, crop/mask and map
```{r}
gee_heatwave<- rast(here("Data", "Climate", "ERA5_Heatwave.tif"))
gee_heatwave
plot(gee_heatwave)

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

heatwave_crop<- terra::crop(gee_heatwave, vect(cerrado))
heatwave_mask<- terra::mask(heatwave_crop, vect(cerrado))
remove(heatwave_crop)
writeRaster(heatwave_mask, here("Outputs", "OtherVariables", "Climate", "heatwaves_2002_2021.tif"))

map_hw<-   
  tm_shape(cerrado)+ tm_fill(title="")+
  tm_shape (heatwave_mask)+
  tm_raster(n=5, style = "quantile",palette = "Reds", title="#Heatwaves" ) 
tmap_save(map_hw, here("Outputs", "OtherVariables", "Climate", "heatwaves_2002_2021.png"),
        height = 30, width = 30, units = "cm", dpi=700)

```

#Step 2- trends in heatwaves through time period using number of heatwaves annually (calculated in gee)
```{r}
gee_annual_heatwave <- rast(here("Data", "Climate", "Annual_ERA5_Heatwave.tif"))

annual_heatwave_crop<- terra::crop(gee_annual_heatwave, vect(cerrado))
annual_heatwave_mask<- terra::mask(annual_heatwave_crop, vect(cerrado))
remove(annual_heatwave_crop)
#writeRaster(annual_heatwave_mask, here("Outputs", "OtherVariables", "Climate", "annual_heatwaves_2002_2021.tif"))

#animation
animation_x_time_map<-
   tm_shape(cerrado)+ tm_borders()+
   tm_shape (annual_heatwave_mask)+
   tm_raster(col.scale = tm_scale_intervals(n=6, style= "quantile", values = "OrRd")) +
   tm_facets(nrow=1, ncol=1, free.coords = TRUE) 

tmap_animation(animation_x_time_map, filename = here("Outputs", "OtherVariables", "Climate", "annual_heatwaves_2002_2021.gif"),
                delay = 50, loop = TRUE, restart.delay = 500,
                width = 900, height = 900, dpi = 150)

annual_heatwave_mask<- rast(here("Outputs", "OtherVariables", "Climate", "annual_heatwaves_2002_2021.tif"))

#trends or anomaly calculation
#1- std deviation
Sys.time(); std_heatwaves <- terra::app (annual_heatwave_mask, fun = sd); Sys.time()

#2- linear regression through time
lm_function <- function (x){
  Y<- 1:length(x)
  
  if(all(is.na(x))){
    return (NA)
  } else{
    lm_model<- lm(x ~ Y )
    slope_coefficient<- lm_model$coefficients[2]
    slope_coefficient
  }
}
Sys.time(); lm_slope_heatwaves <- terra::app (annual_heatwave_mask, fun = lm_function); Sys.time()

#3- mann kendal correlation through time
#install.packages("Kendall")
library(Kendall)
MKtau <- function(z){
  if(all(is.na(z))){
    return (NA)
  } else{
    as.numeric(Kendall::MannKendall(z)$tau)
  }
} 
Sys.time(); kendall_tau_heatwaves <- terra::app (annual_heatwave_mask, fun = MKtau); Sys.time()

#4- theilsen regression
library(mblm)
theilsen_function <- function(z){
  Y<- 1:length(z)
  
  if(all(is.na(z))){
    return (NA)
  } else{
    theilsen_model<- mblm::mblm(z ~ Y, repeated = FALSE)
    slope_coefficient <- theilsen_model$coefficients[2]
    slope_coefficient
  }
} 
Sys.time(); theilsen_heatwaves <- terra::app (annual_heatwave_mask, fun = theilsen_function); Sys.time()
writeRaster(theilsen_heatwaves, here("Outputs", "OtherVariables", "Climate", "theilsen_heatwaves.tif"))

#Map making
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

mapping_function <- function (anomoly_raster, title){
  x_map<- tm_shape(cerrado)+ tm_borders()+
    tm_shape (anomoly_raster) + 
    tm_raster(col.scale = tm_scale_intervals(n=5, style = "quantile", values = "OrRd"),
              col.legend = tm_legend(title = title)) +
    tm_layout(legend.text.size = 1, title.size = 4) 
  x_map
}
Sys.time();std_map <- mapping_function (std_heatwaves, "Std"); Sys.time()
Sys.time();lm_map <- mapping_function (lm_slope_heatwaves, "Linear Regression"); Sys.time()
Sys.time();tau_map <- mapping_function (kendall_tau_heatwaves, "Mann-Kendall"); Sys.time()
Sys.time();theilsen_map <- mapping_function (theilsen_heatwaves, "Theil-Sen Regression"); Sys.time()


anomoly_results_map<- tmap_arrange(std_map, lm_map, tau_map, theilsen_map, ncol =4, nrow= 1) 
tmap_save(anomoly_results_map, here("Outputs", "OtherVariables", "Climate", "anomoly_heatwaves_methods_results.png"),
        height = 50, width = 50, units = "cm", dpi=700)

```
