```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I calculate pixel heterogeneity as per White et al 2022- ie shannon's diversity index of land uses and covers using the vegan R package. 
By pixel heterogeniety, I mean natural cover variety but area weighted. So I do include anthropic covers, but I reclassify all anthropic covers to be the same anthropic cover. For example, all ag classes are under agriculture and all man-made covers are under the same class. 

##Data input
```{r}
raster_list <- list.files(path = here("Data", "Mapbiomas_raw"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
raster_list <-lapply(raster_list, rast)

```

##Processing- calculation of pixel heterogeniety
```{r}
heterogeneity_function<- function (raster, diversity_name){  
  annual_years<- 2002:2021
  annual_df<- tibble(Year= annual_years) 

  subset<- function (Year){
    terra::subset(raster, grep(Year, names(raster)))
  }

  annual_df<-annual_df %>% mutate(annual_lulc= map(Year,subset))
  
  message("df created")
#Reclassification - the excel sheet is a classification I manually made using Mapbiomas 9 classificaiton
  reclass_table<- read.csv(here("Data", "mapbiomas_legend_collection9_classification_pixelheterogeniety.csv"))
  reclass_table<- reclass_table %>% dplyr::select(c("Pixel.value", "Reclass"))
  reclassify<- function(raster_tobe_reclassified){
    classified_raster<- terra::classify(raster_tobe_reclassified, reclass_table, others=NA)
  }

  Sys.time(); annual_df<-annual_df %>% mutate(reclassified_lulc= map(annual_lulc, reclassify)); Sys.time() #6min
  
  message("classification completed")
#Percentage of 5x5km pixel covered by each natural cover type
  mat<- rast(here("Outputs", "OtherVariables", "Climate", "mat_1981_2021.tif"))
  prop_aggregate <- function(reclassified_raster) {
    terra::project(
      terra::segregate(reclassified_raster), mat,# segregate()- binary (0/1) of classes 0,3,4,12
      method = "average", res = res(mat)[1]
    )
  }

  Sys.time(); annual_df<-annual_df %>% mutate(coarsened_split_lulc= map(reclassified_lulc, prop_aggregate)); Sys.time()

  message("coarsening complete")
##Shannon's diversity index
  library(vegan)
  shannon_diversity<- function (reclass_coarse_raster){
    x_df<- as.data.frame(reclass_coarse_raster, cells=TRUE, xy=TRUE)
    x_diversity<- vegan::diversity(x_df %>% dplyr::select(-c(cell,x,y)), index="shannon")
    x_diversity<- as.data.frame(x_diversity)
    x_df<- bind_cols(x_df, x_diversity)
    trial_vector<- terra::vect(x_df %>% dplyr::select(c("cell", "x", "y", "x_diversity")), geom=c("x", "y"), crs="epsg:4326")
    trial_raster<- terra::rasterize(trial_vector, reclass_coarse_raster, "x_diversity", fun="max")
    trial_raster
  }
  Sys.time(); annual_df<-annual_df %>% mutate(diversity= map(coarsened_split_lulc, shannon_diversity)); Sys.time()

  message("diversity index complete")
  diversity_stack<- rast(annual_df$diversity)
  names(diversity_stack)<- annual_df$Year
  writeRaster(diversity_stack, here("Outputs", "OtherVariables", "PixelHeterogeniety", paste0(diversity_name, ".tif")))
  message("diversity raster written to disk")
}


Sys.time(); for ( i in 1:length(raster_list)){
  print (paste0("raster # ", i))
  
  heterogeneity_function(raster_list[[i]], paste0("diversity_",i))

}; Sys.time()

```

##Mosaicing, figures 
```{r}
raster_list <- list.files(path = here("Outputs", "OtherVariables","PixelHeterogeniety"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
raster_list <-lapply(raster_list, rast)

tic(); diversity<- do.call(mosaic,raster_list); toc() 
#writeRaster(diversity, here("Outputs", "OtherVariables","PixelHeterogeniety", "mosaiced_diversity.tif"))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
cerrado<- d_trans %>% st_union()

diversity_crop<- terra::crop(diversity, vect(cerrado))
diversity_mask<- terra::mask(diversity_crop, vect(cerrado))
remove(diversity_crop)
#writeRaster(diversity_mask,here("Outputs", "OtherVariables","PixelHeterogeniety", "cerrado_diversity_2002_2021.tif"))


#Anmation map of diversity 2002- 2021
animation_x_time_map<-
  tm_shape(cerrado)+ tm_fill()+
  tm_shape (diversity_mask)+
  tm_raster(n=5, style = "quantile", palette = ("RdPu"),title = "PixelDiversity") +
  tm_layout (legend.position = c("left", "top"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(cerrado)+ tm_borders()
tmap_animation(animation_x_time_map, filename = here("Outputs", "OtherVariables", "PixelHeterogeniety", "diversity_2002_2021.gif"),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150)
remove(animation_x_time_map)



```


