```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I complete preprocessing and EDA of the temporal monthly anisoEVI product (MODIS based), which I processed in GEE. 

##Data input
```{r}
#evi_list <- list.files(path = here("Data", "Indices", "AnisoEVI_Cerrado_GEE"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
#evi_raster_list<-lapply(evi_list, rast)
#tic(); evi<- do.call(mosaic,evi_raster_list); toc() #4.2 hours
#writeRaster(evi, here("Data", "Indices", "AnisoEVI_Cerrado_GEE", "mosaiced_anisoEVI.tif"))

evi<- rast(here("Data", "Indices", "AnisoEVI_Cerrado_GEE", "mosaiced_anisoEVI.tif"))
tic();evi<- evi/10^5;toc() #scaling factor mentioned on the description page of the GEE asset

##changing band names to year and months
colnames<- list()
tic(); for ( i in 1:nlyr(evi)){
  x<- names(evi)[[i]]
  y<- paste0(strsplit(x, "_")[[1]][5], "_",strsplit(x, "_")[[1]][6] )
  colnames[[i]] <-y
}; toc()
colnames<- unlist(colnames)
names(evi)<- colnames

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)

cerrado<- d_trans %>% st_union()
cerrado_area<- terra::expanse(vect(cerrado), unit= "km") #2040045 sqkm which is 204 Mha

buffer<-st_read(here ("Data", "Cattelanetal_Clustering", "buffer_climate_zones.shp"))

tic();evi<-terra::mask(evi, vect(buffer)); toc()
```

##Monthly anisoEVI time series map
```{r}
library(magick) #for animation

neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))

map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()

tic(); agg_anisoevi<- terra::aggregate(evi, fact = 3 , fun= "median");toc()

tmap_mode("plot")
aniso_evi_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape( agg_anisoevi)+
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="anisoEVI") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()
  

tic();tmap_animation(aniso_evi_map, filename = here("Outputs", "Indices","anisoEVI", "monthly_anisoEVI_map2.gif"),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150) ;toc() 


remove(agg_anisoevi, map_extent, neighbors)
```


##Extract mean aniso evi values in each cliamte zone across the time period (monthly aniso evi product 2000-2021)
```{r}
regions<- d_trans$region

tic(); trial<- terra::zonal(aniso_evi, vect(d_trans), func = "mean", na.rm= T); toc() #2 min
names(trial)<- colnames #colnames from 1st chunk
trial<- trial %>% mutate(Region= d_trans$region)

#plot
aniso_evi_pivot<- pivot_longer(trial, cols= 1:261)
anisoevi_region_plot<- ggplot(aniso_evi_pivot, aes(name, value, group= Region, colour= Region)) +
           geom_line(na.rm=TRUE) +
  ylab("Mean anisoEVI") + xlab("Time")+
  scale_color_manual(values= c("#1b9e77", "#d95f02", "#7570b3"))+
  theme_pubr()+theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
ggsave(here("Outputs", "AnisoEVI", "anisoEVI_2000_2021_monthly_byclimatezone.png"),anisoevi_region_plot, dpi=300)

```

## Area estimates- for each month in the time period, calculating how much of 204 Mha has an anisoEVI value. 
Estimating how much of the three climate zones, there is anisoEVI information (because anisoEVI is not wall to wall)
```{r}
##converting all pixels that have an anisoEVI value to be 1 (only area needed, we do not need to know the actual anisoEVI value)
tic(); coverage<- aniso_evi; toc()
tic(); coverage[!is.na(coverage)]<-1;toc()# 7.5 min
##getting area of all pixel values=1
tic(); area<- terra::cellSize(coverage, mask=T, lyrs=T, unit= "km"); toc()
names(area)<- colnames

area_df<- as.tibble(colnames)
area_df<- area_df %>% separate_wider_delim(cols = value, delim = "_", names = c("Year","Month"))
area_df<- area_df %>% mutate(Year_Month=colnames)

area_stack<- function (Year_Month){
  terra::subset(area, grep(Year_Month, names (area)))
}
tic(); area_df<-area_df %>% mutate(monthlyarea = map(Year_Month, area_stack)); toc()

sum_area<- function (area_raster){
  sum(terra::values(area_raster), na.rm=T)
}
tic(); area_df<-area_df %>% mutate(EVI_area_sqkm = map(monthlyarea, sum_area)); toc() #25 min

area_df<- area_df %>% mutate(TotalCerradoArea_sqkm= cerrado_area)
area_df<- area_df %>% dplyr::select(-monthlyarea)
area_df <- area_df %>% mutate(EVI_area_sqkm=unlist(EVI_area_sqkm))
area_df<- area_df %>% mutate(proportion= (EVI_area_sqkm/TotalCerradoArea_sqkm)*100)

x<-ggplot(data=area_df, aes(x=Year_Month, y=proportion)) +
  geom_bar(stat="identity")+  theme_minimal()+
  xlab("Time Step") + ylab("Percentage of total area of Cerrado/n for which there is anisoEVI information")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
ggsave(here("Outputs", "Indices","anisoEVI", "anisoEVI_cerrado_area_proportion.png"),x, dpi=300) 


tic(); coverage[is.na(coverage)]<-0; toc() #6.3 min
coverage<- terra::mask(coverage, d_trans)
tic(); coverage_sum<- terra::app(coverage, fun= "sum"); toc()


#tmap_mode("plot")
#coverage_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
#  tm_shape(d_trans)+ tm_fill()+
#  tm_shape( coverage_sum)+
#  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="Coverage") + 
#              tm_layout (legend.position = c("left", "bottom"))
#tmap_save(coverage_map, here("Outputs", "Indices", "aniso_evi_coverage.png"))

total_cells<-length(terra::cells(coverage_sum)) #2139977
tic(); coverage_df<- terra::as.data.frame(coverage_sum); toc()

pixels_with_value_plot<- ggplot(coverage_df, aes(x=sum)) + 
  geom_histogram(color= "black", fill= "lightblue")+
  xlab("Temporal anisoEVI sum")+ylab("Count")
ggsave(here("Outputs", "Indices", "anisoEVI_count_histo.png"),pixels_with_value_plot, dpi=300)  

```




##anisoEVI range across known veg types (savanna, dry forest and gallery forest from Lewis et al., 2022 &  Cure et al., 2023)
Marina's collegue Marcio Cure sent to me via email the GPS points of the three veg types in their EVI2 Coupling paper.  

I combine/union lewis et al and cure et al points. 

```{r}
cure_points<- read.csv(here("Data", "Cureetal2023_EVI2Coupling_Chapada", "coordenadas_cure_et_al(2023).csv"))
cure_points %>% group_by(Vegetation) %>% summarise(count=n()) #10 points of each vegetation type
#cure_points<- cure_points %>% mutate(PointID= 1:dim(cure_points)[1])
cure_points_shp<- st_as_sf(cure_points, coords = c("x", "y"))
cure_points_shp<- st_set_crs(cure_points_shp, 4326)
#st_write (cure_points_shp, here("Data", "Cureetal2023_EVI2Coupling_Chapada", "curepoints.shp"))

#to match lewis et al., 2022, I rename the Vegetaion column as Formation with savanna as savanna and dry forests as Forest and Woodland
cure_points_shp<- cure_points_shp %>% mutate(Name= case_when( Vegetation=="Savanna"~"Savanna",
                                                              Vegetation=="Dry Forest"~"mata_seca",
                                                              Vegetation=="Gallery Forest"~"mata_de_galeria"))

cure_points_shp<- cure_points_shp %>% mutate(Vegetation= case_when(Vegetation=="Savanna"~"Savanna",
                                                              Vegetation=="Dry Forest"~"Forest and Woodland",
                                                              Vegetation=="Gallery Forest"~"Forest and Woodland"))
  
names(cure_points_shp)[1]<- "Formation"
                                             

lewis_points<- st_read(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "prepped_points.shp"))

points<- bind_rows(cure_points_shp, lewis_points)
points<- points %>% mutate(PointID= 1:dim(points)[1])

remove(cure_points, cure_points_shp, lewis_points)
```



```{r}
tic();anisoEVI_points<- terra::extract(aniso_evi, points, method = "simple", xy=TRUE); toc()
anisoEVI_points<- dplyr::inner_join(anisoEVI_points, (points %>% st_drop_geometry()), by=join_by(ID==PointID) )

anisoEVI_points_veg<- anisoEVI_points %>% filter(Formation != "Agriculture"& Formation != "Water"& Formation != "Plantation"& Formation!="Pasture"& Formation!="Non Veg")

## boxplot showing anisoEVI (through the years) for each point in each in each point in Cure et al 2023
#p<-ggplot(anisoEVI_points_pivot, aes(x=factor(ID), y=value, fill= factor(Formation))) +
#  geom_boxplot() +
#  scale_fill_brewer(palette="Dark2") +
#  theme_classic() + 
#  theme(legend.title = element_blank()) + 
#  ylab("anisoEVI 2000-2021") +xlab("Points from Cure et al., 2023")
#ggsave(here("Outputs", "Indices", "anisoEVI_points.png"), p)  
#remove(p, anisoEVI_cure_pivot)

#Making Fig 4 from anisoEVI paper i.e. time series of anisoEVI but instead of location, I have formation and name type
# For this I will calcualte the mean anisoEVI value for each time period across all points for the three veg types 
fig4_df<- anisoEVI_points_veg %>% dplyr::select(-c(x,y, ID)) %>%
  group_by(Formation) %>% summarise(across(everything(), list(mean)))
fig4_df_pivot<- pivot_longer(fig4_df, cols = 2:262)
q<-ggplot(fig4_df_pivot, aes(x=name, y=value, group= Formation)) +
  geom_line() + 
  theme_classic()+ 
  theme(axis.text.x = element_blank(), axis.ticks = element_blank())+
  facet_wrap(~Formation, ncol = 1, scales = "free")+ ylab("anisoEVI") +xlab("Time in months (2000-2021)")
#ggsave(here("Outputs", "Indices", "anisoEVI_points_formation_throughtime.png"), q)  
remove(q, fig4_df, fig4_df_pivot)
fig4_df<- anisoEVI_points_veg %>% dplyr::select(-c(x,y, ID)) %>%
  group_by(Name) %>% summarise(across(everything(), list(mean)))
fig4_df_pivot<- pivot_longer(fig4_df, cols = 2:262)
a<-ggplot(fig4_df_pivot, aes(x=name, y=value, group= Name)) +
  geom_line() + 
  theme_classic()+ 
  theme(axis.text.x = element_blank(), axis.ticks = element_blank())+
  facet_wrap(~Name, ncol = 1, scales = "free")+ ylab("anisoEVI") +xlab("Time in months (2000-2021)")
ggsave(here("Outputs", "Indices", "anisoEVI_points_physiognomy_throughtime.png"), a) 
remove(fig4_df, fig4_df_pivot, a)

#Making Fig 5 from anisoEVI paper i.e. monthly time series but instead of phenoregions across Amazon, I have formation and name type
# For this I will calculate mean anisoEVI for all Jans, all Febs etc for all points in each vegetation type (with 10 points I should be able to extract 95% CI)
x<- anisoEVI_points_veg %>% pivot_longer(cols = starts_with("20")) %>% separate_wider_delim(cols = name, delim = "_", names = c("Year", "Month"))
formation_month_evi<- x %>% group_by(Formation, Month) %>% summarise(MeananisoEVI=mean(value, na.rm=T),
                                                              SDEVI= sd(value, na.rm=T),
                                                              nEVI= n())%>%
  mutate(seEVI= SDEVI/sqrt(nEVI),
         lower.ci= MeananisoEVI - qt(1 - (0.05 / 2), nEVI - 1) * seEVI,
         upper.ci= MeananisoEVI + qt(1 - (0.05 / 2), nEVI - 1) * seEVI)

p<-ggplot(data=formation_month_evi, aes(x=Month, y=MeananisoEVI, group=Formation, colour= Formation)) + 
  geom_line() + scale_color_brewer(palette = "Dark2")+
  geom_ribbon(aes(ymin=formation_month_evi$lower.ci, ymax=formation_month_evi$upper.ci, fill= Formation), linetype=2, alpha=0.1)+
  scale_fill_brewer(palette = "Dark2") +theme_classic() +theme(legend.title = element_blank())
#ggsave(here("Outputs", "Indices", "anisoEVI_formation_points_month.png"), p) 
remove(p, formation_month_evi)

name_month_evi<- x %>% group_by(Name, Month) %>% summarise(MeananisoEVI=mean(value, na.rm=T),
                                                              SDEVI= sd(value, na.rm=T),
                                                              nEVI= n())%>%
  mutate(seEVI= SDEVI/sqrt(nEVI),
         lower.ci= MeananisoEVI - qt(1 - (0.05 / 2), nEVI - 1) * seEVI,
         upper.ci= MeananisoEVI + qt(1 - (0.05 / 2), nEVI - 1) * seEVI)

p<-ggplot(data=name_month_evi, aes(x=Month, y=MeananisoEVI, group=Name, colour= Name)) + 
  geom_line() + scale_color_brewer(palette = "Set3")+
  geom_ribbon(aes(ymin=name_month_evi$lower.ci, ymax=name_month_evi$upper.ci, fill= Name), linetype=2, alpha=0.1)+
  scale_fill_brewer(palette = "Set3") +theme_classic() +theme(legend.title = element_blank())
#ggsave(here("Outputs", "Indices", "anisoEVI_name_points_month.png"), p) 
remove(p, name_month_evi)


```

