```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I complete preprocessing and EDA of the temporal monthly kNDVI product (MODIS based). 

##Data input
Each imported raster has 12 bands, each being kNDVI for a month of that yearly raster
```{r}
ndvi_list <- list.files(path = here("Data", "Indices", "kNDVI_Cerrado_GEE"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
ndvi_list <-gtools::mixedsort(ndvi_list ) #arranging rasters chronologically
ndvi_raster_list<-lapply(ndvi_list, rast)


d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
plot(d_trans)
d_trans<- st_transform(d_trans, crs = 4326)

kndvi<- rast(ndvi_raster_list)
names(kndvi) #all rasters are arranges chronologically
res(kndvi) #0.0044 deg which is half that of aniso EVI because I exported kndvi with resolution 500m and aniso EVI with resolution=1000m 

```

##Extract mean kNDVI evi values in each cliamte zone across the time period (monthly kNDVI product 2000-2021)
```{r}
kNDVI<- rast(ndvi_raster_list)
regions<- d_trans$region

tic(); trial<- terra::zonal(kNDVI, vect(d_trans), func = "mean", na.rm= T); toc() #10.4 min

trial<- trial %>% mutate(Region= d_trans$region)

#plot
kndvi_pivot<- pivot_longer(trial, cols= 1:274)
kndvi_region_plot<- ggplot(kndvi_pivot, aes(name, value, group= Region, colour= Region)) +
           geom_line(na.rm=TRUE) +
  ylab("Mean kNDVI") + xlab("Time")+
  scale_color_manual(values= c("#1b9e77", "#d95f02", "#7570b3"))+
  theme_pubr()+theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
ggsave(here("Outputs", "Indices", "kndvi_2000_2022_monthly_byclimatezone.png"),kndvi_region_plot, dpi=300)

```

##Monthly kNDVI time series map animation
```{r}
library(magick) #for animation
library(tmaptools)

neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))

map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()

tic(); agg_kndvi<- terra::aggregate(kndvi, fact = 6 , fun= "median");toc() #aggregating factor double that of anisoEVI takes 1 hour 16 min

tmap_mode("plot")
kndvi_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape( agg_kndvi)+
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="kNDVI") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)

tic();tmap_animation(kndvi_map, filename = here("Outputs", "Indices", "monthly_kNDVI_map.gif"),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150) ;toc() 


remove(agg_kndvi, map_extent, neighbors)

```

##Coverage of monthly kNDVI across space
```{r}
##converting all pixels that have an kndvi value to be 1 and every other NA valued pixel to be 0
tic(); coverage<- kndvi; toc()
tic(); coverage[!is.na(coverage)]<-1;toc() #40 min
tic(); coverage[is.na(coverage)]<-0; toc() #11 min
coverage<- terra::mask(coverage, d_trans)
tic(); coverage_sum<- terra::app(coverage, fun= "sum"); toc()

#tmap_mode("plot")
#coverage_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
#  tm_shape(d_trans)+ tm_fill()+
#  tm_shape( coverage_sum)+
#  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="Coverage") + 
#              tm_layout (legend.position = c("left", "bottom"))
#tmap_save(coverage_map, here("Outputs", "Indices", "kndvi_coverage.png"))


total_cells<-length(terra::cells(coverage_sum)) #8532073
tic(); coverage_df<- terra::as.data.frame(coverage_sum); toc()

pixels_with_value_plot<- ggplot(coverage_df, aes(x=sum)) + 
  geom_histogram(color= "black", fill= "lightblue")+
  xlab("Temporal kNDVI sum")+ylab("Count")
ggsave(here("Outputs", "Indices", "kndvi_count_histo.png"),pixels_with_value_plot, dpi=300) 
```

##kNDVI range across known veg types (savanna, dry forest and gallery forest from Cure et al., 2023)
Marina's collegue Marcio Cure sent to me via email the GPS points of the three veg types in their EVI2 Coupling paper.  
```{r}
cure_points<- read.csv(here("Data", "Cureetal2023_EVI2Coupling_Chapada", "coordenadas_cure_et_al(2023).csv"))
cure_points %>% group_by(Vegetation) %>% summarise(count=n()) #10 points of each vegetation type
cure_points<- cure_points %>% mutate(PointID= 1:dim(cure_points)[1])
cure_points_shp<- st_as_sf(cure_points, coords = c("x", "y"))
cure_points_shp<- st_set_crs(cure_points_shp, 4326)

tic();kndvi_cure<- terra::extract(kndvi, cure_points_shp, method = "simple", xy=TRUE); toc()
kndvi_cure<- dplyr::inner_join(kndvi_cure, (cure_points_shp %>% st_drop_geometry()), by=join_by(ID==PointID) )

kndvi_cure_pivot<- kndvi_cure %>% pivot_longer(cols=2:275)
kndvi_cure_pivot<- kndvi_cure_pivot %>% dplyr::select(-c(x,y))
## boxplot showing kNDVI (through the years) for each point in each in each of the vegetation types
p<-ggplot(kndvi_cure %>% pivot_longer(cols=2:275), aes(x=factor(ID), y=value, fill= factor(Vegetation))) +
  geom_boxplot() +
  scale_fill_brewer(palette="Dark2") +
  theme_classic() + 
  theme(legend.title = element_blank()) + 
  ylab("kNDVI 2000-2022") +xlab("Points from Cure et al., 2023")
#ggsave(here("Outputs", "Indices", "kndvi_Cure_points.png"), p)  
remove(p, kndvi_cure_pivot)

#Making Fig 4 from anisoEVI paper i.e. time series of anisoEVI but instead of location, I have vegetation type
# For this I will calcualte the mean anisoEVI value for each time period across all points for the three veg types 
fig4_df<- kndvi_cure %>% dplyr::select(-c(x,y, ID)) %>%
  group_by(Vegetation) %>% summarise(across(everything(), list(mean)))
fig4_df_pivot<- pivot_longer(fig4_df, cols = 2:275)
q<-ggplot(fig4_df_pivot, aes(x=name, y=value, group= Vegetation)) +
  geom_line() + 
  theme_classic()+ 
  theme(axis.text.x = element_blank(), axis.ticks = element_blank())+
  facet_wrap(~Vegetation, ncol = 1, scales = "free")+ ylab("kNDVI") +xlab("Time in months (2000 - 2022)")
#ggsave(here("Outputs", "Indices", "kNDVI_Cure_points_throughtime.png"), q)  
remove(q, fig4_df, fig4_df_pivot)

#Making Fig 5 from anisoEVI paper i.e. monthly time series but instead of phenoregions across Amazon, I have vegetation type
# For this I will calculate mean anisoEVI for all Jans, all Febs etc for all points in each vegetation type (with 10 points I should be able to extract 95% CI)
x<- kndvi_cure %>% pivot_longer(cols = contains("kndvi_2")) %>% separate_wider_delim(cols = name, delim = "_", names = c("kndvi","Year", "Month"))
x<- x %>% dplyr::select(-kndvi)
x_month_kndvi<- x %>% group_by(Vegetation, Month) %>% summarise(Mean_kndvi=mean(value, na.rm=T),
                                                              SDkndvi= sd(value, na.rm=T),
                                                              nkndvi= n())%>%
  mutate(sekndvi= SDkndvi/sqrt(nkndvi),
         lower.ci= Mean_kndvi - qt(1 - (0.05 / 2), nkndvi - 1) * sekndvi,
         upper.ci= Mean_kndvi + qt(1 - (0.05 / 2), nkndvi - 1) * sekndvi)
x_month_kndvi$Month<- as.integer(x_month_kndvi$Month)
x_month_kndvi <- x_month_kndvi %>% arrange(Month)

p<-ggplot(data=x_month_kndvi, aes(x=Month, y=Mean_kndvi, group=Vegetation, colour= Vegetation)) + 
  geom_line() + scale_color_brewer(palette = "Dark2")+
  geom_ribbon(aes(ymin=x_month_kndvi$lower.ci, ymax=x_month_kndvi$upper.ci, fill= Vegetation), linetype=2, alpha=0.1)+
  scale_x_continuous(breaks=seq(1, 12, 1)) +
  scale_fill_brewer(palette = "Dark2") +theme_classic() +theme(legend.title = element_blank())
ggsave(here("Outputs", "Indices", "kndvi_Cure_points_month.png"), p) 
remove(p, x, x_month_kndvi)


```