```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I delineate the study area within the Cerrado (Brazilian savanna) biome. The aim is to 
exclude all anthropic area and focus only on native savannas. However, I need to figure out how to 
delineate anthropic and native areas.

#Data processing
I use the Mapbiomas vegetation product that is commonly used across Brazil and its biomes. 
This product is at 30m spatial resolution, derived from Landsat imagergy. In GEE, I completed 
processing this data such that the resulting product is a multiband image at 1km resolution,
where each band denotes a year (1985-2023) and each pixel value is a proportion. The 
proportion is the % of total Mapbiomas pixels within the 1km pixel that are anthropic (as per
Mapbiomas typology)
https://code.earthengine.google.com/45f7edff88db11aab8a1be5f522e6653

In the first chunk below, I do some EDA and make plots to better understand anthropic area
```{r}
#Border input
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
d_trans<- d_trans %>% mutate(area_km= terra::expanse(vect(d_trans), unit="km"))
cerrado<- d_trans %>% st_union()

#Mapbiomas mask
mapbiomas_anthropic<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask.tif"))
Sys.time(); mapbiomas_crop<- terra::crop(mapbiomas_anthropic, cerrado)
mapbiomas_mask<- terra::mask(mapbiomas_crop, vect(cerrado)); Sys.time()
remove(mapbiomas_anthropic, mapbiomas_crop)

#1. Animation of anthropic areas through time
library(magick) #for animation

neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))

map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()

tmap_mode("plot")
mapbiomas_animation_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape( mapbiomas_mask)+
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="%pixel that is anthropic") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()
  
tic();tmap_animation(mapbiomas_animation_map, filename = here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "Mapbiomass_anthropic_LULC_mask.gif"),
               delay = 50, loop = TRUE, restart.delay = 500,
               width = 900, height = 900, dpi = 150) ;toc() 
remove(neighbors, map_extent, mapbiomas_animation_map)

#2. % of climate zone that is anthropic (any proportion) per year in each climate zone
calc_df<- tibble(year= 1985:2023)
list_rasters<- list()
for(i in 1:nlyr(mapbiomas_mask)){
  list_rasters[[i]]<- mapbiomas_mask[[i]]
}
calc_df<- calc_df %>% mutate(anthropic_raster= list_rasters)
remove(list_rasters,i)

central_cropmask_function<- function (raster){
  x_crop<- terra::crop(raster, vect(d_trans$geometry[[1]]))
  x_mask<- terra::mask(x_crop, vect(d_trans$geometry[[1]]))
  x_mask
}
southern_cropmask_function<- function (raster){
  x_crop<- terra::crop(raster, vect(d_trans$geometry[[2]]))
  x_mask<- terra::mask(x_crop, vect(d_trans$geometry[[2]]))
  x_mask
}
eastern_cropmask_function<- function (raster){
  x_crop<- terra::crop(raster, vect(d_trans$geometry[[3]]))
  x_mask<- terra::mask(x_crop, vect(d_trans$geometry[[3]]))
  x_mask
}
Sys.time(); calc_df<- calc_df %>% mutate(central= purrr::map(anthropic_raster, central_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern= purrr::map(anthropic_raster, southern_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern= purrr::map(anthropic_raster, eastern_cropmask_function)); Sys.time()
remove(central_cropmask_function, southern_cropmask_function, eastern_cropmask_function)

reclass_function<- function(regional_raster){
  reclass_raster <- classify(regional_raster, cbind(0, NA)) #pixels that have 0% anthropic is made NA
}
Sys.time(); calc_df<- calc_df %>% mutate(central_reclass= purrr::map(central, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_reclass= purrr::map(southern, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_reclass= purrr::map(eastern, reclass_function)); Sys.time()
remove(reclass_function)

area_function<- function(reclass_raster){
    terra::cellSize(reclass_raster, mask=T, unit="km")
}
Sys.time(); calc_df<- calc_df %>% mutate(central_anthropic_areakm= purrr::map(central_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_anthropic_areakm= purrr::map(southern_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_anthropic_areakm= purrr::map(eastern_reclass, area_function)); Sys.time()
remove(area_function)

total_area_function<- function (area_raster){
  sum(terra::values(area_raster), na.rm = T)
}
Sys.time(); calc_df<- calc_df %>% mutate(central_anthropic_totalareakm= purrr::map(central_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(southern_anthropic_totalareakm= purrr::map(southern_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(eastern_anthropic_totalareakm= purrr::map(eastern_anthropic_areakm, total_area_function)); Sys.time()
remove(total_area_function)

area_anthropic_df<- calc_df %>% dplyr::select(c(year, central_anthropic_totalareakm, southern_anthropic_totalareakm, eastern_anthropic_totalareakm)) 
area_anthropic_df$central_anthropic_totalareakm<- unlist(area_anthropic_df$ central_anthropic_totalareakm)
area_anthropic_df$southern_anthropic_totalareakm<- unlist(area_anthropic_df$southern_anthropic_totalareakm)
area_anthropic_df$eastern_anthropic_totalareakm<- unlist(area_anthropic_df$eastern_anthropic_totalareakm)

area_anthropic_df<- area_anthropic_df %>% mutate(central_percent_anthropic= (central_anthropic_totalareakm/d_trans$area_km[[1]])*100)
area_anthropic_df<- area_anthropic_df %>% mutate(southern_percent_anthropic= (southern_anthropic_totalareakm/d_trans$area_km[[2]])*100)
area_anthropic_df<- area_anthropic_df %>% mutate(eastern_percent_anthropic= (eastern_anthropic_totalareakm/d_trans$area_km[[3]])*100)


pivot_area_anthropic_df<- pivot_longer(area_anthropic_df %>% dplyr::select(year, central_percent_anthropic, 
                                                                           southern_percent_anthropic, eastern_percent_anthropic), 2:4)
p<-ggplot(pivot_area_anthropic_df, aes(x=factor(year), y=value, group=name)) +
  geom_line(aes(color=name))+
  geom_point(aes(color=name))+
  ylab("% of pixels in region that are anthropic (any proportion)") +
  xlab("Year")
p<- p+scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
ggsave(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "area_anthropic.png"),p, 
       dpi=700, height = 12, width = 12, units = "cm")
remove(p, pivot_area_anthropic_df, area_anthropic_df)

#3. Boxplot of non-zero pixels per year per climate zone
calc_df<- calc_df %>% dplyr::select(-c(central_anthropic_areakm, southern_anthropic_areakm, eastern_anthropic_areakm,
                                       central_anthropic_totalareakm, southern_anthropic_totalareakm, eastern_anthropic_totalareakm))

values_function<- function (reclass_raster){
  x<- terra::values(reclass_raster, dataframe= T, na.rm=T)
  names(x)<- "frac_anthropic"
  x
}
Sys.time(); calc_df <- calc_df %>% mutate(central_values= purrr::map(central_reclass, values_function)); Sys.time()
Sys.time(); calc_df <- calc_df %>% mutate(southern_values= purrr::map(southern_reclass, values_function)); Sys.time()
Sys.time(); calc_df <- calc_df %>% mutate(eastern_values= purrr::map(eastern_reclass, values_function)); Sys.time()
remove(values_function)

boxplot_df<- calc_df %>% dplyr::select(c(year, central_values, southern_values, eastern_values))
pivot_boxplot_df<- pivot_longer(boxplot_df, 2:4)
pivot_boxplot_df<- pivot_boxplot_df %>% unnest(value)
q<-ggplot(pivot_boxplot_df, aes(x=factor(year), y=frac_anthropic, group= factor(year))) +
  geom_boxplot(aes(color=name)) + facet_grid(name~.)+
  ylab("Anthropic Fraction (0% removed)") +xlab("Year")
q<- q+scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
ggsave(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "fraction_anthropic_boxplot.png"),q, 
       dpi=700, height = 20, width = 20, units = "cm")
remove(q, pivot_boxplot_df, boxplot_df)

#4. Histogram of %anthropic through time (animation)
histogram_df<- calc_df %>% dplyr::select(c(year, central_values, southern_values, eastern_values))
pivot_histogram_df<- pivot_longer(histogram_df, 2:4)
pivot_histogram_df<- pivot_histogram_df %>% unnest(value)
#pivot_histogram_df<- pivot_histogram_df %>% filter(frac_anthropic<1) #excluding pixels where 100% is anthropic because
#I will exclude these for sure

library(gganimate)
r<- ggplot(pivot_histogram_df, aes(x=frac_anthropic, color=name, fill=name)) + 
  geom_histogram(fill="white") + 
  #scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2")+
  transition_time(year)+ facet_grid(name~.)+
  ylab("Pixel count") +xlab("%Anthropic")+
  labs(title = "Year: {frame_time}")
r<- r+scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
anim_save(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "histogram_%anthropic_no0&1.gif"),r)
remove(r, histogram_df, pivot_histogram_df)

```

According to Mattos et al., 2023, "we use an 80% valid threshold to avoid small 
sample sizes inside each larger pixel."I interpret this as 80% is the cut-off i.e. 
more than 80% of larger pixel is natural. Hence 20% the limit for anthropic in larger pixel. 

From area_anthropic output figure, >90% of southern region is anthropic, central region was already
about 70% anthropic in 1985 and eastern region was about 55% anthropic in 1985. So choosing a 20%
threshold like Mattos et al will mean that very,very little "native" area will remain. 

I can make a decision, like Mattos et al, I will retain pixels across all 3 zones
such that 20% is the maximum anthropic area. I also use 10% & 30% as a threshold.

```{r}
mapbiomas_mask

#1. Keep layers 2002-2021
names(mapbiomas_mask)
tp_mapbiomas_mask<- mapbiomas_mask[[18:37]]

#2. Exclude all pixels where pixel value in any year >0.1
threshold_function<- function (i){
  ifelse(max(i)>0.3, NA, i) #change threshold to 0.1,0.2....
}
Sys.time(); threshold_mapbiomas_mask_10<- terra::app(tp_mapbiomas_mask, fun= threshold_function); Sys.time()#1min
#writeRaster(threshold_mapbiomas_mask_10, here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "thresholded_mapbiomas_mask_10.tif"))
Sys.time(); threshold_mapbiomas_mask_20<- terra::app(tp_mapbiomas_mask, fun= threshold_function); Sys.time()#1min
#writeRaster(threshold_mapbiomas_mask_20, here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "thresholded_mapbiomas_mask_20.tif"))
Sys.time(); threshold_mapbiomas_mask_30<- terra::app(tp_mapbiomas_mask, fun= threshold_function); Sys.time()#1min
#writeRaster(threshold_mapbiomas_mask_30, here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "thresholded_mapbiomas_mask_30.tif"))
remove(threshold_function)

#3. Map of thresholded rasters
#lines 51- 57
thresh10_mapbiomas_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape( threshold_mapbiomas_mask_10)+
  tm_raster(style = 'cont',palette= get_brewer_pal("Reds"), n=5, breaks = seq(0,0.3,0.05), title="10% threshold") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()

thresh20_mapbiomas_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape( threshold_mapbiomas_mask_20)+
  tm_raster(style = 'cont',palette= get_brewer_pal("Reds"), n=5, breaks = seq(0,0.3,0.05), title="20% threshold") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()


thresh30_mapbiomas_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape( threshold_mapbiomas_mask_30)+
  tm_raster(style = 'cont',palette= get_brewer_pal("Reds"), n=5, breaks = seq(0,0.3,0.05), title="30% threshold") + 
              tm_layout (legend.position = c("left", "bottom"))+
              tm_facets(nrow=1, ncol=1, free.coords = FALSE)+
  tm_shape(d_trans)+ tm_borders()

all_maps<- tmap_arrange(thresh10_mapbiomas_map, thresh20_mapbiomas_map, thresh30_mapbiomas_map)
tic();tmap_save(all_maps, 
filename = here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "threshold_Mapbiomass_anthropic_LULC_mask.png"),
               width = 900, height = 900, dpi = 150) ;toc() 
remove(neighbors, map_extent, thresh10_mapbiomas_map, thresh20_mapbiomas_map, thresh30_mapbiomas_map, all_maps)


#4. Area after excluding pixels with >10%, >20% and 30% anthropic (anytime)
calc_df<- calc_df %>% dplyr::select(-c(central_values, southern_values, eastern_values))
list_threshold10_rasters<- list()
for(i in 1:nlyr(threshold_mapbiomas_mask_10)){
  list_threshold10_rasters[[i]]<- threshold_mapbiomas_mask_10[[i]]
}
calc_df<- calc_df %>% mutate(thresh10_anthropic_raster= list_threshold10_rasters)
remove(list_threshold10_rasters,i)

list_threshold20_rasters<- list()
for(i in 1:nlyr(threshold_mapbiomas_mask_20)){
  list_threshold20_rasters[[i]]<- threshold_mapbiomas_mask_20[[i]]
}
calc_df<- calc_df %>% mutate(thresh20_anthropic_raster= list_threshold20_rasters)
remove(list_threshold20_rasters,i)

list_threshold30_rasters<- list()
for(i in 1:nlyr(threshold_mapbiomas_mask_30)){
  list_threshold30_rasters[[i]]<- threshold_mapbiomas_mask_30[[i]]
}
calc_df<- calc_df %>% mutate(thresh30_anthropic_raster= list_threshold30_rasters)
remove(list_threshold30_rasters,i)

#lines 80-94
Sys.time(); calc_df<- calc_df %>% mutate(threshold10_central= purrr::map(thresh10_anthropic_raster, central_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(threshold10_southern= purrr::map(thresh10_anthropic_raster, southern_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(threshold10_eastern= purrr::map(thresh10_anthropic_raster, eastern_cropmask_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(threshold20_central= purrr::map(thresh20_anthropic_raster, central_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(threshold20_southern= purrr::map(thresh20_anthropic_raster, southern_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(threshold20_eastern= purrr::map(thresh20_anthropic_raster, eastern_cropmask_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(threshold30_central= purrr::map(thresh30_anthropic_raster, central_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(threshold30_southern= purrr::map(thresh30_anthropic_raster, southern_cropmask_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(threshold30_eastern= purrr::map(thresh30_anthropic_raster, eastern_cropmask_function)); Sys.time()
remove(central_cropmask_function, southern_cropmask_function, eastern_cropmask_function)

#lines 100-102
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_central_reclass= purrr::map(threshold10_central, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_southern_reclass= purrr::map(threshold10_southern, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_eastern_reclass= purrr::map(threshold10_eastern, reclass_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(thresh20_central_reclass= purrr::map(threshold20_central, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh20_southern_reclass= purrr::map(threshold20_southern, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh20_eastern_reclass= purrr::map(threshold20_eastern, reclass_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(thresh30_central_reclass= purrr::map(threshold30_central, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh30_southern_reclass= purrr::map(threshold30_southern, reclass_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh30_eastern_reclass= purrr::map(threshold30_eastern, reclass_function)); Sys.time()
remove(reclass_function)

#lines 108-110
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_central_anthropic_areakm= purrr::map(thresh10_central_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_southern_anthropic_areakm= purrr::map(thresh10_southern_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_eastern_anthropic_areakm= purrr::map(thresh10_eastern_reclass, area_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(thresh20_central_anthropic_areakm= purrr::map(thresh20_central_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh20_southern_anthropic_areakm= purrr::map(thresh20_southern_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh20_eastern_anthropic_areakm= purrr::map(thresh20_eastern_reclass, area_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(thresh30_central_anthropic_areakm= purrr::map(thresh30_central_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh30_southern_anthropic_areakm= purrr::map(thresh30_southern_reclass, area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh30_eastern_anthropic_areakm= purrr::map(thresh30_eastern_reclass, area_function)); Sys.time()
remove(area_function)

#lines 116-118
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_central_anthropic_totalareakm= purrr::map(thresh10_central_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_southern_anthropic_totalareakm= purrr::map(thresh10_southern_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh10_eastern_anthropic_totalareakm= purrr::map(thresh10_eastern_anthropic_areakm, total_area_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(thresh20_central_anthropic_totalareakm= purrr::map(thresh20_central_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh20_southern_anthropic_totalareakm= purrr::map(thresh20_southern_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh20_eastern_anthropic_totalareakm= purrr::map(thresh20_eastern_anthropic_areakm, total_area_function)); Sys.time()

Sys.time(); calc_df<- calc_df %>% mutate(thresh30_central_anthropic_totalareakm= purrr::map(thresh30_central_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh30_southern_anthropic_totalareakm= purrr::map(thresh30_southern_anthropic_areakm, total_area_function)); Sys.time()
Sys.time(); calc_df<- calc_df %>% mutate(thresh30_eastern_anthropic_totalareakm= purrr::map(thresh30_eastern_anthropic_areakm, total_area_function)); Sys.time()
remove(total_area_function)

thresh_area_anthropic_df<- calc_df %>% dplyr::select(c(year, thresh10_central_anthropic_totalareakm, 
                                                       thresh10_southern_anthropic_totalareakm, 
                                                       thresh10_eastern_anthropic_totalareakm,
                                                       thresh20_central_anthropic_totalareakm, 
                                                       thresh20_southern_anthropic_totalareakm, 
                                                       thresh20_eastern_anthropic_totalareakm,
                                                       thresh30_central_anthropic_totalareakm,
                                                       thresh30_southern_anthropic_totalareakm,
                                                       thresh30_eastern_anthropic_totalareakm,
                                                       )) 
thresh_area_anthropic_df$thresh10_central_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh10_central_anthropic_totalareakm)
thresh_area_anthropic_df$thresh10_southern_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh10_southern_anthropic_totalareakm)
thresh_area_anthropic_df$thresh10_eastern_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh10_eastern_anthropic_totalareakm)
thresh_area_anthropic_df$thresh20_central_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh20_central_anthropic_totalareakm)
thresh_area_anthropic_df$thresh20_southern_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh20_southern_anthropic_totalareakm)
thresh_area_anthropic_df$thresh20_eastern_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh20_eastern_anthropic_totalareakm)
thresh_area_anthropic_df$thresh30_central_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh30_central_anthropic_totalareakm)
thresh_area_anthropic_df$thresh30_southern_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh30_southern_anthropic_totalareakm)
thresh_area_anthropic_df$thresh30_eastern_anthropic_totalareakm<- unlist(thresh_area_anthropic_df$thresh30_eastern_anthropic_totalareakm)

thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh10_central_percent_anthropic= (thresh10_central_anthropic_totalareakm/d_trans$area_km[[1]])*100)
thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh10_southern_percent_anthropic= (thresh10_southern_anthropic_totalareakm/d_trans$area_km[[2]])*100)
thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh10_eastern_percent_anthropic= (thresh10_eastern_anthropic_totalareakm/d_trans$area_km[[3]])*100)

thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh20_central_percent_anthropic= (thresh20_central_anthropic_totalareakm/d_trans$area_km[[1]])*100)
thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh20_southern_percent_anthropic= (thresh20_southern_anthropic_totalareakm/d_trans$area_km[[2]])*100)
thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh20_eastern_percent_anthropic= (thresh20_eastern_anthropic_totalareakm/d_trans$area_km[[3]])*100)

thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh30_central_percent_anthropic= (thresh30_central_anthropic_totalareakm/d_trans$area_km[[1]])*100)
thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh30_southern_percent_anthropic= (thresh30_southern_anthropic_totalareakm/d_trans$area_km[[2]])*100)
thresh_area_anthropic_df<- thresh_area_anthropic_df %>% 
  mutate(thresh30_eastern_percent_anthropic= (thresh30_eastern_anthropic_totalareakm/d_trans$area_km[[3]])*100)


pivot_thresharea_anthropic_df<- pivot_longer(thresh_area_anthropic_df %>% dplyr::select(year, thresh10_central_percent_anthropic, 
                                                                           thresh10_southern_percent_anthropic, thresh10_eastern_percent_anthropic,
                                                                           thresh20_central_percent_anthropic, thresh20_southern_percent_anthropic,
                                                                           thresh20_eastern_percent_anthropic,
                                                                           thresh30_eastern_percent_anthropic,
                                                                           thresh30_southern_percent_anthropic,
                                                                           thresh30_central_percent_anthropic
                                                                           ), 2:10)

pivot_thresharea_anthropic_df<- pivot_thresharea_anthropic_df %>% 
  separate_wider_delim(cols = name, delim = "_", names = c("Thresh","Zone", "Percent","Anthropic"))
pivot_thresharea_anthropic_df<- pivot_thresharea_anthropic_df %>% dplyr::select(-c(Percent, Anthropic))
pivot_thresharea_anthropic_df<- pivot_wider(pivot_thresharea_anthropic_df, names_from = Thresh, values_from = value)
#The values are repeated every year because that is the exclusion I have done
#If a pixel in ANY year cross the 10%,20% or 30% anthropic threshold, that pixel through
#time is removed. So the remaining area is the final area and is constant
#through time. 


#lines 108-135
pivot_area_anthropic_df<- pivot_area_anthropic_df %>% 
  separate_wider_delim(cols = name, delim = "_", names = c("Zone", "Percent","Anthropic"))
pivot_area_anthropic_df<- pivot_area_anthropic_df %>% dplyr::select(-c(Percent, Anthropic))
names(pivot_area_anthropic_df)[3]<- "value1"

pivot_df<- full_join(pivot_area_anthropic_df, pivot_thresharea_anthropic_df, by=join_by(year==year, Zone==Zone))

s<-ggplot(pivot_df, aes(x=factor(year), group=Zone)) +
  geom_line(aes(y=value1, color=Zone))+
  geom_point(aes(y=value1,color=Zone))+
  geom_line(aes(y=thresh10, color=Zone), linetype= "dashed")+
  geom_line(aes(y=thresh20, color=Zone), linetype="3313")+
  geom_line(aes(y=thresh30, color=Zone))+
  ylab("% of pixels in region that are anthropic (any proportion)") +
  xlab("Year")
s<- s+scale_color_brewer(labels= c("Central", "Eastern", "Southern"), palette = "Dark2") +
  theme_classic()+
  theme(legend.title=element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 
ggsave(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "threshold_before_after_area.png"),s, 
       dpi=700, height = 25, width = 25, units = "cm")

mean(pivot_df$value1[pivot_df$Zone=="central"])# 12.09981
mean(pivot_df$value1[pivot_df$Zone=="southern"]) #7.533759
mean(pivot_df$value1[pivot_df$Zone=="eastern"]) #10.33583

remove(pivot_area_anthropic_df, pivot_df, pivot_thresharea_anthropic_df)
remove(s)
remove(area_anthropic_df, thresh_area_anthropic_df)
```


