```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script I complete the trajectory shape analyses to the STL decomposed timeseries data for the three regions. 

##Trajectory shape analyses function
This function was sourced from Pelisse et al., 2024 and accordingly modified. 
```{r}
library(MuMIn)

class_trajectory_mod <- function (dataset = NULL, interval_size = 0.5) {
  dataset<- dataset %>% mutate(Months= 1:nrow(dataset))
  #Trajectory fitting
  null_mod<- lm(trend ~ 1, data = dataset) ##1. No intercept
  summary(null_mod) #significant pvalue next to intercept  means that the mean is significantly different from 0
  nmrs_null<- sqrt(sum(summary(null_mod)$residuals^2)/length(dataset$trend))/sd(dataset$trend)
  aic_null<- MuMIn::AICc(null_mod)

  linear_mod<- lm(trend ~ Months, data = dataset) ##2.linear
  summary(linear_mod)
  nmrs_lin<- sqrt(sum(summary(linear_mod)$residuals^2)/length(dataset$trend))/sd(dataset$trend)
  aic_lin<- MuMIn::AICc(linear_mod)

  orth_poly_mod<- lm(trend ~ poly(Months,2, raw=F), data = dataset) #raw=F means orthogonal polynomials are used
  summary(orth_poly_mod)
  nmrs_quad<- sqrt(sum(summary(orth_poly_mod)$residuals^2)/length(dataset$trend))/sd(dataset$trend)
  aic_quad<- MuMIn::AICc(orth_poly_mod)

  ## To get relevant values for quadratic output - Directly from Pellise et al., 2024 
  ## https://github.com/matpelissie/abrupt_shifts_ecological_timeseries_classification/blob/main/R/functions_trajclass.R
  
  # After getting Y = gamma*chi + delta*X' + epsilon with orthogonal polynomial,
  # we have to perform a variable change to obtain relevant values in the X interval 
  # for first_order_coefficient, second_order_coefficient and intercept, 
  # knowing that X'= alpha*X + beta and chi = eta*X'^2 + theta

  gammab  <-  orth_poly_mod$coefficients[3]
  delta  <-  orth_poly_mod$coefficients[2]
  epsilon  <-  orth_poly_mod$coefficients[1]

  alpha  <-  lm(orth_poly_mod$model[, 2][, 1] ~ dataset$Months)$coef[2]
  beta  <-  lm(orth_poly_mod$model[, 2][, 1] ~ dataset$Months)$coef[1]

  eta  <-  1/lm((orth_poly_mod$model[, 2][, 1])^2 ~
                orth_poly_mod$model[, 2][, 2])$coef[2]
  theta  <-  (-lm((orth_poly_mod$model[, 2][, 1])^2 ~
                  orth_poly_mod$model[, 2][, 2])$coef[1])*eta

  Y2 <- dataset$trend*(max(dataset$Months)-min(dataset$Months))/(max(dataset$trend)-min(dataset$trend)) 

  # p2 and p3 are relevant when Y and X amplitudes are equivalent,in particular when 
  # studying scaled-to-1 indices, Y and X amplitudes may be very different, so we 
  # scaled the amplitudes to calculate p2 and p3

  polynomial_orthonormal_basis <- lm(Y2~poly(dataset$Months,2, raw=T))$coefficients

  # Quadratic model output:
  classification <-
    data.frame(first_order_coefficient = (delta+2*beta*gammab*eta)*alpha,
               first_order_pvalue =
                 summary(orth_poly_mod)$coefficients[2, 4],
               second_order_coefficient = (alpha^2)*gammab*eta,
               second_order_pvalue =
                 summary(orth_poly_mod)$coefficients[3, 4],
               strd_error=summary(orth_poly_mod)$coefficients[2, 2],
               intercept = epsilon+beta*delta+(beta^2)*gammab*eta+gammab*theta,
               x_m = (dataset$Months[length(dataset$Months)]-dataset$Months[1])/2+dataset$Months[1],
               # points of interest:
               p1 = -(delta+2*beta*gammab*eta)/(2*alpha*gammab*eta),
               p2 = (-polynomial_orthonormal_basis[2]+1)/
                 (2*polynomial_orthonormal_basis[3]),
               p3 = (-polynomial_orthonormal_basis[2]-1)/
                 (2*polynomial_orthonormal_basis[3]),
               aic = aic_quad,
               nrmse = nmrs_quad,
               loc_brk = NA, 
               trend = NA,
               mag = NA,
               rel_chg = NA,
               SDbef = NA,
               SDaft = NA)
  
  # Linear model output:
  classification[2,] <-
    data.frame(first_order_coefficient = delta*alpha,
               first_order_pvalue =
                 summary(orth_poly_mod)$coefficients[2, 4],
               second_order_coefficient = 0,
               second_order_pvalue =
                 summary(orth_poly_mod)$coefficients[3, 4],
               strd_error=summary(orth_poly_mod)$coefficients[2, 2],
               intercept = epsilon+delta*beta,
               x_m = (dataset$Months[length(dataset$Months)]-dataset$Months[1])/2+dataset$Months[1],
               p1 = NA,
               p2 = NA,
               p3 = NA,
               aic = aic_lin,
               nrmse = nmrs_lin,
               loc_brk = NA, 
               trend = NA,
               mag = NA,
               rel_chg = NA,
               SDbef = NA,
               SDaft = NA)

  # No change model output:
  classification[3,] <-
    data.frame(first_order_coefficient = 0,
               first_order_pvalue =
                 summary(orth_poly_mod)$coefficients[2, 4],
               second_order_coefficient = 0,
               second_order_pvalue =
                 summary(orth_poly_mod)$coefficients[3, 4],
               strd_error=summary(orth_poly_mod)$coefficients[2, 2],
               intercept = null_mod$coefficients,
               x_m = (dataset$Months[length(dataset$Months)]-dataset$Months[1])/2+dataset$Months[1],
               p1 = NA,
               p2 = NA,
               p3 = NA,
               aic = aic_null,
               nrmse = nmrs_null,
               loc_brk = NA, 
               trend = NA,
               mag = NA,
               rel_chg = NA,
               SDbef = NA,
               SDaft = NA)


  # Classification of each model fitted:
  for (i in 1:3){

    # Compute the derivative at xm-delta and at xm + delta with delta being
    # half of the input interval size (its 25% of the time interval which is set as 0.5)
      derivative <-
        2*(classification$x_m[i] - (dataset$Months[length(dataset$Months)]-dataset$Months[1])*(interval_size/2))*  
        classification$second_order_coefficient[i] +
        classification$first_order_coefficient[i]
      derivative2 <-
        2*(classification$x_m[i] + (dataset$Months[length(dataset$Months)]-dataset$Months[1])*(interval_size/2))*
        classification$second_order_coefficient[i] +
        classification$first_order_coefficient[i]


      if(sign(derivative) != sign(derivative2) | i==3){
      # non consistent direction around x_m
        classification$derivative[i]  <-  NA
        classification$intercept_derivative[i]  <-  NA

      } else {
      # consistent direction around x_m
        classification$derivative[i] <- mean(c(derivative, derivative2))
        classification$intercept_derivative[i] <-
          (classification$second_order_coefficient[i]*classification$x_m[i]^2+
             classification$first_order_coefficient[i]*classification$x_m[i]+
              classification$intercept[i]) -
          classification$x_m[i]*classification$derivative[i]
      }

    # Compute the derivative of the curvature function to get acceleration:
      classification$derivated_curvature[i] <-
        -12*(classification$second_order_coefficient[i]^2)*
        (2*classification$second_order_coefficient[i]*classification$x_m[i]+
           classification$first_order_coefficient[i])*
        (classification$second_order_coefficient[i]/
           abs(classification$second_order_coefficient[i]))/
        ((1+(2*classification$second_order_coefficient[i]*classification$x_m[i]+
            classification$first_order_coefficient[i])^2)^(2.5))

    # Keep derivated curvature even if not significant for polynomial fit:
      if(classification$second_order_pvalue[i]>0.05 & i != 1){
        classification$derivated_curvature[i] <- NA
      }

    # Classify the direction:
      classification$direction[i] <- NA
        classification$direction[i][which(
          classification$derivative[i] > 0)] <- "increase"
      classification$direction[i][which(
        classification$derivative[i] < 0)] <- "decrease"
      classification$direction[i][which(
        is.na(classification$derivative[i]))] <- "stable"
      classification$direction[i][which(
        as.numeric(classification$first_order_pvalue[i])>0.05 &
          as.numeric(classification$second_order_pvalue[i])>0.05)] <- "stable"

    # Classify the acceleration:
      classification$acceleration[i] <- NA
        classification$acceleration[i][which(
          classification$derivated_curvature[i] < 0)] <- "accelerated"
      classification$acceleration[i][which(
        classification$derivated_curvature[i] > 0)] <- "decelerated"
      classification$acceleration[i][which(
        classification$direction[i] == "stable" &
          classification$second_order_coefficient[i] < 0)] <- "concave"
      classification$acceleration[i][which(
        classification$direction[i] == "stable" &
          classification$second_order_coefficient[i] > 0)] <- "convex"
      classification$acceleration[i][which(
        is.na(classification$derivated_curvature[i]))] <- "constant"

    # Give the final classification combining direction and acceleration:
      classification$shape_class[i] <- paste(classification$direction[i],
                                             classification$acceleration[i],
                                             sep="_")
  }
  
  # Abrupt breakpoint analyses
  chng_fit<- chngpt::chngptm(formula.1=trend~1,
                      formula.2=~Months,
                      family="gaussian", data=dataset,
                      type="step",
                      var.type="bootstrap", weights=NULL)

  pred_chg <- data.frame(timestep = dataset$Months,
                         bp = chng_fit$best.fit$fitted.values)

  chng_nrmse <- sqrt(sum(residuals(chng_fit)^2)/length(dataset$trend))/sd(dataset$trend)

  classification[4, 13] <- chng_fit$chngpt
  classification[4, 11] <- MuMIn::AICc(chng_fit)
  classification[4, 12] <- chng_nrmse
  classification[4, 14] <- ifelse(pred_chg$bp[1] >
                                       pred_chg$bp[length(pred_chg$bp)],
                                     "decrease", "increase")
  classification[4, 15] <-  pred_chg$bp[length(pred_chg$bp)] - pred_chg$bp[1]
  classification[4, 16] <-  (pred_chg$bp[length(pred_chg$bp)] - pred_chg$bp[1]) /
                        max(abs(pred_chg$bp[length(pred_chg$bp)]),abs(pred_chg$bp[1]))
  classification[4, 17] <-  dataset %>%
                                  dplyr::filter(Months<=chng_fit$chngpt) %>%
                                  dplyr::pull(trend) %>%
                                  sd()
  classification[4, 18] <- dataset %>%
                        dplyr::filter(Months>=chng_fit$chngpt) %>%
                        dplyr::pull(trend) %>%
                        sd()
  
  row.names(classification) <- c("Y_pol","Y_lin", "Y_nch", "Y_abpt")
  
  # classification$cell <- unique(dataset$cell)
  classification$x <- unique(dataset$x)
  classification$y <- unique(dataset$y)

  return(classification)
}

```


Applying the above function
```{r}
#Data input
x<-read_rds("/dungbeetle/home/tg505/Trisha/ecothresholds/Outputs/Indices/anisoEVI/STL_Decomposition/swindow_periodic/stl_central11001_end.rds") #changed this manually

chunk2 <- 240*100
n2 <- nrow(x)
r2  <- rep(1:ceiling(n2/chunk2),each=chunk2)[1:(n2)]
split_pivot_df <- split(x, r2)
remove(chunk2, n2,r2)
gc()
message("Big rds file split")

numCores<- 16
library(foreach)
library(doParallel)

my.cluster<- parallel::makeCluster(
  numCores,
  type = "FORK",
  outfile = ""
)

doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered() #should be TRUE
foreach::getDoParWorkers() #should give numCores


rds_list<- rds_list %>% discard(is.null)

##Converting above to nested parallelized foreach loop
Sys.time(); x<- foreach(
  i= 1:length(split_pivot_df),
  .combine= "rbind") %dopar%
    (split_pivot_df[[i]] %>%
      group_by(cell) %>%
      nest() %>%
      mutate(classification_data = purrr::map(data, class_trajectory_mod)) %>%
      select(-data) %>%
      unnest(classification_data)
       ); Sys.time()

write_rds(x, "/dungbeetle/home/tg505/Trisha/ecothresholds/Outputs/Indices/anisoEVI/TrajectoryShapes/swindow_periodic/trajshape_central11001_end.rds")#manually changed

```


##Model selection function
This function chooses the best model per pixel based on Berdugo et al., 2022
```{r}
model_levels <- c("Null", "Lin", "Step", "Quad")

model_select<- function(models_pixel_df){
  message("Adding model name")
  models_pixel_df <- models_pixel_df %>% mutate(model_order=ordered(c("Quad", "Lin", "Null", "Step"),
                                                                    levels = model_levels))

  message("Least AIC model selection")
  models_pixel_df <- models_pixel_df %>%
    mutate(aic_diff= abs(aic) - min(abs(aic))) #wrt model with least AIC
  
  message("Conditional model selection")
  condition_less2<- models_pixel_df %>% filter(aic_diff<=2)
  
  if (dim(condition_less2)[1]==1){
    models_pixel_df<- condition_less2 
  } else{
    models_pixel_df<- condition_less2 %>% 
      filter(model_order==min(model_order))
  }
  models_pixel_df 
}

```

Applying the above function
```{r}
rds_list <- list.files(path = "/dungbeetle/home/tg505/Trisha/ecothresholds/Outputs/Indices/anisoEVI/TrajectoryShapes/swindow_7/", 
                       pattern='trajshape_eastern$', all.files=TRUE, full.names=TRUE) #manually changed to read all files of each region

data_rds_list<- list()
for (i in 1:length(rds_list)){
  data_rds_list[[i]]<- read_rds(rds_list[i])
}

data_rds_list<- data_rds_list %>% discard(is.null)

numCores<- 16
library(foreach)
library(doParallel)

my.cluster<- parallel::makeCluster(
  numCores,
  type = "FORK",
  outfile = ""
)

doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered() #should be TRUE
foreach::getDoParWorkers() #should give numCores


Sys.time(); y<- foreach (
  i=1:length(data_rds_list),
  .combine = "rbind") %dopar% 
  (data_rds_list[[i]] %>%
     rename("trend_name"="trend") %>%
    dplyr::group_by(cell) %>%
    nest() %>%
    mutate(modelselect_data = purrr::map(data, model_select)) %>%
    select(-data) %>%
    unnest(modelselect_data)
  ); Sys.time()

write_rds(y, "/dungbeetle/home/tg505/Trisha/ecothresholds/Outputs/Indices/anisoEVI/TrajectoryShapes/swindow_7/finalshape_eastern.rds") #manually changed to make one file per region
```