```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script I prepare "ground truth" ecological resilience data using meta analyses papers, primarily using Stevens et al., 2017.
In the SI of the paper, there is information about the papers that were used including extraction of key information 
(such as lat/long of plots researched in the paper) from the papers. I manually made a csv by copy pasting from the SI word doc to excel.

7 papers basically have the same lat/long. These plots are in Assis, SP and are mostly Durigan plots. 

I consider the rasters of the final trajectories after excluding anthropic area for all three indices at 1km scale. 

```{r}
stevens_si<- read_csv(here("Data","Stevensetal_woodyencroachment_metaanalyses_data", "si_plotinformation.csv"))
names(stevens_si)[14]<-"y"
names(stevens_si)[15]<-"x"
stevens_si_removeNA<- stevens_si %>% filter(!is.na(x))
stevens_si_removeNA<- stevens_si_removeNA %>% filter(Country!="Venezuela")

#Duplicate plots resulted in multiple papers eg Durigan et al 1987 and 2006 are from the same plots
#So I "merge" rows and take the mean of the "annual change" that was calculated in Stevens et al 
x<-stevens_si_removeNA %>% 
  group_by(x,y) %>%
  summarise(mean_annual_rate= mean(Annual_change)) %>%
           ungroup()

x_join<- stevens_si_removeNA %>% left_join(x)
x_join2<- x_join %>% filter(!duplicated(mean_annual_rate))

stevens_si_vector<- sf::st_as_sf(x_join2, coords = c("x", "y"), crs="epsg:4326")
stevens_si_vector<- stevens_si_vector %>% mutate(uniqueID= 1:nrow(stevens_si_vector))

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)

#tmap_mode("view")
#tm_shape(d_trans)+ tm_fill()+ tm_shape(stevens_si_vector) + tm_dots(size="Annual_change")
#st_write(stevens_si_vector, here("Data","Stevensetal_woodyencroachment_metaanalyses_data", "plotinformation.shp"), overwrite = T)

anisoEVI_map_results<- rast(here("Outputs", "Final_TrajShape_Models", "Annual", "am_trends11_trajresults.tif"))
# 1- Linear decrease, 2- linear increase, 3- no trend, 4- step decrease, 5- step increase
masked_ndvi_results<- rast(here("Outputs", "Final_TrajShape_Models", "Annual_NDVI", "masked_ndvi_amtrends11_trajresults.tif"))
# 1-Linear decrease, 2- linear increase, 3- no trend, 4- step decrease, 5- step_increase, 6- quad decrease accelerated, 7- quad-decrease decelerated
masked_evi_results<- rast(here("Outputs", "Final_TrajShape_Models", "Annual_EVI", "masked_evi_amtrends11_trajresults.tif"))
# same legend as above NDVI

trajectories<- c(anisoEVI_map_results, masked_evi_results, masked_ndvi_results)
names(trajectories)<- c("anisoEVI", "EVI", "NDVI")

## To make a map of where field points fall in our study
threshold<- rast(here("Data", "Masks_to_define_pixelsofinterest", "mapbiomas_anthropic_mask", "thresholded_mapbiomas_mask_20.tif"))
final_trajectories<- terra::mask(trajectories, threshold)

mappalette<- brewer.pal(n=7, "Set1")
mappalette[[3]]<- "darkgrey"
mappalette[[4]]<-  "#984EA3"
mappalette[[5]]<-  "#FF7F00"
mappalette[[6]]<- "#FFFF33"
mappalette[[7]]<- "#A65628"

points_anisotraj_map<- tm_shape(d_trans)+ tm_fill()+ 
  tm_shape(final_trajectories[[1]]) +
  tm_raster(style = "cat", palette = mappalette)+
  tm_shape(stevens_si_vector) + 
  tm_dots(size="Annual_change")
points_evitraj_map<- tm_shape(d_trans)+ tm_fill()+ 
  tm_shape(final_trajectories[[2]]) +
  tm_raster(style = "cat", palette = mappalette)+
  tm_shape(stevens_si_vector) + 
  tm_dots(size="Annual_change")
points_ndvitraj_map<- tm_shape(d_trans)+ tm_fill()+ 
  tm_shape(final_trajectories[[3]]) +
  tm_raster(style = "cat", palette = mappalette)+
  tm_shape(stevens_si_vector) + 
  tm_dots(size="Annual_change")
points_trajectory_map<- tmap_arrange(points_anisotraj_map, points_evitraj_map, points_ndvitraj_map, ncol = 3, nrow = 1)
tmap_save(points_trajectory_map, here("Outputs", "AnalysesResults", "Relationship_to_fieldplots", "points_trajectories_map.jpeg"),
         height = 40, width = 80, units = "cm", dpi=700)

remove(stevens_si, x, x_join, x_join2, mappalette)
remove(points_anisotraj_map, points_evitraj_map, points_ndvitraj_map, points_trajectory_map)
remove(final_trajectories)

```


I then find the closest trajectory shape (distance and type of trajectory) to each of the plots in the Stevens et al paper. 
```{r}
trajectory_segregate <- function(x) {
    x_split<- terra::segregate(x, other=999)
    x_split_crop<- terra::crop(x_split, vect(d_trans))
    x_split_mask<- terra::mask(x_split_crop, vect(d_trans))
    x_split_mask
}

anisoEVItrajectories_split<- trajectory_segregate(trajectories[[1]])
EVItrajectories_split<- trajectory_segregate(trajectories[[2]])
NDVItrajectories_split<- trajectory_segregate(trajectories[[3]])

distance_function<- function(split_trajectory_raster){
  #meandistanthropic<- terra::app(split_trajectory_raster, fun="mean")
  x_dist<- terra::gridDist(split_trajectory_raster, target= 1)
  
}


calculate_distances<- function (split_raster){
  list_distance_raster<- list()
  for (i in 1:nlyr(split_raster)){
    x_distance<- distance_function(split_raster[[i]])
    list_distance_raster[[i]]<- x_distance
}
  distance_raster<- rast(list_distance_raster)
  names(distance_raster)<- 1:nlyr(split_raster)
  distance_raster
}

Sys.time(); anisoEVI_distances<- calculate_distances(anisoEVItrajectories_split); Sys.time() 
Sys.time(); EVI_distances<- calculate_distances(EVItrajectories_split); Sys.time() 
Sys.time(); NDVI_distances<- calculate_distances(NDVItrajectories_split); Sys.time() 

shortestdistance_trajectory<- function (distance_raster){
  distance_df<- terra::extract(distance_raster, stevens_si_vector)
  pivot_dist_df<- distance_df %>% pivot_longer(-1)
  pivot_dist_df2<-pivot_dist_df  %>% group_by(as.factor(ID)) %>% summarise(min(value), name[which.min(value)])
  names(pivot_dist_df2)<- c("uniqueID", "min_dist_m", "respective_trajshape")
  pivot_dist_df2
}

Sys.time(); anisoEVI_closest_trajectory_df<- shortestdistance_trajectory(anisoEVI_distances); Sys.time()
Sys.time(); EVI_closest_trajectory_df<- shortestdistance_trajectory(EVI_distances); Sys.time()
Sys.time(); NDVI_closest_trajectory_df<- shortestdistance_trajectory(NDVI_distances); Sys.time()
```

Seems like for all indices, the majority of the field points (n=17) are either in no trend (3)
or 5 (step increase) with one 4 (step decrease) in NDVI trajectory and a two 2s (linear increase)
in anisoEVI trajectories.

I then calculate the second closest trajectory to each of the 17 points across all 3 indices. 
In the EVI & NDVI based results, the second closest trajectory to field points are majority
step increase, which is good because the field plots show woody encroachment.
```{r}
second_shortestdistance_trajectory<- function (distance_raster){
  distance_df<- terra::extract(distance_raster, stevens_si_vector)
  pivot_dist_df<- distance_df %>% pivot_longer(-1)
  pivot_dist_df2<-pivot_dist_df %>% group_by(ID) %>% arrange(value) %>% slice (2) %>% select(value, name)
  names(pivot_dist_df2)<- c("uniqueID", "min_dist_m", "respective_trajshape")
  pivot_dist_df2
}

Sys.time(); sec_anisoEVI_closest_trajectory_df<- second_shortestdistance_trajectory(anisoEVI_distances); Sys.time()
Sys.time(); sec_EVI_closest_trajectory_df<- second_shortestdistance_trajectory(EVI_distances); Sys.time()
Sys.time(); sec_NDVI_closest_trajectory_df<- second_shortestdistance_trajectory(NDVI_distances); Sys.time()

```

Results presentation as stacked barchart
For closest trajectories, all field plots fall exactly on a pixel ie there
is no distance
```{r}
df_prep<- function (closest_distance_df, index_name){
  x<- closest_distance_df %>% 
    dplyr::select(-c(min_dist_m))
  names(x)[2]<- index_name
x
  }
  
closest_aniso<- df_prep(anisoEVI_closest_trajectory_df, "anisoEVI")
closest_evi<- df_prep(EVI_closest_trajectory_df, "EVI")
closest_ndvi<- df_prep(NDVI_closest_trajectory_df, "NDVI")

closest_df<- left_join(closest_aniso, closest_evi)
closest_df<- left_join (closest_df, closest_ndvi)
closest_df<- closest_df %>% mutate(proximity= "closest")    
closest_df$uniqueID<-as.numeric(closest_df$uniqueID)
pivot_closest_df<- pivot_longer (closest_df, 2:4)
        
second_closest_aniso<- df_prep(sec_anisoEVI_closest_trajectory_df, "anisoEVI")
second_closest_evi<- df_prep(sec_EVI_closest_trajectory_df, "EVI")
second_closest_ndvi<- df_prep(sec_NDVI_closest_trajectory_df, "NDVI")

sec_closest_df<- left_join(second_closest_aniso, second_closest_evi)
sec_closest_df<- left_join (sec_closest_df, second_closest_ndvi)
sec_closest_df<- sec_closest_df %>% mutate(proximity= "second_closest")
pivot_sec_closest_df<- pivot_longer (sec_closest_df, 2:4)

plot_df<- rbind(pivot_closest_df, pivot_sec_closest_df)
stackedbarplot_df2<- plot_df %>% group_by(proximity, name, value) %>% 
  summarise(sum=n())
stackedbarplot_df2<- stackedbarplot_df2 %>% mutate(value = case_when(value == 1~ "Linear decrease",
                                                                     value == 2~ "Linear increase",
                                                                     value == 3~ "No trend",
                                                                     value == 4~ "Step decrease",
                                                                     value == 5~ "Step increase") )


mappalette<- brewer.pal(n=7, "Set1")
mappalette[[3]]<- "darkgrey"
mappalette[[4]]<-  "#984EA3"
mappalette[[5]]<-  "#FF7F00"
mappalette[[6]]<- "#FFFF33"
mappalette[[7]]<- "#A65628"


distance_stackedbarplot<- ggplot(stackedbarplot_df2, aes( x= proximity, y = sum, fill = value)) + 
    geom_bar(stat = 'identity', position = 'stack') + theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_fill_manual (values = mappalette)+ 
  facet_grid(~ name) 

sec_anisoEVI_closest_trajectory_df <- sec_anisoEVI_closest_trajectory_df %>% mutate(index= "anisoEVI")
sec_EVI_closest_trajectory_df <- sec_EVI_closest_trajectory_df %>% mutate(index= "EVI")
sec_NDVI_closest_trajectory_df <- sec_NDVI_closest_trajectory_df %>% mutate(index= "NDVI")

sec_dist_df<- rbind(sec_anisoEVI_closest_trajectory_df, sec_EVI_closest_trajectory_df, sec_NDVI_closest_trajectory_df)
sec_dist_df<- sec_dist_df %>% mutate(respective_trajshape= case_when(respective_trajshape == 1~ "Linear decrease",
                                                      respective_trajshape == 2~ "Linear increase",
                                                      respective_trajshape == 3~ "No trend",
                                                      respective_trajshape == 4~ "Step decrease",
                                                      respective_trajshape == 5~ "Step increase"))


distance_barplot<- ggplot(sec_dist_df, aes(x= respective_trajshape, y= min_dist_m, fill= respective_trajshape)) +
  geom_boxplot()+ theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_fill_manual (values = mappalette)+
  facet_grid(~index) 

distance<- ggarrange(distance_stackedbarplot, distance_barplot)
ggsave(here("Outputs", "AnalysesResults", "Relationship_to_fieldplots", "fieldplots_trajectoryshape_results.png"),
       distance,dpi = 700, height = 30, width=30, units = "cm")



```













