```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script, I processes different masks to finalize pixels of interest at various spatial scales. 

#MODIS LULC mask
The first mask I work on is to finalize pixels within which I analyse trends in anisoEVI. Since anisoEVI is a MODIS derived product, I use the MODIS 500m annual LULC product on GEE. And with this product, I chose the plant functional types classification (see GEE) and exclude all water, unclassified and urban areas in GEE. I exported the data as a multiband image (each band is each year of LULC) in the original MODIS sinusoidal projection and native resolution (~500m). So in this script I mask to the cerrado border and change projection to match anisoEVI crs and scale. 

```{r}
evi<- rast(here("Data", "Indices", "AnisoEVI_Cerrado_GEE", "mosaiced_anisoEVI.tif"))
tic();evi<- evi/10^5;toc() #scaling factor mentioned on the description page of the GEE asset

##changing band names to year and months
colnames<- list()
tic(); for ( i in 1:nlyr(evi)){
  x<- names(evi)[[i]]
  y<- paste0(strsplit(x, "_")[[1]][5], "_",strsplit(x, "_")[[1]][6] )
  colnames[[i]] <-y
}; toc()
colnames<- unlist(colnames)
names(evi)<- colnames

########################################## BELOW PROJECTION ANALYSES IS COMPLETE, DO NOT REPEAT
#modis_mask<- rast(here("Data", "Masks_to_define_pixelsofinterest", "MODIS_LULC_mask.tif"))
#modis_mask # all NA pixels i.e LULC excluded got exported as 0 from GEE. Also since I used a 'region' in the export to drive function, 
# there are 0 pixel values outside the cerrado boundary which need to be masked. 
# Also I am interested in using this only as a mask i.e. I do not need to know which LULC I have kept i.e do not need to keep pixel values 
# of LULC retained. SO I reclassify all 0 values to NA and all non 0 values to 1. 
#tic(); modis_mask<- terra::classify(modis_mask, cbind(0,NA), others = 1); toc()

#tic(); modis_mask_proj<- terra::project(modis_mask, evi, method = "near", mask= T, align=T); toc() #nearest neighbor becuse LULC is catergorical pixel values
#tic(); modis_mask_proj2<- terra::crop(modis_mask_proj, evi)
#writeRaster(modis_mask_proj2, here("Data", "Masks_to_define_pixelsofinterest", "proj_MODIS_LULC_mask.tif"))
########################################## ABOVE PROJECTION ANALYSES IS COMPLETE, DO NOT REPEAT
modis_mask<- rast(here("Data", "Masks_to_define_pixelsofinterest", "proj_MODIS_LULC_mask.tif"))

modis_mask_names<- seq(from = 2001, to=2022, by=1)
tic(); for ( i in 1:length(modis_mask_names)){
  x<- paste0("mask_", modis_mask_names[[i]])
  modis_mask_names[[i]] <-x
}; toc()
names(modis_mask)<- modis_mask_names

############################## Masking 2001 anisoEVI with 2001 MODIS LULC.......... till 2022
#anisoEVI is from March 2000 to Dec 2021. So to match MODIS annual product which starts at 2001, I will get rid of the 2000 anisoEVI products
evi<- evi[[11:261]]
colnames<- colnames[11:261]

#Also, I have only annual masks, so all monthly anisoEVI for 2001, 2002 (12 rasters each) need to be masked by one mask
mask_df<- as.tibble(colnames)
mask_df<- mask_df %>% separate_wider_delim(cols = value, delim = "_", names = c("Year","Month"))
mask_df<- mask_df %>% mutate(Year_Month=colnames)

evi_stack<- function (Year_Month){
  terra::subset(evi, grep(Year_Month, names (evi)))
}
tic(); mask_df<-mask_df %>% mutate(monthlyanisoEVI = map(Year_Month, evi_stack)); toc()

mask_stack<- function (Year){
  terra::subset(modis_mask, grep(Year, names (modis_mask)))
}
tic(); mask_df<-mask_df %>% mutate(mask = map(Year, mask_stack)); toc() #12 sec

masking_function<- function (anisoEVI_raster, mask_raster){
  terra::cover(mask_raster, anisoEVI_raster, values=1)
}
tic(); mask_df<-mask_df %>% mutate(masked_EVI = purrr::map2(monthlyanisoEVI, mask, masking_function)); toc() #took a long time 65183.45 seconds which is 18 hours!!

writeRaster_function<- function (masked_evi_raster, name){
  terra::writeRaster(masked_evi_raster, here("Outputs", "Indices", "Masked_anisoEVI", paste0(name, "_maskedanisoEVI.tif")))
}
tic(); mask_df<-mask_df %>% mutate(write_put = purrr::map2(masked_EVI, Year_Month, writeRaster_function)); toc() #1 min



```
------- ABOVE CHUNK IS COMPLETE AND WRITTEN OUT, START WITH NEXT CHUNK 

Then I find the amount of area remaining i.e total area of pixels within which trends will be anlaysed. I present this as a proportion of the total cerrado area
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)

cerrado<- d_trans %>% st_union()
cerrado_area<- terra::expanse(vect(cerrado), unit= "km") #2040045 sqkm which is 204 Mha

masked_list <- list.files(path = here("Outputs", "Indices", "Masked_anisoEVI"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
masked_list<-gtools::mixedsort(masked_list) #arranging rasters chronologically
masked_raster_list<-lapply(masked_list, rast)
masked_evi<- rast(masked_raster_list) #251 rasters starting 2001_01
names(masked_evi)<- colnames #from line 66 (which means I need to run the previous lines 31-32 & 38-43)
tic(); coverage<- masked_evi; toc()
tic(); coverage[!is.na(coverage)]<-1;toc()# 2.3 min
tic(); area<- terra::cellSize(coverage, mask=T, lyrs=T, unit= "km"); toc() ##2 min getting area of all pixel values=1
names(area)<- colnames

tic(); area2<- terra::crop(area, vect(d_trans), ); toc()
tic(); area2<- terra::mask(area2, vect(d_trans), ); toc()
names(area2)<- colnames


area_mask_df<- as.tibble(colnames)
area_mask_df<- area_mask_df %>% separate_wider_delim(cols = value, delim = "_", names = c("Year","Month"))
area_mask_df<- area_mask_df %>% mutate(Year_Month=colnames)

area_stack<- function (Year_Month){
  terra::subset(area2, grep(Year_Month, names (area2)))
}
tic(); area_mask_df<-area_mask_df %>% mutate(monthlyarea = map(Year_Month, area_stack)); toc()

sum_area<- function (area_raster){
  sum(terra::values(area_raster), na.rm=T)
}
tic(); area_mask_df<-area_mask_df %>% mutate(masked_EVI_area_sqkm = map(monthlyarea, sum_area)); toc() #20 min
area_mask_df<- area_mask_df %>% mutate(TotalCerradoArea_sqkm= cerrado_area)
area_mask_df<- area_mask_df %>% dplyr::select(-monthlyarea)
area_mask_df <- area_mask_df %>% mutate(masked_EVI_area_sqkm=unlist(masked_EVI_area_sqkm))

#Mean monthly ggplot
mean_monthly_area<- area_mask_df %>% group_by(Month) %>% summarise(mean_monthly_masked_area= mean(masked_EVI_area_sqkm))
mean_monthly_area<- mean_monthly_area %>% mutate(TotalCerradoArea_sqkm= cerrado_area)
mean_monthly_area<- mean_monthly_area %>% mutate(proportions= (mean_monthly_masked_area/TotalCerradoArea_sqkm)*100)

x_mean_monthly<-ggplot(data=mean_monthly_area, aes(x=Month, y=proportions)) +
  geom_bar(stat="identity")+
  ylim(0,100) +
  theme_minimal()+
  xlab("Month") + ylab("Percentage of total area of Cerrado included")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
ggsave(here("Outputs", "Indices","Masked_anisoEVI", "mean_monthly_studyarea_proportion.png"),x_mean_monthly, dpi=300) 

#Facet wrap by year
area_mask_df<- area_mask_df %>% mutate(proportion= (masked_EVI_area_sqkm/TotalCerradoArea_sqkm)*100)
x_annual<-ggplot(data=area_mask_df, aes(x=Month, y=proportion, group= Year)) +
  geom_bar(stat="identity")+  theme_minimal()+
  xlab("Month") + ylab("Percentage of total area of Cerrado included")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(~Year, ncol = 2)
ggsave(here("Outputs", "Indices","Masked_anisoEVI", "annual_studyarea_proportion.png"),x_annual, dpi=300) 


```

Then I determine for each pixels how much % (of 251 time steps) there is data for the pixels after masking. The pixels that have been excluded are NA
```{r}
sum_function<- function (i){
  sum(i, na.rm=T)
}
tic(); coverage_sum<- terra::app(coverage, fun= sum_function); toc()
tic(); coverage_sum<- terra::mask(coverage_sum, vect(cerrado)); toc()
tic(); coverage_df<- terra::as.data.frame(coverage_sum); toc()
pixels_with_value_plot<- ggplot(coverage_df, aes(x=lyr.1)) + #0 is pixels that have been masked 
  geom_histogram(aes(y = after_stat(count / sum(count))))+
    scale_y_continuous(labels = scales::percent)+
  xlab("Temporal anisoEVI sum")+ylab("% of total pixels")
ggsave(here("Outputs", "Indices", "Masked_anisoEVI", "masked_anisoEVI_count_histogram.png"),pixels_with_value_plot, dpi=300) 

#Map of above in terms of percentage
neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))

map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()
tmap_mode("plot")
coverage_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape( (coverage_sum/251)*100)+
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="%of time series with anisoEVI value") + 
              tm_layout (legend.position = c("left", "bottom"))
tmap_save(coverage_map, here("Outputs", "Indices","Masked_anisoEVI", "map_masked_anisoEVI_count.png"))

remove(coverage, coverage_df, coverage_map, coverage_sum, pixels_with_value_plot)

```


In order to remove any seasonality using any method (bfast or a stratight STL decomposition), I cannot have too many NAs (which was the point of the previous chunk i.e. to find how many pixels have nonNA values through the time series). Obviously I will remove all pixels that have no anisoEVI value in any of the years 

I dig into the NAs more in the next chunk. Based on conversation with Josh on 21st Feb- I will find (1) in each year, how many consecutive months have NAs for every pixel. 

```{r}
masked_evi

tic(); masked_evi2<- terra::crop(masked_evi, vect(cerrado));toc()
tic(); masked_evi2<- terra::mask(masked_evi2, vect(cerrado)); toc()


years<- seq(from=2001, to=2021, by=1)
annual_masked_evi_stack_list<- list() #making a list of spatrasters, where each spatrasters is annual with 12 months rasters in the stack
for(i in 1:length(years)){
  x<- subset(masked_evi2, grep(years[i], names(masked_evi2)))
  annual_masked_evi_stack_list[[i]]<-x
}

### TRIAL- to find highest number of consecutive NA values
#x<- c(0.80, -0.32, NA, NA, 0.60,  0.13, -0.48, -0.09,  0.69, NA,  0.78,  0.89)
#y<- is.na(x)
#max(rle(y)$lengths[rle(y)$values]) #longest consecutive NA months i.e. 2 consectuive months of NA, 3 consecutive months of NA

no_consecutive_NA<- function (x){
  y<- is.na(x)
  if (sum(y)==0){
    x<-0
  } else {
    x<- max(rle(y)$lengths[rle(y)$values])
  }
}

consecutive_NA_through_year_list<- list()
tic(); for (i in 1:length(annual_masked_evi_stack_list)){
  x<- terra::app(annual_masked_evi_stack_list[[i]], fun= no_consecutive_NA)
  y<- terra::mask(x, vect(cerrado))
  consecutive_NA_through_year_list[i]<-y
  remove(x,y)
}; toc() #1 hour

for (i in 1:length(consecutive_NA_through_year_list)){
  names(consecutive_NA_through_year_list[[i]])<- years[[i]]
  writeRaster(consecutive_NA_through_year_list[[i]],here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", paste0(years[[i]],"_throughtime.tif")))
}

############################### DO NOT RUN ABOVE ######################################

above_list <- list.files(path = here("Outputs", "Indices","Masked_anisoEVI", "ConsecutiveNA_inyear_throughmonths"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
above_list<-gtools::mixedsort(above_list) #arranging rasters chronologically
above_raster_list<-lapply(above_list, rast)
annual_consecutive_NA<- rast(above_raster_list)

#I am not interested in pixels that have NO CONSECUTIVE NAs.I am not interested in pixels that have all NAs. So reclassifying pixels with value 0 and 12 as NA and then making maps
#m<- rbind(c(0,NA), c(12, NA))
#annual_consecutive_NA<- terra::classify(annual_consecutive_NA,m)

#plot above rasters
tmap_mode("plot")

number_consecutive_map_list<- list()
tic();for (i in 1:nlyr(annual_consecutive_NA)){
  number_consecutive_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(annual_consecutive_NA[[i]])+
  tm_raster(style = 'cat',palette= get_brewer_pal("Oranges")) + 
              tm_layout (legend.position = c("left", "bottom"))
  number_consecutive_map_list[[i]]<- number_consecutive_map
}; toc()
x<-tmap_arrange(number_consecutive_map_list[[17]],number_consecutive_map_list[[18]], number_consecutive_map_list[[19]], 
                number_consecutive_map_list[[20]],number_consecutive_map_list[[21]],  nrow=3, ncol=2)
tmap_save(x, here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", "year_consecutiveNA17_21.png"), width = 1000, height = 750, dpi = 300)
#NA_month<- function(x){#in each annual stackrasters, finding which month anisoEVI data is missing and reclassifying that anisoEVI value to be the month number while remaining months are reclassified to be 0
#  na_x<- is.na(x)
#  y<- na_x*(1:length(x))
#}

#month_number_pixelvalue_list<- list()
#tic(); for (i in 1:length(annual_masked_evi_stack_list)){
#  x<- annual_masked_evi_stack_list[[i]]
#  x<- terra::app(x, fun= NA_month)
#  x<- terra::mask(x, vect(cerrado))
#  month_number_pixelvalue_list[i]<-x
#}; toc()

```


Then I look at the time series considering consecutive years but same month i.e. are there consecutive NA anisoEVI values for the month of Jan for 2 (3 or 4....) consecutive years. It is repeat of previous chunk except that the way the rasters are organized in a stack is such that the stack is a month
```{r}
masked_evi
month_colnames<- list()
tic(); for ( i in 1:length(evi_list)){
  x<- paste0(strsplit(strsplit(evi_list[i], "/")[[1]][10], "_")[[1]][6], "_", strsplit(strsplit(evi_list[i], "/")[[1]][10], "_")[[1]][7])
  year_number<- as.integer((strsplit(x, "_"))[[1]][1])
  month_number<- as.integer((strsplit(x, "_"))[[1]][2])
  if (month_number==1){
    month_colnames[[i]]<- paste0(year_number, "_Jan")
  } else if (month_number==2){
    month_colnames[[i]]<- paste0(year_number, "_Feb")
  } else if (month_number==3){
    month_colnames[[i]]<- paste0(year_number, "_Mar")
  } else if (month_number==4){
    month_colnames[[i]]<- paste0(year_number, "_Apr")
  } else if (month_number==5){
    month_colnames[[i]]<- paste0(year_number, "_May")
  } else if (month_number==6){
    month_colnames[[i]]<- paste0(year_number, "_Jun")
  } else if (month_number==7){
    month_colnames[[i]]<- paste0(year_number, "_Jul")
  } else if (month_number==8){
    month_colnames[[i]]<- paste0(year_number, "_Aug")
  } else if (month_number==9){
    month_colnames[[i]]<- paste0(year_number, "_Sep")
  } else if (month_number==10){
    month_colnames[[i]]<- paste0(year_number, "_Oct")
  } else if (month_number==11){
    month_colnames[[i]]<- paste0(year_number, "_Nov")
  } else {
    month_colnames[[i]]<- paste0(year_number, "_Dec")
  }
}; toc()
month_colnames<- unlist(month_colnames)
month_colnames<- month_colnames[11:261]
names(masked_evi)<- month_colnames

months<- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
month_masked_evi_stack_list<- list() #making a list of spatrasters, where each spatrasters is all the Jans, all the Febs.. all the Dec
for(i in 1:length(months)){
  x<- subset(masked_evi, grep(months[i], names(masked_evi)))
  month_masked_evi_stack_list[[i]]<-x
}

consecutive_NA_through_month_list<- list()
tic(); for (i in 1:length(month_masked_evi_stack_list)){
  x<- terra::app(month_masked_evi_stack_list[[i]], fun= no_consecutive_NA)
  y<- terra::mask(x, vect(cerrado))
  consecutive_NA_through_month_list[i]<-y
  remove(x,y)
}; toc() #1.5 hours

for (i in 1:length(consecutive_NA_through_month_list)){
  names(consecutive_NA_through_month_list[[i]])<- months[[i]]
  writeRaster(consecutive_NA_through_month_list[[i]],here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inmonth_throughyear", paste0(months[[i]],"_throughtime.tif")))
}

############################### DO NOT RUN ABOVE ######################################

above_list <- list.files(path = here("Outputs", "Indices","Masked_anisoEVI", "ConsecutiveNA_inmonth_throughyear"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
above_list<-gtools::mixedsort(above_list) #arranging rasters chronologically
above_raster_list<-lapply(above_list, rast)
month_consecutive_NA<- rast(above_raster_list)

#I am not interested in pixels that have NO CONSECUTIVE NAs.I am not interested in pixels that have all NAs. So reclassifying pixels with value 0 and 21 as NA and then making maps
m<- rbind(c(0,NA), c(21, NA))
month_consecutive_NA<- terra::classify(month_consecutive_NA,m)


#plot above rasters
tmap_mode("plot")

time_number_consecutive_map_list<- list()
tic();for (i in 1:nlyr(month_consecutive_NA)){
  number_consecutive_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(month_consecutive_NA[[i]])+
  tm_raster(style = 'cat',palette= get_brewer_pal("Oranges")) + 
              tm_layout (legend.position = c("left", "bottom"))
  time_number_consecutive_map_list[[i]]<- number_consecutive_map
}; toc()
x<-tmap_arrange(time_number_consecutive_map_list[[11]],time_number_consecutive_map_list[[12]])
tmap_save(x, here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inmonth_throughyear", "month_consecutiveNA11_12.png"), width = 1000, height = 750, dpi = 300)

```

