```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script, I processes different masks to finalize pixels of interest at various spatial scales. 

#MODIS LULC mask
The first mask I work on is to finalize pixels within which I analyse trends in anisoEVI. Since anisoEVI is a MODIS derived product, I use the MODIS 500m annual LULC product on GEE. And with this product, I chose the plant functional types classification (see GEE) and exclude all water, unclassified and urban areas in GEE. I exported the data as a multiband image (each band is each year of LULC) in the original MODIS sinusoidal projection and native resolution (~500m). So in this script I mask to the cerrado border and change projection to match anisoEVI crs and scale. 

```{r}
evi<- rast(here("Data", "Indices", "AnisoEVI_Cerrado_GEE", "mosaiced_anisoEVI.tif"))
tic();evi<- evi/10^5;toc() #scaling factor mentioned on the description page of the GEE asset

##changing band names to year and months
colnames<- list()
tic(); for ( i in 1:nlyr(evi)){
  x<- names(evi)[[i]]
  y<- paste0(strsplit(x, "_")[[1]][5], "_",strsplit(x, "_")[[1]][6] )
  colnames[[i]] <-y
}; toc()
colnames<- unlist(colnames)
names(evi)<- colnames

########################################## BELOW PROJECTION ANALYSES IS COMPLETE, DO NOT REPEAT
#modis_mask<- rast(here("Data", "Masks_to_define_pixelsofinterest", "MODIS_LULC_mask.tif"))
#modis_mask # all NA pixels i.e LULC excluded got exported as 0 from GEE. Also since I used a 'region' in the export to drive function, 
# there are 0 pixel values outside the cerrado boundary which need to be masked. 
# Also I am interested in using this only as a mask i.e. I do not need to know which LULC I have kept i.e do not need to keep pixel values 
# of LULC retained. SO I reclassify all 0 values to NA and all non 0 values to 1. 
#tic(); modis_mask<- terra::classify(modis_mask, cbind(0,NA), others = 1); toc()

#tic(); modis_mask_proj<- terra::project(modis_mask, evi, method = "near", mask= T, align=T); toc() #nearest neighbor becuse LULC is catergorical pixel values
#tic(); modis_mask_proj2<- terra::crop(modis_mask_proj, evi)
#writeRaster(modis_mask_proj2, here("Data", "Masks_to_define_pixelsofinterest", "proj_MODIS_LULC_mask.tif"))
########################################## ABOVE PROJECTION ANALYSES IS COMPLETE, DO NOT REPEAT
modis_mask<- rast(here("Data", "Masks_to_define_pixelsofinterest", "proj_MODIS_LULC_mask.tif"))

modis_mask_names<- seq(from = 2001, to=2022, by=1)
tic(); for ( i in 1:length(modis_mask_names)){
  x<- paste0("mask_", modis_mask_names[[i]])
  modis_mask_names[[i]] <-x
}; toc()
names(modis_mask)<- modis_mask_names

############################## Masking 2001 anisoEVI with 2001 MODIS LULC.......... till 2022
#anisoEVI is from March 2000 to Dec 2021. So to match MODIS annual product which starts at 2001, I will get rid of the 2000 anisoEVI products
evi<- evi[[11:262]]
colnames<- colnames[11:262]

#Also, I have only annual masks, so all monthly anisoEVI for 2001, 2002 (12 rasters each) need to be masked by one mask
mask_df<- as.tibble(colnames)
mask_df<- mask_df %>% separate_wider_delim(cols = value, delim = "_", names = c("Year","Month"))
mask_df<- mask_df %>% mutate(Year_Month=colnames)

evi_stack<- function (Year_Month){
  terra::subset(evi, grep(Year_Month, names (evi)))
}
tic(); mask_df<-mask_df %>% mutate(monthlyanisoEVI = map(Year_Month, evi_stack)); toc()

mask_stack<- function (Year){
  terra::subset(modis_mask, grep(Year, names (modis_mask)))
}
tic(); mask_df<-mask_df %>% mutate(mask = map(Year, mask_stack)); toc() #12 sec

masking_function<- function (anisoEVI_raster, mask_raster){
  terra::cover(mask_raster, anisoEVI_raster, values=1)
}
tic(); mask_df<-mask_df %>% mutate(masked_EVI = purrr::map2(monthlyanisoEVI, mask, masking_function)); toc() #took a long time 65183.45 seconds which is 18 hours!!


writeRaster_function<- function (masked_evi_raster, name){
  terra::writeRaster(masked_evi_raster, here("Outputs", "Indices", "Masked_anisoEVI", paste0(name, "_maskedanisoEVI.tif")))
}
tic(); mask_df<-mask_df %>% mutate(write_put = purrr::map2(masked_EVI, Year_Month, writeRaster_function)); toc() #1 min



```
------- ABOVE CHUNK IS COMPLETE AND WRITTEN OUT, START WITH NEXT CHUNK 

#EDA
Then I find the amount of area remaining i.e total area of pixels within which trends will be anlaysed. I present this as a proportion of the total cerrado area
```{r}
d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)

cerrado<- d_trans %>% st_union()
cerrado_area<- terra::expanse(vect(cerrado), unit= "km") #2040045 sqkm which is 204 Mha

masked_list <- list.files(path = here("Outputs", "Indices", "Masked_anisoEVI"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
masked_list<-gtools::mixedsort(masked_list) #arranging rasters chronologically
masked_raster_list<-lapply(masked_list, rast)
masked_evi<- rast(masked_raster_list) #252 rasters starting 2001_01
names(masked_evi)<- colnames #from line 66 (which means I need to run the previous lines 31-32 & 38-43)
tic(); coverage<- masked_evi; toc()
tic(); coverage[!is.na(coverage)]<-1;toc()# 2.3 min
tic(); area<- terra::cellSize(coverage, mask=T, lyrs=T, unit= "km"); toc() ##2 min getting area of all pixel values=1
names(area)<- colnames

tic(); area2<- terra::crop(area, vect(d_trans), ); toc()
tic(); area2<- terra::mask(area2, vect(d_trans), ); toc()
names(area2)<- colnames


area_mask_df<- as.tibble(colnames)
area_mask_df<- area_mask_df %>% separate_wider_delim(cols = value, delim = "_", names = c("Year","Month"))
area_mask_df<- area_mask_df %>% mutate(Year_Month=colnames)

area_stack<- function (Year_Month){
  terra::subset(area2, grep(Year_Month, names (area2)))
}
tic(); area_mask_df<-area_mask_df %>% mutate(monthlyarea = map(Year_Month, area_stack)); toc()

sum_area<- function (area_raster){
  sum(terra::values(area_raster), na.rm=T)
}
tic(); area_mask_df<-area_mask_df %>% mutate(masked_EVI_area_sqkm = map(monthlyarea, sum_area)); toc() #20 min
area_mask_df<- area_mask_df %>% mutate(TotalCerradoArea_sqkm= cerrado_area)
area_mask_df<- area_mask_df %>% dplyr::select(-monthlyarea)
area_mask_df <- area_mask_df %>% mutate(masked_EVI_area_sqkm=unlist(masked_EVI_area_sqkm))

#Mean monthly ggplot
mean_monthly_area<- area_mask_df %>% group_by(Month) %>% summarise(mean_monthly_masked_area= mean(masked_EVI_area_sqkm))
mean_monthly_area<- mean_monthly_area %>% mutate(TotalCerradoArea_sqkm= cerrado_area)
mean_monthly_area<- mean_monthly_area %>% mutate(proportions= (mean_monthly_masked_area/TotalCerradoArea_sqkm)*100)

x_mean_monthly<-ggplot(data=mean_monthly_area, aes(x=Month, y=proportions)) +
  geom_bar(stat="identity")+
  ylim(0,100) +
  theme_minimal()+
  xlab("Month") + ylab("Percentage of total area of Cerrado included")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
ggsave(here("Outputs", "Indices","Masked_anisoEVI", "mean_monthly_studyarea_proportion.png"),x_mean_monthly, dpi=300) 

#Facet wrap by year
area_mask_df<- area_mask_df %>% mutate(proportion= (masked_EVI_area_sqkm/TotalCerradoArea_sqkm)*100)
x_annual<-ggplot(data=area_mask_df, aes(x=Month, y=proportion, group= Year)) +
  geom_bar(stat="identity")+  theme_minimal()+
  xlab("Month") + ylab("Percentage of total area of Cerrado included")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(~Year, ncol = 2)
ggsave(here("Outputs", "Indices","Masked_anisoEVI", "annual_studyarea_proportion.png"),x_annual, dpi=300) 


```

Then I determine for each pixels how much % (of 251 time steps) there is data for the pixels after masking. The pixels that have been excluded are NA
```{r}
sum_function<- function (i){
  sum(i, na.rm=T)
}
tic(); coverage_sum<- terra::app(coverage, fun= sum_function); toc()
tic(); coverage_sum<- terra::mask(coverage_sum, vect(cerrado)); toc()
tic(); coverage_df<- terra::as.data.frame(coverage_sum); toc()
pixels_with_value_plot<- ggplot(coverage_df, aes(x=lyr.1)) + #0 is pixels that have been masked 
  geom_histogram(aes(y = after_stat(count / sum(count))))+
    scale_y_continuous(labels = scales::percent)+
  xlab("Temporal anisoEVI sum")+ylab("% of total pixels")
ggsave(here("Outputs", "Indices", "Masked_anisoEVI", "masked_anisoEVI_count_histogram.png"),pixels_with_value_plot, dpi=300) 

#Map of above in terms of percentage
neighbors<- st_read(here("Data", "Admin", "brazil_neighbors.shp"))

map_extent<- st_bbox(c(xmin=-77.59, xmax=-31.09,
                       ymin=6.42, ymax=-33.49), crs=4326) %>% st_as_sfc()
tmap_mode("plot")
coverage_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape( (coverage_sum/251)*100)+
  tm_raster(style = 'cont',palette= get_brewer_pal("YlGnBu"), title="%of time series with anisoEVI value") + 
              tm_layout (legend.position = c("left", "bottom"))
tmap_save(coverage_map, here("Outputs", "Indices","Masked_anisoEVI", "map_masked_anisoEVI_count.png"))

remove(coverage, coverage_df, coverage_map, coverage_sum, pixels_with_value_plot)

```


In order to remove any seasonality using any method (bfast or a stratight STL decomposition), I cannot have too many NAs (which was the point of the previous chunk i.e. to find how many pixels have nonNA values through the time series). Obviously I will remove all pixels that have no anisoEVI value in any of the years 

I dig into the NAs more in the next chunk. Based on conversation with Josh on 21st Feb- I will find (1) in each year, how many consecutive months have NAs for every pixel. 

```{r}
masked_evi

tic(); masked_evi2<- terra::crop(masked_evi, vect(cerrado));toc()
tic(); masked_evi2<- terra::mask(masked_evi2, vect(cerrado)); toc()


years<- seq(from=2001, to=2021, by=1)
annual_masked_evi_stack_list<- list() #making a list of spatrasters, where each spatrasters is annual with 12 months rasters in the stack
for(i in 1:length(years)){
  x<- subset(masked_evi2, grep(years[i], names(masked_evi2)))
  annual_masked_evi_stack_list[[i]]<-x
}

### TRIAL- to find highest number of consecutive NA values
#x<- c(0.80, -0.32, NA, NA, 0.60,  0.13, -0.48, -0.09,  0.69, NA,  0.78,  0.89)
#y<- is.na(x)
#max(rle(y)$lengths[rle(y)$values]) #longest consecutive NA months i.e. 2 consectuive months of NA, 3 consecutive months of NA

no_consecutive_NA<- function (x){
  y<- is.na(x)
  if (sum(y)==0){
    x<-0
  } else {
    x<- max(rle(y)$lengths[rle(y)$values])
  }
}

consecutive_NA_through_year_list<- list()
tic(); for (i in 1:length(annual_masked_evi_stack_list)){
  x<- terra::app(annual_masked_evi_stack_list[[i]], fun= no_consecutive_NA)
  y<- terra::mask(x, vect(cerrado))
  consecutive_NA_through_year_list[i]<-y
  remove(x,y)
}; toc() #1 hour

for (i in 1:length(consecutive_NA_through_year_list)){
  names(consecutive_NA_through_year_list[[i]])<- years[[i]]
  writeRaster(consecutive_NA_through_year_list[[i]],here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", paste0(years[[i]],"_throughtime.tif")))
}

############################### DO NOT RUN ABOVE ######################################

above_list <- list.files(path = here("Outputs", "Indices","Masked_anisoEVI", "ConsecutiveNA_inyear_throughmonths"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
above_list<-gtools::mixedsort(above_list) #arranging rasters chronologically
above_raster_list<-lapply(above_list, rast)
annual_consecutive_NA<- rast(above_raster_list)

#I am not interested in pixels that have NO CONSECUTIVE NAs.I am not interested in pixels that have all NAs. So reclassifying pixels with value 0 and 12 as NA and then making maps
m<- rbind(c(0,NA), c(12, NA))
annual_consecutive_NA<- terra::classify(annual_consecutive_NA,m)

#plot above rasters
tmap_mode("plot")

number_consecutive_map_list<- list()
tic();for (i in 1:nlyr(annual_consecutive_NA)){
  number_consecutive_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(annual_consecutive_NA[[i]])+
  tm_raster(style = 'cat',palette= get_brewer_pal("Oranges")) + 
              tm_layout (legend.position = c("left", "bottom"))
  number_consecutive_map_list[[i]]<- number_consecutive_map
}; toc()
x<-tmap_arrange(number_consecutive_map_list[[17]],number_consecutive_map_list[[18]], number_consecutive_map_list[[19]], 
                number_consecutive_map_list[[20]], number_consecutive_map_list[[21]],  nrow=3, ncol=2)
tmap_save(x, here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", "reclass_year_consecutiveNA17_21.png"), width = 1000, height = 750, dpi = 300)
#NA_month<- function(x){#in each annual stackrasters, finding which month anisoEVI data is missing and reclassifying that anisoEVI value to be the month number while remaining months are reclassified to be 0
#  na_x<- is.na(x)
#  y<- na_x*(1:length(x))
#}

#month_number_pixelvalue_list<- list()
#tic(); for (i in 1:length(annual_masked_evi_stack_list)){
#  x<- annual_masked_evi_stack_list[[i]]
#  x<- terra::app(x, fun= NA_month)
#  x<- terra::mask(x, vect(cerrado))
#  month_number_pixelvalue_list[i]<-x
#}; toc()

```


Then I look at the time series considering consecutive years but same month i.e. are there consecutive NA anisoEVI values for the month of Jan for 2 (3 or 4....) consecutive years. It is repeat of previous chunk except that the way the rasters are organized in a stack is such that the stack is a month
```{r}
masked_evi
month_colnames<- list()
tic(); for ( i in 1:length(masked_list)){
  x<- strsplit(masked_list[i], "/")[[1]][10]
  year_number<- as.integer((strsplit(x, "_"))[[1]][1])
  month_number<- as.integer((strsplit(x, "_"))[[1]][2])
  if (month_number==1){
    month_colnames[[i]]<- paste0(year_number, "_Jan")
  } else if (month_number==2){
    month_colnames[[i]]<- paste0(year_number, "_Feb")
  } else if (month_number==3){
    month_colnames[[i]]<- paste0(year_number, "_Mar")
  } else if (month_number==4){
    month_colnames[[i]]<- paste0(year_number, "_Apr")
  } else if (month_number==5){
    month_colnames[[i]]<- paste0(year_number, "_May")
  } else if (month_number==6){
    month_colnames[[i]]<- paste0(year_number, "_Jun")
  } else if (month_number==7){
    month_colnames[[i]]<- paste0(year_number, "_Jul")
  } else if (month_number==8){
    month_colnames[[i]]<- paste0(year_number, "_Aug")
  } else if (month_number==9){
    month_colnames[[i]]<- paste0(year_number, "_Sep")
  } else if (month_number==10){
    month_colnames[[i]]<- paste0(year_number, "_Oct")
  } else if (month_number==11){
    month_colnames[[i]]<- paste0(year_number, "_Nov")
  } else {
    month_colnames[[i]]<- paste0(year_number, "_Dec")
  }
}; toc()
month_colnames<- unlist(month_colnames)
names(masked_evi)<- month_colnames

months<- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
month_masked_evi_stack_list<- list() #making a list of spatrasters, where each spatrasters is all the Jans, all the Febs.. all the Dec
for(i in 1:length(months)){
  x<- subset(masked_evi, grep(months[i], names(masked_evi)))
  month_masked_evi_stack_list[[i]]<-x
}

consecutive_NA_through_month_list<- list()
tic(); for (i in 1:length(month_masked_evi_stack_list)){
  x<- terra::app(month_masked_evi_stack_list[[i]], fun= no_consecutive_NA)
  y<- terra::mask(x, vect(cerrado))
  consecutive_NA_through_month_list[i]<-y
  remove(x,y)
}; toc() #26 min

for (i in 1:length(consecutive_NA_through_month_list)){
  names(consecutive_NA_through_month_list[[i]])<- months[[i]]
  writeRaster(consecutive_NA_through_month_list[[i]],here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inmonth_throughyear", paste0(months[[i]],"_throughtime.tif")))
}

############################### DO NOT RUN ABOVE ######################################

above_list <- list.files(path = here("Outputs", "Indices","Masked_anisoEVI", "ConsecutiveNA_inmonth_throughyear"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
above_list<-gtools::mixedsort(above_list) #arranging rasters chronologically
above_raster_list<-lapply(above_list, rast)
month_consecutive_NA<- rast(above_raster_list)

#I am not interested in pixels that have NO CONSECUTIVE NAs.I am not interested in pixels that have all NAs. So reclassifying pixels with value 0 and 21 as NA and then making maps
m<- rbind(c(0,NA), c(21, NA))
month_consecutive_NA<- terra::classify(month_consecutive_NA,m)


#plot above rasters
tmap_mode("plot")

time_number_consecutive_map_list<- list()
tic();for (i in 1:nlyr(month_consecutive_NA)){
  number_consecutive_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(month_consecutive_NA[[i]])+
  tm_raster(style = 'cat',palette= get_brewer_pal("Oranges")) + 
              tm_layout (legend.position = c("left", "bottom"))
  time_number_consecutive_map_list[[i]]<- number_consecutive_map
}; toc()
x<-tmap_arrange(time_number_consecutive_map_list[[12]])
tmap_save(x, here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inmonth_throughyear", "reclass_month_consecutiveNA_Sep.png"), width = 1000, height = 750, dpi = 300)

remove(x_arrange, consecutiveNAmonth_sameyear_list, consecutiveNAmonth_sameyear_raster_list)
```


I still need to collapse the above NA maps (for each month through the year and all months in each year) to a format that can help me decide which pixels to keep (based on how many NA values are in the timeseries)

Considering each year and the number of NAs in all months, I collapse all the years by finding the max number oif consecutive NAs (so, if a pixel has value 3 in 2005 and 10 in 2020, then resulting pixel will have value of 10 ie max consecutive months in a year with NAs across all years)
```{r}
consecutiveNAmonth_sameyear_list<- list.files(path = here("Outputs", "Indices", "Masked_anisoEVI", "ConsecutiveNA_inyear_throughmonths"), pattern='.tif$', all.files=TRUE, full.names=TRUE)

consecutiveNAmonth_sameyear_list<-gtools::mixedsort(consecutiveNAmonth_sameyear_list) #arranging rasters chronologically
consecutiveNAmonth_sameyear_raster_list<-lapply(consecutiveNAmonth_sameyear_list, rast)
consecutiveNAmonth_sameyear<- rast(consecutiveNAmonth_sameyear_raster_list) #21 

tic(); max_consecutiveNAs<- terra::app(consecutiveNAmonth_sameyear, fun= "max"); toc()
writeRaster(max_consecutiveNAs,here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", "maxconsecutiveNA.tif"))

map_maxconsecutiveNAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(max_consecutiveNAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max consecutive NA (in each year)") + 
              tm_layout (legend.position = c("left", "bottom"))

#reclassifying such that all pixels that have all months of the year NA, those pixels are excluded completely 
allNAremove_max_consecutiveNAs<- terra::classify(max_consecutiveNAs,cbind(12,NA))
map_allNAremove_max_consecutiveNAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(allNAremove_max_consecutiveNAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max consecutive NA (in each year)") + 
              tm_layout (legend.position = c("left", "bottom"))

x_arrange<- tmap_arrange(map_maxconsecutiveNAs,map_allNAremove_max_consecutiveNAs)
tmap_save(x_arrange, here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", "map_maxconsecutiveNA.png"), width = 1000, height = 750, dpi = 300)

```


Apart from consecutive NAs across months in year, I also want to know how many NA months exist in a year (non-concecutive) and what is the highest number of NA months across years.
```{r}
#reran lines 32, 36-42, 69, 111-115, 206-207, 210-215
no_NA<- function (x){
  y<- is.na(x)
  x<- sum(y)
}

NA_through_year_list<- list()
tic(); for (i in 1:length(annual_masked_evi_stack_list)){
  x<- terra::app(annual_masked_evi_stack_list[[i]], fun= no_NA)
  y<- terra::mask(x, vect(cerrado))
  NA_through_year_list[i]<-y
  remove(x,y)
}; toc() #10min

for (i in 1:length(NA_through_year_list)){
  names(NA_through_year_list[[i]])<- years[[i]]
  writeRaster(NA_through_year_list[[i]],here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", paste0(years[[i]],"_throughtime.tif")))
}

############################### DO NOT RUN ABOVE ######################################

above_list <- list.files(path = here("Outputs", "Indices","Masked_anisoEVI", "TotalNA_inyear_throughmonths"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
above_list<-gtools::mixedsort(above_list) #arranging rasters chronologically
above_raster_list<-lapply(above_list, rast)
annual_NA<- rast(above_raster_list)

#plot above rasters
tmap_mode("plot")

NA_map_list<- list()
tic();for (i in 1:nlyr(annual_NA)){
  NA_map<-tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(annual_NA[[i]])+
  tm_raster(style = 'cat',palette= get_brewer_pal("Oranges")) + 
              tm_layout (legend.position = c("left", "bottom"))
  NA_map_list[[i]]<- NA_map
}; toc()
x<-tmap_arrange(NA_map_list[[17]],NA_map_list[[18]], NA_map_list[[19]], 
                NA_map_list[[20]],NA_map_list[[21]],  nrow=3, ncol=2)
tmap_save(x, here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "NAyear17_21.png"), width = 1000, height = 750, dpi = 300)


tic(); max_NAs<- terra::app(annual_NA, fun= "max"); toc()
writeRaster(max_NAs,here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "maxNA.tif"))

map_maxNAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(max_NAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max NA (in each year)") + 
              tm_layout (legend.position = c("left", "bottom"))

#reclassifying such that all pixels that have all months of the year NA, those pixels are excluded completely 
allNAremove_max_NAs<- terra::classify(max_NAs,cbind(12,NA))
map_allNAremove_max_NAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(allNAremove_max_NAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max NA (in each year)") + 
              tm_layout (legend.position = c("left", "bottom"))

x_arrange<- tmap_arrange(map_maxNAs,map_allNAremove_max_NAs)
tmap_save(x_arrange, here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "map_maxNA.png"), width = 1000, height = 750, dpi = 300)


x_arrange<- tmap_arrange(map_allNAremove_max_consecutiveNAs, map_allNAremove_max_NAs)
tmap_save(x_arrange, here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "comparison_consecutiveNA_totalNA.png"), width = 1000, height = 750, dpi = 300)


```

It looks like a lot of the NAs, specifically the total number of NAs in a year in a pixel, is because of the year 2001. 2001-2002 waws when Terra and Aqua of MODIS were being launched; only Terra launched in 2001, so the reflectance data was not very complete and hence the anisoEVI calcualted for 2001 might show a lot of NAs. So excluding year 2001 and calculating max consecutive and non-consecutive NAs and for each pixels 2002-2021.

```{r}
#rerun lines 378-382
consecutiveNAmonth_sameyear
consecutiveNAmonth_sameyear<- consecutiveNAmonth_sameyear[[2:21]]#excluding 1st raster from stack which is the 2001 consecutive months NA raster

tic(); no2001_max_consecutiveNAs<- terra::app(consecutiveNAmonth_sameyear, fun= "max"); toc()
writeRaster(no2021_max_consecutiveNAs,here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", "no2001_maxconsecutiveNA.tif"))

map_no2001_maxconsecutiveNAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(no2001_max_consecutiveNAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max consecutive NA (in each year, excluding 2001)") + 
              tm_layout (legend.position = c("left", "bottom"))

#reclassifying such that all pixels that have all months of the year NA, those pixels are excluded completely 
no2001_allNAremove_max_consecutiveNAs<- terra::classify(no2001_max_consecutiveNAs,cbind(12,NA))
map_no2001_allNAremove_max_consecutiveNAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(no2001_allNAremove_max_consecutiveNAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max consecutive NA (in each year, excluding 2001)") + 
              tm_layout (legend.position = c("left", "bottom"))

x_arrange<- tmap_arrange(map_no2001_maxconsecutiveNAs,map_no2001_allNAremove_max_consecutiveNAs)
tmap_save(x_arrange, here("Outputs","Indices", "Masked_anisoEVI","ConsecutiveNA_inyear_throughmonths", "map_no2001_maxconsecutiveNA.png"), width = 1000, height = 750, dpi = 300)

#rerun lines 430-433
annual_NA<- annual_NA[[2:21]]

tic(); no2001_max_NAs<- terra::app(annual_NA, fun= "max"); toc()
writeRaster(no2001_max_NAs,here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "no2001_maxNA.tif"))

map_no2001_maxNAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(no2001_max_NAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max NA (in each year, excluding 2001)") + 
              tm_layout (legend.position = c("left", "bottom"))

#reclassifying such that all pixels that have all months of the year NA, those pixels are excluded completely 
no2001_allNAremove_max_NAs<- terra::classify(no2001_max_NAs,cbind(12,NA))
map_no2001_allNAremove_max_NAs<- tm_shape(neighbors, bbox = map_extent)+ tm_borders()+
  tm_shape(d_trans)+ tm_fill()+
  tm_shape(no2001_allNAremove_max_NAs)+
  tm_raster(style = 'cat',palette= get_brewer_pal("Set3"), title="Max NA (in each year, excluding 2001)") + 
              tm_layout (legend.position = c("left", "bottom"))

x_arrange<- tmap_arrange(map_no2001_maxNAs,map_no2001_allNAremove_max_NAs)
tmap_save(x_arrange, here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "map_no2001_maxNA.png"), width = 1000, height = 750, dpi = 300)

x_arrange<- tmap_arrange(map_no2001_allNAremove_max_consecutiveNAs, map_no2001_allNAremove_max_NAs)
tmap_save(x_arrange, here("Outputs","Indices", "Masked_anisoEVI","TotalNA_inyear_throughmonths", "comparison_consecutiveNA_totalNA_no2001.png"), width = 1000, height = 750, dpi = 300)

tic();df_no2001_allNAremove_max_NAs<- as.data.frame(no2001_allNAremove_max_NAs);toc()
percent_no2001_allNAremove_max_NAs<- df_no2001_allNAremove_max_NAs %>% group_by(max) %>% summarise(count=n())
percent_no2001_allNAremove_max_NAs<- percent_no2001_allNAremove_max_NAs %>% mutate(percent=(count/sum(count))*100)
names(percent_no2001_allNAremove_max_NAs)<- c("NACount", "Count_Max","PercentMax")

tic();df_no2001_allNAremove_max_consecutiveNAs<- as.data.frame(no2001_allNAremove_max_consecutiveNAs);toc()
percent_no2001_allNAremove_consecutive_NAs<- df_no2001_allNAremove_max_consecutiveNAs %>% group_by(max) %>% summarise(count=n())
percent_no2001_allNAremove_consecutive_NAs<- percent_no2001_allNAremove_consecutive_NAs %>% mutate(percent=(count/sum(count))*100)
names(percent_no2001_allNAremove_consecutive_NAs)<- c("NACount", "Count_MaxConsecutive","PercentMaxConsecutive")

both_df<- full_join(percent_no2001_allNAremove_max_NAs,percent_no2001_allNAremove_consecutive_NAs)
both_df<- both_df %>% dplyr::select(c(NACount, PercentMax, PercentMaxConsecutive))
both_df<- pivot_longer(both_df, 2:3)

ggplot_both<- ggplot(both_df, aes(x=as.factor(NACount), y=value, fill=name)) + geom_bar(stat="identity",position=position_dodge())+
  theme_minimal() +ylab("Percentage of total pixels")+ xlab("Number of NAs")+scale_fill_manual(values=c('#999999','#E69F00'))

ggsave(here("Outputs", "Indices","Masked_anisoEVI","TotalNA_inyear_throughmonths", "comparison_consecutiveNA_totalNA_percentarea_no2001.png"),ggplot_both, dpi=300) 

```






#################### Final criteria to retain pixels of interest

#Decision 1- Exclude all pixels that have ALL anisoEVI values through time are missing/are NA
```{r}
#rerun lines 32, 36-42, 111-115

#trial
#trial<- c(NA, NA, NA, NA, NA)

numberoflayers<- nlyr(masked_evi)
exclude_NA<- function (x){
  y<- is.na(x)
  if (sum(y)==numberoflayers){ 
    x<-NA
  } else {
    x<- 1
  }
}
tic(); exclude_all_NA_pixels<- terra::app(masked_evi, fun= exclude_NA);toc()
writeRaster(exclude_all_NA_pixels, here("Outputs", "Pixels_of_interest", "allNAs_excluded_mask.tif"))

remove(numberoflayers)
tic(); excludeallNA_evi<- terra::mask(masked_evi, exclude_all_NA_pixels); toc()
```


