```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.5, tempdir = "C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//ecothresholds//Scratch", progress=10)
```

##Introduction
In this script I complete EDA of 'transition' zones across the cerrado. 

## EDA-1 Cerrado climatalogical zones intersected with the Brazil subjuridisctions
Marina's colleague has created climatalogical zones in the Cerrado by doing a clustering analyses using MAP, seasonality, DSL etc. I am trying to use the 3 resulting zones as the basis within which we see threshold changes of some X variable (%tc, NDVI etc). In this way, the 3 zones sort of control for the climate and the vegetaiton that has adapted. Colleague has sent geojson file which I convert to shp, to then upload to GEE in the chunk below
```{r}
d<- read_sf("C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//ecothresholds//Data//Cattelanetal_Clustering//regions1.geojson")
d #projection is SIRGAS 2000. I will reproject to EPSG 4326 because that is the default CRS in GEE

d_trans<- st_transform(d, crs= "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")


#st_write(d_trans, "C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//Data//Cattelanetal_Clustering//cerrado_climate_zones.shp")

```

## EDA-2 Transitions analyses 
I have processed Mapbiomas 1985-2022 and obtained 'transition' pixels from GEE, such that 
Value 1= always savanna
Value 2= switched between savanna and forest
Value 3= switched between savann and other native non forest veg
Value 4= savanna switched between all non savanna native veg
All pixels that were non-native veg i.e. belonged to other Mapbiomas classes- non-forest formation, farming, non-vegetetated areas, water were completely excluded at this stage. 

First, I calculate total area of above native vegetation types (1985-2022). The area was calculated on GEE and exported out as a csv, so I make a plot here in below chunk 

```{r}
transition_area<- read.csv("C://Users//tg505//OneDrive - University of Exeter//Ecological Resilience//ecothresholds//Outputs//Transitions//transitions_area.csv")
transition_area<- transition_area %>% dplyr::select("X1", "X2", "X3", "X4", "zone")
names(transition_area)<- c("Always savanna", "Savanna forest switch", "Savanna other native veg switch", "Multiple switches", "Zone")
transition_area_pivot<- transition_area %>% pivot_longer(1:4)
transition_area_pivot<- transition_area_pivot %>% replace(is.na(.),0)
transition_area_pivot<- transition_area_pivot %>% group_by(Zone) %>%
    mutate(percentage= (value/sum(value))*100)


transition_area_plot<-ggplot(transition_area_pivot, aes(x=Zone,y=percentage, fill=name)) +
  geom_bar(stat="identity", color= "black", position = position_dodge())+ theme_minimal()+
  theme(legend.title=element_blank())+
  xlab("Climate Zones") + ylab("Proportion of total transition\n area (%)")


transition_area_plot<-transition_area_plot + scale_fill_manual(values=c('#7fc97f','#beaed4','#fdc086','#ffff99'))
ggsave(here("Outputs", "Transitions", "overall_transitions.png"),transition_area_plot, dpi=300)
```

Second, I decided that it would be good to break down above plot into 5 year composites i.e. changes in area of the 4 types of transitions every 5 years. I extracted the area information from GEE and exported out the table csvs for each of the 5 year periods

```{r}
area_list <- list.files(path = here("Outputs", "Transitions", "fiveyear_transition_areas"), pattern='.csv$', all.files=TRUE, full.names=TRUE)
area_list<-gtools::mixedsort(area_list) #arranging
area_list_df<-lapply(area_list, read.csv)

area_function<- function(area_df){
  area_df<- area_df %>% dplyr::select("X1", "X2", "X3", "X4", "zone")
  names(area_df)<- c("Always savanna", "Savanna-forest switch", "Savanna-other native veg switch", "Multiple switches", "Zone")
  area_df_pivot<- area_df %>% pivot_longer(1:4)
  area_df_pivot<- area_df_pivot %>% replace(is.na(.),0)
  area_df_pivot<- area_df_pivot %>% group_by(Zone) %>%
    mutate(percentage= (value/sum(value))*100)
  area_df_pivot
}

tic(); area_list_for_plots<- purrr::map(area_list_df, area_function); toc()

#manually adding a column with the time period for each df because R did not read the files in order as I have named the files funnily. I get the order from area_list
area_list_for_plots[[1]]<- area_list_for_plots[[1]] %>% mutate(TimePeriod="2000-2004")
area_list_for_plots[[2]]<- area_list_for_plots[[2]] %>% mutate(TimePeriod="2005-2009")
area_list_for_plots[[3]]<- area_list_for_plots[[3]] %>% mutate(TimePeriod="2010-2014")
area_list_for_plots[[4]]<- area_list_for_plots[[4]] %>% mutate(TimePeriod="2015-2019")
area_list_for_plots[[5]]<- area_list_for_plots[[5]] %>% mutate(TimePeriod="2020-2022")
area_list_for_plots[[6]]<- area_list_for_plots[[6]] %>% mutate(TimePeriod="1985-1989")
area_list_for_plots[[7]]<- area_list_for_plots[[7]] %>% mutate(TimePeriod="1990-1994")
area_list_for_plots[[8]]<- area_list_for_plots[[8]] %>% mutate(TimePeriod="1995-1999")

area_df<- do.call("rbind", area_list_for_plots)

composite_transition_area_plot<-ggplot(area_df, aes(x=Zone,y=percentage, fill=name)) +
  geom_bar(stat="identity", color= "black", position = position_dodge())+ theme_minimal()+
  theme(legend.title=element_blank())+
  xlab("Climate Zones") + ylab("Proportion of total transition\n area (%)") + facet_wrap(~TimePeriod, ncol = 2)
composite_transition_area_plot<-composite_transition_area_plot + scale_fill_manual(values=c('#7fc97f','#beaed4','#fdc086','#ffff99'))
ggsave(here("Outputs", "Transitions", "composite_transitions.png"),composite_transition_area_plot, dpi=300)

```




Second, across each of the transition types, I determined the %tc using WRI 10m 2020 tc data.  In the peer- reviewed paper for the WRI product, there are 2 processing details mentioned-
1- they create a binary map of there is/there is no tree cover using a threshold of 35% probability
2- they take the average of the probability of tree cover when aggregating to a biger pixel,and that average pixel value is the %tc for that bigger pixel. 
I proceeded with doing (2) i.e. aggregate 10m pixel to 30m by taking the average of the 10m pixel values in GEE and exported seperate rasters for the 4 tyeps of transitions. I extract the data from each of the rasters and make the histogram
```{r}

###################################### Reading in exported rasters from GEE downloaded onto computer and mosaicing
mosaic_function<- function(input_path, output_path){
  input_list <- list.files(path= input_path , pattern = '.tif$', full.names = T)
  rasters_list <- lapply(input_list, rast)
  raster_mosaic<- do.call(terra::mosaic, rasters_list)
  terra::writeRaster(raster_mosaic, output_path)
}

Sys.time(); mosaic_function((here("Outputs", "Transitions","tc2020_savannaforest")), (here("Outputs", "Transitions", "mosaic_tc2020_savannaforest.tif"))); Sys.time()
Sys.time(); mosaic_function((here("Outputs", "Transitions","tc2020_alwayssavanna")), (here("Outputs", "Transitions", "mosaic_tc2020_alwayssavanna.tif"))); Sys.time()
Sys.time(); mosaic_function((here("Outputs", "Transitions","tc2020_savanna_nativeveg")), (here("Outputs", "Transitions", "mosaic_tc2020_savannanativeveg.tif"))); Sys.time()
Sys.time(); mosaic_function((here("Outputs", "Transitions","tc2020_multipleswitches")), (here("Outputs", "Transitions", "mosaic_tc2020_savannamultipleswitches.tif"))); Sys.time()

##### I deleted the individual tiles from the computer after mosaicing above.
```


