```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```


#Introduction 
In this script I process the anisoEVI (monthly and annual) & NDVI (monthly) from MODIS, obtained from GEE for the Caatinga. The wider goal is to check if the trajectory classification and model selection is robust. Doing the analyses for the Caatinga because then the results can be compared to Berdugo et al., 2022 PNAS paper. 

I also mask out all areas masked by Berdugo et al., in this script.

#1. Caatinga shp prep
```{r}
caatinga_border<- st_read(here("Data", "Admin", "Caatinga_border", "caatinga_border.shp"))
caatinga_border
caatinga_border_proj<- st_transform(caatinga_border, crs = 4326)
st_write(caatinga_border_proj, here("Data", "Admin", "Caatinga_border", "caatinga_border_WGSprojection.shp"))
```

#2. Mean annual MODIS based anisoEVI & NDVI and monthly anisoEVI & NDVI (prepared on GEE) prep
```{r}
meanannualanisoEVI<- rast(here("Data", "Indices", "Caatinga_meanannual_anisoEVI_2000_2021.tif")) #2000- 2021
monthly_anisoEVI<- rast(here("Data", "Indices", "Caatinga_monthly_anisoEVI.tif")) #April 2000 to Dec 2021

########################## DO NOT RUN BELOW- ALL RUN COMPLETE AND WRITTEN TO DISK
#meanannualNDVI<- rast(here("Data", "Indices", "Caatinga_meanannual_NDVI_2000_2022.tif")) #2000-2021
#monthly_NDVI<- rast(here("Data", "Indices", "Caatinga_monthly_NDVI.tif")) #Feb 2000 to Dec 2021
#changing projection of the NDVI product from default MODIS Siusoidal to EPSG 4326 ie WGS 1984
#tic(); proj_meanannualNDVI<- terra::project(meanannualNDVI, meanannualanisoEVI, method = "bilinear", align = T); toc() #5s
#names(proj_meanannualNDVI)<- names(meanannualNDVI)
#writeRaster(proj_meanannualNDVI, here("Data", "Indices", "projected_meanannualNDVI.tif"))
#tic(); proj_monthlyNDVI<- terra::project(monthly_NDVI, monthly_anisoEVI, method = "bilinear", align = T); toc() #2.5 min
#names(proj_monthlyNDVI)<- names(monthly_NDVI)
#writeRaster(proj_monthlyNDVI, here("Data", "Indices", "projected_monthlyNDVI.tif"))

#remove(meanannualNDVI, monthly_NDVI)
########################## RUN FROM BELOW- ALL RUN COMPLETE AND WRITTEN TO DISK
proj_meanannualNDVI<- rast(here("Data", "Indices", "projected_meanannualNDVI.tif"))
proj_monthlyNDVI<- rast(here("Data", "Indices", "projected_monthlyNDVI.tif"))

#changing the names of the rasters in stack for easy manipulation later on
years<- rep(seq(from= 2001, to =2021), each=12)
months<- rep(seq(from=1, to=12), 21)
month_names<- rep(c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec"), 21)

meanannualanisoEVI<- meanannualanisoEVI[[2:22]]
proj_meanannualNDVI<- proj_meanannualNDVI[[2:22]]
monthly_anisoEVI<- monthly_anisoEVI[[11:262]]
proj_monthlyNDVI<- proj_monthlyNDVI[[12:263]]

years_months<- paste0(years, "_", month_names)
names(monthly_anisoEVI)<- years_months
names(proj_monthlyNDVI)<- years_months


Sys.time(); proj_monthlyNDVI<- terra::resample(proj_monthlyNDVI, monthly_anisoEVI, threads = T); Sys.time() #half minute
Sys.time(); proj_meanannualNDVI<- terra::resample(proj_meanannualNDVI, monthly_anisoEVI, threads = T); Sys.time() #few seconds
```

Discrepancy with Berdugo when looking at NDVI specifically- Berdugo seemed to have calcualted mean NDVI from the original MODIS reflectance data (250m resolution). Specifically, they had a bunch of criteria pertaining to how many biweekly data was available in a year (MODIS is every 16 days)

#3. Preparing LULC Mask- DO NOT RUN AS OUTPUTS SAVED TO DISK
```{r}
########################################## BELOW PROJECTION ANALYSES IS COMPLETE, DO NOT REPEAT
modis_mask<- rast(here("Data", "Masks_to_define_pixelsofinterest", "Caatinga_MODIS_LULC_mask.tif")) #2001-2002
modis_mask 
#  I am interested in using this only as a mask i.e. I do not need to know which LULC 
#I have kept i.e do not need to keep pixel values 
# of LULC retained. SO I reclassify all 0 values to NA and all non 0 values to 1. 
tic(); modis_mask<- terra::classify(modis_mask, cbind(0,NA), others = 1); toc()

tic(); modis_mask_proj<- terra::project(modis_mask, meanannualanisoEVI, 
                                        method = "near", mask= T, align=T); toc() #nearest neighbor becuse LULC is catergorical pixel values
tic(); modis_mask_proj2<- terra::crop(modis_mask_proj, meanannualanisoEVI)
#writeRaster(modis_mask_proj2, here("Data", "Masks_to_define_pixelsofinterest", "Caatinga_proj_MODIS_LULC_mask.tif"))
remove(modis_mask, modis_mask_proj)

########################################## ABOVE PROJECTION ANALYSES IS COMPLETE, DO NOT REPEAT
#modis_mask_proj2<- rast(here("Data", "Masks_to_define_pixelsofinterest", "Caatinga_proj_MODIS_LULC_mask.tif"))

modis_mask_proj2<- modis_mask_proj2[[1:21]]
modis_mask_names<- seq(from = 2001, to=2021, by=1)
tic(); for ( i in 1:length(modis_mask_names)){
  x<- paste0("mask_", modis_mask_names[[i]])
  modis_mask_names[[i]] <-x
}; toc()
names(modis_mask_proj2)<- modis_mask_names

for (i in 1:nlyr(modis_mask_proj2)){
  writeRaster(modis_mask_proj2[[i]], here("Data", "Masks_to_define_pixelsofinterest","Caatinga_annual_mask" ,paste0("Caatinga_",modis_mask_names[[i]], ".tif")))
}


```


#4. Applying above prepared LULC Mask
```{r}
mask_list <- list.files(path = here("Data", "Masks_to_define_pixelsofinterest", "Caatinga_annual_mask"), 
                       pattern='.tif$', all.files=TRUE, full.names=TRUE) 

mask_rasters_list<- list()
for ( i in 1:length(mask_list)){
  x<- rast(mask_list[[i]])
  mask_rasters_list[[i]] <- x
}
modis_lulc<- rast(mask_rasters_list)
remove(x,i, mask_list, mask_rasters_list)

mask_df<- tibble(years, months)

annualEVI_annual_stack<- function (Year){
  terra::subset(meanannualanisoEVI, grep(Year, names (meanannualanisoEVI)))
}
tic(); mask_df<-mask_df %>% mutate(annual_anisoEVI = map(years, annualEVI_annual_stack)); toc()
remove(annualEVI_annual_stack)

annualNDVI_annual_stack<- function (Year){
  terra::subset(proj_meanannualNDVI, grep(Year, names (proj_meanannualNDVI)))
}
tic(); mask_df<-mask_df %>% mutate(annual_NDVI = map(years, annualNDVI_annual_stack)); toc()
remove(annualNDVI_annual_stack)

mask_df<- mask_df %>% mutate(Years_Months= paste0(years,"_", month_names))

monthlyNDVI_annual_stack<- function (Year_Months){
  terra::subset(proj_monthlyNDVI, grep(Year_Months, names (proj_monthlyNDVI)))
}
tic(); mask_df<-mask_df %>% mutate(monthly_NDVI = map(Years_Months, monthlyNDVI_annual_stack)); toc()
remove(monthlyNDVI_annual_stack)

monthlyEVI_annual_stack<- function (Year_Months){
  terra::subset(monthly_anisoEVI, grep(Year_Months, names (monthly_anisoEVI)))
}
tic(); mask_df<-mask_df %>% mutate(monthly_EVI = map(Years_Months, monthlyEVI_annual_stack)); toc()
remove(monthlyEVI_annual_stack)

modis_lulc_asign<- function(years){
  terra::subset(modis_lulc, grep(years, names (modis_lulc)))
}
tic(); mask_df<-mask_df %>% mutate(annual_LULC = map(years, modis_lulc_asign)); toc()
remove(modis_lulc_asign)
write_rds(mask_df, here("Data",  "Masks_to_define_pixelsofinterest", "caatinga_mask_df.rds"))

masking_function<- function (mask, indice_raster){
  terra::cover(mask, indice_raster, values=1)
}

tic(); mask_df<-mask_df %>% mutate(masked_annualEVI = purrr::map2(annual_anisoEVI, annual_LULC, masking_function)); toc() #took a long time 65183.45 seconds which is 18 hours!!
tic(); mask_df<-mask_df %>% mutate(masked_annualNDVI = purrr::map2(annual_NDVI, annual_LULC, masking_function)); toc()
tic(); mask_df<-mask_df %>% mutate(masked_monthlyEVI = purrr::map2(monthly_EVI, annual_LULC, masking_function)); toc()
tic(); mask_df<-mask_df %>% mutate(masked_annualEVI = purrr::map2(monthly_NDVI, annual_LULC, masking_function)); toc()


```

Discrepancy with Berdugo about LULC masking data used- Berdugo vaguely mentions that anthropogenic classes were removed and water bodies were removed using FAO LULC maps. They do not specify the classes and whether the exclusions were done annually. I googled the FAO Global LULC product and it is available in GEE included in the MODIS annual LULC product (as LC_Prop1 Class and LC_Prop2 Class), so do not know what Berdugo used. 

I am going to go ahead and use the MODIS PFT annual product for masking, which is what I have chosen for this project. This LULC cannot be very different from the FAO Global LULC, especially for urban, ag classes. 

Discrepancy with Berdugo about years- Berdugo's time period is 2000-2020. While for this analyses I am going to do 2001- 2021. I do not think that the +-1 year should make any difference to overall results.


#4. Any other data processing like STL decomposition
Berdugo et al., do not decompose the long term trend of the 20 year NDVI series. So I also do not do this for 