```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I process extract summary statistics of each index 
in each climate zone (processed in 1_indices_preparation.Rmd).
(1) Read in all data of each idex for each zone processed in 1_indices_preparation.Rmd
(2) Calculate summary statistics per zone

#Step 1- reading in all data written in 1_indices_preparation.Rmd
```{r}
ndvi_files <- list.files(path = here("Outputs", "Indices", "Climate_Zones_Indices"), pattern= "*NDVI*", all.files=TRUE, full.names=TRUE)
ndvi_list <- lapply(ndvi_files, rast)
evi_files <- Sys.glob(here("Outputs", "Indices", "Climate_Zones_Indices", "EVI_scaled_projected_*.tif"))
evi_list <- lapply(evi_files, rast)
aniso_files <- list.files(path = here("Outputs", "Indices", "Climate_Zones_Indices"), pattern= "*aniso*", all.files=TRUE, full.names=TRUE)
aniso_list <- lapply(aniso_files, rast)

remove(ndvi_files, evi_files, aniso_files)
```


#Step 2a- summary statistics 1- mean index value of all pixels through time and zones
```{r}
calc_df<- tibble(index= rep(c("NDVI", "EVI", "anisoEVI"), time=3))
calc_df<- calc_df %>% mutate(zone= rep(c("central", "southern", "eastern"), each=3))
list_rasters<- list(ndvi_list[[1]], evi_list[[1]], aniso_list[[1]],
                    ndvi_list[[2]], evi_list[[2]], aniso_list[[2]],
                    ndvi_list[[3]], evi_list[[3]], aniso_list[[3]])

calc_df<- calc_df %>% mutate(raster_stack= list_rasters)
remove(list_rasters)

mean_raster_function<- function (raster_stack){
   mean_of_raster<- terra::global(raster_stack, fun="mean", na.rm=T)
   mean_of_raster #df
}
Sys.time(); calc_df<- calc_df %>% mutate(mean_zone=purrr::map(raster_stack, mean_raster_function)); Sys.time() 

add_timestamp<- function (df){
  x<- tibble(year= rep(2000:2021, each=12))
  x<- x %>% mutate(month= rep(seq(from=1, to=12),times=22))
  x<- x %>% filter(year!=2000 | (month!=1 & month!=2))
  df<- cbind(df,x)
}
Sys.time(); calc_df<- calc_df %>% mutate(mean_zone_timestamp=purrr::map(mean_zone, add_timestamp)); Sys.time() 

mean_df<- calc_df %>% dplyr::select(-c(raster_stack, mean_zone)) %>% unnest(c(mean_zone_timestamp))
mean_df<- mean_df %>% mutate(Time= paste0(year, "_", month, "_01"))
mean_df<- mean_df %>% mutate(Time= as_date(Time))

p<- ggplot(data=mean_df, aes(x=Time, y=mean, group=index)) +
  geom_line(aes(color= index))+ 
  ylab("mean value of index") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  facet_grid(zone~.)
p<- p+scale_color_brewer( palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()
#ggsave(here("Outputs", "Indices", "Climate_Zones_Indices", "mean_indices_throughtime_zone.png"),p, 
#       dpi=700, height = 25, width = 25, units = "cm")

remove(mean_raster_function, add_timestamp)
remove( )
```

#Step 2b- summary statistics 2- seasonal mean index value through time and zones
```{r}
seasonal_df<- mean_df %>% group_by(month, index, zone) %>% summarise(mean_of_mean= mean(mean))

q<- ggplot(data=seasonal_df, aes(x=as.factor(month), y=mean_of_mean, group=index)) +
  geom_line(aes(color= index))+ 
  ylab("mean value of index") + xlab("Month")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  facet_grid(zone~.)
q<- q+scale_color_brewer( palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()
#ggsave(here("Outputs", "Indices", "Climate_Zones_Indices", "seasonal_mean_indices_throughmonths_zone.png"),q, 
#       dpi=700, height = 25, width = 25, units = "cm")

remove(q)
remove(seasonal_df, mean_df)
```