```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(RColorBrewer)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```
##Introduction
In this script, I extract the time series of all indices for ground truth points 
of different physiognomies and land uses/cover from Lewis et al., 2020 as part of
EDA
(1) Read in all rasters of each index for each zone and ground truth points from 
Lewis et al. Note Lewis et al., data was processed in ancilliary_lewisetal2020_points_processing.R
(2) Preparation of Lewis et al points for further analyses
(3) Extracting time series of all indices for all Lewis et al points 
(4) Preparation of df for plotting
(5) Plots

##Step 1- reading all files
```{r}
ndvi_files <- list.files(path = here("Outputs", "Indices", "Climate_Zones_Indices"), pattern= "*NDVI*", all.files=TRUE, full.names=TRUE)
ndvi_list <- lapply(ndvi_files, rast)
evi_files <- Sys.glob(here("Outputs", "Indices", "Climate_Zones_Indices", "EVI_scaled_projected_*.tif"))
evi_list <- lapply(evi_files, rast)
aniso_files <- list.files(path = here("Outputs", "Indices", "Climate_Zones_Indices"), pattern= "*aniso*", all.files=TRUE, full.names=TRUE)
aniso_list <- lapply(aniso_files, rast)

lewis_points<- st_read(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "prepped_points.shp")) 

remove(ndvi_files, evi_files, aniso_files)

```

##Step 2- preparation of Lewis et al data
Note- I exclude ground truth points in Lewis et al that are "water" and "nonveg". The former becuase theoretically
there should be no vegetation index values for "water" points. The latter because "nonveg" points include covers/uses
such as narrow roads which are not of interest. Lastly, the "water" points will have a respective veg index value
because the "water" points in Lewis et al are narrow streams, which have not been excluded from the veg indices 
as the water features are within the 1km resolution. 
```{r}
lewis_points <- lewis_points %>% filter(Name!="nonveg" & Name!="water")

d_trans<- st_read(here ("Data", "Cattelanetal_Clustering", "cerrado_climate_zones.shp"))
d_trans<- st_transform(d_trans, crs = 4326)
#Climate zone vector number- 1= Central; 2 = Southern; 3 = Eastern

region_lewis_points <- st_intersection(lewis_points, d_trans)
eastern_lewis_points <- region_lewis_points %>% filter(region =="Eastern")
central_lewis_points <- region_lewis_points %>% filter(region =="Central")

n_obs_name<- lewis_points %>% count(Name) %>% st_drop_geometry()

remove(lewis_points, d_trans)
```

##Step 3- extraction of time series for Lewis et al points. Note time series is from Mar 2000 to Dec 2021
```{r}
index_extraction <- function (index_central, index_eastern){
  x_points_central<- terra::extract(index_central, central_lewis_points, cells= TRUE, xy=TRUE, bind=TRUE)
  x_points_eastern<- terra::extract(index_eastern, eastern_lewis_points, cells= TRUE, xy=TRUE, bind=TRUE)
  
  df_central<- terra::as.data.frame(x_points_central)
  df_eastern<- terra::as.data.frame(x_points_eastern)
  
  extracted_df<- bind_rows(df_central, df_eastern)
  extracted_df
}

Sys.time(); ndvi <- index_extraction (ndvi_list[[1]], ndvi_list[[3]]); Sys.time()
Sys.time(); evi <- index_extraction (evi_list[[1]], evi_list[[3]]); Sys.time()
Sys.time(); aniso <- index_extraction (aniso_list[[1]], aniso_list[[3]]); Sys.time()

year<- rep(2000:2021, each=12)
month<- rep(seq(from=1, to=12),times=22)
col_names <- vector()
for (i in 1:length(year)){
  col_names[[i]] <- paste0(year[[i]], "_", month[[i]])
}
col_names<- col_names[3:264]

names(ndvi)<- c("Name", "Formation", "ID", "region", col_names, "cell", "x", "y")
names(evi)<- c("Name", "Formation", "ID", "region", col_names, "cell", "x", "y")
names(aniso)<- c("Name", "Formation", "ID", "region", col_names, "cell", "x", "y")

remove(year, month, col_names, i)
remove (index_extraction)
```


##Step 4- preparation of df for plotting
```{r}
pivot_function<- function (time_series_points){
  x_pivot <- time_series_points %>% pivot_longer(5:266)
  
  x_pivot <- x_pivot %>%  separate_wider_delim(cols = name, delim = "_", names = c("Year","Month"))
  x_pivot
}

Sys.time(); pivot_ndvi <- pivot_function (ndvi); Sys.time()
Sys.time(); pivot_evi <- pivot_function (evi); Sys.time()
Sys.time(); pivot_aniso <- pivot_function (aniso); Sys.time()

summary_function <- function (pivoted_index_df){
  summary_df <- pivoted_index_df %>% 
    group_by(Name, Year, Month) %>%
    summarise(mean_acrosspoints = mean(value, na.rm = T), sd_acrosspoints = sd(value, na.rm = T))
  summary_df
}

Sys.time(); summary_ndvi <- summary_function (pivot_ndvi); Sys.time()
names(summary_ndvi) <- c("Name", "Year", "Month", "ndvi_mean", "ndvi_sd")
Sys.time(); summary_evi <- summary_function (pivot_evi); Sys.time()
names(summary_evi) <- c("Name", "Year", "Month", "evi_mean", "evi_sd")
Sys.time(); summary_aniso <- summary_function (pivot_aniso); Sys.time()
names(summary_aniso) <- c("Name", "Year", "Month", "aniso_mean", "aniso_sd")

summary_df<- full_join(summary_ndvi, summary_evi)
summary_df<- full_join(summary_df, summary_aniso)
summary_df<- full_join(summary_df, n_obs_name, by=join_by(Name))
summary_df<- summary_df %>% mutate(Time= paste0(Year, "_", Month, "_01"))
summary_df<- summary_df %>% mutate(Time= as_date(Time))

remove(pivot_function, summary_function)
```

##Step 5a- plotting -1 Line graph comparing all indices through time by "name" in Lewis et al
```{r}
pivot_summary_df1<- summary_df %>% 
  dplyr::select(c(Name, Year, Month, Time, ndvi_mean, evi_mean, aniso_mean)) %>%
  pivot_longer(5:7)

pivot_summary_df1<- pivot_summary_df1 %>% mutate(name=case_when(name=="ndvi_mean"~"NDVI",
                                                                name=="evi_mean"~"EVI",
                                                                TRUE~"anisoEVI"))

a<-ggplot(pivot_summary_df1, aes(x=Time, y=value, group= name, color=name)) +
  geom_line() + 
  theme_classic()+ 
  theme(axis.text.x = element_blank(), axis.ticks = element_blank())+
  facet_wrap(~Name, ncol = 1, scales = "free")+ ylab("Mean index across all points") +xlab("Time in months (2002-2021)")
a<- a+scale_color_brewer( palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2")+
  theme_classic()+ theme(legend.title=element_blank()) 
ggsave(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "indices_acrosspoints_name_throughtime.png"),a, 
       dpi=700, height = 25, width = 25, units = "cm")

remove(a, pivot_summary_df1)
remove(summary_df, summary_ndvi, summary_evi, summary_aniso)
```

##Step 5b- plotting-2 Line graph comparing index for all Jans, all Febs etc by "name" in Lewis et al data with CI ribbon
```{r}
summary_ndvi<- pivot_ndvi %>% 
  group_by(Name, Month) %>%
  summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm=T)) %>% mutate(Index="NDVI")

summary_evi<- pivot_evi %>% 
  group_by(Name, Month) %>%
  summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm=T)) %>% mutate(Index="EVI")

summary_aniso<- pivot_aniso %>% 
  group_by(Name, Month) %>%
  summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm=T)) %>% mutate(Index="anisoEVI")


summary_df<- bind_rows(summary_ndvi, summary_evi, summary_aniso)
summary_df<- full_join(summary_df, n_obs_name, by=join_by(Name))
summary_df<- summary_df %>% mutate(se= sd/sqrt(n),
         lower.ci= mean- qt(1 - (0.05 / 2), n - 1) * se,
         upper.ci= mean + qt(1 - (0.05 / 2), n - 1) * se)
summary_df$Month <- as.numeric(summary_df$Month)

p<-ggplot(data = summary_df, aes(x=Month, y=mean, group=Index, colour= Index)) + 
  geom_line() + scale_color_brewer(palette = "Dark2") +
  geom_ribbon(aes(ymin = summary_df$lower.ci, ymax = summary_df$upper.ci, fill= Index), linetype=2, alpha=0.1)+
  scale_fill_brewer(palette = "Dark2") + theme_classic() + theme(legend.title = element_blank())+
  facet_wrap(~Name, scales = "fixed")
ggsave(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "indices_forallJansFebsacrossYears_byname.png"),p, 
       dpi=700, height = 25, width = 25, units = "cm")

remove(p, summary_ndvi, summary_evi, summary_aniso, summary_df)
```

##Step 5c- plotting 3- Line graph comparing anisoEVI & NDVI for all Jans, all Febs by "name" in Lewisetal data such that the data is normalized 
```{r}
summary_ndvi<- pivot_ndvi %>% 
  group_by(Name, Month) %>%
  summarise(mean = mean(value, na.rm=T)) %>% mutate(Index="NDVI")

summary_evi<- pivot_evi %>% 
  group_by(Name, Month) %>%
  summarise(mean = mean(value, na.rm=T)) %>% mutate(Index="EVI")

summary_aniso<- pivot_aniso %>% 
  group_by(Name, Month) %>%
  summarise(mean = mean(value, na.rm=T)) %>% mutate(Index="anisoEVI")

summary_df<- bind_rows(summary_ndvi, summary_evi, summary_aniso)
summary_df<- full_join(summary_df, n_obs_name, by=join_by(Name))

summary_df<- summary_df %>% 
  group_by(Name, Index) %>%
  mutate(mean_annual= mean(mean))
summary_df<- summary_df %>% 
  mutate(percentage_change= ((mean-mean_annual)/mean)*100)
summary_df$Month <- as.numeric(summary_df$Month)

s<-ggplot(data = summary_df, aes(x=Month, y=percentage_change, group=Index, colour= Index)) + 
  geom_line() + scale_color_brewer(palette = "Dark2")+
  scale_fill_brewer(palette = "Dark2") +theme_classic() +theme(legend.title = element_blank())+
  facet_wrap(~Name, scales = "fixed")
ggsave(here("Data", "Lewisetal2022_ChapadaNP_PhysiognomyMaps", "percentage_change_indices_forallJansFebsacrossYears_byname.png"),s, 
       dpi=700, height = 25, width = 25, units = "cm")
```


