```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)


library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script, I do structural equation modeling to understand causal relationships between 
environmental drives and stability at the biome scale.

##Data input
```{r}
#response- stable area from all three indices
#aniso_stable<- rast(here("Outputs", "Final_TrajShape_Models", "Annual", "agg_stability_am_11_trajresults.tif"))
ndvi_unstable<- rast(here("Outputs", "TrendsResults", "results_rasters", "monthly" , "unstablearea_5km", "monthlyswin11_ndvi_unstablearea.tif"))
evi_unstable<- rast(here("Outputs", "TrendsResults", "results_rasters", "monthly", "unstablearea_5km", "monthlyswin11_evi_unstablearea.tif"))

#drivers
anthropic_dist<- rast(here("Outputs", "OtherVariables", "AnthropicDist", "5km_meandist_20thresholdsanthropicpixel.tif"))
burnedarea<- rast(here("Outputs", "OtherVariables", "Fire", "theilsen_burnedarea_5km.tif"))
hand<- rast(here("Outputs", "OtherVariables","HAND", "cerrado_hand_5km.tif"))
re<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_re.tif"))
mat<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_at_5km.tif"))
spei<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_spei.tif"))
heatwave<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_heatwaves.tif"))
savannapercentage <- rast(here("Outputs", "OtherVariables", "Formations_Heterogeniety", "theilsen_savannapercentage.tif"))


```

##Preprocessing of all data- making appropriate data structure 

```{r}
drivers<- c(anthropic_dist, burnedarea, hand, re, mat, heatwave, spei, savannapercentage)
drivers
names(drivers)<- c("anthropic_dist", "burnedarea", "hand", "re", "mat", "heatwave", "spei", "savannapercentage")

instability<- c(ndvi_unstable, evi_unstable)
names(instability)<- c("NDVI_instability", "EVI_instability")

all_data<- c(instability, drivers)
analyses_df<- terra::as.data.frame(all_data)
remove( evi_unstable, ndvi_unstable, anthropic_dist, burnedarea, hand, re, mat, heatwave, spei, savannapercentage)
remove(drivers, instability)
```


##SEM
#EDA
First because the sample size is the population, I know that all test statistics will
be unreliable, especially statistics like chi sq that are dependent on sample size. 
Hence I sample a % of the total data. I then do EDA because SEM usually
is run on Gaussian distributed data
```{r}
analyses_df
summary(analyses_df)
analyses_df_NA <- analyses_df %>% drop_na() #Removing rows with NA values i.e. applying anthropic mask

set.seed(1000)
sample_df<- analyses_df_NA %>% slice_sample(prop=0.3, replace= FALSE) #20% randomly sampled
summary(sample_df)

#install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
chart.Correlation(sample_df, histogram = TRUE, pch= 19)

#Removing outliers considering distribution of HAND
no_outliers_sample_df <- sample_df %>% 
          mutate (IQR = IQR(hand),
                  O_upper = quantile (hand, probs= c(0.75), na.rm= FALSE)+ 1.5*IQR,
                  O_lower = quantile(hand, probs= c(0.25), na.rm= FALSE) - 1.5*IQR
                  ) %>% 
          filter(O_lower <= hand & hand <= O_upper) %>%
          dplyr::select(-c(IQR, O_upper, O_lower))

chart.Correlation(no_outliers_sample_df, histogram = TRUE, pch= 19)

no_outliers_sample_df

analyses_ndviinstability<- no_outliers_sample_df %>% dplyr::select(-c(EVI_instability))
scaled_analyses_ndviinstability<- data.frame(scale(analyses_ndviinstability, center= TRUE, scale= TRUE)) 
chart.Correlation(scaled_analyses_ndviinstability, histogram = TRUE, pch= 19)

#plot
pivot_df <- scaled_analyses_ndviinstability  %>%
  pivot_longer(1:9)
distribution_plot <- ggplot(data = pivot_df, aes(value)) +
  geom_histogram() +facet_wrap(~name, scales = "free", labeller = label_wrap_gen(4)) +
  theme_classic(base_size = 16)
#analyses_evistability<- no_outliers_sample_df %>% dplyr::select(-c(NDVI_stability))
#scaled_analyses_evistability<- data.frame(scale(analyses_evistability, center= TRUE, scale= TRUE)) 

```
All response variables of stable area by index are right skewed (because most of the Cerrado is stable).
Heatwaves# is normally distributed, while the remaining variables are not. I excluded outliers beyond the 
interquartile range for hand. After exclusion of outliers, most variables are still not normally distributed. 

#Analyses
I first consider on NDVI based stable area response variable
```{r}
scaled_analyses_ndviinstability
#scaled_analyses_evistability
#install.packages("cSEM")
library(cSEM)

comp_trial2<- '
  #Structural model
  INSTABILITY ~ BACKGROUND_CLIMATE + PULSE_STRESSORS + WATER_ACCESS + VEGETATION + ANTHROPIC_PRESSURES + FIRE
  VEGETATION  ~ PULSE_STRESSORS  + BACKGROUND_CLIMATE + ANTHROPIC_PRESSURES + FIRE
  FIRE ~ PULSE_STRESSORS + BACKGROUND_CLIMATE + ANTHROPIC_PRESSURES
  WATER_ACCESS ~ ANTHROPIC_PRESSURES
  
  #Composite variables with equal weighting of indicators
  BACKGROUND_CLIMATE <~ re + mat
  PULSE_STRESSORS <~  heatwave + spei
  
  #Measurement model (Formative)
  INSTABILITY =~ NDVI_instability
  WATER_ACCESS =~ hand
  VEGETATION =~ savannapercentage 
  FIRE =~ burnedarea
  ANTHROPIC_PRESSURES =~ anthropic_dist
'
#Sys.time(); modeltrial2<- cSEM::csem(.data =  scaled_analyses_ndvistability %>% dplyr::select(-c(forestpercentage, grasspercentage)), 
#                                     .model = comp_trial2); Sys.time()
#Sys.time(); modeltrial3<- cSEM::csem(.data =  scaled_analyses_ndvistability %>% dplyr::select(-c(forestpercentage, grasspercentage)), 
#                                     .model = comp_trial2, 
#                                     .resample_method = "bootstrap"); Sys.time()
Sys.time(); modeltrial4<- cSEM::csem(.data =  scaled_analyses_ndviinstability, 
                                     .model = comp_trial2, 
                                     .approach_weights = "PLS-PM",
                                     .resample_method = "bootstrap",
                                     .R =1000,
                                     .seed= 1000); Sys.time()
cSEM::summarize(modeltrial4)
x<-cSEM::assess(modeltrial4) #model fit statistics look good - R2 ffor structural regression equations from here
cSEM::verify(modeltrial4) # model admissability looks good
modelfit<- cSEM::testOMF(modeltrial4) #reject null hypothesis, so model fit to observed data is poor

```

VBerification of model fit ie the acceptable range of values of various model fit statistics is from 
https://www.smartpls.com/documentation/functionalities/thresholds
which is the thresholds used by SmartPLS software. 
