```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(RColorBrewer)
library(lattice)
library(tidyverse)
library(here)
library(tictoc)

library(ggplot2)
library(ggpubr)
library(sf)
library(terra)

library(tmap)
library(tmaptools)
library(viridis)

terraOptions(memfrac=0.8, tempdir = here("Scratch"), progress=10)
```

##Introduction
In this script, I do GAMs to understand the relationship between instability (from NDVI) and 
various predictor variables. 

#Step 1- data input
```{r}
#response- stable area from all three indices
#aniso_stable<- rast(here("Outputs", "Final_TrajShape_Models", "Annual", "agg_stability_am_11_trajresults.tif"))
monthly_ndvi_unstable<- rast(here("Outputs", "TrendsResults", "results_rasters", "monthly" ,"monthlyswin11_ndvi_unstablearea.tif"))
annual_ndvi_unstable<- rast(here("Outputs", "TrendsResults", "results_rasters", "annual" ,"annualswin11_evi_unstablearea.tif"))


#drivers
anthropic_dist<- rast(here("Outputs", "OtherVariables", "AnthropicDist", "5km_meandist_20thresholdsanthropicpixel.tif"))
burnedarea<- rast(here("Outputs", "OtherVariables", "Fire", "theilsen_burnedarea.tif"))
hand<- rast(here("Outputs", "OtherVariables","HAND", "cerrado_hand_5km.tif"))
re<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_re.tif"))
mat<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_at.tif"))
#spei<- rast(here("Outputs", "OtherVariables", "Climate", "annualspei_2002_2021.tif"))
heatwave<- rast(here("Outputs", "OtherVariables", "Climate", "theilsen_heatwaves.tif"))
forestpercentage <- rast(here("Outputs", "OtherVariables", "Formations_Heterogeniety", "theilsen_forestpercentage.tif"))
savannapercentage <- rast(here("Outputs", "OtherVariables", "Formations_Heterogeniety", "theilsen_savannapercentage.tif"))
grasspercentage <- rast(here("Outputs", "OtherVariables", "Formations_Heterogeniety", "theilsen_grasspercentage.tif"))


```

#Step 2- data structuring and applying anthropic threshold mask to predictors
```{r}
drivers<- c(anthropic_dist, burnedarea, hand, re, mat, heatwave, forestpercentage, savannapercentage, grasspercentage)

drivers_masked <- terra::mask(drivers, monthly_ndvi_unstable)

all_data<- c(monthly_ndvi_unstable, annual_ndvi_unstable, drivers_masked)
names(all_data)<- c("ndvi_instability", "annual_ndvi_instability" ,"anthropic_dist", "burnedarea", "hand", "re", "annualtemp", "heatwave", "forestpercentage", "savannapercentage", "grasspercentage")

plot(all_data)

analyses_df<- terra::as.data.frame(all_data, xy = TRUE)

remove(monthly_ndvi_unstable, annual_ndvi_unstable, anthropic_dist, burnedarea, hand, re, mat, heatwave, forestpercentage, savannapercentage, grasspercentage)
remove(drivers_masked)
```

#Step 3- eda - understanding the correlations and distributions of all the variables 
```{r}
analyses_df
summary(analyses_df)
analyses_df_noNA <- analyses_df %>% drop_na() #Removing rows with NA values

library(PerformanceAnalytics)
chart.Correlation(analyses_df_noNA, histogram = TRUE, method = "spearman", pch= 19)


#Removing outliers considering distribution of HAND
no_outliers_sample_df <- analyses_df_noNA %>% 
          mutate (IQR = IQR(hand),
                  O_upper = quantile (hand, probs= c(0.75), na.rm= FALSE)+ 1.5*IQR,
                  O_lower = quantile(hand, probs= c(0.25), na.rm= FALSE) - 1.5*IQR
                  ) %>% 
          filter(O_lower <= hand & hand <= O_upper) %>%
          dplyr::select(-c(IQR, O_upper, O_lower))

chart.Correlation(no_outliers_sample_df %>% dplyr::select(-c(forestpercentage, grasspercentage)), histogram = TRUE, method = "spearman",  pch= 19)

no_outliers_sample_df <- no_outliers_sample_df %>% mutate(ndvi_stability = 1- ndvi_instability)
no_outliers_sample_df2 <- no_outliers_sample_df %>% dplyr::select(-c(forestpercentage, grasspercentage, ndvi_instability))
```
Firstly, the response variable instability % area (from NDVI) is right skewed. There are significant large correlations (>0.7) between predictor variables anthropic distance and trends in annual temperature, trends in burned area% and forest area%m HAND and RE, HAND and MAT, RE and trends in heatwaves, RE and trends in grass area%, trends in annual temperature and heatwaves, trends in forest area% and grass area%. 


#Step 4- data partitioning to testing and training (OPTIONAL FOR NOW- DO NOT RUN)
```{r}
#set.seed(1234)
###1 Data partition
#fractionTraining   <- 0.80
#fractionTest       <- 0.20

###2 Compute sample sizes.
#sampleSizeTraining_all   <- floor(fractionTraining   * nrow(no_outliers_sample_df))
#sampleSizeTest_all       <- floor(fractionTest       * nrow(no_outliers_sample_df))

###3 Create the randomly-sampled indices for the dataframe. Use setdiff() to
# avoid overlapping subsets of indices.
#indicesTraining_all    <- sort(sample(seq_len(nrow(no_outliers_sample_df)), size=sampleSizeTraining_all))
#indicesTest_all <- setdiff(seq_len(nrow(no_outliers_sample_df)), indicesTraining_all)

###4 Finally, output the dataframes for training, validation and test.
#set.seed(12345)
#dfTraining_all   <- no_outliers_sample_df[indicesTraining_all, ] 
#dfTest_all       <- no_outliers_sample_df[indicesTest_all, ]

#remove(fractionTraining, fractionTest, indicesTraining_all, indicesTest_all, sampleSizeTraining_all, sampleSizeTest_all)
```


#Step 5- regression analyses
The distribution family to use is beta because my response variable is a proportion i.e.
proportion of total area of a 5x5km pixel that is unstable. However, this response variable
is skewed to the left skewed i.e. there are a lot of 5x5km pixels that are completely unstable i.e
100% of many 5x5km pixels are unstable. So it is a zero and one inflated beta family that I need to 
use. 

I could convert the instability to stability (1- instability) and then use gamma distribution
because exponential distribution is a subclass of gamma distribution (which is supported in mgcv),
however there would still be lots of 0s i.e. 0% of 5x5km pixels are stable and 0 inflated data
is not appropriate for gamma distribition

mgcv package does not have above distribution family or functionality, which is my 
understanding with trials in next chunk.
```{r}
library(mgcv)
#basic model
Sys.time(); model1 <- gam(ndvi_stability ~ s(re) +
                   s(annualtemp) +
                   s(heatwave) +
                   s(burnedarea) +
                   s(anthropic_dist) +
                   s(savannapercentage) +
                     te(x,y),
                 data = no_outliers_sample_df%>% 
                   dplyr::select(-c(forestpercentage, grasspercentage, ndvi_instability, annual_ndvi_instability)),
                 method = "REML",
                 family = betar(link = "logit")); Sys.time()

summary(model1)
plot(model1, pages = 1, all.terms = TRUE, rug = TRUE, shade = TRUE)
gam.check(model1)

#interaction between background conditions
Sys.time(); model2 <- gam(ndvi_instability ~ s(re) + s(annualtemp) + ti(re, annualtemp) +
                   s(heatwave) +
                   s(burnedarea) +
                   s(anthropic_dist) +
                   s(savannapercentage) +
                     te(x,y),
                 data = bound_no_outliers_sample_df2 %>% dplyr::select(-c(forestpercentage, grasspercentage)),
                 method = "REML",
                 family = betar(link = "logit")); Sys.time()

summary(model2)
plot(model2, pages = 1, scheme =1)

#interaction between re and other terms
Sys.time(); model3 <- gam(ndvi_instability ~ s(re) + s(annualtemp) + ti(re, annualtemp) +
                   te(heatwave, re) +
                   te(burnedarea, re) +
                   te(anthropic_dist, re) +
                   te(savannapercentage, re) +
                     te(x,y),
                 data = bound_no_outliers_sample_df2 %>% dplyr::select(-c(forestpercentage, grasspercentage)),
                 method = "REML",
                 family = betar(link = "logit")); Sys.time()

summary(model3)
plot(model3, pages = 1, scheme =1)

Sys.time(); model4 <- gam(ndvi_instability ~ s(re) + s(annualtemp) + ti(re, annualtemp) +
                   s(heatwave) + ti(heatwave, re) +
                   s(burnedarea) + ti(burnedarea, re) +
                   s(anthropic_dist) + ti(anthropic_dist, re) +
                   s(savannapercentage) + ti(savannapercentage, re) +
                     te(x,y),
                 data = bound_no_outliers_sample_df2 %>% dplyr::select(-c(forestpercentage, grasspercentage)),
                 method = "REML", 
                 family = betar(link = "logit")); Sys.time()

summary(model4)
plot(model4, pages = 1, scheme =1)

Sys.time(); model5 <- gam(ndvi_instability ~ s(re, bs= "cs") + s(annualtemp, bs= "cs") + ti(re, annualtemp, bs= "cs") +
                   s(heatwave, bs= "cs") + ti(heatwave, re, bs= "cs") +
                   s(burnedarea, bs= "cs") + ti(burnedarea, re, bs= "cs") +
                   s(anthropic_dist, bs= "cs") + ti(anthropic_dist, re, bs= "cs") +
                   s(savannapercentage, bs= "cs") + ti(savannapercentage, re, bs= "cs") +
                     te(x,y, bs= "cs"),
                 data = bound_no_outliers_sample_df2 %>% dplyr::select(-c(forestpercentage, grasspercentage)),
                 method = "REML", select = TRUE, 
                 family = betar(link = "logit")); Sys.time()

summary(model5)
plot(model5, pages = 1, scheme =1)




```

I choose to use gamlss r package to do regression modelling. However, this type
of regression modellign does not allow interactions between predictor variables. I need
to be able to do interactions considering x,y and also considering re. 

So I settled on doing bayesian style GAM using the brms pacakge. 


```{r}
library(brms)

priors <- c(
  set_prior("student_t(3, 0, 2.5)", class = "Intercept") +
  set_prior("normal(0, 1)", class = "b") +
  set_prior("logistic(0, 1", class = "Intercept", dpar = "zi")
)

#without priors
Sys.time(); model1 <- brm (bf(ndvi_stability ~ s(re) +
                   s(annualtemp) +
                   s(heatwave) +
                   s(burnedarea) +
                   s(anthropic_dist) +
                   s(savannapercentage) +
                     t2(x,y)),
                   phi ~ 1,
                   zi ~ 1,
               data = no_outliers_sample_df2,
               family = zero_inflated_beta(),
               cores = 4, seed = 17, iter = 2000, warmup = 1000, thin = 10 ); Sys.time()
  
  
  

m2 <- brm(bf(accel ~ s(times)),
          data = mcycle, family = gaussian(), cores = 4, seed = 17,
          iter = 4000, warmup = 1000, thin = 10, refresh = 0,
          control = list(adapt_delta = 0.99))



```
